### 2015-09-11 docs/paper/out/frontiers.tex
set up LaTeX infrastructure for the Frontiers submission
There is now an `out.sh` script which (assuming you have
LaTeX and pandoc installed) will generate a  nice pdf in
the `out` folder. Bibliography is still not configured.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.1 Generated 2015/22/05 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square}
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\linenumbers

% fix for pandoc 1.14
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Alexey Alekhin {et~al.}} %use et al only if is more than 1 author
\def\Authors{
$for(authors)$ $authors.name$\,$$^{1}$$ $endfor$ and $correspondingAuthor.name$ $$^{1,*}$$
}% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$$^{1}$$Oh no sequences! Research Group, Era7 Bioinformatics, Granada, Spain}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Corresponding Author}
\def\corrAddress{Oh no sequences! Research Group, Era7 Bioinformatics, Plaza Campo Verde 3, Granada, 18001, Spain}
\def\corrEmail{eparejatobes@ohnosequences.com}

\begin{document}
\onecolumn
\firstpage{1}

\title[MG7: Configurable and scalable 16S metagenomics data analysis]{$title$}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

$abstract$

\tiny
 \keyFont{ \section{Keywords:} $keywords$ } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

$body$

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{refs.bib}

\end{document}

### 2015-09-11 docs/paper/out/paper.tex
set up LaTeX infrastructure for the Frontiers submission
There is now an `out.sh` script which (assuming you have
LaTeX and pandoc installed) will generate a  nice pdf in
the `out` folder. Bibliography is still not configured.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.1 Generated 2015/22/05 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square}
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\linenumbers

% fix for pandoc 1.14
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Alexey Alekhin {et~al.}} %use et al only if is more than 1 author
\def\Authors{
 Alexey Alekhin\,$^{1}$  Evdokim Kovach\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
}% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Oh no sequences! Research Group, Era7 Bioinformatics, Granada, Spain}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Corresponding Author}
\def\corrAddress{Oh no sequences! Research Group, Era7 Bioinformatics, Plaza Campo Verde 3, Granada, 18001, Spain}
\def\corrEmail{eparejatobes@ohnosequences.com}

\begin{document}
\onecolumn
\firstpage{1}

\title[MG7: Configurable and scalable 16S metagenomics data analysis]{MG7: Configurable and scalable 16S metagenomics data analysis -- new
methods optimized for massive cloud computing}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

No abstract yet. Will be here.

\tiny
 \keyFont{ \section{Keywords:} Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
Cloud computing, NGS, Genomic big data } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{1. Introduction}\label{introduction}

Metagenomics data analysis is growing at exponential rate during the
last years. The increasing throughput of massively parallel sequencing
technologies, the derived decreasing cost, and the high impact of
metagenomics studies, especially in human health (diagnostics,
treatments, drug response, prevention), are crucial reasons responsible
for this growth of Metagenomics. There is a growing interest in
sequencing all kind of microbiomes (gut, mouth, skin, urinary tract,
airway, milk, bladder), in different conditions of health and disease,
or after different treatments. Metagenomics is also impacting
environmental sciences, crop sciences, agrifood sector and biotechnology
in general. This new possibilities for exploring the diversity of
micro-organisms in the most diverse environments is opening many new
research areas but, due to this wide interest, it is expected that the
amount of data will be overwhelming in the short time
{[}@Stephens-2015{]}.

Genome researchers have raised the alarm over big data in the past
\href{http://www.nature.com/news/genome-researchers-raise-alarm-over-big-data-1.17912}{nature
news add ref} but even a more serious challenge might be faced with the
metagenomics boom/ upswing. If we compare metagenomics data with other
genomics data used in clinical genotyping we find a differential
feature: the key role of time. Thus, for example, in some longitudinal
studies, serial sampling of the same patient along several weeks (or
years) is being used for the follow up of some intestinal pathologies,
for studying the evolution of gut microbiome after antibiotic treatment,
or for colon cancer early detection {[}@Zeller-2014{]}. This need of
sampling across time adds more complexity to metagenomics data storage
and demands adapted algorithms to detect state variations across time as
well as idiosyncratic commonalities of the microbiome of each individual
{[}@Franzosa-2015{]}. In addition to the intra-individual sampling-time
dependence, metagenomic clinical test results vary depending on the
specific region of extraction of the clinical specimen. This local
variability adds complexity to the analysis since different
localizations (different tissues, different anatomical regions, healthy
or tumour tissues) are required to have a sufficiently complete
landscape of the human microbiome. Moreover, reanalysis of old samples
using new tools and better reference databases might be also demanded
from time to time.

During the last years other sciences as astronomy or particle physics
are facing the big data challenge but, at least, these science have
standards for data processing {[}@Stephens-2015{]}. Global standards for
converting raw sequence data into processed data are not yet well
defined in metagenomics and there are shortcomings derived from the fact
that many bioinformatics methodologies currently used for metagenomics
data analysis were designed for a scenario very different that the
current one. These are some of the aspects that have suffered crucial
changes and advances with a direct impact in metagenomics data analysis.
i. The first aspect is related to the sequences to be analyzed: the
reads are larger, the sequencing depth and the number of samples of each
project are considerably bigger. The first metagenomics studies were
very local projects, while nowadays the most fruitful studies are done
at a global level (international, continental, national). This kind of
global studies has yielded the discovery of clinical biomarkers for
diseases of the importance of cancer, obesity or inflammatory bowel
diseases and has allowed exploring the biodiversity in many earth
environments ii. The second aspect derives from the impressive genomics
explosion, its effect being felt in this case in the reference
sequences. The immense amount of sequences available in public
repositories demands new approaches in curation, update and storage for
metagenomics reference databases: current models will or already have
problems to face the future avalanche of metagenomic sequences. iii. The
third aspect to consider for metagenomics data analysis is related to
the appearance of new models for massive computation and storage and to
the new programming methodologies (Scala, \ldots{}) and new cloud models
and resources. The immense new possibilities that these advances offer
must have a direct impact in the metagenomics data analysis. iv. And
finally the new social manner to do science, and especially genomic
science is the fourth aspect to consider. Metagenomics evolves in a
social and global scenario following a science democratization trend in
which many small research groups from distant countries share a common
big metagenomics project. This global cooperation demands systems
allowing following exactly the same pipelines using equivalent cloud
resources to modularly execute the analysis in an asynchronous way of
working between different groups. This definitively new scenario demands
new methods and tools to handle the current and future volume of
metagenomic data with the sufficient speed of analysis. Considering all
these aspects we have designed a new open source methodology for
analyzing metagenomics data that exploits the new possibilities that
cloud computing offers to get a system robust, programmatically
configurable, modular, distributed, flexible, scalable and traceable in
which the biological databases of reference sequences can be easily
updated and/or frequently substituted by new ones or by databases
specifically designed for focused projects.

\section{2. Materials and Methods}\label{materials-and-methods}

\subsection{2.x Amazon Web Services}\label{x-amazon-web-services}

\subsection{2.x Scala}\label{x-scala}

\href{http://www.scala-lang.org/}{Scala} is a hybrid object-functional
programming language which runs on Java Virtual Machine. It has support
for type-level programming, type-dependent types (through type members)
and singleton types, which permits a restricted form of dependent types
where types can depend essentially on values determined at compile time
(through their corresponding singleton types). Conversely, through
implicits one can retrieve the value corresponding to a singleton type.

The other key feature for us is Java interoperability, which let us
build on the vast number of existing Java libraries; we take advantage
of this when using Bio4j as an API for the NCBI taxonomy.

MG7 itself and all the libraries used are written in Scala
\texttt{2.11}.

\subsection{2.x Statika}\label{x-statika}

\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
library developed by \textbf{so and so} which serves as a way of
defining and composing machine behaviors statically. The main component
are \textbf{bundles}. Each bundle declares a sequence of computations
(its behavior) which will be executed in an \textbf{environment}. A
bundle can \emph{depend} on other bundles, and when being executed by an
environment, its DAG of dependencies is linearized and run in sequence.
In our use, bundles correspond to what an EC2 instance should do and an
environment to an image (AMI: \textbf{A}mazon \textbf{M}achine
\textbf{I}mage) which prepares the basic configuration, downloads the
Scala code and runs it.

\subsection{2.x Datasets}\label{x-datasets}

\href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
library developed by \textbf{so and so} to declare datasets and their
locations. \textbf{Data} is represented as type-indexed fields: Keys are
modeled as singleton types, and values correspond to what could be
called a denotation of the key: a value of type \texttt{Location} tagged
with the key type. Then a \textbf{Dataset} is essentially a collection
of data, which are guaranteed statically to be different through
type-level predicates, making use of the value ↔ type correspondence
which can be established through singleton types and implicits. A
dataset location is then just a list of locations formed by locations of
each data member of that dataset.

Data keys can further have a reference to a \textbf{data type}, which,
as the name hints at, can help in providing information about the type
of data we are working with. For example, when declaring Illumina reads
as a data, a data type containing information about the read length,
insert size or end type (single or paired) is used.

A \textbf{location} can be, for example, an S3 object or a local file;
by leaving the location type used to denote particular data free we can
work with different ``physical'' representations, while keeping track of
to which logical data they are a representation of. Thus, a process can
generate locally a \texttt{.fastq} file representing the merged reads,
while another can put it in S3 with the fact that they all correspond to
the ``same'' merged reads is always present, as the data that those
``physical'' representations denote.

\subsection{2.x Loquat}\label{x-loquat}

\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
developed by \textbf{so and so} designed for the execution of
embarrassingly parallel tasks using S3, SQS and EC2.

A \textbf{loquat} executes a process with explicit input and output
datasets (declared using the \emph{Datasets} library described above).
Workers (EC2 instances) read from an SQS queue the S3 locations for both
input and output data; then they download the input to local files, and
pass these file locations to the process to be executed. The output is
then put in the corresponding S3 locations.

A manager instance is used to monitor workers, provide initial data to
be put in the SQS queue and optionally release resources depending on a
set of configurable conditions.

Both worker and manager instances are Statika bundles. In the case of
the worker, it can declare any dependencies needed to perform its task:
other tools, libraries, or data.

All configuration such as the number of workers or the instance types is
declared statically, the specification of a loquat being ultimately a
Scala object. There are deploy and resource management methods, making
it easy to use an existing loquat either as a library or from (for
example) a Scala REPL.

The input and output (and their locations) being defined statically has
several critical advantages. First, composing different loquats is easy
and safe; just use the output types and locations of the first one as
input for the second one. Second, data and their types help in not
mixing different resources when implementing a process, while serving as
a safe and convenient mechanism for writing generic processing tasks.
For example, merging paired-end Illumina reads generically is easy as
the data type includes the relevant information (insert size, read
length, etc) to pass to a tool such as FLASH.

\subsection{2.x Type-safe DSLs for BLAST and
FLASH}\label{x-type-safe-dsls-for-blast-and-flash}

We developed our own type-safe DSLs (Domain Specific Language) for
\href{https://github.com/ohnosequences/flash}{FLASH} and
\href{https://github.com/ohnosequences/blast}{BLAST} expressions and
their execution.

\subsubsection{2.x.a BLAST DSL}\label{x.a-blast-dsl}

In the case of BLAST we use a model for expressions where we can
guarantee for each BLAST command expression at compile time

\begin{itemize}
\tightlist
\item
  all required arguments are provided
\item
  only valid options are provided
\item
  correct types for each option value
\item
  valid output record specification
\end{itemize}

Generic type-safe parsers returning an heterogeneous record of BLAST
output fields are also available, together with output data defined
using \emph{Datasets} which have a reference to the exact BLAST command
options which yielded that output. This let us provide generic parsers
for BLAST output which are guaranteed to be correct, for example.

\subsubsection{2.x.b FLASH DSL}\label{x.b-flash-dsl}

In the same spirit as for BLAST,

\subsection{2.x Bio4j}\label{x-bio4j}

\href{https://github.com/bio4j/bio4j}{Bio4j} is a data platform
integrating data from different resources such as UniProt or GO in a
graph data paradigm. We use the module containing the NCBI Taxonomy, and
the use their Java API from Scala in the assignment phase.

\section{3. Results}\label{results}

\subsection{3.1 Overview}\label{overview}

To tackle the challenges posed by metagenomics big data analysis
outlined in the Introduction,

­- AWS resources in Scala (??) - A new approach to data analysis
specification, management and specification based on working with it in
exactly the same way as for a software project, together with the
extensive use of compile-time structures and checks. - Parallelization
and distributed analysis based on AWS, with on-demand infrastructure as
the basic paradigm ­- fully automated processes, data and cloud
resources management. - Static reproducible specification of
dependencies and behavior of the different components using
\emph{Statika} and \emph{Datasets} - Definition of complex pipelines
using \emph{Loquat} a composable system for scaling/parallelizing
stateless computations especially designed for Amazon Web Services (AWS)
- Modeling of the taxonomy tree using the new paradigm of graph
databases (Bio4j). It facilitates the taxonomic assignment tasks and the
calculation of the taxa abundance values considering the hierarchic
structure of taxonomy tree (cumulative values). - per-read assignment
(??)

\subsection{3.x 16S Reference Database
Construction}\label{x-16s-reference-database-construction}

Our 16S Reference Database is a curated subset of sequences from NCBI
nucleotide database \textbf{nt}. This subset of 16S sequences was
selected by similarity with the bacterial and archaeal reference
sequences downloaded from RDP database {[}@Cole-2014{]}. RDP unaligned
sequences were used to capture new16S sequences from nt using BLAST
similarity strategies and, then, performing additional curation steps to
remove sequences with poor taxonomic assignments to taxonomic nodes
close to the root of the taxonomic tree. All the nucleotide sequences
included in nt database has a taxonomic assignment provided by the
genbank sequence submitter. NCBI provides a table (available at
ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
based on a submitter-maintained taxonomic annotation system for
reference sequences that supposes a sustainable system able to face the
expected number of reference sequences that will populate the public
global nucleotide databases in the near future. Another advantageous
point is that we are based on NCBI taxonomy, the de facto standard
taxonomic classification for biomolecular data {[}@Cochrane-2010{]}.
NCBI taxonomy is, without any doubt, the most used taxonomy all over the
world and the most similar to the official taxonomies of each specific
field. This is a crucial point because all the type-culture and tissue
databanks follow this official taxonomical classification and, in
addition, all the knowledge accumulated is referred to this taxonomy. In
addition NCBI provides a direct connection between taxonomical formal
names and the physical specimens that serve as exemplars for the species
{[}@Federhen-2015{]}.

If metagenomics results are easily integrated with the theoretical and
experimental knowledge of each specific area, the impact of metagenomics
will be higher that if metagenomics progress in a disconnected research
branch. This strategy for building our database allows substituting the
16S database by any other subset of nt, even by the complete nt database
if it would needed, for example, for analyzing shotgun metagenomics
data.

\subsection{3.x Bio4j and Graph
Databases}\label{x-bio4j-and-graph-databases}

\subsection{3.x MG7 Pipeline
Description}\label{x-mg7-pipeline-description}

\subsection{3.x Taxonomic Assignment
Algorithms}\label{x-taxonomic-assignment-algorithms}

\subsubsection{3.x.y Lowest Common Ancestor based Taxonomic
Assignment}\label{x.y-lowest-common-ancestor-based-taxonomic-assignment}

For each read:

­1. Select only one BLASTN alignment (HSP) per reference sequence (the
HSP with lowest e value) 2. Filter all the HSPs with bitscore below a
defined BLASTN bitscore threshold s\_0 3. Find the best bitscore value S
in the set of BLASTN HSPs corresponding to hits of that read 4. Filter
all the alignments with bitscore below p * S (where p is a fixed by the
user coefficient to define the bitscore required, e.g.~if p=0.9 and
S=700 the required bitscore threshold would be 630) 5. Select all the
taxonomic nodes to which map the reference sequences involved in the
selected HSPs: - If all the selected taxonomic nodes forms a line in the
taxonomy tree (are located in a not branched lineage to the tree root)
we should choose the most specific taxID as the final assignment for
that read - If not, we should search for the (sensu stricto) Lowest
Common Ancestor (LCA) of all the selected taxonomic nodes (See Figure X)

In this approach the value used for evaluating the similarity is the
bitscore that is a value that increases when similarity is higher and
depends a lot on the length of the HSP

\subsubsection{3.x.z Best BLAST hit taxonomic
assignment}\label{x.z-best-blast-hit-taxonomic-assignment}

We have maintained the simpler method of Best BLAST Hit (BBH) taxonomic
assignment because, in some cases, it can provide information about the
sequences that can be more useful than the obtained using LCA algorithm.
Using LCA algorithm when some reference sequences with BLAST alignments
over the required thresholds map to a not sufficiently specific taxID,
the read can be assigned to an unspecific taxon near to the root. If the
BBH reference sequence maps to a more specific taxa this method, in that
case, gives us useful information.

\subsection{3.x Other}\label{x-other}

General approach. An analysis is defined as a software project. It can
evolve in the same way. We can run the analysis in a test phase, review
configuration and changes, etc. Key advantages of this approach are

\begin{itemize}
\tightlist
\item
  \textbf{Reproducibility} the same analysis can be run again with
  exactly the same configuration in a trivial way.
\item
  \textbf{Versioning} The analysis is a software project so it goes
  through the same stages, there can be different versions, stable
  releases, etc.
\item
  \textbf{Reuse} we can build standard configurations on top of this and
  reuse them for subsequent data analysis.
\item
  \textbf{Decoupling} We can start working on the analysis
  specification, without any need for data in a much easier way.
\item
  \textbf{Expresiveness and safety} choose only from valid Illumina read
  types, build default FLASH command based on that, \ldots{}
\end{itemize}

\subsection{3.x Using MG7 with some example
data-sets}\label{x-using-mg7-with-some-example-data-sets}

We selected the datasets described in {[}Kennedy-2014{]} (??)

\subsection{3.7 MG7 availability}\label{mg7-availability}

MG7 is open source, available at https://github.com/ohnosequences/mg7
under an \href{http://www.gnu.org/licenses/agpl-3.0.en.html}{AGPLv3}
license.

\section{4. Discussion}\label{discussion}

\subsection{4.1 Novelty points of MG7}\label{novelty-points-of-mg7}

The most innovative ideas and developments integrated in MG7 are:

­- The management dependencies checking their correctness before
compilation using Scala type system - The automation of cloud resources
and processes (parallelization management) - The cloud-oriented
development of the system including a modeling AWS resources based on
the powerful data typing of Scala - The use of the Graph databases
paradigm to store and manage the taxonomy tree to obtain the taxonomic
assignments and the cumulative frequencies - MG7 provides a sustainable
model for updating the database of reference sequences appropriate to
face the challenging amount of sequences that are generating the new
high throughput technologies of sequencing

\subsection{4.2 Designed for future
challenges}\label{designed-for-future-challenges}

Other possible uses of the general schema: statika, loquat, \ldots{}

\subsection{4.3 MG7 Future developments}\label{mg7-future-developments}

\subsubsection{4.3.1 Comparison of groups of
samples}\label{comparison-of-groups-of-samples}

\subsubsection{4.3.2 Interactive visualizations using the output files
of MG7 (Biographika
project)}\label{interactive-visualizations-using-the-output-files-of-mg7-biographika-project}

\section{5 Acknowledgements}\label{acknowledgements}

INTERCROSSING (Grant 289974)

\section{6 References}\label{references}

\section{7 Tables and Figures}\label{tables-and-figures}

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{refs.bib}

\end{document}

### 2015-09-11 docs/paper/out/out.sh

@@ -1,0 +1,6 @@
+#!/bin/sh
+set -x
+ME=`dirname $0`
+pandoc --template=${ME}/frontiers.tex ${ME}/../paper.md --smart -s -o ${ME}/paper.tex
+pdflatex ${ME}/paper.tex
+pdflatex ${ME}/paper.tex

### 2015-09-12 docs/paper/out/out.sh
fix scripts for generating the paper pdf

@@ -2,5 +2,6 @@
 set -x
 ME=`dirname $0`
 pandoc --template=${ME}/frontiers.tex ${ME}/../paper.md --smart -s -o ${ME}/paper.tex
-pdflatex ${ME}/paper.tex
-pdflatex ${ME}/paper.tex
+cd ${ME}
+pdflatex -halt-on-error paper.tex
+pdflatex -halt-on-error paper.tex

### 2015-09-12 docs/paper/out/watch.sh

@@ -1,0 +1,4 @@
+#!/bin/sh
+set -x
+ME=`dirname $0`
+ls ${ME}/../paper.md ${ME}/../refs.bib | entr -c ${ME}/out.sh

### 2015-09-12 docs/paper/out/paper.tex
generate pdf, refs working -- paper

@@
 amount of data will be overwhelming in the short time
-{[}@Stephens-2015{]}.
+\citep{Stephens-2015}.
 
@@
 for studying the evolution of gut microbiome after antibiotic treatment,
-or for colon cancer early detection {[}@Zeller-2014{]}. This need of
-sampling across time adds more complexity to metagenomics data storage
-and demands adapted algorithms to detect state variations across time as
-well as idiosyncratic commonalities of the microbiome of each individual
-{[}@Franzosa-2015{]}. In addition to the intra-individual sampling-time
-dependence, metagenomic clinical test results vary depending on the
-specific region of extraction of the clinical specimen. This local
-variability adds complexity to the analysis since different
+or for colon cancer early detection \citep{zeller2014potential}. This
+need of sampling across time adds more complexity to metagenomics data
+storage and demands adapted algorithms to detect state variations across
+time as well as idiosyncratic commonalities of the microbiome of each
+individual \citep{Franzosa-2015}. In addition to the intra-individual
+sampling-time dependence, metagenomic clinical test results vary
+depending on the specific region of extraction of the clinical specimen.
+This local variability adds complexity to the analysis since different
 localizations (different tissues, different anatomical regions, healthy
@@
 are facing the big data challenge but, at least, these science have
-standards for data processing {[}@Stephens-2015{]}. Global standards for
-converting raw sequence data into processed data are not yet well
+standards for data processing \citep{Stephens-2015}. Global standards
+for converting raw sequence data into processed data are not yet well
 defined in metagenomics and there are shortcomings derived from the fact
@@
 selected by similarity with the bacterial and archaeal reference
-sequences downloaded from RDP database {[}@Cole-2014{]}. RDP unaligned
+sequences downloaded from RDP database \citep{Cole-2014}. RDP unaligned
 sequences were used to capture new16S sequences from nt using BLAST
@@
 point is that we are based on NCBI taxonomy, the de facto standard
-taxonomic classification for biomolecular data {[}@Cochrane-2010{]}.
+taxonomic classification for biomolecular data \citep{Cochrane-2010}.
 NCBI taxonomy is, without any doubt, the most used taxonomy all over the
@@
 names and the physical specimens that serve as exemplars for the species
-{[}@Federhen-2015{]}.
+\citep{Federhen-2015}.
 
@@
 
-\section{7 Tables and Figures}\label{tables-and-figures}
-
 \bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
 %\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
-\bibliography{refs.bib}
+\bibliography{refs}
 

### 2015-09-12 docs/paper/out/out.sh

@@ -1,7 +1,8 @@
 #!/bin/sh
 set -x
 ME=`dirname $0`
-pandoc --template=${ME}/frontiers.tex ${ME}/../paper.md --smart -s -o ${ME}/paper.tex
+pandoc --natbib --template=${ME}/frontiers.tex ${ME}/../paper.md --smart -s -o ${ME}/paper.tex
 cd ${ME}
 pdflatex -halt-on-error paper.tex
+bibtex paper
 pdflatex -halt-on-error paper.tex

### 2015-09-12 docs/paper/out/watch.sh

@@ -1,4 +1,4 @@
 #!/bin/sh
 set -x
 ME=`dirname $0`
-ls ${ME}/../paper.md ${ME}/../refs.bib | entr -c ${ME}/out.sh
+ls ${ME}/../paper.md ${ME}/refs.bib | entr -c ${ME}/out.sh

### 2015-09-12 docs/paper/out/paper.tex
fix citations and generate pdf -- paper

@@
 amount of data will be overwhelming in the short time
-\citep{Stephens-2015}.
+\citep{stephens2015big}.
 
@@
 time as well as idiosyncratic commonalities of the microbiome of each
-individual \citep{Franzosa-2015}. In addition to the intra-individual
-sampling-time dependence, metagenomic clinical test results vary
-depending on the specific region of extraction of the clinical specimen.
-This local variability adds complexity to the analysis since different
-localizations (different tissues, different anatomical regions, healthy
-or tumour tissues) are required to have a sufficiently complete
-landscape of the human microbiome. Moreover, reanalysis of old samples
-using new tools and better reference databases might be also demanded
-from time to time.
+individual \citep{franzosa2015identifying}. In addition to the
+intra-individual sampling-time dependence, metagenomic clinical test
+results vary depending on the specific region of extraction of the
+clinical specimen. This local variability adds complexity to the
+analysis since different localizations (different tissues, different
+anatomical regions, healthy or tumour tissues) are required to have a
+sufficiently complete landscape of the human microbiome. Moreover,
+reanalysis of old samples using new tools and better reference databases
+might be also demanded from time to time.
 
@@
 are facing the big data challenge but, at least, these science have
-standards for data processing \citep{Stephens-2015}. Global standards
+standards for data processing \citep{stephens2015big}. Global standards
 for converting raw sequence data into processed data are not yet well
@@
 selected by similarity with the bacterial and archaeal reference
-sequences downloaded from RDP database \citep{Cole-2014}. RDP unaligned
-sequences were used to capture new16S sequences from nt using BLAST
-similarity strategies and, then, performing additional curation steps to
-remove sequences with poor taxonomic assignments to taxonomic nodes
-close to the root of the taxonomic tree. All the nucleotide sequences
-included in nt database has a taxonomic assignment provided by the
-genbank sequence submitter. NCBI provides a table (available at
+sequences downloaded from RDP database \citep{cole2013ribosomal}. RDP
+unaligned sequences were used to capture new16S sequences from nt using
+BLAST similarity strategies and, then, performing additional curation
+steps to remove sequences with poor taxonomic assignments to taxonomic
+nodes close to the root of the taxonomic tree. All the nucleotide
+sequences included in nt database has a taxonomic assignment provided by
+the genbank sequence submitter. NCBI provides a table (available at
 ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
@@
 point is that we are based on NCBI taxonomy, the de facto standard
-taxonomic classification for biomolecular data \citep{Cochrane-2010}.
+taxonomic classification for biomolecular data \citep{cochrane20102010}.
 NCBI taxonomy is, without any doubt, the most used taxonomy all over the
@@
 names and the physical specimens that serve as exemplars for the species
-\citep{Federhen-2015}.
+\citep{federhen2014type}.
 

### 2015-09-12 docs/paper/out/paper.tex
remove duplicate numbering in sections -- paper

@@
 
-\section{1. Introduction}\label{introduction}
+\section{Introduction}\label{introduction}
 
@@
 
-\section{2. Materials and Methods}\label{materials-and-methods}
+\section{Materials and Methods}\label{materials-and-methods}
 
-\subsection{2.x Amazon Web Services}\label{x-amazon-web-services}
+\subsection{Amazon Web Services}\label{amazon-web-services}
 
-\subsection{2.x Scala}\label{x-scala}
+\subsection{Scala}\label{scala}
 
@@
 
-\subsection{2.x Statika}\label{x-statika}
+\subsection{Statika}\label{statika}
 
@@
 
-\subsection{2.x Datasets}\label{x-datasets}
+\subsection{Datasets}\label{datasets}
 
@@
 
-\subsection{2.x Loquat}\label{x-loquat}
+\subsection{Loquat}\label{loquat}
 
@@
 
-\subsection{2.x Type-safe DSLs for BLAST and
-FLASH}\label{x-type-safe-dsls-for-blast-and-flash}
+\subsection{Type-safe DSLs for BLAST and
+FLASH}\label{type-safe-dsls-for-blast-and-flash}
 
@@
 
-\subsubsection{2.x.a BLAST DSL}\label{x.a-blast-dsl}
+\subsubsection{BLAST DSL}\label{blast-dsl}
 
@@
 
-\subsubsection{2.x.b FLASH DSL}\label{x.b-flash-dsl}
+\subsubsection{FLASH DSL}\label{flash-dsl}
 
@@
 
-\subsection{2.x Bio4j}\label{x-bio4j}
+\subsection{Bio4j}\label{bio4j}
 
@@
 
-\section{3. Results}\label{results}
+\section{Results}\label{results}
 
-\subsection{3.1 Overview}\label{overview}
+\subsection{Overview}\label{overview}
 
@@
 
-\subsection{3.x 16S Reference Database
-Construction}\label{x-16s-reference-database-construction}
+\subsection{16S Reference Database
+Construction}\label{s-reference-database-construction}
 
@@
 
-\subsection{3.x Bio4j and Graph
-Databases}\label{x-bio4j-and-graph-databases}
+\subsection{Bio4j and Graph Databases}\label{bio4j-and-graph-databases}
 
-\subsection{3.x MG7 Pipeline
-Description}\label{x-mg7-pipeline-description}
+\subsection{MG7 Pipeline Description}\label{mg7-pipeline-description}
 
-\subsection{3.x Taxonomic Assignment
-Algorithms}\label{x-taxonomic-assignment-algorithms}
+\subsection{Taxonomic Assignment
+Algorithms}\label{taxonomic-assignment-algorithms}
 
-\subsubsection{3.x.y Lowest Common Ancestor based Taxonomic
-Assignment}\label{x.y-lowest-common-ancestor-based-taxonomic-assignment}
+\subsubsection{Lowest Common Ancestor based Taxonomic
+Assignment}\label{lowest-common-ancestor-based-taxonomic-assignment}
 
@@
 
-\subsubsection{3.x.z Best BLAST hit taxonomic
-assignment}\label{x.z-best-blast-hit-taxonomic-assignment}
+\subsubsection{Best BLAST hit taxonomic
+assignment}\label{best-blast-hit-taxonomic-assignment}
 
@@
 
-\subsection{3.x Other}\label{x-other}
+\subsection{Other}\label{other}
 
@@
 
-\subsection{3.x Using MG7 with some example
-data-sets}\label{x-using-mg7-with-some-example-data-sets}
+\subsection{Using MG7 with some example
+data-sets}\label{using-mg7-with-some-example-data-sets}
 
@@
 
-\subsection{3.7 MG7 availability}\label{mg7-availability}
+\subsection{MG7 availability}\label{mg7-availability}
 
@@
 
-\section{4. Discussion}\label{discussion}
+\section{Discussion}\label{discussion}
 
-\subsection{4.1 Novelty points of MG7}\label{novelty-points-of-mg7}
+\subsection{Novelty points of MG7}\label{novelty-points-of-mg7}
 
@@
 
-\subsection{4.2 Designed for future
+\subsection{Designed for future
 challenges}\label{designed-for-future-challenges}
@@
 
-\subsection{4.3 MG7 Future developments}\label{mg7-future-developments}
+\subsection{MG7 Future developments}\label{mg7-future-developments}
 
-\subsubsection{4.3.1 Comparison of groups of
+\subsubsection{Comparison of groups of
 samples}\label{comparison-of-groups-of-samples}
 
-\subsubsection{4.3.2 Interactive visualizations using the output files
-of MG7 (Biographika
+\subsubsection{Interactive visualizations using the output files of MG7
+(Biographika
 project)}\label{interactive-visualizations-using-the-output-files-of-mg7-biographika-project}
 
-\section{5 Acknowledgements}\label{acknowledgements}
+\section{Acknowledgements}\label{acknowledgements}
 
 INTERCROSSING (Grant 289974)
-
-\section{6 References}\label{references}
 

### 2015-09-12 src/main/scala/bundles/bio4jTaxonomy.scala
Project def improvements; fixed bug with the titan config

@@ -41,12 +41,17 @@ case object bio4jTaxonomy extends Bundle() {
 
   lazy val conf: BaseConfiguration = {
     val base = new BaseConfiguration()
-    base.setProperty("storage.directory", location)
+    base.setProperty("storage.directory", location.getCanonicalPath)
     base.setProperty("storage.backend", "berkeleyje")
+    base.setProperty("storage.batch-loading", "false")
+    base.setProperty("storage.transactions", "true")
+    base.setProperty("query.fast-property", "false")
+    base.setProperty("schema.default", "none")
     base
   }
 
   // the graph; its only (direct) use is for indexes
+  // FIXME: this works but still with errors, should be fixed (something about transactions)
   lazy val graph: TitanNCBITaxonomyGraph =
     new TitanNCBITaxonomyGraph(
       new DefaultTitanGraph(TitanFactory.open(conf))

### 2015-09-13 docs/paper/out/paper.tex
add new version of 16S db from @rtobes -- paper
See https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-139846932

@@
 Our 16S Reference Database is a curated subset of sequences from NCBI
-nucleotide database \textbf{nt}. This subset of 16S sequences was
-selected by similarity with the bacterial and archaeal reference
-sequences downloaded from RDP database \citep{cole2013ribosomal}. RDP
-unaligned sequences were used to capture new16S sequences from nt using
-BLAST similarity strategies and, then, performing additional curation
-steps to remove sequences with poor taxonomic assignments to taxonomic
-nodes close to the root of the taxonomic tree. All the nucleotide
-sequences included in nt database has a taxonomic assignment provided by
-the genbank sequence submitter. NCBI provides a table (available at
+nucleotide database \textbf{nt}. The sequences included were selected by
+similarity with the bacterial and archaeal reference sequences
+downloaded from the \textbf{RDP database} \citep{cole2013ribosomal}. RDP
+unaligned sequences were used to capture new 16S sequences from
+\textbf{nt} using BLAST similarity search strategies and then,
+performing additional curation steps to remove sequences with poor
+taxonomic assignments to taxonomic nodes close to the root of the
+taxonomy tree. All the nucleotide sequences included in \textbf{nt}
+database has a taxonomic assignment provided by the \textbf{Genbank}
+sequence submitter. NCBI provides a table (available at
 ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
 Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
-based on a submitter-maintained taxonomic annotation system for
-reference sequences that supposes a sustainable system able to face the
-expected number of reference sequences that will populate the public
+based on a crowdsourced submitter-maintained taxonomic annotation system
+for reference sequences. It supposes a sustainable system able to face
+the expected number of reference sequences that will populate the public
 global nucleotide databases in the near future. Another advantageous
-point is that we are based on NCBI taxonomy, the de facto standard
-taxonomic classification for biomolecular data \citep{cochrane20102010}.
-NCBI taxonomy is, without any doubt, the most used taxonomy all over the
-world and the most similar to the official taxonomies of each specific
-field. This is a crucial point because all the type-culture and tissue
-databanks follow this official taxonomical classification and, in
-addition, all the knowledge accumulated is referred to this taxonomy. In
-addition NCBI provides a direct connection between taxonomical formal
-names and the physical specimens that serve as exemplars for the species
+point is that we are based on NCBI taxonomy, the \emph{de facto}
+standard taxonomic classification for biomolecular data
+\citep{cochrane20102010}. NCBI taxonomy is, undoubtedly, the most used
+taxonomy all over the world and the most similar to the official
+taxonomies of each specific field. This is a crucial point because all
+the type-culture and tissue databanks follow this official taxonomical
+classification and, in addition, all the knowledge accumulated during
+last decades is referred to this taxonomy. In addition NCBI provides a
+direct connection between taxonomical formal names and the physical
+specimens that serve as exemplars for the species
 \citep{federhen2014type}.
 
-If metagenomics results are easily integrated with the theoretical and
-experimental knowledge of each specific area, the impact of metagenomics
-will be higher that if metagenomics progress in a disconnected research
-branch. This strategy for building our database allows substituting the
-16S database by any other subset of nt, even by the complete nt database
-if it would needed, for example, for analyzing shotgun metagenomics
-data.
+Certainly, if metagenomics results are easily integrated with the
+theoretical and experimental knowledge of each specific area, the impact
+of metagenomics will be higher that if metagenomics progresses as a
+disconnected research branch. Considering that metagenomics data
+interoperability, which is especially critical in clinical environments,
+requires a stable taxonomy to be used as reference, we decided to rely
+on the most widely used taxonomy: the NCBI taxonomy. In addition, the
+biggest global sequence database GenBank follows this taxonomy to
+register the origin of all their submitted sequences. Our 16S database
+building strategy allows the substitution of the 16S database by any
+other subset of \textbf{nt}, even by the complete \textbf{nt} database
+if it would be needed, for example, for analyzing shotgun metagenomics
+data. This possibility of changing the reference database provides
+flexibility to the system enabling it for easy updating and
+project-driven personalization.
 

### 2015-09-13 docs/paper/out/paper.tex
reorganize Discussion, add sections -- paper

@@
 
-\subsection{Other}\label{other}
+\subsection{Using MG7 with some example
+data-sets}\label{using-mg7-with-some-example-data-sets}
+
+We selected the datasets described in {[}Kennedy-2014{]} (??)
+
+\subsection{MG7 availability}\label{mg7-availability}
+
+MG7 is open source, available at https://github.com/ohnosequences/mg7
+under an \href{http://www.gnu.org/licenses/agpl-3.0.en.html}{AGPLv3}
+license.
+
+\section{Discussion}\label{discussion}
+
+\subsection{What MG7 brings}\label{what-mg7-brings}
+
+We could summarize the most innovative ideas and developments in MG7:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Treat data analysis as a software project. This makes for radical
+  improvements in \emph{reproducibility}, \emph{reuse},
+  \emph{versioning}, \emph{safety}, \emph{automation} and
+  \emph{expressiveness}
+\item
+  input and output data, their locations and type are expressible and
+  checked at compile-time using our Scala library \emph{datasets} ­3.
+  management of dependencies and machine configurations using our Scala
+  library \emph{Statika}
+\item
+  automation of AWS cloud resources and processes, including
+  distribution and parallelization through the use of \emph{Loquat}
+\item
+  taxonomic data and related operations are treated natively as what
+  they are: graphs, through the use of \emph{Bio4j}
+\item
+  MG7 provides a sustainable model for taxonomic assignment, appropriate
+  to face the challenging amount of data that high throughput sequencing
+  technologies generate
+\end{enumerate}
+
+\subsection{A new approach to data
+analysis}\label{a-new-approach-to-data-analysis}
 
@@
 
-\subsection{Using MG7 with some example
-data-sets}\label{using-mg7-with-some-example-data-sets}
+\subsection{Inputs, outputs, data: compile-time, expressive,
+composable}\label{inputs-outputs-data-compile-time-expressive-composable}
 
-We selected the datasets described in {[}Kennedy-2014{]} (??)
+\subsection{Tools, data, dependencies and machine
+configurations}\label{tools-data-dependencies-and-machine-configurations}
 
-\subsection{MG7 availability}\label{mg7-availability}
+\subsection{Parallel cloud execution ??}\label{parallel-cloud-execution}
 
-MG7 is open source, available at https://github.com/ohnosequences/mg7
-under an \href{http://www.gnu.org/licenses/agpl-3.0.en.html}{AGPLv3}
-license.
+\subsection{Taxonomy and Bio4j}\label{taxonomy-and-bio4j}
 
-\section{Discussion}\label{discussion}
+\subsection{Future-proof}\label{future-proof}
 
-\subsection{Novelty points of MG7}\label{novelty-points-of-mg7}
-
-The most innovative ideas and developments integrated in MG7 are:
-
-­- The management dependencies checking their correctness before
-compilation using Scala type system - The automation of cloud resources
-and processes (parallelization management) - The cloud-oriented
-development of the system including a modeling AWS resources based on
-the powerful data typing of Scala - The use of the Graph databases
-paradigm to store and manage the taxonomy tree to obtain the taxonomic
-assignments and the cumulative frequencies - MG7 provides a sustainable
-model for updating the database of reference sequences appropriate to
-face the challenging amount of sequences that are generating the new
-high throughput technologies of sequencing
-
-\subsection{Designed for future
-challenges}\label{designed-for-future-challenges}
+\subsection{MG7 Future developments}\label{mg7-future-developments}
 
 Other possible uses of the general schema: statika, loquat, \ldots{}
-
-\subsection{MG7 Future developments}\label{mg7-future-developments}
 

### 2015-09-13 docs/paper/out/paper.tex
add @rtobes lines about bio4j and taxonomy
see https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-139849162

@@
 
+The hierarchic structure of the taxonomy of the living organisms is a
+tree, and, hence, is also a graph in which each node, with the exception
+of the root node, has a unique parent node. It led us to model the
+taxonomy tree as a graph using the graph database paradigm. Previously
+we developed Bio4j \textbf{{[}Pareja-Tobes-2015{]}}, a platform for the
+integration of semantically rich biological data using typed graph
+models. It integrates most publicly available data linked with sequences
+into a set of interdependent graphs to be used for bioinformatics
+analysis and especially for biological data.
+
 \subsection{Future-proof}\label{future-proof}

### 2015-09-13 docs/paper/out/paper.tex
add mg7 for shotgun as a future development -- paper
See https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-139857648

@@
 
+\subsubsection{Shotgun metagenomics}\label{shotgun-metagenomics}
+
+It is certainly possible to adapt MG7 to work with shotgun metagenomics
+data. Simply changing the reference database to include whole genome
+sequence data could yield interesting results. This could also be
+refined by restricting reference sequences according to all sort of
+criteria, like biological function or taxonomy. Bio4j would be an
+invaluable tool here, thanks to its ability to express express complex
+predicates on sequences using all the information linked with them (GO
+annotations, UniProt data, NCBI taxonomy, etc).
+
 \subsubsection{Comparison of groups of

### 2015-09-13 docs/paper/out/paper.tex
improve and expand the idea of data analysis as software -- paper

@@
 
-General approach. An analysis is defined as a software project. It can
-evolve in the same way. We can run the analysis in a test phase, review
-configuration and changes, etc. Key advantages of this approach are
+MG7 proposes to define and work with a particular data analysis task as
+a software project, using Scala. The idea is that \emph{everything}:
+data description, their location, configuration parameters, the
+infrastructure used, \ldots{} should be expressed as Scala code, and
+treated in the same way as any (well-managed) software project. This
+includes, among other things, using version control systems
+(\texttt{git} in our case), writing tests, making stable releases
+following \href{http://semver.org/}{semantic versioning} or publishing
+artifacts to a repository.
+
+What we see as key advantages of this approach (when coupled with
+compile-time specification and checking), are
 
@@
 \item
-  \textbf{Versioning} The analysis is a software project so it goes
-  through the same stages, there can be different versions, stable
-  releases, etc.
+  \textbf{Versioning} as in any software project, there can be different
+  versions, stable releases, etc.
 \item
   \textbf{Reuse} we can build standard configurations on top of this and
-  reuse them for subsequent data analysis.
+  reuse them for subsequent data analysis. A particular data analysis
+  \emph{task} can be used as a \emph{library} in further analysis.
 \item
   \textbf{Decoupling} We can start working on the analysis
-  specification, without any need for data in a much easier way.
+  specification, without any need for available data in a much easier
+  way.
 \item
-  \textbf{Expresiveness and safety} choose only from valid Illumina read
-  types, build default FLASH command based on that, \ldots{}
+  \textbf{Documentation} We can take advantage of all the effort put
+  into software documentation tools and practices, such as in our case
+  Scaladoc or literate programming. As documentation, analysis processes
+  and data specification live together in the files, it is much easier
+  to keep coherence between them.
+\item
+  \textbf{Expresiveness and safety} For example in our case we can
+  choose only from valid Illumina read types, and then build a default
+  FLASH command based on that. The output locations, being declared
+  statically, are also available for use in further analysis.
 \end{itemize}

### 2015-09-13 docs/paper/out/paper.tex
add who did what to libraries -- paper

@@
 \href{https://github.com/ohnosequences/statika}{Statika} is a Scala
-library developed by \textbf{so and so} which serves as a way of
+library developed by the first and last authors which serves as a way of
 defining and composing machine behaviors statically. The main component
@@
 \href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
-library developed by \textbf{so and so} to declare datasets and their
-locations. \textbf{Data} is represented as type-indexed fields: Keys are
-modeled as singleton types, and values correspond to what could be
-called a denotation of the key: a value of type \texttt{Location} tagged
-with the key type. Then a \textbf{Dataset} is essentially a collection
-of data, which are guaranteed statically to be different through
-type-level predicates, making use of the value ↔ type correspondence
-which can be established through singleton types and implicits. A
-dataset location is then just a list of locations formed by locations of
-each data member of that dataset.
+library developed by the first and last authors to declare datasets and
+their locations. \textbf{Data} is represented as type-indexed fields:
+Keys are modeled as singleton types, and values correspond to what could
+be called a denotation of the key: a value of type \texttt{Location}
+tagged with the key type. Then a \textbf{Dataset} is essentially a
+collection of data, which are guaranteed statically to be different
+through type-level predicates, making use of the value ↔ type
+correspondence which can be established through singleton types and
+implicits. A dataset location is then just a list of locations formed by
+locations of each data member of that dataset.
 
@@
 \href{https://github.com/ohnosequences/loquat}{Loquat} is a library
-developed by \textbf{so and so} designed for the execution of
-embarrassingly parallel tasks using S3, SQS and EC2.
+developed by the first, second and last authors designed for the
+execution of embarrassingly parallel tasks using S3, SQS and EC2.
 

### 2015-09-13 docs/paper/out/paper.tex
add specific versions for libs with link to releases -- paper
WARNING: some of them do not exist yet, but they should be the ones
that we will use.

@laughedelic please review.

@@
 
+MG7 uses
+\href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
+2.0.0}.
+
 \subsection{Datasets}\label{datasets}
@@
 ``physical'' representations denote.
+
+MG7 uses
+\href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
+0.2.0}.
 
@@
 
-\subsection{Type-safe DSLs for BLAST and
-FLASH}\label{type-safe-dsls-for-blast-and-flash}
+MG7 uses
+\href{https://github.com/ohnosequences/loquat/releases/tag/v?.?.?}{ohnosequences/loquat
+?.?.?}.
 
-We developed our own type-safe DSLs (Domain Specific Language) for
+\subsection{Type-safe EDSLs for BLAST and
+FLASH}\label{type-safe-edsls-for-blast-and-flash}
+
+We developed our own Scala-based type-safe EDSLs (Embedded Domain
+Specific Language) for
 \href{https://github.com/ohnosequences/flash}{FLASH} and
@@
 
-\subsubsection{BLAST DSL}\label{blast-dsl}
+\subsubsection{BLAST EDSL}\label{blast-edsl}
 
@@
 
-\subsubsection{FLASH DSL}\label{flash-dsl}
+MG7 uses
+\href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
+0.2.0}.
 
-In the same spirit as for BLAST,
+\subsubsection{FLASH EDSL}\label{flash-edsl}
+
+In the same spirit as for BLAST, we implemented a type-safe EDSL for
+FLASH expressions and their execution, sporting features equivalent to
+those outlined for the BLAST EDSL.
+
+MG7 uses
+\href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
+0.1.0}.
 
@@
 the use their Java API from Scala in the assignment phase.
+
+MG7 uses
+\href{https://github.com/bio4j/bio4j/releases/tag/v0.12.0-RC3}{bio4j/bio4j
+0.12.0-RC3} and
+\href{https://github.com/bio4j/bio4j-titan/releases/tag/v0.4.0-RC2}{bio4j/bio4j-titan
+0.4.0-RC2}.
 

### 2015-09-13 docs/paper/out/paper.tex
set loquat version used to 2.0.0 -- paper
cite records, mention cosas -- paper

@@
 implicits. A dataset location is then just a list of locations formed by
-locations of each data member of that dataset.
+locations of each data member of that dataset. All this is based on what
+could be described as an embedding in Scala of an extensible record
+system with concatenation on disjoint labels, in the spirit of
+\citep[\citet{harper1991record}]{harper1990extensible}. For that
+\emph{Datasets} uses
+\href{https://github.com/ohnosequences/cosas/releases/tag/v0.7.0}{ohnosequences/cosas
+0.7.0}.
 
@@
 MG7 uses
-\href{https://github.com/ohnosequences/loquat/releases/tag/v?.?.?}{ohnosequences/loquat
-?.?.?}.
+\href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
+2.0.0}.
 

### 2015-09-13 src/main/scala/metagenomica/bundles/flash.scala
Struggling with type not-inference

@@ -1,0 +1,6 @@
+package ohnosequences.metagenomica.bundles
+
+import ohnosequencesBundles.statika.Flash
+
+// TODO: it should be in era7bio/bundles
+case object flash extends Flash("1.2.11")

### 2015-09-13 src/main/scala/metagenomica/configuration.scala

@@ -1,0 +1,39 @@
+package ohnosequences.metagenomica
+
+import ohnosequences.cosas._, typeSets._, types._
+
+import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
+
+import ohnosequences.flash.api._
+import ohnosequences.flash.data._
+
+import java.io.File
+
+case object configuration {
+
+  // TODO: move it to datasets
+  implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
+    new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
+
+  trait MetagenomicaData {
+
+    // Paired end reads as input
+    type ReadsType <: AnyReadsType { type EndType = pairedEndType }
+    val  readsType: ReadsType
+
+    case object reads1 extends PairedEnd1Fastq(readsType, "reads1.fastq.gz"); type Reads1 = reads1.type
+    case object reads2 extends PairedEnd2Fastq(readsType, "reads2.fastq.gz"); type Reads2 = reads2.type
+
+    // TODO: leave free additional modifications
+    lazy val flashOptions = flash.defaults update (
+      read_len(readsType.length.toInt)    :~:
+      max_overlap(readsType.length.toInt) :~: ∅
+    )
+
+    case object mergedReads extends MergedReads(readsType, reads1, reads2, flashOptions)
+
+    case object mergedReadsStats extends MergedReadsStats(mergedReads)
+
+  }
+
+}

### 2015-09-13 src/main/scala/metagenomica/loquats/flash/config.scala

@@ -23,68 +23,70 @@ import java.io.File
 
 
 trait AnyFlashConfig extends Era7LoquatConfig { config =>
+  type Data <: AnyFlashData
+  // val  data: Data
 
-  type DataProcessing <: AnyFlashDataProcessing
+  type DataProcessing <: FlashDataProcessing[Data]
   val  dataProcessing: DataProcessing
 
   type FlashDataMappring = AnyDataMapping { type DataProcessing = config.DataProcessing }
   val  dataMappings: List[FlashDataMappring]
 }
 
-abstract class FlashConfig[D <: AnyFlashDataProcessing](val dataProcessing: D) extends AnyFlashConfig {
-
-  type DataProcessing = D
-}
+// abstract class FlashConfig[D <: AnyFlashData](val dataProcessing: D) extends AnyFlashConfig {
+//
+//   type DataProcessing = D
+// }
 
 //////////////////////////////////////////////////////////////////////////////////////
 
-case object flashTest {
-
-  // TODO: move it to datasets
-  implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
-    new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
-
-
-  case object testData extends FlashData(illumina.PairedEnd(bp300, InsertSize(3000)))
-
-  case object testDataProcessing extends FlashDataProcessing(testData)
-
-  case object testConfig extends FlashConfig(testDataProcessing) {
-
-    val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
-
-    val managerConfig = ManagerConfig(
-      instanceType = m3_medium,
-      purchaseModel = SpotAuto
-    )
-
-    val workersConfig = WorkersConfig(
-      instanceType = m3_medium,
-      purchaseModel = SpotAuto,
-      groupSize = WorkersGroupSize(0, 1, 10)
-    )
-
-    val terminationConfig = TerminationConfig(
-      terminateAfterInitialDataMappings = true
-    )
-
-    val dataMappings: List[FlashDataMappring] =
-      List(
-        DataMapping(
-          "ERR567374_1",
-          dataProcessing
-        )(remoteInput =
-            dataProcessing.data.reads1.atS3(ObjectAddress("resources.ohnosequences.com", "16s/public-datasets/PRJEB6592/reads/ERR567374_1.fastq.gz")) :~:
[…]

### 2015-09-13 src/main/scala/metagenomica/loquats/flash/dataProcessing.scala

@@ -1,31 +1,37 @@
 package ohnosequences.metagenomica.loquats.flash
 
+import ohnosequences.metagenomica.configuration._
+import ohnosequences.metagenomica.bundles
+
 import ohnosequences.loquat._, utils._, dataProcessing._
-import ohnosequences.statika._, bundles._, instructions._
-import ohnosequences.flash._, api._, data._
+
+import ohnosequences.statika.bundles._
+import ohnosequences.statika.instructions._
+
+import ohnosequences.flash._
+import ohnosequences.flash.api._
+import ohnosequences.flash.data._
+
 import ohnosequences.cosas._, typeSets._, types._
 import ops.typeSets._
+
 import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-import ohnosequencesBundles.statika.Flash
 import java.io.File
 
-case object flashBundle extends Flash("1.2.11")
 
+class FlashDataProcessing[D <: AnyFlashData](val data: D)(implicit
+  parseInputFiles: ParseDenotations[(D#Reads1 :^: D#Reads2 :^: DNil)#LocationsAt[FileDataLocation], File],
+  outputFilesToMap: ToMap[(D#Merged :^: D#Stats :^: DNil)#LocationsAt[FileDataLocation], AnyData, FileDataLocation]
+  // parseInputFiles: ParseDenotations[(data.reads1.type :^: data.reads2.type :^: DNil)#LocationsAt[FileDataLocation], File],
+  // outputFilesToMap: ToMap[(data.merged.type :^: data.stats.type :^: DNil)#LocationsAt[FileDataLocation], AnyData, FileDataLocation]
+) extends DataProcessingBundle(bundles.flash)(
+  input = (data.reads1: D#Reads1) :^: (data.reads2: D#Reads2) :^: DNil,
+  output = (data.merged: D#Merged) :^: (data.stats: D#Stats) :^: DNil
+  // input = (data.reads1 :^: data.reads2 :^: DNil): (D#Reads1 :^: D#Reads2 :^: DNil),
+  // output = (data.merged :^: data.stats :^: DNil): (D#Merged :^: D#Stats :^: DNil)
+)(parseInputFiles, outputFilesToMap) {
 
-trait AnyFlashDataProcessing extends AnyDataProcessingBundle {
-
-  type Data <: AnyFlashData
-  val  data: Data
-
[…]

### 2015-09-13 docs/paper/out/paper.tex
put Materials & Methods *after* Discussion -- paper
seriously restructure the paper structure -- paper
Much better! it starts to look like a paper.

@@
 current one. These are some of the aspects that have suffered crucial
-changes and advances with a direct impact in metagenomics data analysis.
-i. The first aspect is related to the sequences to be analyzed: the
-reads are larger, the sequencing depth and the number of samples of each
+changes and advances with a direct impact in metagenomics data analysis:
+
+\begin{enumerate}
+\def\labelenumi{\roman{enumi}.}
+\tightlist
+\item
+  The first aspect is related to the sequences to be analyzed: the reads
+  are larger, the sequencing depth and the number of samples of each
   project are considerably bigger. The first metagenomics studies were
@@
   diseases and has allowed exploring the biodiversity in many earth
-environments ii. The second aspect derives from the impressive genomics
-explosion, its effect being felt in this case in the reference
-sequences. The immense amount of sequences available in public
-repositories demands new approaches in curation, update and storage for
-metagenomics reference databases: current models will or already have
-problems to face the future avalanche of metagenomic sequences. iii. The
-third aspect to consider for metagenomics data analysis is related to
-the appearance of new models for massive computation and storage and to
-the new programming methodologies (Scala, \ldots{}) and new cloud models
-and resources. The immense new possibilities that these advances offer
-must have a direct impact in the metagenomics data analysis. iv. And
-finally the new social manner to do science, and especially genomic
-science is the fourth aspect to consider. Metagenomics evolves in a
-social and global scenario following a science democratization trend in
-which many small research groups from distant countries share a common
-big metagenomics project. This global cooperation demands systems
-allowing following exactly the same pipelines using equivalent cloud
-resources to modularly execute the analysis in an asynchronous way of
-working between different groups. This definitively new scenario demands
+  environments
+\item
+  The second aspect derives from the impressive genomics explosion, its
+  effect being felt in this case in the reference sequences. The immense
+  amount of sequences available in public repositories demands new
+  approaches in curation, update and storage for metagenomics reference
+  databases: current models will or already have problems to face the
+  future avalanche of metagenomic sequences.
+\item
+  The third aspect to consider for metagenomics data analysis is related
+  to the appearance of new models for massive computation and storage
+  such as the so-called cloud, or the widespread adoption of programming
+  methodologies like functional programming, or, more speculatively,
+  dependently typed programming. The immense new possibilities that
+  these advances offer must have a direct impact in metagenomics data
+  analysis.
+\item
+  And finally the new social manner to do science, and especially
+  genomic science is the fourth aspect to consider. Metagenomics evolves
+  in a social and global scenario following a science democratization
+  trend in which many small research groups from distant countries share
+  a common big metagenomics project. This global cooperation demands
+  systems allowing following exactly the same pipelines using equivalent
+  cloud resources to modularly execute the analysis in an asynchronous
+  way of working between different groups. This new scenario calls for
   new methods and tools to handle the current and future volume of
-metagenomic data with the sufficient speed of analysis. Considering all
-these aspects we have designed a new open source methodology for
-analyzing metagenomics data that exploits the new possibilities that
-cloud computing offers to get a system robust, programmatically
-configurable, modular, distributed, flexible, scalable and traceable in
-which the biological databases of reference sequences can be easily
-updated and/or frequently substituted by new ones or by databases
-specifically designed for focused projects.
+  metagenomic data with the sufficient speed of analysis.
+\end{enumerate}
 
-\section{Materials and Methods}\label{materials-and-methods}
+Considering all these aspects we have designed a new open source
+methodology for analyzing metagenomics data that exploits the new
+possibilities that cloud computing offers to get a system robust,
+programmatically configurable, modular, distributed, flexible, scalable
+and traceable in which the biological databases of reference sequences
+can be easily updated and/or frequently substituted by new ones or by
+databases specifically designed for focused projects.
 
-\subsection{Amazon Web Services}\label{amazon-web-services}
+\section{Results}\label{results}
 
-\subsection{Scala}\label{scala}
+\subsection{Overview}\label{overview}
 
-\href{http://www.scala-lang.org/}{Scala} is a hybrid object-functional
-programming language which runs on Java Virtual Machine. It has support
-for type-level programming, type-dependent types (through type members)
-and singleton types, which permits a restricted form of dependent types
-where types can depend essentially on values determined at compile time
-(through their corresponding singleton types). Conversely, through
-implicits one can retrieve the value corresponding to a singleton type.
+To tackle the challenges posed by metagenomics big data analysis
+outlined in the Introduction
 
-The other key feature for us is Java interoperability, which let us
-build on the vast number of existing Java libraries; we take advantage
-of this when using Bio4j as an API for the NCBI taxonomy.
+\begin{itemize}
+\tightlist
+\item
+  Static reproducible specification of dependencies and behavior of the
+  different components using \emph{Statika} and \emph{Datasets}
+\item
+  Parallelization and distributed analysis based on AWS, with on-demand
+  infrastructure as the basic paradigm
+\item
+  Definition of complex pipelines using \emph{Loquat}, a composable
+  system for scaling/parallelizing stateless computations especially
+  designed for Amazon Web Services (AWS)
+\item
+  A new approach to data analysis specification, management and
+  specification based on working with it in exactly the same way as for
+  a software project, together with the extensive use of compile-time
+  structures and checks
+\item
+  Modeling of the taxonomy tree using the new paradigm of graph
+  databases (Bio4j). It facilitates the taxonomic assignment tasks and
+  the calculation of the taxa abundance values considering the
+  hierarchic structure of taxonomy tree (cumulative values)
+\item
+  per-read assignment (??)
+\end{itemize}
 
-MG7 itself and all the libraries used are written in Scala
-\texttt{2.11}.
+\subsection{Libraries and resources}\label{libraries-and-resources}
 
-\subsection{Statika}\label{statika}
+In this section we describe the resources and Scala libraries developed
+by the authors on top of which MG7 is built.
+
+\href{http://www.scala-lang.org/}{Scala}, a hybrid object-functional
+programming language, was chosen based on the possibility of using
+certain advanced programming styles, and Java interoperability, which
+let us build on the vast number of existing Java libraries; we take
+advantage of this when using Bio4j as an API for the NCBI taxonomy. It
+has support for type-level programming, type-dependent types (through
+type members) and singleton types, which permits a restricted form of
+dependent types where types can depend essentially on values determined
+at compile time (through their corresponding singleton types).
+Conversely, through implicits one can retrieve the value corresponding
+to a singleton type.
+
+\subsubsection{\texorpdfstring{\emph{Statika}: machine configuration and
+behavior}{Statika: machine configuration and behavior}}\label{statika-machine-configuration-and-behavior}
 
@@
 
-MG7 uses
-\href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
-2.0.0}.
-
-\subsection{Datasets}\label{datasets}
+\subsubsection{\texorpdfstring{\emph{Datasets}: a mini-language for
+data}{Datasets: a mini-language for data}}\label{datasets-a-mini-language-for-data}
 
 \href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
-library developed by the first and last authors to declare datasets and
-their locations. \textbf{Data} is represented as type-indexed fields:
-Keys are modeled as singleton types, and values correspond to what could
-be called a denotation of the key: a value of type \texttt{Location}
-tagged with the key type. Then a \textbf{Dataset} is essentially a
-collection of data, which are guaranteed statically to be different
-through type-level predicates, making use of the value ↔ type
-correspondence which can be established through singleton types and
-implicits. A dataset location is then just a list of locations formed by
-locations of each data member of that dataset. All this is based on what
-could be described as an embedding in Scala of an extensible record
-system with concatenation on disjoint labels, in the spirit of
+library developed by the first and last authors with the goal of being a
+Scala-embedded mini-language for datasets and their locations.
+\textbf{Data} is represented as type-indexed fields: Keys are modeled as
+singleton types, and values correspond to what could be called a
+denotation of the key: a value of type \texttt{Location} tagged with the
+key type. Then a \textbf{Dataset} is essentially a collection of data,
+which are guaranteed statically to be different through type-level
+predicates, making use of the value ↔ type correspondence which can be
+established through singleton types and implicits. A dataset location is
+then just a list of locations formed by locations of each data member of
+that dataset. All this is based on what could be described as an
+embedding in Scala of an extensible record system with concatenation on
+disjoint labels, in the spirit of
 \citep[\citet{harper1991record}]{harper1990extensible}. For that
 \emph{Datasets} uses
-\href{https://github.com/ohnosequences/cosas/releases/tag/v0.7.0}{ohnosequences/cosas
-0.7.0}.
+\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}.
 
@@
 
-MG7 uses
-\href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
-0.2.0}.
-
-\subsection{Loquat}\label{loquat}
+\subsubsection{\texorpdfstring{\emph{Loquat}: Parallel data processing
+with
+AWS}{Loquat: Parallel data processing with AWS}}\label{loquat-parallel-data-processing-with-aws}
 
@@
 
-MG7 uses
-\href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
-2.0.0}.
-
-\subsection{Type-safe EDSLs for BLAST and
+\subsubsection{Type-safe eDSLs for BLAST and
 FLASH}\label{type-safe-edsls-for-blast-and-flash}
 
-We developed our own Scala-based type-safe EDSLs (Embedded Domain
-Specific Language) for
+We developed our own Scala-based type-safe eDSLs (\textbf{e}mbedded
+\textbf{D}omain \textbf{S}pecific \textbf{L}anguage) for
 \href{https://github.com/ohnosequences/flash}{FLASH} and
@@
 their execution.
-
-\subsubsection{BLAST EDSL}\label{blast-edsl}
 
@@
 
-MG7 uses
-\href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
-0.2.0}.
-
-\subsubsection{FLASH EDSL}\label{flash-edsl}
-
 In the same spirit as for BLAST, we implemented a type-safe EDSL for
@@
 
-MG7 uses
-\href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
-0.1.0}.
+\subsubsection{Bio4j and Graph
+Databases}\label{bio4j-and-graph-databases}
 
-\subsection{Bio4j}\label{bio4j}
+\citep[Bio4j][]{pareja2015bio4j} is a data platform integrating data
+from different resources such as UniProt or GO in a graph data paradigm.
+In the assignment phase we use a subgraph containing the NCBI Taxonomy,
+wrapping in Scala its Java API in a tree algebraic data type.
 
-\href{https://github.com/bio4j/bio4j}{Bio4j} is a data platform
-integrating data from different resources such as UniProt or GO in a
-graph data paradigm. We use the module containing the NCBI Taxonomy, and
-the use their Java API from Scala in the assignment phase.
-
-MG7 uses
-\href{https://github.com/bio4j/bio4j/releases/tag/v0.12.0-RC3}{bio4j/bio4j
-0.12.0-RC3} and
-\href{https://github.com/bio4j/bio4j-titan/releases/tag/v0.4.0-RC2}{bio4j/bio4j-titan
-0.4.0-RC2}.
-
-\section{Results}\label{results}
-
-\subsection{Overview}\label{overview}
-
-To tackle the challenges posed by metagenomics big data analysis
-outlined in the Introduction,
-
-­- AWS resources in Scala (??) - A new approach to data analysis
-specification, management and specification based on working with it in
-exactly the same way as for a software project, together with the
-extensive use of compile-time structures and checks. - Parallelization
-and distributed analysis based on AWS, with on-demand infrastructure as
-the basic paradigm ­- fully automated processes, data and cloud
-resources management. - Static reproducible specification of
-dependencies and behavior of the different components using
-\emph{Statika} and \emph{Datasets} - Definition of complex pipelines
-using \emph{Loquat} a composable system for scaling/parallelizing
-stateless computations especially designed for Amazon Web Services (AWS)
-- Modeling of the taxonomy tree using the new paradigm of graph
-databases (Bio4j). It facilitates the taxonomic assignment tasks and the
-calculation of the taxa abundance values considering the hierarchic
-structure of taxonomy tree (cumulative values). - per-read assignment
-(??)
-
-\subsection{16S Reference Database
+\subsubsection{16S Reference Database
 Construction}\label{s-reference-database-construction}
@@
 
-\subsection{Bio4j and Graph Databases}\label{bio4j-and-graph-databases}
+\subsection{Pipeline Description}\label{pipeline-description}
 
-\subsection{MG7 Pipeline Description}\label{mg7-pipeline-description}
-
-\subsection{Taxonomic Assignment
+\subsubsection{Taxonomic Assignment
 Algorithms}\label{taxonomic-assignment-algorithms}
 
-\subsubsection{Lowest Common Ancestor based Taxonomic
+\paragraph{Lowest Common Ancestor based Taxonomic
 Assignment}\label{lowest-common-ancestor-based-taxonomic-assignment}
@@
 
-\subsubsection{Best BLAST hit taxonomic
+\paragraph{Best BLAST hit taxonomic
 assignment}\label{best-blast-hit-taxonomic-assignment}
@@
 
-\subsection{MG7 availability}\label{mg7-availability}
+\subsection{Availability}\label{availability}
 
@@
   input and output data, their locations and type are expressible and
-  checked at compile-time using our Scala library \emph{datasets} ­3.
-  management of dependencies and machine configurations using our Scala
-  library \emph{Statika}
+  checked at compile-time using \emph{Datasets} ­3. management of
+  dependencies and machine configurations using \emph{Statika}
 \item
@@
 \end{enumerate}
+
+We will expand on each item in the following sections.
 
@@
 
-Other possible uses of the general schema: statika, loquat, \ldots{}
-
 \subsubsection{Shotgun metagenomics}\label{shotgun-metagenomics}
@@
 
+\section{Materials and Methods}\label{materials-and-methods}
+
+\subsection{Amazon Web Services}\label{amazon-web-services}
+
+We use EC2, S3 and SQS through a Scala wrapper of the official
+\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK},
+\href{https://github.com/ohnosequences/aws-scala-tools/???}{ohnosequences/aws-scala-tools
+0.13.2}. This uses version \texttt{1.10.9} of the AWS Java SDK.
+
+\subsection{Scala}\label{scala}
+
+MG7 itself and all the libraries used are written in Scala
+\texttt{2.11}.
+
+\subsection{Statika}\label{statika}
+
+MG7 uses
+\href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
+2.0.0} for specifying the configuration and behavior of EC2 instances.
+
+\subsection{Datasets}\label{datasets}
+
+MG7 uses
+\href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
+0.2.0} for specifying input and output data, their type and their
+location.
+
+\subsection{Loquat}\label{loquat}
+
+MG7 uses
+\href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
+2.0.0} for the specification of data processing tasks and their
+execution using AWS resources.
+
+\subsection{BLAST eDSL}\label{blast-edsl}
+
+MG7 uses
+\href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
+0.2.0}. The BLAST version used is \texttt{2.2.31+}
+
+\subsection{FLASH eDSL}\label{flash-edsl}
+
+MG7 uses
+\href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
+0.1.0}. The FLASH version used is \texttt{?.?.?}
+
+\subsection{Bio4j}\label{bio4j}
+
+MG7 uses
+\href{https://github.com/bio4j/bio4j/releases/tag/v0.12.0-RC3}{bio4j/bio4j
+0.12.0-RC3} and
+\href{https://github.com/bio4j/bio4j-titan/releases/tag/v0.4.0-RC2}{bio4j/bio4j-titan
+0.4.0-RC2} as an API for the NCBI taxonomy.
+
 \section{Acknowledgements}\label{acknowledgements}
 
-INTERCROSSING (Grant 289974)
+The two first authors are funded by INTERCROSSING (Grant 289974).
 

### 2015-09-13 src/main/scala/metagenomica/configuration.scala
Moved flash data declaratinos to the global data config

@@ -15,25 +15,75 @@ case object configuration {
   implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
     new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
 
-  trait MetagenomicaData {
+  trait AnyMetagenomicaData {
 
     // Paired end reads as input
     type ReadsType <: AnyReadsType { type EndType = pairedEndType }
-    val  readsType: ReadsType
+    val readsType: ReadsType
 
-    case object reads1 extends PairedEnd1Fastq(readsType, "reads1.fastq.gz"); type Reads1 = reads1.type
-    case object reads2 extends PairedEnd2Fastq(readsType, "reads2.fastq.gz"); type Reads2 = reads2.type
+    type Reads1 <: AnyPairedEnd1Fastq { type DataType = ReadsType }
+    val reads1: Reads1
 
-    // TODO: leave free additional modifications
+    type Reads2 <: AnyPairedEnd2Fastq { type DataType = ReadsType }
+    val reads2: Reads2
+
+    type FlashInput = Reads1 :^: Reads2 :^: DNil
+    val  flashInput: FlashInput
+
+    type Merged <: MergedReads[ReadsType, Reads1, Reads2]
+    val  merged: Merged
+
+    type Stats <: MergedReadsStats[Merged]
+    val stats: Stats
+
+    type FlashOutput = Merged :^: Stats :^: DNil
+    val  flashOutput: FlashOutput
+
+  }
+
+  class MetagenomicaData [
+    RT <: AnyReadsType { type EndType = pairedEndType }
+  ](val readsType: RT) extends AnyMetagenomicaData {
+
+    type ReadsType = RT
+
+    case object reads1_ extends PairedEnd1Fastq(readsType, "reads1.fastq.gz")
+    case object reads2_ extends PairedEnd2Fastq(readsType, "reads2.fastq.gz")
+
+    type Reads1 = reads1_.type
+    val  reads1 = reads1_
+
+    type Reads2 = reads2_.type
+    val  reads2 = reads2_
+
+    val  flashInput = reads1 :^: reads2 :^: DNil
+
     lazy val flashOptions = flash.defaults update (
-      read_len(readsType.length.toInt)    :~:
+      read_len(readsType.length.toInt)   :~:
       max_overlap(readsType.length.toInt) :~: ∅
     )
 
[…]

### 2015-09-13 src/main/scala/metagenomica/loquats/flash/config.scala

@@ -22,16 +22,16 @@ import java.io.File
 
 
 
-trait AnyFlashConfig extends Era7LoquatConfig { config =>
-  type Data <: AnyFlashData
-  // val  data: Data
-
-  type DataProcessing <: FlashDataProcessing[Data]
-  val  dataProcessing: DataProcessing
-
-  type FlashDataMappring = AnyDataMapping { type DataProcessing = config.DataProcessing }
-  val  dataMappings: List[FlashDataMappring]
-}
+// trait AnyFlashConfig extends Era7LoquatConfig { config =>
+//   type Data <: AnyFlashData
+//   // val  data: Data
+//
+//   type DataProcessing <: FlashDataProcessing[Data]
+//   val  dataProcessing: DataProcessing
+//
+//   type FlashDataMappring = AnyDataMapping { type DataProcessing = config.DataProcessing }
+//   val  dataMappings: List[FlashDataMappring]
+// }
 
 // abstract class FlashConfig[D <: AnyFlashData](val dataProcessing: D) extends AnyFlashConfig {
 //

### 2015-09-13 src/main/scala/metagenomica/loquats/flash/data.scala

@@ -1,60 +1,0 @@
-package ohnosequences.metagenomica.loquats.flash
-
-import ohnosequences.loquat._, utils._, dataProcessing._
-import ohnosequences.statika._, bundles._, instructions._
-import ohnosequences.flash.api._
-import ohnosequences.flash.data._
-import ohnosequences.cosas._, typeSets._, types._
-import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-import ohnosequencesBundles.statika.Flash
-
-
-trait AnyFlashData {
-
-  // Paired end reads as input
-  type ReadsType <: AnyReadsType { type EndType = pairedEndType }
-  val readsType: ReadsType
-
-  type Reads1 <: AnyPairedEnd1Fastq { type DataType = ReadsType }
-  val reads1: Reads1
-
-  type Reads2 <: AnyPairedEnd2Fastq { type DataType = ReadsType }
-  val reads2: Reads2
-
-  type Merged <: MergedReads[ReadsType, Reads1, Reads2]
-  val  merged: Merged
-
-  type Stats <: MergedReadsStats[Merged]
-  val stats: Stats
-
-  lazy val flashOptions = flash.defaults update (
-    read_len(readsType.length.toInt)   :~:
-    max_overlap(readsType.length.toInt) :~: ∅
-  )
-
-}
-
-
-class FlashData [
-  RT <: AnyReadsType { type EndType = pairedEndType }
-](val readsType: RT) extends AnyFlashData {
-
-  type ReadsType = RT
-
-  case object reads1_ extends PairedEnd1Fastq(readsType, "reads1.fastq.gz")
-  case object reads2_ extends PairedEnd2Fastq(readsType, "reads2.fastq.gz")
-
-  type Reads1 = reads1_.type
-  val  reads1 = reads1_
-
-  type Reads2 = reads2_.type
-  val  reads2 = reads2_
-
-  case object merged_ extends MergedReads(readsType, reads1, reads2, flashOptions)
-  type Merged = merged_.type
-  val  merged = merged_
-
-  case object stats_ extends MergedReadsStats[Merged](merged)
-  type Stats = stats_.type
-  val  stats = stats_
-}

### 2015-09-13 src/main/scala/metagenomica/loquats/flash/dataProcessing.scala

@@ -19,19 +19,26 @@ import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads.
 import java.io.File
 
 
-class FlashDataProcessing[D <: AnyFlashData](val data: D)(implicit
-  parseInputFiles: ParseDenotations[(D#Reads1 :^: D#Reads2 :^: DNil)#LocationsAt[FileDataLocation], File],
-  outputFilesToMap: ToMap[(D#Merged :^: D#Stats :^: DNil)#LocationsAt[FileDataLocation], AnyData, FileDataLocation]
-  // parseInputFiles: ParseDenotations[(data.reads1.type :^: data.reads2.type :^: DNil)#LocationsAt[FileDataLocation], File],
-  // outputFilesToMap: ToMap[(data.merged.type :^: data.stats.type :^: DNil)#LocationsAt[FileDataLocation], AnyData, FileDataLocation]
+class FlashDataProcessing[
+  RT <: AnyReadsType { type EndType = pairedEndType },
+  R1 <: AnyPairedEnd1Fastq { type DataType = RT },
+  R2 <: AnyPairedEnd2Fastq { type DataType = RT },
+  M  <: MergedReads[RT, R1, R2],
+  S  <: MergedReadsStats[M]
+](val readsType: RT,
+  val reads1: R1,
+  val reads2: R2,
+  val merged: M,
+  val stats: S
+)(implicit
+  parseInputFiles: ParseDenotations[(R1 :^: R2 :^: DNil)#LocationsAt[FileDataLocation], File],
+  outputFilesToMap: ToMap[(M :^: S :^: DNil)#LocationsAt[FileDataLocation], AnyData, FileDataLocation]
 ) extends DataProcessingBundle(bundles.flash)(
-  input = (data.reads1: D#Reads1) :^: (data.reads2: D#Reads2) :^: DNil,
-  output = (data.merged: D#Merged) :^: (data.stats: D#Stats) :^: DNil
-  // input = (data.reads1 :^: data.reads2 :^: DNil): (D#Reads1 :^: D#Reads2 :^: DNil),
-  // output = (data.merged :^: data.stats :^: DNil): (D#Merged :^: D#Stats :^: DNil)
+  input = reads1 :^: reads2 :^: DNil,
+  output = merged :^: stats :^: DNil
 )(parseInputFiles, outputFilesToMap) {
 
-  def instructions: AnyInstructions = say("I'll be fast a a flash!")
[…]

### 2015-09-13 docs/paper/out/paper.tex
add the dagger of equal contribution -- paper

@@
 \def\Authors{
- Alexey Alekhin\,$^{1}$  Evdokim Kovach\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
+ Alexey Alekhin\(^\dagger\)\,$^{1}$  Evdokim Kovach\(^\dagger\)\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
 }% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 No abstract yet. Will be here.
+
+\(^\dagger\) The first and second authors contributed equally to this
+work
 

### 2015-09-13 src/main/scala/metagenomica/configuration.scala
WIP

@@ -7,6 +7,8 @@ import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads.
 import ohnosequences.flash.api._
 import ohnosequences.flash.data._
 
+import ohnosequences.blast._, api._, data._, outputFields._
+
 import java.io.File
 
 case object configuration {
@@ -17,7 +19,8 @@ case object configuration {
 
   trait AnyMetagenomicaData {
 
-    // Paired end reads as input
+    // FLASH
+
     type ReadsType <: AnyReadsType { type EndType = pairedEndType }
     val readsType: ReadsType
 
@@ -39,9 +42,23 @@ case object configuration {
     type FlashOutput = Merged :^: Stats :^: DNil
     val  flashOutput: FlashOutput
 
+
+    // BLAST
+
+    type BlastRecord <: AnyBlastOutputRecord
+    val  blastRecord: BlastRecord
+
+
+    type BlastOutput <: AnyBlastOutput
+    // {
+    //   type BlastExpressionType <: AnyBlastExpressionType {
+    //     type OutputRecord = BlastRecord
+    //   }
+    // }
+    val  blastOutput: BlastOutput
   }
 
-  class MetagenomicaData [
+  abstract class MetagenomicaData [
     RT <: AnyReadsType { type EndType = pairedEndType }
   ](val readsType: RT) extends AnyMetagenomicaData {
 
@@ -72,18 +89,14 @@ case object configuration {
     val  stats = stats_
 
     val  flashOutput = merged :^: stats :^: DNil
+
+    // case object blastExprType extends BlastExpressionType(blastn)(blastRecord)
+    // case object blastOutputType extends BlastOutputType(blastExprType, "blastn.out")
+    // case object blastOutput_ extends data.BlastOutput(blastOutputType, "blast.out.csv")
+    //
+    // type BlastOutput = blastOutput_.type
+    // val  blastOutput = blastOutput_
   }
 
-
-  // example
-  case object dt extends MetagenomicaData(illumina.PairedEnd(bp300, InsertSize(3000)))
-
-  case object flashProcessor extends loquats.flash.FlashDataProcessing(
-    dt.readsType,
-    dt.reads1,
-    dt.reads2,
-    dt.merged,
-    dt.stats
-  )
 
 }

### 2015-09-13 src/main/scala/metagenomica/loquats/assignment/config.scala

@@ -1,91 +1,91 @@
-package ohnosequences.metagenomica.loquats.assignment
-
-import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
-
-import ohnosequences.cosas._, typeSets._, types._, properties._, records._
-import ohnosequences.cosas.ops.typeSets._
-
-import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
-
-import ohnosequences.blast._
-import ohnosequences.blast.api._, outputFields._
-import ohnosequences.blast.data._
-
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.aws.api._
-import ohnosequences.statika.aws.amazonLinuxAMIs._
-
-import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
-import ohnosequences.awstools.autoscaling._
-
-import era7.project.loquats._
-
-import java.io.File
-
-import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
-
-
-abstract class AnyAssignmentConfig extends Era7LoquatConfig { config =>
-
-  // type DataProcessing = assignmentDataProcessing.type
-  // val  dataProcessing = assignmentDataProcessing: DataProcessing
-
-  type AssignmentDataMapping = AnyDataMapping { type DataProcessing = assignmentDataProcessing.type }
-  val  dataMappings: List[AssignmentDataMapping]
-}
-
-// abstract class AssignmentConfig[D <: AnyAssignmentDataProcessing](val dataProcessing: D) extends AnyAssignmentConfig {
+// package ohnosequences.metagenomica.loquats.assignment
 //
-//   type DataProcessing = D
+// import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
+//
+// import ohnosequences.cosas._, typeSets._, types._, properties._, records._
+// import ohnosequences.cosas.ops.typeSets._
+//
+// import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
+//
+// import ohnosequences.blast._
+// import ohnosequences.blast.api._, outputFields._
+// import ohnosequences.blast.data._
+//
+// import ohnosequences.statika.bundles._
+// import ohnosequences.statika.aws.api._
+// import ohnosequences.statika.aws.amazonLinuxAMIs._
+//
[…]

### 2015-09-13 src/main/scala/metagenomica/loquats/assignment/dataProcessing.scala

@@ -1,157 +1,157 @@
-package ohnosequences.metagenomica.loquats.assignment
-
-import  ohnosequences.metagenomica._
-
-import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
-
-import ohnosequences.loquat.dataProcessing._
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.instructions._
-import ohnosequencesBundles.statika.Blast
-import ohnosequences.blast._, api._, data._
-import ohnosequences.cosas._, typeSets._, types._, properties._
-import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-import java.io.{ BufferedWriter, FileWriter, File }
-import ohnosequences.fastarious._, fasta._, fastq._
-import ohnosequences.blast._, api._, data._, outputFields._
-
-import scala.util.Try
-
-
-case object dataProcessing {
-
-  // TODO: move it to the config
-  case object CSVDataType extends AnyDataType { val label = "fastq" }
-  case object lcaCSV extends Data(CSVDataType, "lca.csv")
-  case object bbhCSV extends Data(CSVDataType, "bbh.csv")
-
-  // type BlastOutput = loquats.blast.blastTest.blastOutput.type
-  val  blastOutput = loquats.blast.blastTest.blastOutput
-
-  type BlastRecord = loquats.blast.blastDataProcessing.outRec.type
-  val blastRecord: BlastRecord = loquats.blast.blastDataProcessing.outRec
-
-  val headers: Seq[String] = blastRecord.properties.mapToList(typeLabel)
-
-  // TODO: move it somewhere up for global use
-  type ID = String
-
-  type GI = ID
-  type TaxID = ID
-  type ReadID = ID
-  type NodeID = ID
-
-  type LCA = Option[NodeID]
-  type BBH = Int
-
-  // TODO: move it somewhere up for global use
-  def parseInt(str: String): Option[Int] = Try(str.toInt).toOption
-
-  // FIXME: this thing should be used globally (ideally it should be in datasets)
-  implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
-    new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
-
-
-  case object assignmentDataProcessing extends DataProcessingBundle(
-    bundles.bio4jTaxonomy,
[…]

### 2015-09-13 src/main/scala/metagenomica/loquats/blast/config.scala

@@ -1,95 +1,95 @@
-package ohnosequences.metagenomica.loquats.blast
-
-import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
-
-import ohnosequences.cosas._, typeSets._, types._, properties._, records._
-import ohnosequences.cosas.ops.typeSets._
-
-import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
-
-import ohnosequences.blast._
-import ohnosequences.blast.api._, outputFields._
-import ohnosequences.blast.data._
-
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.aws.api._
-import ohnosequences.statika.aws.amazonLinuxAMIs._
-
-import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
-import ohnosequences.awstools.autoscaling._
-
-import era7.project.loquats._
-
-import java.io.File
-
-import blastDataProcessing._
-
-
-trait AnyBlastConfig extends Era7LoquatConfig { config =>
-
-  type DataProcessing <: AnyBlastDataProcessing
-  val  dataProcessing: DataProcessing
-
-  type BlastDataMapping = AnyDataMapping { type DataProcessing = config.DataProcessing }
-  val  dataMappings: List[BlastDataMapping]
-}
-
-abstract class BlastConfig[D <: AnyBlastDataProcessing](val dataProcessing: D) extends AnyBlastConfig {
-
-  type DataProcessing = D
-}
-
-//////////////////////////////////////////////////////////////////////////////////////
-
-case object blastTest {
-
-  // TODO: move it to datasets
-  implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
-    new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
-
-  case object FastqDataType extends AnyDataType
-  case object fastqInput extends Data(FastqDataType, "foo.fastq")
-
-
-  case object blastOutput extends BlastOutput(blastOutputType, "blast.out.csv")
-
-
-  case object testDataProcessing extends BlastDataProcessing(fastqInput, blastOutput)
-
-  case object testConfig extends BlastConfig(testDataProcessing) {
-
-    val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
-
-    val managerConfig = ManagerConfig(
[…]

### 2015-09-13 src/main/scala/metagenomica/loquats/blast/dataProcessing.scala

@@ -1,12 +1,12 @@
 package ohnosequences.metagenomica.loquats.blast
 
-import  ohnosequences.metagenomica._, bundles._
+import ohnosequences.metagenomica.configuration._
+import ohnosequences.metagenomica.bundles
 
 import ohnosequences.loquat._, dataProcessing._
 
 import ohnosequences.statika.bundles._
 import ohnosequences.statika.instructions._
-import ohnosequencesBundles.statika.Blast
 
 import ohnosequences.blast._, api._, data._, outputFields._
 
@@ -23,17 +23,6 @@ import sys.process._
 
 
 case object blastDataProcessing {
-
-  case object blastBundle extends Blast("2.2.31")
-
-  import ohnosequences.statika.aws._, api._, amazonLinuxAMIs._
-  import ohnosequences.awstools.regions.Region._
-
-  case object blastCompat extends Compatible(
-    amzn_ami_64bit(Ireland, Virtualization.HVM)(1),
-    blastBundle,
-    generated.metadata.Metagenomica
-  )
 
   // TODO with great power comes great responsibility. Move to conf
   case object outRec extends BlastOutputRecord(
@@ -90,20 +79,31 @@ case object blastDataProcessing {
 
     def instructions: AnyInstructions = say("Let the blasting begin!")
 
-    // this is defined in the constructor
-    // val bundleDependencies: List[AnyBundle] = List[AnyBundle](blastBundle, blast16s)
+    val bundleDependencies: List[AnyBundle] = List[AnyBundle](
+      bundles.blast,
+      bundles.blast16s
+    )
 
+    // TODO: more precise type
     type FastqInput <: AnyData
     val  fastqInput: FastqInput
+
+
+    // FIXME: we should use this instead of blastCmd, but .cmd requires implicits "/
+    // type BlastExpr <: AnyBlastExpression
+    // val  blastExpr: BlastExpr
+    // val blastCmd: Seq[String]
 
     type BlastOutput <: AnyBlastOutput
     val  blastOutput: BlastOutput
 
-    type Input  <: FastqInput :^: DNil
-    type Output <: BlastOutput :^: DNil
 
-    // What's this?
-    // private def blastOutputFile(context: Context): File = (context / "mapping.out").javaFile
+    type Input = FastqInput :^: DNil
+    val  input = fastqInput :^: DNil: Input
+
+    type Output = BlastOutput :^: DNil
[…]

### 2015-09-13 src/main/scala/loquats/counting/dataProcessing.scala
First working version of the counting loquat; tested

@@ -1,0 +1,93 @@
+package ohnosequences.metagenomica.loquats.counting
+
+import ohnosequences.metagenomica._
+import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
+
+import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
+
+import ohnosequences.loquat.dataProcessing._
+import ohnosequences.statika.bundles._
+import ohnosequences.statika.instructions._
+import ohnosequencesBundles.statika.Blast
+import ohnosequences.blast._, api._, data._
+import ohnosequences.cosas._, typeSets._, types._, properties._
+import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
+import java.io.{ BufferedWriter, FileWriter, File }
+import ohnosequences.fastarious._, fasta._, fastq._
+import ohnosequences.blast._, api._, data._, outputFields._
+
+import scala.util.Try
+
+
+case object dataProcessing {
+
+  case object directCountsCSV extends Data(CSVDataType, "directCounts.csv")
+  case object accumulatedCountsCSV extends Data(CSVDataType, "accumulatedCounts.csv")
+
+
+  case object countingDataProcessing extends DataProcessingBundle(
+    bundles.bio4jTaxonomy
+  )(input = lcaCSV :^: DNil,
+    output = directCountsCSV :^: accumulatedCountsCSV :^: DNil
+  ) {
+
+    def instructions: AnyInstructions = say("I'm counting you!")
+
+    def processData(
+      dataMappingId: String,
+      context: Context
+    ): Instructions[OutputFiles] = {
+
+      import com.github.tototoshi.csv._
+
+      val lcaReader: CSVReader = CSVReader.open( context.file(lcaCSV).javaFile )
+
+      val directCounts: Map[TaxID, Int] = lcaReader.iterator.toStream
+        .foldLeft(Map[TaxID, Int]()) { (acc, row) =>
+          // TODO: use csv api?
+          val taxId = row(1)
+          val current: Int = acc.get(taxId).getOrElse(0)
+          acc.updated(taxId, current + 1)
+        }
+
+      lcaReader.close
+
+      // TODO: figure out some more effective algorithm
+      // TODO: split it and test separately
+      val accumulatedCounts: Map[TaxID, Int] = directCounts
+        .foldLeft(Map[TaxID, Int]()) { case (acc, (taxId, count)) =>
[…]

### 2015-09-13 src/main/scala/loquats/counting/test.scala

@@ -1,0 +1,76 @@
+package ohnosequences.metagenomica.loquats.counting
+
+import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
+
+import ohnosequences.cosas._, typeSets._, types._, properties._, records._
+import ohnosequences.cosas.ops.typeSets._
+
+import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
+
+import ohnosequences.blast._
+import ohnosequences.blast.api._, outputFields._
+import ohnosequences.blast.data._
+
+import ohnosequences.statika.bundles._
+import ohnosequences.statika.aws.api._
+import ohnosequences.statika.aws.amazonLinuxAMIs._
+
+import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
+import ohnosequences.awstools.autoscaling._
+
+import era7.project.loquats._
+
+import java.io.File
+
+import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
+import ohnosequences.metagenomica.loquats.counting.dataProcessing._
+
+
+case object countingTest {
+
+  case object testConfig extends Era7LoquatConfig {
+
+    val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
+
+    val managerConfig = ManagerConfig(
+      instanceType = m3_medium,
+      purchaseModel = SpotAuto
+    )
+
+    val workersConfig = WorkersConfig(
+      instanceType = m3_medium,
+      purchaseModel = SpotAuto,
+      groupSize = WorkersGroupSize(0, 1, 10)
+    )
+
+    val terminationConfig = TerminationConfig(
+      terminateAfterInitialDataMappings = true
+    )
+
+    val dataMappings: List[AnyDataMapping] =
+      List(
+        DataMapping(
+          "ERR567374_1",
+          countingDataProcessing
+        )(remoteInput =
+            lcaCSV.atS3(ObjectAddress(
+              "resources.ohnosequences.com",
+              "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.lca.csv"
+            )) :~: ∅,
+          remoteOutput =
+            directCountsCSV.atS3(ObjectAddress(
+              "resources.ohnosequences.com",
+              "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.counts.direct.csv"
+            )) :~:
[…]

### 2015-09-14 src/main/scala/loquats/assignment/dataProcessing.scala
Fixed the bbh bug; tested

@@ -42,10 +42,14 @@ case object dataProcessing {
   type NodeID = ID
 
   type LCA = Option[NodeID]
-  type BBH = Int
+  type BBH = Option[NodeID]
 
   // TODO: move it somewhere up for global use
   def parseInt(str: String): Option[Int] = Try(str.toInt).toOption
+
+  // this method looks up particular column by its header
+  def column(row: Seq[String], header: AnyOutputField): Option[String] =
+    headers.zip(row).toMap.get(header.label)
 
   // FIXME: this thing should be used globally (ideally it should be in datasets)
   implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
@@ -60,27 +64,6 @@ case object dataProcessing {
   ) {
 
     def instructions: AnyInstructions = say("Let's see who is who!")
-
-    // val bundleDependencies: List[AnyBundle] = List[AnyBundle](
-    //   bundles.bio4jTaxonomy,
-    //   bundles.filteredGIs
-    // )
-
-    // TODO: this should be setup from the global config
-    // type BlastOutput <: AnyBlastOutput {
-    //   type BlastExpressionType <: AnyBlastExpressionType {
-    //     type OutputRecord = BlastRecord
-    //   }
-    // }
-    // val  blastOutput: BlastOutput
-
-    // TODO: these thing should be setup from the global config
-    // type Input = blastOutput.type :^: DNil
-    // val  input = blastOutput :^: DNil
-    //
-    // type Output = lcaCSV.type :^: bbhCSV.type :^: DNil
-    // val  output = lcaCSV :^: bbhCSV :^: DNil
-
 
     def processData(
       dataMappingId: String,
@@ -106,17 +89,19 @@ case object dataProcessing {
         case (None, _) => None
         case (Some(readId), hits) => {
 
-          // this method chooses particular column by its header
-          def column(header: AnyOutputField): List[String] =
-            hits.toList.flatMap { columns: Seq[String] =>
-              headers.zip(columns).toMap.get(header.label)
+          val bbh: BBH =
+            // this shouldn't happen, but let's be careful
+            if (hits.isEmpty) None
+            else {
+              // best blast score is just a maximum in the `bitscore` column
+              val maxRow: Seq[String] = hits.maxBy { row: Seq[String] =>
+                column(row, bitscore).flatMap(parseInt).getOrElse(0)
+              }
[…]

### 2015-09-14 src/main/scala/loquats/assignment/test.scala

@@ -25,25 +25,9 @@ import java.io.File
 import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
 
 
-abstract class AnyAssignmentConfig extends Era7LoquatConfig { config =>
-
-  // type DataProcessing = assignmentDataProcessing.type
-  // val  dataProcessing = assignmentDataProcessing: DataProcessing
-
-  type AssignmentDataMapping = AnyDataMapping { type DataProcessing = assignmentDataProcessing.type }
-  val  dataMappings: List[AssignmentDataMapping]
-}
-
-// abstract class AssignmentConfig[D <: AnyAssignmentDataProcessing](val dataProcessing: D) extends AnyAssignmentConfig {
-//
-//   type DataProcessing = D
-// }
-
-//////////////////////////////////////////////////////////////////////////////////////
-
 case object assignmenTest {
 
-  case object testConfig extends AnyAssignmentConfig {
+  case object testConfig extends Era7LoquatConfig {
 
     val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
 
@@ -62,7 +46,7 @@ case object assignmenTest {
       terminateAfterInitialDataMappings = true
     )
 
-    val dataMappings: List[AssignmentDataMapping] =
+    val dataMappings: List[AnyDataMapping] =
       List(
         DataMapping(
           "ERR567374_1",

### 2015-09-14 src/main/scala/loquats/counting/dataProcessing.scala
Fixed counting; added bbh; should work but not tested due to some amazon issues

@@ -6,6 +6,8 @@ import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
 import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
 
 import ohnosequences.loquat.dataProcessing._
+import ohnosequences.loquat.utils._
+
 import ohnosequences.statika.bundles._
 import ohnosequences.statika.instructions._
 import ohnosequencesBundles.statika.Blast
@@ -21,14 +23,54 @@ import scala.util.Try
 
 case object dataProcessing {
 
-  case object directCountsCSV extends Data(CSVDataType, "directCounts.csv")
-  case object accumulatedCountsCSV extends Data(CSVDataType, "accumulatedCounts.csv")
+  case object lcaCountsCSV extends Data(CSVDataType, "lca.counts.csv")
+  case object bbhCountsCSV extends Data(CSVDataType, "bbh.counts.csv")
 
+  // returns count of the given id and a filtered list (without that id)
+  def count(id: ID, list: List[ID]): (Int, List[ID]) =
+    list.foldLeft( (0, List[ID]()) ) { case ((count, rest), next) =>
+      if (next == id) (count + 1, rest)
+      else (count, next :: rest)
+    }
+
+  // TODO: make it tailrec
+  def directCounts(taxIds: List[TaxID]): Map[TaxID, Int] = {
+
+    @scala.annotation.tailrec
+    def rec(list: List[TaxID], acc: Map[TaxID, Int]): Map[TaxID, Int] =
+      list match{
+        case Nil => acc
+        case h :: t => {
+          val (n, rest) = count(h, t)
+          rec(rest, acc.updated(h, n + 1))
+        }
+      }
+
+    rec(taxIds, Map[TaxID, Int]())
+  }
+
+  // TODO: figure out some more effective algorithm
+  // Caution: it uses bio4j bundle!
+  def accumulatedCounts(counts: Map[TaxID, Int]): Map[TaxID, (Int, Int)] = {
+    counts.foldLeft(
+      Map[TaxID, (Int, Int)]()
+    ) { case (acc, (id, count)) =>
+      val node: Option[TitanTaxonNode] = titanTaxonNode(bundles.bio4jTaxonomy.graph, id)
+      val ancestors: Seq[AnyTaxonNode] = node.map{ n => pathToTheRoot(n, Seq()) }.getOrElse(Seq())
+
+      ancestors.foldLeft(
+        acc.updated(id, (count, 0))
+      ) { (acc, node) =>
[…]

### 2015-09-14 src/main/scala/loquats/counting/test.scala

@@ -56,15 +56,19 @@ case object countingTest {
             lcaCSV.atS3(ObjectAddress(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.lca.csv"
+            )) :~:
+            bbhCSV.atS3(ObjectAddress(
+              "resources.ohnosequences.com",
+              "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.bbh.csv"
             )) :~: ∅,
           remoteOutput =
-            directCountsCSV.atS3(ObjectAddress(
+            lcaCountsCSV.atS3(ObjectAddress(
               "resources.ohnosequences.com",
-              "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.counts.direct.csv"
+              "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.lca.counts.csv"
             )) :~:
-            accumulatedCountsCSV.atS3(ObjectAddress(
+            bbhCountsCSV.atS3(ObjectAddress(
               "resources.ohnosequences.com",
-              "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.counts.accumulated.csv"
+              "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.bbh.counts.csv"
             )) :~:
             ∅
         )

### 2015-09-14 docs/paper/out/paper.tex
remove section about sample data -- paper
add workflow description from @rtobes -- paper
See https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-140018675

@@
 
-\subsection{Pipeline Description}\label{pipeline-description}
+\subsection{Workflow Description}\label{workflow-description}
+
+The MG7 analysis workflow is summarized in Figure X. The input files for
+MG7 are the FASTQ files resulting from a paired-end NGS sequencing
+experiment.
+
+\subsubsection{Joining reads of each pair using
+FLASH}\label{joining-reads-of-each-pair-using-flash}
+
+In the first step the paired-end reads, designed with an insert size
+that yields pairs of reads with an overlapping region between them, are
+assembled using FLASH \citep{magovc2011flash}. FLASH is designed to
+merge pairs of reads when the original DNA fragments are shorter than
+twice the length of reads. Thus, the sequence obtained after joining the
+2 reads of each pair is larger and has better quality since the sequence
+at the ends of the reads is refined merging both ends in the assembly.
+To have a larger and improved sequence is crucial to do more precise the
+inference of the bacterial origin based on similarity with reference
+sequences.
+
+\subsubsection{Parallelized BLASTN of each read against the
+16S-DB7}\label{parallelized-blastn-of-each-read-against-the-16s-db7}
+
+The second step is to search for similar 16S sequences in our 16S-DB7
+database. The taxonomic assignment for each read is based on BLASTN of
+each read against the 16S database. Assignment based on direct
+similarity of each read one by one compared against a sufficiently wide
+database is a very exhaustive method for assignment
+\citep{segata2013computational} . Some methods of assignment compare the
+sequences only against the available complete bacterial genomes or avoid
+computational cost clustering or binning the sequences first, and then
+doing the assignments only for the representative sequence of each
+cluster. MG7 carries out an exhaustive comparison of all the reads under
+analysis and it does not applies any binning strategy. Every read is
+specifically compared with all the sequences of the 16S database. We
+select the best BLAST hits (10 hits by default) obtained for each read
+to do the taxonomic assignment.
 
@@
 
+All the reads are assigned under two different algorithms of assignment:
+i. Lowest Common Ancestor based taxonomic assignment (LCA) and ii. Best
+BLAST Hit based taxonomic assignment (BBH). Figure X displays
+schematically the LCA algorithm applied sensu stricto (left panel) and
+the called `in line' exception (right panel) designed in order to gain
+specificity in the assignments in the cases in which the topology of the
+taxonomical nodes corresponding to the BLAST hits support sufficiently
+the assignment to the most specific taxon.
+
 \paragraph{Lowest Common Ancestor based Taxonomic
@@
 
-For each read:
-
-­1. Select only one BLASTN alignment (HSP) per reference sequence (the
-HSP with lowest e value) 2. Filter all the HSPs with bitscore below a
-defined BLASTN bitscore threshold s\_0 3. Find the best bitscore value S
-in the set of BLASTN HSPs corresponding to hits of that read 4. Filter
-all the alignments with bitscore below p * S (where p is a fixed by the
-user coefficient to define the bitscore required, e.g.~if p=0.9 and
-S=700 the required bitscore threshold would be 630) 5. Select all the
-taxonomic nodes to which map the reference sequences involved in the
-selected HSPs: - If all the selected taxonomic nodes forms a line in the
+For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
+over a threshold of similarity (by default \(evalue \leq e^{-15}\))
+filtering those hits that are not sufficiently good comparing them with
+the best one. We select the best HSP (High Similarity Pair) per
+reference sequence and then choose the best HSP (that with lowest e
+value) between all the selected ones. The bitscore of this best HSP
+(called S) is used as reference to filter the rest of HSPs. All the HSPs
+with bitscore below p x S are filtered. p is a coefficient fixed by the
+user to define the bitscore required, e.g.~if p=0.9 and S=700 the
+required bitscore threshold would be 630. Once we have the definitive
+HSPs selected, we obtain their corresponding taxonomic nodes using the
+taxonomic assignments that NCBI provides for all the nt database
+sequences. Now we have to analyze the topological distribution of these
+nodes in the taxonomy tree: i. If all the nodes forms a line in the
 taxonomy tree (are located in a not branched lineage to the tree root)
 we should choose the most specific taxID as the final assignment for
-that read - If not, we should search for the (sensu stricto) Lowest
-Common Ancestor (LCA) of all the selected taxonomic nodes (See Figure X)
-
-In this approach the value used for evaluating the similarity is the
-bitscore that is a value that increases when similarity is higher and
-depends a lot on the length of the HSP
+that read. We call to this kind of assignment the `in line' exception
+(see Figure X right panel). ii. If not, we should search for the
+\emph{sensu stricto} Lowest Common Ancestor (LCA) of all the selected
+taxonomic nodes (See Figure X left panel). In this approach we decided
+to use the bitscore for evaluating the similarity because it is a value
+that increases when similarity is higher and depends a lot on the length
+of the HSP. Some reads could not find sequences with enough similarity
+in the database and then they would be classified as reads with no hits.
+Advanced metagenomics analysis approaches \citep{huson2013microbial}
+have adopted LCA assignment algorithms because it provides fine and
+trusted taxonomical assignment.
 
@@
 
-We have maintained the simpler method of Best BLAST Hit (BBH) taxonomic
-assignment because, in some cases, it can provide information about the
-sequences that can be more useful than the obtained using LCA algorithm.
-Using LCA algorithm when some reference sequences with BLAST alignments
-over the required thresholds map to a not sufficiently specific taxID,
-the read can be assigned to an unspecific taxon near to the root. If the
-BBH reference sequence maps to a more specific taxa this method, in that
-case, gives us useful information.
+We decided to maintain the simpler method of Best BLAST Hit (BBH) for
+taxonomic assignment because, in some cases, it can provide information
+about the sequences that adds information to that obtained using LCA
+algorithm. Using LCA algorithm, when some reference sequences with BLAST
+alignments over the required thresholds map to a not sufficiently
+specific taxID, the read can be assigned to an unspecific taxon near to
+the root of the taxonomy tree. If the BBH reference sequence maps to
+more specific taxa, this method, in that case, gives us useful
+information.
 
-\subsection{Using MG7 with some example
-data-sets}\label{using-mg7-with-some-example-data-sets}
+\subsubsection{Output for LCA and BBH
+assignments}\label{output-for-lca-and-bbh-assignments}
 
-We selected the datasets described in {[}Kennedy-2014{]} (??)
+MG7 provides independent results for the 2 different approaches, LCA and
+BBH. The output files include, for each taxonomy node (with some read
+assigned), abundance values for direct assignment and cumulative
+assignment. The abundances are provided in counts (absolute values) and
+in percentage normalized to the number of reads of each sample. Direct
+assignments are calculated counting reads specifically assigned to a
+taxonomic node, not including the reads assigned to the descendant nodes
+in the taxonomy tree. Cumulative assignments are calculated including
+the direct assignments and also the assignments of the descendant nodes.
+For each sample MG7 provides 8 kinds of abundance values: LCA direct
+counts, LCA cumu. counts, LCA direct \%, LCA cumu. \%, BBH direct
+counts, BBH cumu. counts, BBH direct \%, BBH cumu. \%.
 

### 2015-09-14 docs/paper/out/paper.tex
add Results Overview from @rtobes -- paper
see https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-140027234

@@
 
-Considering all these aspects we have designed a new open source
-methodology for analyzing metagenomics data that exploits the new
-possibilities that cloud computing offers to get a system robust,
-programmatically configurable, modular, distributed, flexible, scalable
-and traceable in which the biological databases of reference sequences
-can be easily updated and/or frequently substituted by new ones or by
-databases specifically designed for focused projects.
-
 \section{Results}\label{results}
@@
 
-To tackle the challenges posed by metagenomics big data analysis
-outlined in the Introduction
+Considering the current new metagenomics scenario and to tackle the
+challenges posed by metagenomics big data analysis outlined in the
+Introduction we have designed a new open source methodology for
+analyzing metagenomics data. It exploits the new possibilities that
+cloud computing offers to get a system robust, programmatically
+configurable, modular, distributed, flexible, scalable and traceable in
+which the biological databases of reference sequences can be easily
+updated and/or frequently substituted by new ones or by databases
+specifically designed for focused projects.
+
+These are some of the more innovative MG7 features:
 
@@
 \item
-  Definition of complex pipelines using \emph{Loquat}, a composable
+  Definition of complex workflows using \emph{Loquat}, a composable
   system for scaling/parallelizing stateless computations especially
@@
 \item
-  per-read assignment (??)
+  Exhaustive per-read taxonomic assignment using two complementary
+  assignment algorithms Lowest Common Ancestor and Best BLAST Hit
+\item
+  Using a new 16S database of reference sequences (16S-DB7) with a
+  flexible and sustainable system of updating and project-driven
+  customization
 \end{itemize}

### 2015-09-14 docs/paper/out/paper.tex
correct wording in Results/Scala -- paper
it was basically unreadable, a bit better now

@@
 
-In this section we describe the resources and Scala libraries developed
-by the authors on top of which MG7 is built.
-
+In this section we describe the resources and libraries developed by the
+authors on top of which MG7 is built. All MG7 code is written in
 \href{http://www.scala-lang.org/}{Scala}, a hybrid object-functional
-programming language, was chosen based on the possibility of using
+programming language. Scala was chosen based on the possibility of using
 certain advanced programming styles, and Java interoperability, which

### 2015-09-14 docs/paper/out/paper.tex
add analysis data software project section, fix acknowledgements -- paper
add ideas for data analysis as code -- paper
help needed! see https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-140128079

@@
 
+\subsection{Data analysis as a software
+project}\label{data-analysis-as-a-software-project}
+
+\textbf{IDEAS}
+
+\begin{itemize}
+\tightlist
+\item
+  the user needs to write code, a repo would be nice, AWS account
+\item
+  there's some sort of conf required for the AWS account (add users,
+  roles, permissions, whatever)
+\item
+  Why this is a good thing (or should this just be in Discussion)
+\item
+  something else?
+\end{itemize}
+
 \subsection{Availability}\label{availability}
@@
 
-The two first authors are funded by INTERCROSSING (Grant 289974).
+\emph{Funding:} The two first authors are funded by INTERCROSSING (Grant
+289974).
 

### 2015-09-14 docs/paper/out/paper.tex
reword first part of intro -- paper

@@
 
-Metagenomics data analysis is growing at exponential rate during the
-last years. The increasing throughput of massively parallel sequencing
-technologies, the derived decreasing cost, and the high impact of
+During the past few years, metagenomics data analysis is growing
+exponentially. Some of the reasons behind this are the increasing
+throughput of massively parallel sequencing technologies (with the
+derived decrease in sequencing costs), and the wide impact of
 metagenomics studies, especially in human health (diagnostics,
-treatments, drug response, prevention), are crucial reasons responsible
-for this growth of Metagenomics. There is a growing interest in
-sequencing all kind of microbiomes (gut, mouth, skin, urinary tract,
-airway, milk, bladder), in different conditions of health and disease,
-or after different treatments. Metagenomics is also impacting
-environmental sciences, crop sciences, agrifood sector and biotechnology
-in general. This new possibilities for exploring the diversity of
-micro-organisms in the most diverse environments is opening many new
-research areas but, due to this wide interest, it is expected that the
-amount of data will be overwhelming in the short time
-\citep{stephens2015big}.
+treatments, drug response or prevention). We should also mention what
+could be called the microbiome explosion: all kind of microbiomes (gut,
+mouth, skin, urinary tract, airway, milk, bladder) are now routinely
+sequenced in different conditions of health and disease, or after
+different treatments. The impact of Metagenomics is also being felt in
+environmental sciences, crop sciences, the agrifood sector and
+biotechnology in general. These new possibilities for exploring the
+diversity of micro--organisms in the most varied environments are
+opening new research areas, and drastically changing the existing ones.
+
+As a consequence, the challenge is thus moving (as in other fields) from
+data acquisition to data analysis: the amount of data is expected to be
+overwhelming in a very short time \citep{stephens2015big}.
 

### 2015-09-14 docs/paper/out/paper.tex
fix Nature News ref in intro -- paper
make minor edits in first part of intro -- paper

@@
 
-During the past few years, metagenomics data analysis is growing
+During the past decade, metagenomics data analysis is growing
 exponentially. Some of the reasons behind this are the increasing
@@
 Genome researchers have raised the alarm over big data in the past
-\href{http://www.nature.com/news/genome-researchers-raise-alarm-over-big-data-1.17912}{nature
-news add ref} but even a more serious challenge might be faced with the
-metagenomics boom/ upswing. If we compare metagenomics data with other
-genomics data used in clinical genotyping we find a differential
+\citep{hayden2015genome}, but even a more serious challenge might be
+faced with the metagenomics boom. If we compare metagenomics data with
+other genomics data used in clinical genotyping we find a differential
 feature: the key role of time. Thus, for example, in some longitudinal
@@
 years) is being used for the follow up of some intestinal pathologies,
-for studying the evolution of gut microbiome after antibiotic treatment,
-or for colon cancer early detection \citep{zeller2014potential}. This
-need of sampling across time adds more complexity to metagenomics data
-storage and demands adapted algorithms to detect state variations across
-time as well as idiosyncratic commonalities of the microbiome of each
-individual \citep{franzosa2015identifying}. In addition to the
-intra-individual sampling-time dependence, metagenomic clinical test
-results vary depending on the specific region of extraction of the
-clinical specimen. This local variability adds complexity to the
-analysis since different localizations (different tissues, different
-anatomical regions, healthy or tumour tissues) are required to have a
-sufficiently complete landscape of the human microbiome. Moreover,
-reanalysis of old samples using new tools and better reference databases
-might be also demanded from time to time.
+for studying the evolution of the gut microbiome after antibiotic
+treatment, or for colon cancer early detection
+\citep{zeller2014potential}. This need of sampling across time adds more
+complexity to metagenomics data storage and demands adapted algorithms
+to detect state variations across time as well as idiosyncratic
+commonalities of the microbiome of each individual
+\citep{franzosa2015identifying}. In addition to the intra-individual
+sampling-time dependence, metagenomic clinical test results vary
+depending on the specific region of extraction of the clinical specimen.
+This local variability adds complexity to the analysis since different
+localizations (different tissues, different anatomical regions, healthy
+or tumour tissues) are required to have a sufficiently complete
+landscape of the human microbiome. Moreover, reanalysis of old samples
+using new tools and better reference databases might be also demanded
+from time to time.
 

### 2015-09-14 docs/paper/out/paper.tex
add my fixes/enhancements to the Introduction -- paper

@@
 or tumour tissues) are required to have a sufficiently complete
-landscape of the human microbiome. Moreover, reanalysis of old samples
+landscape of the human microbiome. Moreover, re--analysis of old samples
 using new tools and better reference databases might be also demanded
@@
 
-During the last years other sciences as astronomy or particle physics
-are facing the big data challenge but, at least, these science have
-standards for data processing \citep{stephens2015big}. Global standards
-for converting raw sequence data into processed data are not yet well
-defined in metagenomics and there are shortcomings derived from the fact
-that many bioinformatics methodologies currently used for metagenomics
-data analysis were designed for a scenario very different that the
-current one. These are some of the aspects that have suffered crucial
-changes and advances with a direct impact in metagenomics data analysis:
+Other disciplines such as astronomy or particle physics have faced the
+big data challenge before. A key difference is the existence of
+standards for data processing \citep{stephens2015big}; in metagenomics
+global standards for converting raw sequence data into processed data
+are not yet well defined, and there are shortcomings derived from the
+fact that most bioinformatics methodologies used for metagenomics data
+analysis were designed for scenarios very different from the current
+one. These are some of the aspects that have suffered crucial changes
+and advances with a direct impact in metagenomics data analysis:
 
@@
 \item
-  The first aspect is related to the sequences to be analyzed: the reads
-  are larger, the sequencing depth and the number of samples of each
-  project are considerably bigger. The first metagenomics studies were
-  very local projects, while nowadays the most fruitful studies are done
-  at a global level (international, continental, national). This kind of
-  global studies has yielded the discovery of clinical biomarkers for
-  diseases of the importance of cancer, obesity or inflammatory bowel
-  diseases and has allowed exploring the biodiversity in many earth
-  environments
+  \textbf{Sequence data} the reads are larger, the sequencing depth and
+  the number of samples of each project are considerably bigger. The
+  first metagenomics studies were very local projects, while nowadays
+  the most fruitful studies are done at a global level (international,
+  continental, national). This kind of global studies has yielded the
+  discovery of clinical biomarkers for diseases of the importance of
+  cancer, obesity or inflammatory bowel diseases and has allowed
+  exploring the biodiversity of varied earth environments.
 \item
-  The second aspect derives from the impressive genomics explosion, its
-  effect being felt in this case in the reference sequences. The immense
-  amount of sequences available in public repositories demands new
-  approaches in curation, update and storage for metagenomics reference
-  databases: current models will or already have problems to face the
-  future avalanche of metagenomic sequences.
+  \textbf{The genomics explosion} its effect being felt in this case in
+  the reference sequences. The immense amount of sequences available in
+  public repositories demands new strategies for curation, update and
+  storage of metagenomics reference databases: current models will
+  (already) have problems to face the future avalanche of metagenomic
+  sequence data.
 \item
-  The third aspect to consider for metagenomics data analysis is related
-  to the appearance of new models for massive computation and storage
-  such as the so-called cloud, or the widespread adoption of programming
-  methodologies like functional programming, or, more speculatively,
-  dependently typed programming. The immense new possibilities that
-  these advances offer must have a direct impact in metagenomics data
-  analysis.
+  \textbf{Cloud computing} the appearance of new models for massive
+  computation and storage such as the so-called cloud, or the widespread
+  adoption of programming methodologies like functional programming, or,
+  more speculatively, dependently typed programming. The immense new
+  possibilities that these advances offer must have a direct impact in
+  metagenomics data analysis.
 \item
-  And finally the new social manner to do science, and especially
-  genomic science is the fourth aspect to consider. Metagenomics evolves
-  in a social and global scenario following a science democratization
-  trend in which many small research groups from distant countries share
-  a common big metagenomics project. This global cooperation demands
-  systems allowing following exactly the same pipelines using equivalent
-  cloud resources to modularly execute the analysis in an asynchronous
-  way of working between different groups. This new scenario calls for
-  new methods and tools to handle the current and future volume of
-  metagenomic data with the sufficient speed of analysis.
+  \textbf{Open science} the new social manner to do science,
+  particularly so in the case of genomics, brings its own set of
+  requirements. Metagenomics evolves in a social and global scenario
+  following a science democratization trend in which many small research
+  groups from distant countries share a common big metagenomics project;
+  this global cooperation demands systems allowing for reproducible data
+  analysis, data interoperability, and tools and practices for
+  asynchronous collaboration between different groups.
 \end{enumerate}

### 2015-09-14 docs/paper/out/paper.tex
add : to def lists -- paper

@@
 
-No abstract yet. Will be here.
+The exponential growth of metagenomics is adding a significant plus of
+complexity to the big data problem in genomics. In this new scenario
+impacted by the wide scale and scope of the projects and by the
+explosion of sequence data to be analyzed is especially opportune the
+use of new possibilities that cloud computing approaches, new functional
+and dependently typed programming languages and new database paradigms
+as graph databases offer. To tackle the challenges of big data analysis
+in this work we have used these new means to design and develop a new
+open source methodology for analyzing metagenomics data, MG7. It
+exploits the new possibilities that cloud computing offers to get a
+system robust, programmatically configurable, modular, distributed,
+flexible, scalable and traceable in which the biological databases of
+reference sequences can be easily updated and / or frequently
+substituted by new ones or by databases specifically designed for
+focused projects. MG7 uses parallelization and distributed analysis
+based on AWS, with on-demand infrastructure as the basic paradigm and
+allow the definition of complex workflows using a composable system for
+scaling/parallelizing stateless computations especially designed for
+Amazon Web Services (AWS) that counts with a static reproducible
+specification of dependencies and behavior of the different components.
+The modelling of the taxonomy tree is done using the new paradigm of
+graph databases of Bio4j that facilitates the taxonomic assignment tasks
+and the calculation of the taxa abundance values considering the
+hierarchic structure of taxonomy tree. MG7 includes the new 16S database
+16S-DB7 built with a flexible and sustainable system of updating and
+project-driven personalization.
 
@@
 \begin{enumerate}
-\def\labelenumi{\roman{enumi}.}
+\def\labelenumi{\arabic{enumi}.}
 \tightlist
 \item
-  \textbf{Sequence data} the reads are larger, the sequencing depth and
+  \textbf{Sequence data:} the reads are larger, the sequencing depth and
   the number of samples of each project are considerably bigger. The
@@
 \item
-  \textbf{The genomics explosion} its effect being felt in this case in
+  \textbf{The genomics explosion:} its effect being felt in this case in
   the reference sequences. The immense amount of sequences available in
@@
 \item
-  \textbf{Cloud computing} the appearance of new models for massive
+  \textbf{Cloud computing:} the appearance of new models for massive
   computation and storage such as the so-called cloud, or the widespread
@@
 \item
-  \textbf{Open science} the new social manner to do science,
-  particularly so in the case of genomics, brings its own set of
-  requirements. Metagenomics evolves in a social and global scenario
-  following a science democratization trend in which many small research
-  groups from distant countries share a common big metagenomics project;
-  this global cooperation demands systems allowing for reproducible data
-  analysis, data interoperability, and tools and practices for
-  asynchronous collaboration between different groups.
+  \textbf{Open science:} the new social manner to do science,
+  particularly so in genomics, brings its own set of requirements.
+  Metagenomics evolves in a social and global scenario following a
+  science democratization trend in which many small research groups from
+  distant countries share a common big metagenomics project; this global
+  cooperation demands systems allowing for reproducible data analysis,
+  data interoperability, and tools and practices for asynchronous
+  collaboration between different groups.
 \end{enumerate}

### 2015-09-14 docs/paper/out/paper.tex
remove much unneeded section in discussion -- paper
fix the bio4j subsec in discussion -- paper

@@
 
-\subsection{What MG7 brings}\label{what-mg7-brings}
-
 We could summarize the most innovative ideas and developments in MG7:
@@
 The hierarchic structure of the taxonomy of the living organisms is a
-tree, and, hence, is also a graph in which each node, with the exception
-of the root node, has a unique parent node. It led us to model the
-taxonomy tree as a graph using the graph database paradigm. Previously
-we developed Bio4j \textbf{{[}Pareja-Tobes-2015{]}}, a platform for the
-integration of semantically rich biological data using typed graph
-models. It integrates most publicly available data linked with sequences
-into a set of interdependent graphs to be used for bioinformatics
-analysis and especially for biological data.
+tree, and, hence, is also a graph in which every node but the root has a
+unique parent node. It led us to model the taxonomy tree as a graph. For
+that, Bio4j \citep{pareja2015bio4j}, a platform for the integration of
+semantically rich biological data using typed graph models, provides a
+perfect match. It integrates most publicly available data linked with
+sequences into a set of interdependent graphs to be used for
+bioinformatics analysis.
 

### 2015-09-14 docs/paper/out/paper.tex
fix references to figures -- paper

@@
 
-The MG7 analysis workflow is summarized in Figure X. The input files for
+The MG7 analysis workflow is summarized in Figure 1. The input files for
 MG7 are the FASTQ files resulting from a paired-end NGS sequencing
@@
 i. Lowest Common Ancestor based taxonomic assignment (LCA) and ii. Best
-BLAST Hit based taxonomic assignment (BBH). Figure X displays
+BLAST Hit based taxonomic assignment (BBH). Figure 2 displays
 schematically the LCA algorithm applied sensu stricto (left panel) and
@@
 that read. We call to this kind of assignment the `in line' exception
-(see Figure X right panel). ii. If not, we should search for the
+(see Figure 2 right panel). ii. If not, we should search for the
 \emph{sensu stricto} Lowest Common Ancestor (LCA) of all the selected
-taxonomic nodes (See Figure X left panel). In this approach we decided
+taxonomic nodes (See Figure 2 left panel). In this approach we decided
 to use the bitscore for evaluating the similarity because it is a value

### 2015-09-14 docs/paper/out/paper.tex

--- 2015-09-14 docs/paper/out/paper.tex
@@
 bundle can \emph{depend} on other bundles, and when being executed by an
-environment, its DAG of dependencies is linearized and run in sequence.
-In our use, bundles correspond to what an EC2 instance should do and an
-environment to an image (AMI: \textbf{A}mazon \textbf{M}achine
-\textbf{I}mage) which prepares the basic configuration, downloads the
-Scala code and runs it.
+environment, its DAG (Directed Acyclic Graph) of dependencies is
+linearized and run in sequence. In our use, bundles correspond to what
+an EC2 instance should do and an environment to an AMI (Amazon Machine
+Image) which prepares the basic configuration, downloads the Scala code
+and runs it.
 
@@
 developed by the first, second and last authors designed for the
-execution of embarrassingly parallel tasks using S3, SQS and EC2.
+execution of embarrassingly parallel tasks using S3, SQS and EC2 Amazon
+services.
 
-A \textbf{loquat} executes a process with explicit input and output
+A \emph{loquat} executes a process with explicit input and output
 datasets (declared using the \emph{Datasets} library described above).
@@
 
-Both worker and manager instances are Statika bundles. In the case of
-the worker, it can declare any dependencies needed to perform its task:
-other tools, libraries, or data.
+Both worker and manager instances are \emph{Statika} bundles. In the
+case of the worker, it can declare any dependencies needed to perform
+its task: other tools, libraries, or data.
 
@@
 
-Generic type-safe parsers returning an heterogeneous record of BLAST
+Generic type-safe parsers returning a heterogeneous record of BLAST
 output fields are also available, together with output data defined
@@
 options which yielded that output. This let us provide generic parsers
-for BLAST output which are guaranteed to be correct, for example.
+for BLAST output which are guaranteed to be correct.
 
-In the same spirit as for BLAST, we implemented a type-safe EDSL for
-FLASH expressions and their execution, sporting features equivalent to
-those outlined for the BLAST EDSL.
+In the same spirit as for BLAST, we implemented a type-safe eDSL for
+FLASH expressions and their execution, supporting features equivalent to
+those outlined for the BLAST eDSL.
 
@@
 \href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK},
-\href{https://github.com/ohnosequences/aws-scala-tools/???}{ohnosequences/aws-scala-tools
-0.13.2}. This uses version \texttt{1.10.9} of the AWS Java SDK.
+\href{https://github.com/ohnosequences/aws-scala-tools/releases/tag/v0.13.2}{ohnosequences/aws-scala-tools
+\texttt{0.13.2}}. This uses version \texttt{1.9.25} of the AWS Java SDK.
 
@@
 \href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
-2.0.0} for specifying the configuration and behavior of EC2 instances.
+\texttt{2.0.0}} for specifying the configuration and behavior of EC2
+instances.
 
@@
 \href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
-0.2.0} for specifying input and output data, their type and their
-location.
+\texttt{0.2.0}} for specifying input and output data, their type and
+their location.
 
@@
 \href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
-2.0.0} for the specification of data processing tasks and their
+\texttt{2.0.0}} for the specification of data processing tasks and their
 execution using AWS resources.
@@
 \href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
-0.2.0}. The BLAST version used is \texttt{2.2.31+}
+\texttt{0.2.0}}. The BLAST version used is \texttt{2.2.31+}
 
@@
 \href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
-0.1.0}. The FLASH version used is \texttt{?.?.?}
+\texttt{0.1.0}}. The FLASH version used is \texttt{1.2.11}
 

### 2015-09-14 docs/paper/out/paper.tex

@@
 \section{Discussion}\label{discussion}
-
-\subsection{What MG7 brings}\label{what-mg7-brings}
 

### 2015-09-14 docs/paper/out/paper.tex
add moar refs, minor edits -- paper

--- 2015-09-14 docs/paper/out/paper.tex
@@
 allow the definition of complex workflows using a composable system for
-scaling/parallelizing stateless computations especially designed for
-Amazon Web Services (AWS) that counts with a static reproducible
-specification of dependencies and behavior of the different components.
-The modelling of the taxonomy tree is done using the new paradigm of
-graph databases of Bio4j that facilitates the taxonomic assignment tasks
-and the calculation of the taxa abundance values considering the
-hierarchic structure of taxonomy tree. MG7 includes the new 16S database
-16S-DB7 built with a flexible and sustainable system of updating and
+scaling/parallelizing stateless computations designed for Amazon Web
+Services (AWS) that counts with a static reproducible specification of
+dependencies and behavior of the different components. The modelling of
+the taxonomy tree is done using the new paradigm of graph databases of
+Bio4j that facilitates the taxonomic assignment tasks and the
+calculation of the taxa abundance values considering the hierarchic
+structure of taxonomy tree. MG7 includes the new 16S database 16S-DB7
+built with a flexible and sustainable system of updating and
 project-driven personalization.
@@
 derived decrease in sequencing costs), and the wide impact of
-metagenomics studies, especially in human health (diagnostics,
-treatments, drug response or prevention). We should also mention what
-could be called the microbiome explosion: all kind of microbiomes (gut,
-mouth, skin, urinary tract, airway, milk, bladder) are now routinely
-sequenced in different conditions of health and disease, or after
-different treatments. The impact of Metagenomics is also being felt in
-environmental sciences, crop sciences, the agrifood sector and
-biotechnology in general. These new possibilities for exploring the
-diversity of micro--organisms in the most varied environments are
-opening new research areas, and drastically changing the existing ones.
+metagenomics studies \citep{oulas2015metagenomics}, especially in human
+health \citep{bikel2015combining}. We should also mention what could be
+called the microbiome explosion: all kind of microbiomes (gut, mouth,
+skin, urinary tract, airway, milk, bladder) are now routinely sequenced
+in different conditions of health and disease, or after different
+treatments. The impact of Metagenomics is also being felt in
+environmental sciences \citep{ufarte2015discovery}, crop sciences, the
+agrifood sector \citep{coughlan2015biotechnological} and biotechnology
+in general \citep[\citet{kodzius2015marine}]{cowan2015metagenomics}.
+These new possibilities for exploring the diversity of micro--organisms
+in the most varied environments are opening new research areas, and
+drastically changing the existing ones.
 
@@
 feature: the key role of time. Thus, for example, in some longitudinal
-studies, serial sampling of the same patient along several weeks (or
-years) is being used for the follow up of some intestinal pathologies,
-for studying the evolution of the gut microbiome after antibiotic
-treatment, or for colon cancer early detection
-\citep{zeller2014potential}. This need of sampling across time adds more
-complexity to metagenomics data storage and demands adapted algorithms
-to detect state variations across time as well as idiosyncratic
-commonalities of the microbiome of each individual
+studies, serial sampling from the same patient
+\citep{faust2015metagenomics} along several weeks (or years) is being
+used for the follow up of some intestinal pathologies, for studying the
+evolution of the gut microbiome after antibiotic treatment, or for colon
+cancer early detection
+\citep[\citet{garrett2015cancer}]{zeller2014potential}. This need of
+sampling across time adds more complexity to metagenomics data storage
+and demands adapted algorithms to detect state variations across time as
+well as idiosyncratic commonalities of the microbiome of each individual
 \citep{franzosa2015identifying}. In addition to the intra-individual
@@
 
-Both worker and manager instances are \emph{Statika} bundles. In the
-case of the worker, it can declare any dependencies needed to perform
-its task: other tools, libraries, or data.
+Both worker and manager instances are \emph{Statika} bundles. The worker
+can declare any dependencies needed to perform its task: other tools,
+libraries, or data.
 
@@
 declared statically, the specification of a loquat being ultimately a
-Scala object. There are deploy and resource management methods, making
-it easy to use an existing loquat either as a library or from (for
-example) a Scala REPL.
+Scala object. Deploy and resource management methods make easy to use an
+existing loquat either as a library or from (for example) a Scala REPL.
 

### 2015-09-14 docs/paper/out/paper.tex
various fixes, add refs etc -- paper

@@
 database is a very exhaustive method for assignment
-\citep{segata2013computational} . Some methods of assignment compare the
-sequences only against the available complete bacterial genomes or avoid
-computational cost clustering or binning the sequences first, and then
-doing the assignments only for the representative sequence of each
-cluster. MG7 carries out an exhaustive comparison of all the reads under
-analysis and it does not applies any binning strategy. Every read is
-specifically compared with all the sequences of the 16S database. We
-select the best BLAST hits (10 hits by default) obtained for each read
-to do the taxonomic assignment.
+\citep[\citet{morgan2012chapter}]{segata2013computational}. Some methods
+of assignment compare the sequences only against the available complete
+bacterial genomes or avoid computational cost clustering or binning the
+sequences first, and then doing the assignments only for the
+representative sequence of each cluster. MG7 carries out an exhaustive
+comparison of all the reads under analysis and it does not applies any
+binning strategy. Every read is specifically compared with all the
+sequences of the 16S database. We select the best BLAST hits (10 hits by
+default) obtained for each read to do the taxonomic assignment.
 
@@
 
-\subsection{Future-proof}\label{future-proof}
-
-\subsection{MG7 Future developments}\label{mg7-future-developments}
+\subsection{Future developments}\label{future-developments}
 
@@
 
-\subsubsection{Comparison of groups of
-samples}\label{comparison-of-groups-of-samples}
+\subsubsection{Comparing groups of
+samples}\label{comparing-groups-of-samples}
 
-\subsubsection{Interactive visualizations using the output files of MG7
-(Biographika
-project)}\label{interactive-visualizations-using-the-output-files-of-mg7-biographika-project}
+\subsubsection{Interactive visualizations based on
+Biographika}\label{interactive-visualizations-based-on-biographika}
 

### 2015-09-14 docs/paper/out/paper.tex
minor fixes in intro -- paper

@@
 metagenomics studies \citep{oulas2015metagenomics}, especially in human
-health \citep{bikel2015combining}. We should also mention what could be
-called the microbiome explosion: all kind of microbiomes (gut, mouth,
-skin, urinary tract, airway, milk, bladder) are now routinely sequenced
-in different conditions of health and disease, or after different
+health (diagnostics, treatments, drug response or prevention)
+\citep{bikel2015combining}. We should also mention what could be called
+the microbiome explosion: all kind of microbiomes (gut, mouth, skin,
+urinary tract, airway, milk, bladder) are now routinely sequenced in
+different conditions of health and disease, or after different
 treatments. The impact of Metagenomics is also being felt in

### 2015-09-18 docs/paper/out/paper.tex
make wording changes in using MG7, review #35 -- paper

@@
 
-\textbf{IDEAS}
+The MG7 16 data analysis workflow is indeed a set of tasks, all of them
+based in \emph{Loquat}. For each task, a set of inputs and outputs as
+well as configuration parameters must be statically defined. The user is
+also free to leave the reasonable defaults for configuration, needing
+only to define the input and output of the whole workflow. The
+definition of this configuration is Scala code and the way of starting
+an MG7 analysis is compiling the project code and launching it from the
+Scala interactive console.
 
-\begin{itemize}
-\tightlist
-\item
-  the user needs to write code, a repo would be nice, AWS account
-\item
-  there's some sort of conf required for the AWS account (add users,
-  roles, permissions, whatever)
-\item
-  Why this is a good thing (or should this just be in Discussion)
-\item
-  something else?
-\end{itemize}
+Code compilation prior to launching any analysis assures that no AWS
+resources are launched if the analysis is not well--defined, avoiding
+expenses not leading to any analysis. Besides compile--time checks,
+runtime checks are made before launch to ensure existence of input data
+and availability of resources.
+
+An MG7 analysis is then a Scala project where the user only needs to set
+certain variables at the code level (input, output and parameters),
+compile the code and run it. To facilitate the process of setting up the
+Scala project, a template with sensible defaults is provided.
+
+In order to be able to exploit Amazon Web Services infrastructure for
+the MG7 analysis, the user needs to set up an AWS account with certain
+IAM (Identity and Access Management) permission policies that will grant
+access to the resources used in the workflow.
 
@@
 
-\subsection{Inputs, outputs, data: compile-time, expressive,
-composable}\label{inputs-outputs-data-compile-time-expressive-composable}
+\subsection{Input and output data
+declaration}\label{input-and-output-data-declaration}
 
-\subsection{Tools, data, dependencies and machine
-configurations}\label{tools-data-dependencies-and-machine-configurations}
+An important aspect of the MG7 workflow is the way it deals with data
+resources. All the data that is going to be used in the analysis or
+produced as an output is described as Scala code using rich types from
+the \emph{Datasets} language. This allows user to specify all the
+information about the type of the data that can be utilized then by the
+tools analyzing this data. For example, we can specify that for the
+first part of the MG7 workflow running FLASH in parallel, requires
+Illumina paired end reads and produces joined reads.
 
-\subsection{Parallel cloud execution ??}\label{parallel-cloud-execution}
+On one hand, specification of the input data allows us to restrict its
+type and force users to be conscious about what they pass as an input.
+On the other hand specification of the output data helps to build a
+workflow as a \emph{composition} of several parts: we can ensure on the
+Scala code type level that the output of one component fits as an input
+of the next component. This can be crucial, as often the way the
+analysis works depends a lot on the particular structure of the data.
+For instance, in the MG7 workflow, using BLAST eDSL, we can describe
+exactly which format will the output of the BLAST step have, which
+information it will include, and then in the next step we can reuse this
+description to easily parse BLAST output and retrieve the part of the
+information needed for the taxonomy assignment analysis. Having data
+structure described statically as Scala code allows us to be sure that
+we won't have parsing problems or other issues with incompatible data
+passed between components of the workflow.
+
+All this doesn't compromise flexibility of the way user works with data
+in MG7. On the contrary, having static data declarations as a part of
+the configuration allows user to reuse component of analysis and modify
+it easily according to particular needs. Besides that, an important
+advantage of the type-level control is the additional insurance from
+unsuccessful analysis launches, which may lead to the lost of time and
+as a consequence finance spent on the cloud resources.
+
+\subsection{Tools, data, dependencies and automated
+deployment}\label{tools-data-dependencies-and-automated-deployment}
+
+Bioinformatics software often has a complicated installation process and
+requires various dependencies with unclear versions. This makes the
+deployment of the bioinformatics tools an involved task and resolving it
+manually is not a solution in the context of cloud computations. To face
+this problem, one needs an automated system of managing tools and
+resources, which will allow an expressive way for describing
+dependencies between parts of a pipeline and provide a reproducible
+procedure of its deployment. We have developed \emph{Statika} for this
+purpose and successfully use it in MG7.
+
+Every external tool involved in the workflow is represented as a
+\emph{Statika} bundle, which is essentially a Scala project describing
+the installation process of this tool and declaring dependencies on
+other bundles which will be installed prior to the considered tool
+itself. Describing relationships between bundles on the code level
+allows us to track the directed acyclic graph of their dependencies and
+linearize them to automatically install them sequentially in the right
+order. Meanwhile describing installation process on the code level
+allows user to utilize wide range of Scala and Java APIs, making
+installation a well-defined sequence of steps rather than an unreliable
+script dependent on the certain environment. This way \emph{Statika}
+provides an easy way to make deployment an automated reproducible
+process.
+
+Besides the bioinformatics tools like BLAST and FLASH, \emph{Statika}
+bundles are used for wrapping data dependencies and all inner components
+of the system that require cloud deployment. In particular, all
+components of \emph{Loquat} are bundles, which allows user to define
+which components are needed for the parallel processing on each
+computation unit in an expressive way, declaring them as bundle
+dependencies of the loquat ``worker'' bundle. This modularization is
+also important for the matter of making components of the system easily
+reusable for different projects and liberating user from most of the
+tasks related to their deployment.
+
+\subsection{Parallel computations in the
+cloud}\label{parallel-computations-in-the-cloud}
+
+The MG7 workflow consists of certain steps, each of which perform some
+work in parallel, using the cloud infrastructure managed by
+\emph{Loquat}. It is important to notice the horizontal scalability of
+this approach. Irrespectively of how much data is needed to be
+proccessed, MG7 will easily handle it, by splitting data on chunks and
+performing the analysis on multiple computaion units. The Amazon Elastic
+Compute Cloud (EC2) provides a transparent way of managing computation
+infrastructure, called autoscaling groups. User can easily set up MG7
+configuration parameters, adjusting the amount of EC2 instances they
+want to occupy for each task and their type. When using Amazon Web
+Services for parallel computations, it is easy to scale the computation
+resources, because the you pay for the time, not the amount of
+computation units.
 
@@
 
-We use EC2, S3 and SQS through a Scala wrapper of the official
-\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK},
+MG7 uses various Amazon Web Services:
+
+\begin{itemize}
+\tightlist
+\item
+  EC2 (Elastic Compute Cloud) autoscaling groups for launching and
+  managing computation units
+\item
+  S3 (Simple Storage Service) for storing input and output data
+\item
+  SQS (Simple Queue Service) for communication between different
+  components of the system
+\item
+  SNS (Simple Notification Service) for notifying user about the
+  progress
+\end{itemize}
+
+These services are used through a Scala wrapper of the official
+\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK
+\texttt{1.9.25}},
 \href{https://github.com/ohnosequences/aws-scala-tools/releases/tag/v0.13.2}{ohnosequences/aws-scala-tools
-\texttt{0.13.2}}. This uses version \texttt{1.9.25} of the AWS Java SDK.
+\texttt{0.13.2}}.
 

### 2015-09-18 docs/paper/out/paper.tex
make wording changes in AWS M&M, review #35 -- paper
correct – to - --paper

@@
 in general \citep[\citet{kodzius2015marine}]{cowan2015metagenomics}.
-These new possibilities for exploring the diversity of micro--organisms
+These new possibilities for exploring the diversity of micro-organisms
 in the most varied environments are opening new research areas, and
@@
 or tumour tissues) are required to have a sufficiently complete
-landscape of the human microbiome. Moreover, re--analysis of old samples
+landscape of the human microbiome. Moreover, re-analysis of old samples
 using new tools and better reference databases might be also demanded
@@
 Code compilation prior to launching any analysis assures that no AWS
-resources are launched if the analysis is not well--defined, avoiding
-expenses not leading to any analysis. Besides compile--time checks,
+resources are launched if the analysis is not well-defined, avoiding
+expenses not leading to any analysis. Besides compile-time checks,
 runtime checks are made before launch to ensure existence of input data
@@
 
-MG7 uses various Amazon Web Services:
+MG7 uses the following Amazon Web Services:
 
@@
 \href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK
-\texttt{1.9.25}},
+\texttt{1.9.25}}:
 \href{https://github.com/ohnosequences/aws-scala-tools/releases/tag/v0.13.2}{ohnosequences/aws-scala-tools

### 2015-09-18 docs/paper/out/paper.tex
make minor wording corrections on Discussion -- paper #35

@@
 Scala code type level that the output of one component fits as an input
-of the next component. This can be crucial, as often the way the
-analysis works depends a lot on the particular structure of the data.
-For instance, in the MG7 workflow, using BLAST eDSL, we can describe
-exactly which format will the output of the BLAST step have, which
-information it will include, and then in the next step we can reuse this
-description to easily parse BLAST output and retrieve the part of the
-information needed for the taxonomy assignment analysis. Having data
+for the next component. This is crucial, as obviously the way a data
+analysis task works depends a lot on the particular structure of the
+data. For instance, in the MG7 workflow, using BLAST eDSL, we can
+precisely describe which format will have the output of the BLAST step,
+which information it will include, and then in the next step we can
+reuse this description to parse BLAST output and retrieve the part of
+the information needed for the taxonomy assignment analysis. Having data
 structure described statically as Scala code allows us to be sure that
 we won't have parsing problems or other issues with incompatible data
-passed between components of the workflow.
+passed between workflow components.
 
-All this doesn't compromise flexibility of the way user works with data
-in MG7. On the contrary, having static data declarations as a part of
-the configuration allows user to reuse component of analysis and modify
-it easily according to particular needs. Besides that, an important
-advantage of the type-level control is the additional insurance from
-unsuccessful analysis launches, which may lead to the lost of time and
-as a consequence finance spent on the cloud resources.
+All this does not compromise flexibility in how the user works with data
+in MG7: having static data declarations as a part of the configuration
+allows the user to reuse analysis components, or modify them according
+to particular needs. Besides that, an important advantage of the
+type-level control is the added protection from the execution (and
+deployment) of a wrongly configured analysis task, which may lead to
+significant costs in both time and money.
 
@@
 of the system that require cloud deployment. In particular, all
-components of \emph{Loquat} are bundles, which allows user to define
-which components are needed for the parallel processing on each
-computation unit in an expressive way, declaring them as bundle
-dependencies of the loquat ``worker'' bundle. This modularization is
-also important for the matter of making components of the system easily
-reusable for different projects and liberating user from most of the
-tasks related to their deployment.
+components of \emph{Loquat} are bundles; the user can then define which
+components are needed for the parallel processing on each computation
+unit in an expressive way, declaring them as bundle dependencies of the
+loquat ``worker'' bundle. This modularization is also important for the
+matter of making components of the system reusable for different
+projects and liberating the user from most of the tasks related to their
+deployment.
 
@@
 \emph{Loquat}. It is important to notice the horizontal scalability of
-this approach. Irrespectively of how much data is needed to be
-proccessed, MG7 will easily handle it, by splitting data on chunks and
-performing the analysis on multiple computaion units. The Amazon Elastic
-Compute Cloud (EC2) provides a transparent way of managing computation
-infrastructure, called autoscaling groups. User can easily set up MG7
-configuration parameters, adjusting the amount of EC2 instances they
-want to occupy for each task and their type. When using Amazon Web
-Services for parallel computations, it is easy to scale the computation
-resources, because the you pay for the time, not the amount of
-computation units.
+this approach. Irrespectively of how much data needs to be processed,
+MG7 will easily handle it, by splitting data on chunks and performing
+the analysis on multiple computation units. The Amazon Elastic Compute
+Cloud (EC2) service provides a transparent way of managing computation
+infrastructure, called autoscaling groups. The User can set MG7
+configuration parameters, adjusting for each task the amount and
+hardware characteristics of the EC2 instances they want to occupy for
+it. When using Amazon Web Services for parallel computations, it is easy
+to scale the computation resources, because the you pay for the time,
+not the amount of computation units.
 
@@
 \item
-  EC2 (Elastic Compute Cloud) autoscaling groups for launching and
-  managing computation units
+  \href{https://aws.amazon.com/ec2}{EC2} (Elastic Compute Cloud)
+  autoscaling groups for launching and managing computation units
 \item
-  S3 (Simple Storage Service) for storing input and output data
+  \href{https://aws.amazon.com/s3}{S3} (Simple Storage Service) for
+  storing input and output data
 \item
-  SQS (Simple Queue Service) for communication between different
-  components of the system
+  \href{https://aws.amazon.com/sqs}{SQS} (Simple Queue Service) for
+  communication between different components of the system
 \item
-  SNS (Simple Notification Service) for notifying user about the
-  progress
+  \href{https://aws.amazon.com/sns}{SNS} (Simple Notification Service)
+  for e-mail notifications
 \end{itemize}

### 2015-09-21 docs/paper/template/frontiers.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.1 Generated 2015/22/05 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square}
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\linenumbers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Sample {et~al.}} %use et al only if is more than 1 author
\def\Authors{First Author\,$^{1,*}$, Co-Author\,$^{2}$ and Co-Author\,$^2$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X \\
$^{2}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X  }
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Corresponding Author}
\def\corrAddress{Laboratory X, Institute X, Department X, Organization X, Street X, City X , State XX (only USA, Canada and Australia), Zip Code, X Country X}
\def\corrEmail{email@uni.edu}

\begin{document}
\onecolumn
\firstpage{1}

\title[Running Title]{Article Title}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% The sections below are for reference only.
%%%
%%% For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the
%%% following sections or their equivalents for your field:
%%% Abstract, Introduction, Material and Methods, Results, and Discussion.
%%% Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
%%%
%%%For information about Clinical Trial Registration, please go to http://www.frontiersin.org/about/AuthorGuidelines#ClinicalTrialRegistration
%%%
%%% For Clinical Case Studies the following sections are mandatory: Abstract, Introduction, Background, Discussion, and Concluding Remarks.
%%%
%%% For all other article types there are no mandatory sections.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

%%% Leave the Abstract empty if your article falls under any of the following categories: Editorial Book Review, Commentary, Field Grand Challenge, Opinion or specialty Grand Challenge.
\section{}
%As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract.
For full guidelines regarding your manuscript please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines}{Author Guidelines} \\ or \textbf{Table \ref{Tab:01}} for a summary according to article type.

\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
%
% For Clinical Case Studies the Introduction should include symptoms at presentation, physical exams and lab results.
%
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.

%\begin{methods}
\section{Material \& Methods}

Text Text Text Text Text Text  Text Text Text Text Text Text \cite{conference} Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{patent} might want to know about  text text text text Text Text Text Text  Text Text Text Text Text Text  Text Text. \citep{article} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{book} might want to know about  text text text text \cite{chapter}

\begin{table}[!t]
\textbf{\refstepcounter{table}\label{Tab:01} Table \arabic{table}.}{ Maximum size of the Manuscript }

\processtable{ }
{\begin{tabular}{lllll}\toprule
 & Abstract max. legth (incl. spaces) & Figures or tables & Manuscript max. length \\\midrule
Clinical Case Study & & & &\\
Clinical Trial & & & &\\
Hypothesis and Theory & & & &\\
Methods & 2000 characters  & 15 & 12000 words \\
Original Research & & & &\\
Review & & & &\\
Technology Report & & & &\\\midrule
Focused Review & 2000 characters & 5 & 5000 words \\\midrule
CPC &  1250 characters& 6 & 2500 words  \\\midrule
Perspective & 1250 characters & 2 & 3000 words  \\
Mini Review & & & &\\\midrule
Data Report & None & 2 & 3000 words\\\midrule
Classification & 1250 characters & 10 & 2000 words \\\midrule
Editorial & None & None & 1000 words  \\\midrule
Frontiers Commentary  & & &\\
General Commentary & None & 1 & 1000 words\\
Book review & & & \\\midrule
Opinion   & & &\\
Specialty Grand Challenge & None & 1 & 2000 words\\
Field Grand Challenge & & & &\\\botrule
\end{tabular}}{}
\end{table}

Please note that large tables covering several pages cannot be included in the final PDF for formatting reasons. These tables will be published as supplementary material on the online article abstract page at the time of acceptance. The author will notified during the typesetting of the final article if this is the case. A link in the final PDF will direct to the online material.

\subsection{Original Research Articles, Clinical Trial Articles, and Technology Reports}

For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the following sections or their equivalents for your field:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Succinct, with no subheadings.
\item Materials and Methods: This section may be divided by subheadings. This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
\item Results: This section may be divided by subheadings. Footnotes should not be used and have to be transferred into the main text.
\item Discussion: This section may be divided by subheadings. Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
\end{itemize}

Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\subsection{Clinical Case Studies}

For Clinical Case Studies the following sections are mandatory:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Include symptoms at presentation, physical exams and lab results.
\item Background: This section may be divided by subheadings. Include history and review of similar cases.
\item Results: This section may be divided by subheadings. Include diagnosis and treatment.
\item Concluding Remarks
\end{itemize}

%\end{methods}

\section{Results}

\subsection{Figures}
Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript. Figures will then be automatically embedded at the bottom of the submitted manuscript. Kindly ensure that each table and figure is mentioned in the text and in numerical order. Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions. Figures which are not according to the guidelines will cause substantial delay during the production process.

\begin{table}[!t]
\textbf{\refstepcounter{table}\label{Tab:02} Table \arabic{table}.}{ Resolution Requirements for the figures}

\processtable{}
{\begin{tabular}{lllll}\toprule
Image Type & Description & Format & Color Mode & Resolution\\\midrule
Line Art & An image composed of lines and text,  & TIFF, JPEG & RGB, Bitmap & 900 - 1200 dpi\\
           & which does not contain tonal or shaded areas.& & &\\
           Halftone & A continuous tone photograph, which contains no text. & TIFF, EPS, JPEG & RGB, Grayscale & 300 dpi\\
Combination & Image contains halftone + text or line art elements. & TIFF, JPEG & RGB,Grayscale & 600 - 900 dpi\\\botrule
\end{tabular}}{}
\end{table}

\textbf{Table \ref{Tab:02}} shows the resolution requirements for the figures. The figures must be legible:
\begin{enumerate}
\item The smallest visible text is no less than 8 points in height, when viewed at actual size.
\item Solid lines are not broken up.
\item Image areas are not pixelated or stair stepped.
\item Text is legible and of high quality.
\item Any lines in the graphic are no smaller than 2 points width.
\item The actual size of the figure must be of at least 8.5 cm.
\end{enumerate}

\subsection{Nomenclature}
\begin{itemize}
\item The use of abbreviations should be kept to a minimum. Non-standard abbreviations should be avoided unless they appear at least four times, and defined upon first use in the main text. Consider also giving a list of non-standard abbreviations at the end, immediately before the Acknowledgments.
\item Gene symbols should be italicized; protein products are not italicized.
\item Chemical compounds and biomolecules should be referred to using systematic nomenclature, preferably using the recommendations by IUPAC.
\item We encourage the use of Standard International Units in all manuscripts.
\item To take part in the Resource Identification Initiative, please cite antibodies, genetically modified organisms, software tools, data, databases and services using the corresponding catalog number and RRID in your current manuscript. For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
\end{itemize}

\begin{equation}
\sum x+ y =Z\label{eq:01}
\end{equation}

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text.
Additional Requirements:
\subsection{Corrections}

If you need to communicate important changes to a published article please submit a General Commentary. Submit the article with the title “Corrigendum: Original Title of Article”.

\subsection{Commentaries on Articles}

At the beginning of your Commentary, please provide the citation of the article commented on. Rebuttals may be submitted in response to Commentaries; our limit in place is one commentary and one response. Rebuttals should also be submitted as General Commentary articles.

\subsection{Human Search and Animal Research}

All experiments on live vertebrates or higher invertebrates must be performed in accordance with relevant institutional and national guidelines and regulations. In the manuscript, authors must identify the committee approving the experiments and must confirm that all experiments conform to the relevant regulatory standards. For manuscripts reporting experiments on human subjects, authors must identify the committee approving the experiments and must also include a statement confirming that informed consent was obtained from all subjects. In Original Research Articles and Clinical Trial Articles these statements should appear in the Materials and Methods section.

\subsection{Clinical Trial Registration}

Clinical trials should be registered in a public trials registry in order to become the object of a publication at Frontiers. Trials must be registered at or before the start of patient enrollment. A clinical trial is defined as "any research study that prospectively assigns human participants or groups of humans to one or more health-related interventions to evaluate the effects on health outcomes."(\href{www.who.int/ictrp/en}{www.who.int/ictrp/en}). A list of acceptable registries can be found at \href{www.who.int/ictrp/en}{www.who.int/ictrp/en} and \href{www.icmje.org}{www.icmje.org}.

\subsection{Inclusion of Proteomics Data}

Authors should provide relevant information relating to how the peptide/protein matches were undertaken, including methods used to process and analyze data, false discovery rates (FDR) for large-scale studies and threshold or cut-off rates for peptide and protein matches. Further information could include software used, mass spectrometer type, sequence database and version, number of sequences in database, processing methods, mass tolerances used for matching, variable/fixed modifications, allowable missed cleavages, etc.

Authors should provide as supplementary material information used to identify proteins and/or peptides. This should include information such as accession numbers, observed mass (m/z), charge, delta mass, matched mass, peptide/protein scores, peptide modification, miscleavages, peptide sequence, match rank, matched species (for cross species matching), number of peptide matches, ambiguous protein/peptide matches should be indicated, etc.
For quantitative proteomics analyses authors should provide information to justify the statistical significance including biological replicates, statistical methods, estimates of uncertainty and the methods used for calculating error.

For peptide matches with biologically relevant post-translational modifications (PTM) and for any protein match that has occurred using a single mass spectrum, authors should include this information as raw data, annotated spectra or submit data to an online repository (recommended option).
Authors are encouraged to submit raw or matched data and 2-DE images to public proteomics repositories. Submission codes and/or links to data should be provided within the manuscript.

\subsection{Data Sharing}

Frontiers supports the policy of data sharing, and authors are advised to make freely available any materials and information described in their article, and any data relevant to the article (while not compromising confidentiality in the context of human-subject research) that may be reasonably requested by others for the purpose of academic and non-commercial research. In regards to deposition of data and data sharing through databases, Frontiers urges authors to comply with the current best practices within their discipline.

\section*{Disclosure/Conflict-of-Interest Statement}
%Frontiers follows the recommendations by the International Committee of Medical Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require that all financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest. When disclosing the potential conflict of interest, the authors need to address the following points:
%•	Did you or your institution at any time receive payment or services from a third party for any aspect of the submitted work?
%•	Please declare financial relationships with entities that could be perceived to influence, or that give the appearance of potentially influencing, what you wrote in the submitted work.
%•	Please declare patents and copyrights, whether pending, issued, licensed and/or receiving royalties relevant to the work.
%•	Please state other relationships or activities that readers could perceive to have influenced, or that give the appearance of potentially influencing, what you wrote in the submitted work.

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}
%When determining authorship the following criteria should be observed:
%•	Substantial contributions to the conception or design of the work; or the acquisition, analysis, or interpretation of data for the work; AND
%•	Drafting the work or revising it critically for important intellectual content; AND
%•	Final approval of the version to be published ; AND
%•	Agreement to be accountable for all aspects of the work in ensuring that questions related to the accuracy or integrity of any part of the work are appropriately investigated and resolved.
%Contributors who meet fewer than all 4 of the above criteria for authorship should not be listed as authors, but they should be acknowledged. (http://www.icmje.org/roles_a.html)

The statement about the authors and contributors can be up to several sentences long, describing the tasks of individual authors referred to by their initials and should be included at the end of the manuscript before the References section.

\section*{Acknowledgments}
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\textit{Funding\textcolon} Text Text Text Text Text Text  Text Text.

\section*{Supplemental Data}
Supplementary Material should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{test}

%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file

\section*{Figures}

%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.jpg and logo2.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{logo1}% This is a *.jpg file
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[width=10cm]{logo2}% This is an *.eps file
%\end{center}
%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
%\end{figure}

%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.

%%% Frontiers will add the figures at the end of the provisional pdf automatically %%%

%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

\end{document}

### 2015-09-21 docs/paper/out/paper.tex
Updated tex/pdf

@@
 hardware characteristics of the EC2 instances they want to occupy for
-it. When using Amazon Web Services for parallel computations, it is easy
-to scale the computation resources, because the you pay for the time,
-not the amount of computation units.
+it. But it's important to note that as each workflow step is not very
+resource demanding, it is not needed to hire EC2 instances with some
+advanced hardware, instead an average type will work and you can win
+time by simply scaling the number of the hired instances.
 

### 2015-09-22 docs/paper/out/paper.tex
Added v- prefix for the versions; some lost dots

@@
 flexible, scalable and traceable in which the biological databases of
-reference sequences can be easily updated and / or frequently
-substituted by new ones or by databases specifically designed for
-focused projects. MG7 uses parallelization and distributed analysis
-based on AWS, with on-demand infrastructure as the basic paradigm and
+reference sequences can be easily updated and/or frequently substituted
+by new ones or by databases specifically designed for focused projects.
+MG7 uses parallelization and distributed analysis based on AWS (Amazon
+Web Services), with on-demand infrastructure as the basic paradigm and
 allow the definition of complex workflows using a composable system for
-scaling/parallelizing stateless computations designed for Amazon Web
-Services (AWS) that counts with a static reproducible specification of
-dependencies and behavior of the different components. The modelling of
-the taxonomy tree is done using the new paradigm of graph databases of
-Bio4j that facilitates the taxonomic assignment tasks and the
-calculation of the taxa abundance values considering the hierarchic
-structure of taxonomy tree. MG7 includes the new 16S database 16S-DB7
-built with a flexible and sustainable system of updating and
-project-driven personalization.
+scaling/parallelizing stateless computations designed for AWS that
+counts with a static reproducible specification of dependencies and
+behavior of the different components. The modelling of the taxonomy tree
+is done using the new paradigm of graph databases of Bio4j that
+facilitates the taxonomic assignment tasks and the calculation of the
+taxa abundance values considering the hierarchic structure of taxonomy
+tree. MG7 includes the new 16S-DB7 database built with a flexible and
+sustainable system of updating and project-driven personalization.
 
@@
   adoption of programming methodologies like functional programming, or,
-  more speculatively, dependently typed programming. The immense new
+  more speculatively, dependently typed programming. The new
   possibilities that these advances offer must have a direct impact in
@@
 Scala-embedded mini-language for datasets and their locations.
-\textbf{Data} is represented as type-indexed fields: Keys are modeled as
+\textbf{Data} is represented as type-indexed fields: keys are modeled as
 singleton types, and values correspond to what could be called a
@@
 which are guaranteed statically to be different through type-level
-predicates, making use of the value ↔ type correspondence which can be
+predicates, making use of the value--type correspondence which can be
 established through singleton types and implicits. A dataset location is
-then just a list of locations formed by locations of each data member of
-that dataset. All this is based on what could be described as an
-embedding in Scala of an extensible record system with concatenation on
-disjoint labels, in the spirit of
-\citep[\citet{harper1991record}]{harper1990extensible}. For that
-\emph{Datasets} uses
-\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}.
+then just a list of locations formed by locations of each dataset key.
+All this is based on what could be described as an embedding in Scala of
+an extensible record system with concatenation on disjoint labels, in
+the spirit of \citep[\citet{harper1991record}]{harper1990extensible}.
+For that \emph{Datasets} uses
+\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
+library.
 
@@
 
-We developed our own Scala-based type-safe eDSLs (\textbf{e}mbedded
-\textbf{D}omain \textbf{S}pecific \textbf{L}anguage) for
+We developed our own Scala-based type-safe eDSLs (embedded Domain
+Specific Languages) for
 \href{https://github.com/ohnosequences/flash}{FLASH} and
@@
 
-In the case of BLAST we use a model for expressions where we can
-guarantee for each BLAST command expression at compile time
+In the case of BLAST we use a model where we can guarantee for each
+BLAST command expression at compile time that
 
@@
 using \emph{Datasets} which have a reference to the exact BLAST command
-options which yielded that output. This let us provide generic parsers
+options which yielded that output. This lets us provide generic parsers
 for BLAST output which are guaranteed to be correct.
@@
 theoretical and experimental knowledge of each specific area, the impact
-of metagenomics will be higher that if metagenomics progresses as a
-disconnected research branch. Considering that metagenomics data
-interoperability, which is especially critical in clinical environments,
-requires a stable taxonomy to be used as reference, we decided to rely
-on the most widely used taxonomy: the NCBI taxonomy. In addition, the
-biggest global sequence database GenBank follows this taxonomy to
-register the origin of all their submitted sequences. Our 16S database
-building strategy allows the substitution of the 16S database by any
-other subset of \textbf{nt}, even by the complete \textbf{nt} database
-if it would be needed, for example, for analyzing shotgun metagenomics
-data. This possibility of changing the reference database provides
-flexibility to the system enabling it for easy updating and
-project-driven personalization.
+of metagenomics will be higher than if it progresses as a disconnected
+research branch. Considering that metagenomics data interoperability,
+which is especially critical in clinical environments, requires a stable
+taxonomy to be used as reference, we decided to rely on the most widely
+used taxonomy: the NCBI taxonomy. In addition, the biggest global
+sequence database GenBank follows this taxonomy to register the origin
+of all their submitted sequences. Our 16S database building strategy
+allows the substitution of the 16S database by any other subset of
+\textbf{nt}, even by the complete \textbf{nt} database if it would be
+needed, for example, for analyzing shotgun metagenomics data. This
+possibility of changing the reference database provides flexibility to
+the system enabling it for easy updating and project-driven
+personalization.
 
@@
 a software project, using Scala. The idea is that \emph{everything}:
-data description, their location, configuration parameters, the
-infrastructure used, \ldots{} should be expressed as Scala code, and
-treated in the same way as any (well-managed) software project. This
-includes, among other things, using version control systems
-(\texttt{git} in our case), writing tests, making stable releases
-following \href{http://semver.org/}{semantic versioning} or publishing
-artifacts to a repository.
+data description, their location, configuration parameters and the
+infrastructure used should be expressed as Scala code, and treated in
+the same way as any (well-managed) software project. This includes,
+among other things, using version control systems (\texttt{git} in our
+case), writing tests, making stable releases following
+\href{http://semver.org/}{semantic versioning} or publishing artifacts
+to a repository.
 
@@
 These services are used through a Scala wrapper of the official
-\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK
-\texttt{1.9.25}}:
+\href{https://aws.amazon.com/sdk-for-java/}{AWS Java SDK v1.9.25}:
 \href{https://github.com/ohnosequences/aws-scala-tools/releases/tag/v0.13.2}{ohnosequences/aws-scala-tools
-\texttt{0.13.2}}.
+v0.13.2}.
 
@@
 
-MG7 itself and all the libraries used are written in Scala
-\texttt{2.11}.
+MG7 itself and all the libraries used are written in Scala v2.11.
 
@@
 \href{https://github.com/statika/statika/releases/tag/v0.2.0}{ohnosequences/statika
-\texttt{2.0.0}} for specifying the configuration and behavior of EC2
-instances.
+v2.0.0} for specifying the configuration and behavior of EC2 instances.
 
@@
 \href{https://github.com/ohnosequences/datasets/releases/tag/v0.2.0}{ohnosequences/datasets
-\texttt{0.2.0}} for specifying input and output data, their type and
-their location.
+v0.2.0} for specifying input and output data, their type and their
+location.
 
@@
 \href{https://github.com/ohnosequences/loquat/releases/tag/v2.0.0}{ohnosequences/loquat
-\texttt{2.0.0}} for the specification of data processing tasks and their
+v2.0.0} for the specification of data processing tasks and their
 execution using AWS resources.
@@
 \href{https://github.com/ohnosequences/blast/releases/tag/v0.2.0}{ohnosequences/blast
-\texttt{0.2.0}}. The BLAST version used is \texttt{2.2.31+}
+v0.2.0}. The BLAST version used is v2.2.31+.
 
@@
 \href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
-\texttt{0.1.0}}. The FLASH version used is \texttt{1.2.11}
+v0.1.0}. The FLASH version used is v1.2.11.
 
@@
 \href{https://github.com/bio4j/bio4j/releases/tag/v0.12.0-RC3}{bio4j/bio4j
-0.12.0-RC3} and
+v0.12.0-RC3} and
 \href{https://github.com/bio4j/bio4j-titan/releases/tag/v0.4.0-RC2}{bio4j/bio4j-titan
-0.4.0-RC2} as an API for the NCBI taxonomy.
+v0.4.0-RC2} as an API for the NCBI taxonomy.
 

### 2015-09-22 docs/paper/out/paper.tex
Updated tex/pdf

@@
 counts with a static reproducible specification of dependencies and
-behavior of the different components. The modelling of the taxonomy tree
+behavior of the different components. The modeling of the taxonomy tree
 is done using the new paradigm of graph databases of Bio4j that
@@
 localizations (different tissues, different anatomical regions, healthy
-or tumour tissues) are required to have a sufficiently complete
-landscape of the human microbiome. Moreover, re-analysis of old samples
-using new tools and better reference databases might be also demanded
-from time to time.
+or tumor tissues) are required to have a sufficiently complete landscape
+of the human microbiome. Moreover, re-analysis of old samples using new
+tools and better reference databases might be also demanded from time to
+time.
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
@@
   \textbf{Cloud computing:} the appearance of new models for massive
-  computation and storage such as the so-called cloud, or the widespread
-  adoption of programming methodologies like functional programming, or,
-  more speculatively, dependently typed programming. The new
-  possibilities that these advances offer must have a direct impact in
-  metagenomics data analysis.
+  computation and storage such as the cloud-based platforms, or the
+  widespread adoption of programming methodologies like functional
+  programming, or, more speculatively, dependently typed programming.
+  The new possibilities that these advances offer must have a direct
+  impact in metagenomics data analysis.
 \item

### 2015-09-22 docs/paper/out/paper.tex
add @marina-manrique corrections, some of my own

@@
 \def\Authors{
- Alexey Alekhin\(^\dagger\)\,$^{1}$  Evdokim Kovach\(^\dagger\)\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
+ Alexey Alekhin\(^\dagger\)\,$^{1}$  Evdokim Kovach\(^\dagger\)\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja-Tobes\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
 }% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 % Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Corresponding Author}
+\def\corrAuthor{Eduardo Pareja-Tobes}
 \def\corrAddress{Oh no sequences! Research Group, Era7 Bioinformatics, Plaza Campo Verde 3, Granada, 18001, Spain}
@@
 by new ones or by databases specifically designed for focused projects.
-MG7 uses parallelization and distributed analysis based on AWS (Amazon
-Web Services), with on-demand infrastructure as the basic paradigm and
+MG7 uses parallelization and distributed analysis based on Amazon Web
+Services (AWS), with on-demand infrastructure as the basic paradigm and
 allow the definition of complex workflows using a composable system for
@@
   system for scaling/parallelizing stateless computations especially
-  designed for Amazon Web Services (AWS)
+  designed for AWS
 \item
@@
 the data type includes the relevant information (insert size, read
-length, etc) to pass to a tool such as FLASH.
+length, etc) to pass to a tool such as FLASh.
 
 \subsubsection{Type-safe eDSLs for BLAST and
-FLASH}\label{type-safe-edsls-for-blast-and-flash}
+FLASh}\label{type-safe-edsls-for-blast-and-flash}
 
@@
 Specific Languages) for
-\href{https://github.com/ohnosequences/flash}{FLASH} and
+\href{https://github.com/ohnosequences/flash}{FLASh} and
 \href{https://github.com/ohnosequences/blast}{BLAST} expressions and
@@
 In the same spirit as for BLAST, we implemented a type-safe eDSL for
-FLASH expressions and their execution, supporting features equivalent to
+FLASh expressions and their execution, supporting features equivalent to
 those outlined for the BLAST eDSL.
@@
 downloaded from the \textbf{RDP database} \citep{cole2013ribosomal}. RDP
-unaligned sequences were used to capture new 16S sequences from
+unaligned sequences were used to capture new 16S RNA sequences from
 \textbf{nt} using BLAST similarity search strategies and then,
@@
 \subsubsection{Joining reads of each pair using
-FLASH}\label{joining-reads-of-each-pair-using-flash}
+FLASh}\label{joining-reads-of-each-pair-using-flash}
 
@@
 that yields pairs of reads with an overlapping region between them, are
-assembled using FLASH \citep{magovc2011flash}. FLASH is designed to
+assembled using FLASh \citep{magovc2011flash}. FLASh is designed to
 merge pairs of reads when the original DNA fragments are shorter than
@@
 For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
-over a threshold of similarity (by default \(evalue \leq e^{-15}\))
+over a BLAST expect value threshold (by default \(evalue \leq e^{-15}\))
 filtering those hits that are not sufficiently good comparing them with
 the best one. We select the best HSP (High Similarity Pair) per
-reference sequence and then choose the best HSP (that with lowest e
-value) between all the selected ones. The bitscore of this best HSP
+reference sequence and then choose the best HSP (that with lowest
+E-value) between all the selected ones. The bitscore of this best HSP
 (called S) is used as reference to filter the rest of HSPs. All the HSPs
-with bitscore below p x S are filtered. p is a coefficient fixed by the
-user to define the bitscore required, e.g.~if p=0.9 and S=700 the
-required bitscore threshold would be 630. Once we have the definitive
-HSPs selected, we obtain their corresponding taxonomic nodes using the
-taxonomic assignments that NCBI provides for all the nt database
-sequences. Now we have to analyze the topological distribution of these
-nodes in the taxonomy tree: i. If all the nodes forms a line in the
-taxonomy tree (are located in a not branched lineage to the tree root)
-we should choose the most specific taxID as the final assignment for
-that read. We call to this kind of assignment the `in line' exception
-(see Figure 2 right panel). ii. If not, we should search for the
-\emph{sensu stricto} Lowest Common Ancestor (LCA) of all the selected
-taxonomic nodes (See Figure 2 left panel). In this approach we decided
-to use the bitscore for evaluating the similarity because it is a value
-that increases when similarity is higher and depends a lot on the length
-of the HSP. Some reads could not find sequences with enough similarity
-in the database and then they would be classified as reads with no hits.
-Advanced metagenomics analysis approaches \citep{huson2013microbial}
-have adopted LCA assignment algorithms because it provides fine and
-trusted taxonomical assignment.
+with bitscore below the product \(p S\) are filtered. p is a coefficient
+fixed by the user to define the bitscore required, e.g.~if \(p=0.9\) and
+\(S=700\) the required bitscore threshold would be \(630\). Once we have
+the definitive HSPs selected, we obtain their corresponding taxonomic
+nodes using the taxonomic assignments that NCBI provides for all the nt
+database sequences. Now we have to analyze the topological distribution
+of these nodes in the taxonomy tree: i. If all the nodes forms a line in
+the taxonomy tree (are located in a not branched lineage to the tree
+root) we should choose the most specific taxID as the final assignment
+for that read. We call to this kind of assignment the `in line'
+exception (see Figure 2 right panel). ii. If not, we should search for
+the \emph{sensu stricto} Lowest Common Ancestor (LCA) of all the
+selected taxonomic nodes (See Figure 2 left panel). In this approach we
+decided to use the bitscore for evaluating the similarity because it is
+a value that increases when similarity is higher and depends a lot on
+the length of the HSP. Some reads could not find sequences with enough
+similarity in the database and then they would be classified as reads
+with no hits. Advanced metagenomics analysis approaches
+\citep{huson2012microbial} have adopted LCA assignment algorithms
+because it provides fine and trusted taxonomical assignment.
 
@@
 
-In order to be able to exploit Amazon Web Services infrastructure for
-the MG7 analysis, the user needs to set up an AWS account with certain
-IAM (Identity and Access Management) permission policies that will grant
-access to the resources used in the workflow.
+In order to be able to exploit AWS infrastructure for the MG7 analysis,
+the user needs to set up an AWS account with certain IAM (Identity and
+Access Management) permission policies that will grant access to the
+resources used in the workflow.
 
@@
   input and output data, their locations and type are expressible and
-  checked at compile-time using \emph{Datasets} ­3. management of
-  dependencies and machine configurations using \emph{Statika}
+  checked at compile-time using \emph{Datasets}
+\item
+  management of dependencies and machine configurations using
+  \emph{Statika}
 \item
@@
   choose only from valid Illumina read types, and then build a default
-  FLASH command based on that. The output locations, being declared
+  FLASh command based on that. The output locations, being declared
   statically, are also available for use in further analysis.
@@
 produced as an output is described as Scala code using rich types from
-the \emph{Datasets} language. This allows user to specify all the
-information about the type of the data that can be utilized then by the
-tools analyzing this data. For example, we can specify that for the
-first part of the MG7 workflow running FLASH in parallel, requires
+the \emph{Datasets} language. This allows The user can specify
+information about types of data, information that can then be utilized
+by tools analyzing this data. For example, we can specify that for the
+first part of the MG7 workflow running FLASh in parallel, requires
 Illumina paired end reads and produces joined reads.
@@
 structure described statically as Scala code allows us to be sure that
-we won't have parsing problems or other issues with incompatible data
+we will not have parsing problems or other issues with incompatible data
 passed between workflow components.
@@
 linearize them to automatically install them sequentially in the right
-order. Meanwhile describing installation process on the code level
-allows user to utilize wide range of Scala and Java APIs, making
-installation a well-defined sequence of steps rather than an unreliable
-script dependent on the certain environment. This way \emph{Statika}
-provides an easy way to make deployment an automated reproducible
-process.
+order. Meanwhile describing the installation process on the code level
+allows the user to utilize the wide range of available Scala and Java
+APIs and tools, making installation a well-defined sequence of steps
+rather than an unreliable script, dependent on a certain environment.
+\emph{Statika} offers an easy path towards making deployment an
+automated, reproducible process.
 
-Besides the bioinformatics tools like BLAST and FLASH, \emph{Statika}
+Besides bioinformatics tools like BLAST and FLASh, \emph{Statika}
 bundles are used for wrapping data dependencies and all inner components
@@
 
-The MG7 workflow consists of certain steps, each of which perform some
+The MG7 workflow consists of certain steps, each of which performs some
 work in parallel, using the cloud infrastructure managed by
@@
 this approach. Irrespectively of how much data needs to be processed,
-MG7 will easily handle it, by splitting data on chunks and performing
-the analysis on multiple computation units. The Amazon Elastic Compute
-Cloud (EC2) service provides a transparent way of managing computation
+MG7 will handle it, by splitting data into chunks and performing the
+analysis on multiple computation units. The Amazon Elastic Compute Cloud
+(EC2) service provides a transparent way of managing computation
 infrastructure, called autoscaling groups. The User can set MG7
 configuration parameters, adjusting for each task the amount and
-hardware characteristics of the EC2 instances they want to occupy for
-it. But it's important to note that as each workflow step is not very
+hardware characteristics of the EC2 instances they want to use for it.
+But it is important to note that, as each workflow step is not very
 resource demanding, it is not needed to hire EC2 instances with some
-advanced hardware, instead an average type will work and you can win
-time by simply scaling the number of the hired instances.
+advanced hardware, instead an average type will work and you can reduce
+execution time by simply scaling out the number of instances.
 
@@
 criteria, like biological function or taxonomy. Bio4j would be an
-invaluable tool here, thanks to its ability to express express complex
+invaluable tool here, thanks to its ability to express complex
 predicates on sequences using all the information linked with them (GO
@@
 
-\subsection{FLASH eDSL}\label{flash-edsl}
+\subsection{FLASh eDSL}\label{flash-edsl}
 
@@
 \href{https://github.com/ohnosequences/flash/releases/tag/v0.1.0}{ohnosequences/flash
-v0.1.0}. The FLASH version used is v1.2.11.
+v0.1.0}. The FLASh version used is v1.2.11.
 

### 2015-09-24 docs/paper/out/paper.tex
add BLAST ref, minor fixes in wording -- paper

@@
  \keyFont{ \section{Keywords:} Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
-Cloud computing, NGS, Genomic big data } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+Cloud computing, NGS, Genomic big data, Microbiome, Environmental, 16S
+Database } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \end{abstract}
@@
 Specific Languages) for
-\href{https://github.com/ohnosequences/flash}{FLASh} and
-\href{https://github.com/ohnosequences/blast}{BLAST} expressions and
-their execution.
+\href{https://github.com/ohnosequences/flash}{FLASh}
+\citep{magovc2011flash} and
+\href{https://github.com/ohnosequences/blast}{BLAST}
+\citep{camacho2009blast} expressions and their execution.
 
@@
 database. The taxonomic assignment for each read is based on BLASTN of
-each read against the 16S database. Assignment based on direct
-similarity of each read one by one compared against a sufficiently wide
-database is a very exhaustive method for assignment
-\citep[\citet{morgan2012chapter}]{segata2013computational}. Some methods
-of assignment compare the sequences only against the available complete
-bacterial genomes or avoid computational cost clustering or binning the
-sequences first, and then doing the assignments only for the
-representative sequence of each cluster. MG7 carries out an exhaustive
-comparison of all the reads under analysis and it does not applies any
-binning strategy. Every read is specifically compared with all the
-sequences of the 16S database. We select the best BLAST hits (10 hits by
-default) obtained for each read to do the taxonomic assignment.
+each read against the 16S database.Assignment based on direct similarity
+of each read one by one compared against a sufficiently wide database is
+considered in different reviews of metagenomics analysis methodologies
+\citep[\citet{morgan2012chapter}]{segata2013computational} as a very
+exhaustive method for assignment. Some methods of assignment compare the
+sequences only against the 16S genes from available complete bacterial
+genomes or avoid computational cost clustering or binning the sequences
+first, and then doing the assignments only for the representative
+sequence of each cluster. MG7 carries out an exhaustive comparison of
+all the reads under analysis and it does not applies any binning
+strategy. Every read is specifically compared with all the sequences of
+the 16S database. We select the best BLAST hits (10 hits by default)
+obtained for each read to do the taxonomic assignment.
 
@@
 For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
-over a BLAST expect value threshold (by default \(evalue \leq e^{-15}\))
-filtering those hits that are not sufficiently good comparing them with
-the best one. We select the best HSP (High Similarity Pair) per
-reference sequence and then choose the best HSP (that with lowest
-E-value) between all the selected ones. The bitscore of this best HSP
-(called S) is used as reference to filter the rest of HSPs. All the HSPs
-with bitscore below the product \(p S\) are filtered. p is a coefficient
-fixed by the user to define the bitscore required, e.g.~if \(p=0.9\) and
-\(S=700\) the required bitscore threshold would be \(630\). Once we have
-the definitive HSPs selected, we obtain their corresponding taxonomic
-nodes using the taxonomic assignments that NCBI provides for all the nt
-database sequences. Now we have to analyze the topological distribution
-of these nodes in the taxonomy tree: i. If all the nodes forms a line in
-the taxonomy tree (are located in a not branched lineage to the tree
-root) we should choose the most specific taxID as the final assignment
-for that read. We call to this kind of assignment the `in line'
-exception (see Figure 2 right panel). ii. If not, we should search for
-the \emph{sensu stricto} Lowest Common Ancestor (LCA) of all the
-selected taxonomic nodes (See Figure 2 left panel). In this approach we
-decided to use the bitscore for evaluating the similarity because it is
-a value that increases when similarity is higher and depends a lot on
-the length of the HSP. Some reads could not find sequences with enough
-similarity in the database and then they would be classified as reads
-with no hits. Advanced metagenomics analysis approaches
-\citep{huson2012microbial} have adopted LCA assignment algorithms
-because it provides fine and trusted taxonomical assignment.
+over a threshold of similarity. To evaluate similarity for this first
+filtering of hits we use the Expect value (by default
+\(evalue \leq e^{-15}\)) that describes the number of hits one can
+``expect'' to see by chance when searching a database of a particular
+size. In a second filtering step we filtering those hits that are not
+sufficiently good comparing them with the best one. We select the best
+HSP (High Similarity Pair) per reference sequence and then choose the
+best HSP (that with lowest E-value) between all the selected ones. The
+bitscore of this best HSP (called S) is used as reference to filter the
+rest of HSPs. All the HSPs with bitscore below the product \(p S\) are
+filtered. p is a coefficient fixed by the user to define the bitscore
+required, e.g.~if \(p=0.9\) and \(S=700\) the required bitscore
+threshold would be \(630\). Once we have the definitive HSPs selected,
+we obtain their corresponding taxonomic nodes using the taxonomic
+assignments that NCBI provides for all the nt database sequences. Now we
+have to analyze the topological distribution of these nodes in the
+taxonomy tree: i. If all the nodes forms a line in the taxonomy tree
+(are located in a not branched lineage to the tree root) we should
+choose the most specific taxID as the final assignment for that read. We
+call to this kind of assignment the `in line' exception (see Figure 2
+right panel). ii. If not, we should search for the \emph{sensu stricto}
+Lowest Common Ancestor (LCA) of all the selected taxonomic nodes (See
+Figure 2 left panel). In this approach we decided to use the bitscore
+for evaluating the similarity because it is a value that increases when
+similarity is higher and depends a lot on the length of the HSP. Some
+reads could not find sequences with enough similarity in the database
+and then they would be classified as reads with no hits. Advanced
+metagenomics analysis approaches \citep{huson2012microbial} have adopted
+LCA assignment algorithms because it provides fine and trusted
+taxonomical assignment.
 
@@
 counts, LCA cumu. counts, LCA direct \%, LCA cumu. \%, BBH direct
-counts, BBH cumu. counts, BBH direct \%, BBH cumu. \%.
+counts, BBH cumu. counts, BBH direct \% and BBH cumu. \%.
 
@@
 
-The MG7 16 data analysis workflow is indeed a set of tasks, all of them
+The MG7 16S data analysis workflow is indeed a set of tasks, all of them
 based in \emph{Loquat}. For each task, a set of inputs and outputs as
@@
 \item
-  Treat data analysis as a software project. This makes for radical
+  Treating data analysis as a software project. This makes for radical
   improvements in \emph{reproducibility}, \emph{reuse},
@@
 \item
-  input and output data, their locations and type are expressible and
-  checked at compile-time using \emph{Datasets}
+  Checking at compile-time: input and output data, their locations and
+  type are expressible and checked at compile-time using \emph{Datasets}
 \item
-  management of dependencies and machine configurations using
+  Management of dependencies and machine configurations using
   \emph{Statika}
 \item
-  automation of AWS cloud resources and processes, including
+  Automation of AWS cloud resources and processes, including
   distribution and parallelization through the use of \emph{Loquat}
 \item
-  taxonomic data and related operations are treated natively as what
+  Taxonomic data and related operations are treated natively as what
   they are: graphs, through the use of \emph{Bio4j}
@@
 
-\subsection{A new approach to data
-analysis}\label{a-new-approach-to-data-analysis}
+\subsection{A new approach to data analysis: data analysis as a software
+project and checking at
+compile-time}\label{a-new-approach-to-data-analysis-data-analysis-as-a-software-project-and-checking-at-compile-time}
 
@@
   \textbf{Expresiveness and safety} For example in our case we can
-  choose only from valid Illumina read types, and then build a default
+  choose only from valid illumina read types, and then build a default
   FLASh command based on that. The output locations, being declared
@@
 produced as an output is described as Scala code using rich types from
-the \emph{Datasets} language. This allows The user can specify
+the \emph{Datasets} language. This allows the user to specify
 information about types of data, information that can then be utilized
-by tools analyzing this data. For example, we can specify that for the
-first part of the MG7 workflow running FLASh in parallel, requires
-Illumina paired end reads and produces joined reads.
+by tools analyzing this data. For example, we can specify that, for the
+first part of the MG7 workflow, running FLASh in parallel requires
+illumina paired end reads and produces joined reads.
 
@@
 type and force users to be conscious about what they pass as an input.
-On the other hand specification of the output data helps to build a
+On the other hand, specification of the output data helps to build a
 workflow as a \emph{composition} of several parts: we can ensure on the
 Scala code type level that the output of one component fits as an input
-for the next component. This is crucial, as obviously the way a data
+for the next component. This is crucial as, obviously, the way a data
 analysis task works depends a lot on the particular structure of the
@@
 reuse this description to parse BLAST output and retrieve the part of
-the information needed for the taxonomy assignment analysis. Having data
-structure described statically as Scala code allows us to be sure that
-we will not have parsing problems or other issues with incompatible data
-passed between workflow components.
+the information needed for the taxonomy assignment analysis. Having the
+data structure described statically as Scala code allows us to be sure
+that we will not have parsing problems or other issues with incompatible
+data passed between workflow components.
 
@@
 procedure of its deployment. We have developed \emph{Statika} for this
-purpose and successfully use it in MG7.
+purpose and successfully used it in MG7.
 
@@
 linearize them to automatically install them sequentially in the right
-order. Meanwhile describing the installation process on the code level
+order. Meanwhile, describing the installation process on the code level
 allows the user to utilize the wide range of available Scala and Java
@@
 resource demanding, it is not needed to hire EC2 instances with some
-advanced hardware, instead an average type will work and you can reduce
+advanced hardware. Instead, an average type will work and you can reduce
 execution time by simply scaling out the number of instances.
@@
 into a set of interdependent graphs to be used for bioinformatics
-analysis and especially for biological data.
+analysis and especially for biological data. MG7 works based on the
+Bio4j taxonomy module. It opens the possibility to connect the taxonomic
+profiling data obtained with MG7 to all the biological knowledge
+associated to each taxon. Using the information available in Bio4j for
+all the proteins assigned to each taxon we are connected to all the
+functional data available in Uniprot related with it.
 
@@
 
+The comparison of the taxonomic profiles between different groups of
+samples is a need for many metagenomics studies. Tasks related with this
+group-based analysis, such as the extraction of the minimal tree with
+all the taxa with some direct or accumulated assignment, will be part of
+a new MG7 module, already in development.
+
 \subsubsection{Interactive visualizations based on
 Biographika}\label{interactive-visualizations-based-on-biographika}
+
+New visualization tools for metagenomics results are undoubtedly needed.
+Interactivity is a especially interesting feature for metagenomics data
+visualization, since the expert needs to explore the results in a
+knowledge-driven way. The majority of the available metagenomics data
+visualizations are static. We are working in the \emph{Biographika}
+project \citep{tobes2015biographika}, to provide interactive rich
+visualizations on the web for Bio4j data. The development of
+visualizations specific for MG7 is one of Biographika current goals.
+Biographika is based on D3.js, the de-facto standard JavaScript data
+visualization library, and is open source.
 

### 2015-09-24 docs/paper/out/paper.tex
add author contributions -- paper

@@
 
+\section{Disclosure/Conflict-of-Interest
+Statement}\label{disclosureconflict-of-interest-statement}
+
+All authors work at the \emph{Oh no sequences!} research group, part of
+Era7 Bioinformatics. Era7 offers metagenomics data analysis services
+based on MG7. MG7 is open source, available under the OSI-approved
+AGPLv3 license.
+
+\section{Author Contributions}\label{author-contributions}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{AA} developed \emph{MG7}, \emph{Statika}, \emph{Datasets}, and
+  \emph{aws-scala-tools}; wrote the paper;
+\item
+  \textbf{EK} developed \emph{nispero} (a prototype for \emph{Loquat}
+  \citep{kovach2014nispero}) and \emph{aws-scala-tools}.
+\item
+  \textbf{MM} MG7 workflow design; curation and design for the 16S-DB7
+  reference database; wrote the paper.
+\item
+  \textbf{PPT} curation, design, data extraction code for the 16S-DB7
+  reference database.
+\item
+  \textbf{EP} MG7 workflow design; wrote the paper.
+\item
+  \textbf{RT} MG7 workflow design, assignment strategy; curation and
+  design for the 16S-DB7 reference database; wrote the paper.
+\item
+  \textbf{EPT} developed \emph{MG7}, \emph{Statika}, \emph{Datasets},
+  \emph{FLASh/BLAST eDSLs}; data analysis approach and design; wrote the
+  paper.
+\end{itemize}
+
+All authors have read and approved the final manuscript.
+
 \section{Acknowledgements}\label{acknowledgements}

### 2015-09-24 docs/paper/out/paper.tex
minor formatting edits in author contributions -- paper

@@
 \item
-  \textbf{MM} MG7 workflow design; curation and design for the 16S-DB7
-  reference database; wrote the paper.
+  \textbf{MM} \emph{MG7} workflow design; curation and design of the
+  \emph{16S-DB7} reference database; wrote the paper.
 \item
-  \textbf{PPT} curation, design, data extraction code for the 16S-DB7
-  reference database.
+  \textbf{PPT} curation, design, data extraction code of the
+  \emph{16S-DB7} reference database.
 \item
-  \textbf{EP} MG7 workflow design; wrote the paper.
+  \textbf{EP} \emph{MG7} workflow design; wrote the paper.
 \item
-  \textbf{RT} MG7 workflow design, assignment strategy; curation and
-  design for the 16S-DB7 reference database; wrote the paper.
+  \textbf{RT} \emph{MG7} workflow design, assignment strategy; curation
+  and design of the \emph{16S-DB7} reference database; wrote the paper.
 \item

### 2015-09-24 src/main/scala/loquats/assignment/test.scala
Updated code to compile with the new aws-scala-tools

@@ -15,7 +15,8 @@ import ohnosequences.statika.bundles._
 import ohnosequences.statika.aws.api._
 import ohnosequences.statika.aws.amazonLinuxAMIs._
 
-import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
+import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
+import ohnosequences.awstools.s3._
 import ohnosequences.awstools.autoscaling._
 
 import era7.project.loquats._
@@ -32,12 +33,12 @@ case object assignmenTest {
     val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
 
     val managerConfig = ManagerConfig(
-      instanceType = m3_medium,
+      instanceType = m3.medium,
       purchaseModel = SpotAuto
     )
 
     val workersConfig = WorkersConfig(
-      instanceType = m3_medium,
+      instanceType = m3.medium,
       purchaseModel = SpotAuto,
       groupSize = WorkersGroupSize(0, 1, 10)
     )
@@ -52,16 +53,16 @@ case object assignmenTest {
           "ERR567374_1",
           assignmentDataProcessing
         )(remoteInput =
-            blastOutput.atS3(ObjectAddress(
+            blastOutput.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/blast-test/ERR567374_1.blast.partial.csv"
             )) :~: ∅,
           remoteOutput =
-            lcaCSV.atS3(ObjectAddress(
+            lcaCSV.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.lca.csv"
             )) :~:
-            bbhCSV.atS3(ObjectAddress(
+            bbhCSV.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.bbh.csv"
             )) :~:

### 2015-09-24 src/main/scala/loquats/blast/config.scala

@@ -15,7 +15,8 @@ import ohnosequences.statika.bundles._
 import ohnosequences.statika.aws.api._
 import ohnosequences.statika.aws.amazonLinuxAMIs._
 
-import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
+import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
+import ohnosequences.awstools.s3._
 import ohnosequences.awstools.autoscaling._
 
 import era7.project.loquats._
@@ -61,12 +62,12 @@ case object blastTest {
     val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
 
     val managerConfig = ManagerConfig(
-      instanceType = m3_medium,
+      instanceType = m3.medium,
       purchaseModel = SpotAuto
     )
 
     val workersConfig = WorkersConfig(
-      instanceType = m3_medium,
+      instanceType = m3.medium,
       purchaseModel = SpotAuto,
       groupSize = WorkersGroupSize(0, 1, 10)
     )
@@ -81,10 +82,10 @@ case object blastTest {
           "ERR567374_1",
           dataProcessing
         )(remoteInput =
-            fastqInput.atS3(ObjectAddress("resources.ohnosequences.com", "16s/public-datasets/PRJEB6592/flash-test/ERR567374_1.merged.fastq")) :~:
+            fastqInput.inS3Object(S3Object("resources.ohnosequences.com", "16s/public-datasets/PRJEB6592/flash-test/ERR567374_1.merged.fastq")) :~:
             ∅,
           remoteOutput =
-            blastOutput.atS3(ObjectAddress("resources.ohnosequences.com", "16s/public-datasets/PRJEB6592/blast-test/ERR567374_1.merged.blast.csv")) :~:
+            blastOutput.inS3Object(S3Object("resources.ohnosequences.com", "16s/public-datasets/PRJEB6592/blast-test/ERR567374_1.merged.blast.csv")) :~:
             ∅
         )
       )

### 2015-09-24 src/main/scala/loquats/counting/test.scala

@@ -15,7 +15,8 @@ import ohnosequences.statika.bundles._
 import ohnosequences.statika.aws.api._
 import ohnosequences.statika.aws.amazonLinuxAMIs._
 
-import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
+import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
+import ohnosequences.awstools.s3._
 import ohnosequences.awstools.autoscaling._
 
 import era7.project.loquats._
@@ -33,12 +34,12 @@ case object countingTest {
     val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
 
     val managerConfig = ManagerConfig(
-      instanceType = m3_medium,
+      instanceType = m3.medium,
       purchaseModel = SpotAuto
     )
 
     val workersConfig = WorkersConfig(
-      instanceType = m3_medium,
+      instanceType = m3.medium,
       purchaseModel = SpotAuto,
       groupSize = WorkersGroupSize(0, 1, 10)
     )
@@ -53,20 +54,20 @@ case object countingTest {
           "ERR567374_1",
           countingDataProcessing
         )(remoteInput =
-            lcaCSV.atS3(ObjectAddress(
+            lcaCSV.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.lca.csv"
             )) :~:
-            bbhCSV.atS3(ObjectAddress(
+            bbhCSV.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.bbh.csv"
             )) :~: ∅,
           remoteOutput =
-            lcaCountsCSV.atS3(ObjectAddress(
+            lcaCountsCSV.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.lca.counts.csv"
             )) :~:
-            bbhCountsCSV.atS3(ObjectAddress(
+            bbhCountsCSV.inS3Object(S3Object(
               "resources.ohnosequences.com",
               "16s/public-datasets/PRJEB6592/counting-test/ERR567374_1.bbh.counts.csv"
             )) :~:

### 2015-09-24 src/main/scala/loquats/flash/data.scala

@@ -28,8 +28,8 @@ trait AnyFlashData {
   val stats: Stats
 
   lazy val flashOptions = flash.defaults update (
-    read_len(readsType.length.toInt)   :~:
-    max_overlap(readsType.length.toInt) :~: ∅
+    read_len(readsType.length.length)   :~:
+    max_overlap(readsType.length.length) :~: ∅
   )
 
 }

### 2015-09-24 docs/paper/out/paper.tex
super duper final version

@@
 
-The exponential growth of metagenomics is adding a significant plus of
-complexity to the big data problem in genomics. In this new scenario
-impacted by the wide scale and scope of the projects and by the
-explosion of sequence data to be analyzed is especially opportune the
-use of new possibilities that cloud computing approaches, new functional
-and dependently typed programming languages and new database paradigms
-as graph databases offer. To tackle the challenges of big data analysis
-in this work we have used these new means to design and develop a new
-open source methodology for analyzing metagenomics data, MG7. It
-exploits the new possibilities that cloud computing offers to get a
-system robust, programmatically configurable, modular, distributed,
-flexible, scalable and traceable in which the biological databases of
-reference sequences can be easily updated and/or frequently substituted
-by new ones or by databases specifically designed for focused projects.
-MG7 uses parallelization and distributed analysis based on Amazon Web
-Services (AWS), with on-demand infrastructure as the basic paradigm and
-allow the definition of complex workflows using a composable system for
-scaling/parallelizing stateless computations designed for AWS that
-counts with a static reproducible specification of dependencies and
-behavior of the different components. The modeling of the taxonomy tree
-is done using the new paradigm of graph databases of Bio4j that
-facilitates the taxonomic assignment tasks and the calculation of the
-taxa abundance values considering the hierarchic structure of taxonomy
-tree. MG7 includes the new 16S-DB7 database built with a flexible and
-sustainable system of updating and project-driven personalization.
+As part of the Cambrian explosion of omics data, metagenomics brings to
+the table a specific, defining trait: its social essence. The
+\emph{meta} prefix exerts its influence, with multitudes manifesting
+themselves everywhere; from samples to data analysis, from actors
+involved to (present and future) applications. Of these dimensions, data
+analysis is where needs lay further from what current tools provide. Key
+features are, among others, scalability, reproducibility, data
+provenance and distribution, process identity and versioning. These are
+the goals guiding our work in MG7, a 16S metagenomics data analysis
+system. The basic principle is a new approach to data analysis, where
+configuration, processes, or data locations are static, type-checked and
+subject to the standard evolution of a well-maintained software project.
+Cloud computing, in its Amazon Web Services incarnation, when coupled
+with these ideas, produces a robust, safely configurable, scalable tool.
+Processes, data, machine behaviors and their dependencies are expressed
+using a set of libraries which bring as much as possible checking and
+validation to the type level, without sacrificing expressiveness.
+Together they form a toolkit for defining scalable cloud-based workflows
+composed of stateless computations, with a static reproducible
+specification of dependencies, behavior and wiring of all steps. The
+modeling of taxonomy data is done using Bio4j, where the new paradigm of
+graph databases allows for both a simple expression of taxonomic
+assignment tasks and the calculation of taxa abundance values
+considering the hierarchic structure of the taxonomy tree. MG7 includes
+a new 16S reference database, \emph{16S-DB7}, built with a flexible and
+sustainable update system, and the possibility of project-driven
+personalization.
 
 \(^\dagger\) The first and second authors contributed equally to this
-work
+work.
 
@@
 
-\subsubsection{16S Reference Database
-Construction}\label{s-reference-database-construction}
+\subsubsection{16S-DB7 Reference Database
+Construction}\label{s-db7-reference-database-construction}
 
-Our 16S Reference Database is a curated subset of sequences from NCBI
-nucleotide database \textbf{nt}. The sequences included were selected by
-similarity with the bacterial and archaeal reference sequences
-downloaded from the \textbf{RDP database} \citep{cole2013ribosomal}. RDP
-unaligned sequences were used to capture new 16S RNA sequences from
-\textbf{nt} using BLAST similarity search strategies and then,
-performing additional curation steps to remove sequences with poor
-taxonomic assignments to taxonomic nodes close to the root of the
-taxonomy tree. All the nucleotide sequences included in \textbf{nt}
-database has a taxonomic assignment provided by the \textbf{Genbank}
-sequence submitter. NCBI provides a table (available at
+Our 16S-DB7 Reference Database is a curated subset of sequences from the
+NCBI nucleotide database \textbf{nt}. The sequences included were
+selected by similarity with the bacterial and archaeal reference
+sequences downloaded from the \textbf{RDP database}
+\citep{cole2013ribosomal}. RDP unaligned sequences were used to capture
+new 16S RNA sequences from \textbf{nt} using BLAST similarity search
+strategies and then, performing additional curation steps to remove
+sequences with poor taxonomic assignments to taxonomic nodes close to
+the root of the taxonomy tree. All the nucleotide sequences included in
+\textbf{nt} database has a taxonomic assignment provided by the
+\textbf{Genbank} sequence submitter. NCBI provides a table (available at
 ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
@@
 global nucleotide databases in the near future. Another advantageous
-point is that we are based on NCBI taxonomy, the \emph{de facto}
+point is that we are based on NCBI taxonomy, the \emph{de-facto}
 standard taxonomic classification for biomolecular data

### 2015-09-25 src/main/scala/metagenomica/configuration.scala
Changed flash data config; commented out loquats that are not ready yet

@@ -17,30 +17,36 @@ case object configuration {
   implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
     new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
 
+
   trait AnyMetagenomicaData {
 
     // FLASH
 
     type ReadsType <: AnyReadsType { type EndType = pairedEndType }
-    val readsType: ReadsType
+    val  readsType: ReadsType
 
-    type Reads1 <: AnyPairedEnd1Fastq { type DataType = ReadsType }
-    val reads1: Reads1
+    type Reads1 >: PairedEnd1Fastq[ReadsType]
+                <: PairedEnd1Fastq[ReadsType]
+    lazy val reads1: Reads1 = new PairedEnd1Fastq(readsType, "reads1.fastq.gz")
 
-    type Reads2 <: AnyPairedEnd2Fastq { type DataType = ReadsType }
-    val reads2: Reads2
+    type Reads2 >: PairedEnd2Fastq[ReadsType]
+                <: PairedEnd2Fastq[ReadsType]
+    lazy val reads2: Reads2 = new PairedEnd2Fastq(readsType, "reads2.fastq.gz")
 
-    type FlashInput = Reads1 :^: Reads2 :^: DNil
-    val  flashInput: FlashInput
 
-    type Merged <: MergedReads[ReadsType, Reads1, Reads2]
-    val  merged: Merged
+    // TODO: make it free
+    lazy val flashOptions = flash.defaults update (
+      read_len(readsType.length.length)   :~:
+      max_overlap(readsType.length.length) :~: ∅
+    )
 
-    type Stats <: MergedReadsStats[Merged]
-    val stats: Stats
+    type Merged >: MergedReads[ReadsType, Reads1, Reads2]
+                <: MergedReads[ReadsType, Reads1, Reads2]
+    lazy val merged: Merged = new MergedReads(readsType, reads1, reads2, flashOptions)
 
-    type FlashOutput = Merged :^: Stats :^: DNil
-    val  flashOutput: FlashOutput
+    type Stats >: MergedReadsStats[Merged]
+               <: MergedReadsStats[Merged]
+    lazy val stats: Stats = new MergedReadsStats(merged)
 
 
     // BLAST
@@ -58,45 +64,40 @@ case object configuration {
     val  blastOutput: BlastOutput
   }
 
-  abstract class MetagenomicaData [
-    RT <: AnyReadsType { type EndType = pairedEndType }
[…]

### 2015-09-25 src/main/scala/metagenomica/loquats/assignment/dataProcessing.scala

@@ -1,140 +1,140 @@
-package ohnosequences.metagenomica.loquats.assignment
-
-import  ohnosequences.metagenomica._
-
-import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
-
-import ohnosequences.loquat.dataProcessing._
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.instructions._
-import ohnosequencesBundles.statika.Blast
-import ohnosequences.blast._, api._, data._
-import ohnosequences.cosas._, typeSets._, types._, properties._
-import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-import java.io.{ BufferedWriter, FileWriter, File }
-import ohnosequences.fastarious._, fasta._, fastq._
-import ohnosequences.blast._, api._, data._, outputFields._
-
-import scala.util.Try
-
-
-case object dataProcessing {
-
-  // TODO: move it to the config
-  case object CSVDataType extends AnyDataType { val label = "fastq" }
-  case object lcaCSV extends Data(CSVDataType, "lca.csv")
-  case object bbhCSV extends Data(CSVDataType, "bbh.csv")
-
-  // type BlastOutput = loquats.blast.blastTest.blastOutput.type
-  val  blastOutput = loquats.blast.blastTest.blastOutput
-
-  type BlastRecord = loquats.blast.blastDataProcessing.outRec.type
-  val blastRecord: BlastRecord = loquats.blast.blastDataProcessing.outRec
-
-  val headers: Seq[String] = blastRecord.properties.mapToList(typeLabel)
-
-  // TODO: move it somewhere up for global use
-  type ID = String
-
-  type GI = ID
-  type TaxID = ID
-  type ReadID = ID
-  type NodeID = ID
-
-  type LCA = Option[NodeID]
-  type BBH = Option[NodeID]
-
-  // TODO: move it somewhere up for global use
-  def parseInt(str: String): Option[Int] = Try(str.toInt).toOption
-
-  // this method looks up particular column by its header
-  def column(row: Seq[String], header: AnyOutputField): Option[String] =
-    headers.zip(row).toMap.get(header.label)
-
-  // FIXME: this thing should be used globally (ideally it should be in datasets)
-  implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
-    new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
-
-
[…]

### 2015-09-25 src/main/scala/metagenomica/loquats/assignment/test.scala

@@ -1,76 +1,76 @@
-package ohnosequences.metagenomica.loquats.assignment
-
-import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
-
-import ohnosequences.cosas._, typeSets._, types._, properties._, records._
-import ohnosequences.cosas.ops.typeSets._
-
-import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
-
-import ohnosequences.blast._
-import ohnosequences.blast.api._, outputFields._
-import ohnosequences.blast.data._
-
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.aws.api._
-import ohnosequences.statika.aws.amazonLinuxAMIs._
-
-import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
-import ohnosequences.awstools.s3._
-import ohnosequences.awstools.autoscaling._
-
-import era7.project.loquats._
-
-import java.io.File
-
-import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
-
-
-case object assignmenTest {
-
-  case object testConfig extends Era7LoquatConfig {
-
-    val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
-
-    val managerConfig = ManagerConfig(
-      instanceType = m3.medium,
-      purchaseModel = SpotAuto
-    )
-
-    val workersConfig = WorkersConfig(
-      instanceType = m3.medium,
-      purchaseModel = SpotAuto,
-      groupSize = WorkersGroupSize(0, 1, 10)
-    )
-
-    val terminationConfig = TerminationConfig(
-      terminateAfterInitialDataMappings = true
-    )
-
-    val dataMappings: List[AnyDataMapping] =
-      List(
-        DataMapping(
-          "ERR567374_1",
-          assignmentDataProcessing
-        )(remoteInput =
-            blastOutput.inS3Object(S3Object(
-              "resources.ohnosequences.com",
-              "16s/public-datasets/PRJEB6592/blast-test/ERR567374_1.blast.partial.csv"
-            )) :~: ∅,
-          remoteOutput =
-            lcaCSV.inS3Object(S3Object(
-              "resources.ohnosequences.com",
-              "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.lca.csv"
-            )) :~:
-            bbhCSV.inS3Object(S3Object(
-              "resources.ohnosequences.com",
[…]

### 2015-09-25 src/main/scala/metagenomica/loquats/counting/dataProcessing.scala

@@ -1,114 +1,114 @@
-package ohnosequences.metagenomica.loquats.counting
-
-import ohnosequences.metagenomica._
-import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
-
-import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
-
-import ohnosequences.loquat.dataProcessing._
-import ohnosequences.loquat.utils._
-
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.instructions._
-import ohnosequencesBundles.statika.Blast
-import ohnosequences.blast._, api._, data._
-import ohnosequences.cosas._, typeSets._, types._, properties._
-import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-import java.io.{ BufferedWriter, FileWriter, File }
-import ohnosequences.fastarious._, fasta._, fastq._
-import ohnosequences.blast._, api._, data._, outputFields._
-
-import scala.util.Try
-
-
-case object dataProcessing {
-
-  case object lcaCountsCSV extends Data(CSVDataType, "lca.counts.csv")
-  case object bbhCountsCSV extends Data(CSVDataType, "bbh.counts.csv")
-
-  // returns count of the given id and a filtered list (without that id)
-  def count(id: ID, list: List[ID]): (Int, List[ID]) =
-    list.foldLeft( (0, List[ID]()) ) { case ((count, rest), next) =>
-      if (next == id) (count + 1, rest)
-      else (count, next :: rest)
-    }
-
-  // TODO: make it tailrec
-  def directCounts(taxIds: List[TaxID]): Map[TaxID, Int] = {
-
-    @scala.annotation.tailrec
-    def rec(list: List[TaxID], acc: Map[TaxID, Int]): Map[TaxID, Int] =
-      list match{
-        case Nil => acc
-        case h :: t => {
-          val (n, rest) = count(h, t)
-          rec(rest, acc.updated(h, n + 1))
-        }
-      }
-
-    rec(taxIds, Map[TaxID, Int]())
-  }
-
-  // TODO: figure out some more effective algorithm
-  // Caution: it uses bio4j bundle!
-  def accumulatedCounts(counts: Map[TaxID, Int]): Map[TaxID, (Int, Int)] = {
-    counts.foldLeft(
-      Map[TaxID, (Int, Int)]()
-    ) { case (acc, (id, count)) =>
[…]

### 2015-09-25 src/main/scala/metagenomica/loquats/counting/test.scala

@@ -1,81 +1,81 @@
-package ohnosequences.metagenomica.loquats.counting
-
-import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
-
-import ohnosequences.cosas._, typeSets._, types._, properties._, records._
-import ohnosequences.cosas.ops.typeSets._
-
-import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
-
-import ohnosequences.blast._
-import ohnosequences.blast.api._, outputFields._
-import ohnosequences.blast.data._
-
-import ohnosequences.statika.bundles._
-import ohnosequences.statika.aws.api._
-import ohnosequences.statika.aws.amazonLinuxAMIs._
-
-import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
-import ohnosequences.awstools.s3._
-import ohnosequences.awstools.autoscaling._
-
-import era7.project.loquats._
-
-import java.io.File
-
-import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
-import ohnosequences.metagenomica.loquats.counting.dataProcessing._
-
-
-case object countingTest {
-
-  case object testConfig extends Era7LoquatConfig {
-
-    val metadata: AnyArtifactMetadata = generated.metadata.Metagenomica
-
-    val managerConfig = ManagerConfig(
-      instanceType = m3.medium,
-      purchaseModel = SpotAuto
-    )
-
-    val workersConfig = WorkersConfig(
-      instanceType = m3.medium,
-      purchaseModel = SpotAuto,
-      groupSize = WorkersGroupSize(0, 1, 10)
-    )
-
-    val terminationConfig = TerminationConfig(
-      terminateAfterInitialDataMappings = true
-    )
-
-    val dataMappings: List[AnyDataMapping] =
-      List(
-        DataMapping(
-          "ERR567374_1",
-          countingDataProcessing
-        )(remoteInput =
-            lcaCSV.inS3Object(S3Object(
-              "resources.ohnosequences.com",
-              "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.lca.csv"
-            )) :~:
-            bbhCSV.inS3Object(S3Object(
-              "resources.ohnosequences.com",
-              "16s/public-datasets/PRJEB6592/assignment-test/ERR567374_1.bbh.csv"
-            )) :~: ∅,
-          remoteOutput =
-            lcaCountsCSV.inS3Object(S3Object(
[…]

### 2015-09-25 src/main/scala/metagenomica/loquats/flash/config.scala

@@ -10,7 +10,8 @@ import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessin
 import ohnosequences.statika.bundles._
 import ohnosequences.statika.aws._, api._, amazonLinuxAMIs._
 
-import ohnosequences.awstools._, regions.Region._, s3.ObjectAddress, ec2.InstanceType._
+import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
+import ohnosequences.awstools.s3._
 import ohnosequences.awstools.autoscaling._
 
 import ohnosequences.flash.api._

### 2015-09-25 src/main/scala/metagenomica/loquats/flash/dataProcessing.scala

@@ -19,24 +19,15 @@ import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads.
 import java.io.File
 
 
-class FlashDataProcessing[
-  RT <: AnyReadsType { type EndType = pairedEndType },
-  R1 <: AnyPairedEnd1Fastq { type DataType = RT },
-  R2 <: AnyPairedEnd2Fastq { type DataType = RT },
-  M  <: MergedReads[RT, R1, R2],
-  S  <: MergedReadsStats[M]
-](val readsType: RT,
-  val reads1: R1,
-  val reads2: R2,
-  val merged: M,
-  val stats: S
-)(implicit
-  parseInputFiles: ParseDenotations[(R1 :^: R2 :^: DNil)#LocationsAt[FileDataLocation], File],
-  outputFilesToMap: ToMap[(M :^: S :^: DNil)#LocationsAt[FileDataLocation], AnyData, FileDataLocation]
-) extends DataProcessingBundle(bundles.flash)(
-  input = reads1 :^: reads2 :^: DNil,
-  output = merged :^: stats :^: DNil
-)(parseInputFiles, outputFilesToMap) {
+trait AnyFlashDataProcessing extends AnyDataProcessingBundle {
+
+  type MD <: AnyMetagenomicaData;
+  val md: MD
+
+  val bundleDependencies: List[AnyBundle] = List( bundles.flash )
+
+  type Input = MD#Reads1 :^: MD#Reads2 :^: DNil
+  type Output = MD#Merged :^: MD#Stats :^: DNil
 
   def instructions: AnyInstructions = say("I'll be fast as a flash!")
 
@@ -46,8 +37,8 @@ class FlashDataProcessing[
     context: Context
   ): Instructions[OutputFiles] = {
 
-    val reads1gz: file = context.file(reads1)
-    val reads2gz: file = context.file(reads2)
+    val reads1gz: file = context.file(md.reads1: MD#Reads1)
+    val reads2gz: file = context.file(md.reads2: MD#Reads2)
 
     val reads1fastq: file = reads1gz.rename( _.stripSuffix(".gz") )
     val reads2fastq: file = reads2gz.rename( _.stripSuffix(".gz") )
@@ -61,18 +52,13 @@ class FlashDataProcessing[
     // define output
     lazy val flashOutput = FlashOutputAt(context / "output", prefix = "")
 
-    lazy val flashOptions = flash.defaults update (
[…]

### 2015-09-25 metapasta/metapastatest/src/main/scala/configuration.scala
remove legacy metapasta code

@@ -1,71 +1,0 @@
-package metapastatest
-
-import ohnosequences.awstools.s3.ObjectAddress
-import ohnosequences.nisperon.bundles.NisperonMetadataBuilder
-import ohnosequences.metapasta._
-import ohnosequences.awstools.autoscaling._
-import ohnosequences.nisperon._
-import ohnosequences.awstools.ec2._
-
-object mockSamples {
-  val testBucket = "metapasta-test"
-
-  val ss2 = "supermock3"
-  val s2 = PairedSample(ss2, ObjectAddress(testBucket, "mock/" + ss2 + ".fastq"), ObjectAddress(testBucket, "mock/" + ss2 + ".fastq"))
-
-  val samples = List(s2)
-}
-
-object configuration extends BlastConfiguration (
-  metadataBuilder = new NisperonMetadataBuilder(new generated.metadata.metapastatest()),
-  email = "museeer@gmail.com",
-  mappingWorkers = Group(size = 1, max = 20, instanceType = InstanceType.T1Micro, purchaseModel = OnDemand),
-  uploadWorkers = None,
-  samples = mockSamples.samples,
-  logging = true,
-  database = NTDatabase,
-  xmlOutput = true
-)
-
-object metapastatest extends Metapasta(configuration) {
-
-  def check(test: Boolean, description: String) {
-    if (!test) {
-      throw new Error(description)
-    }
-  }
-
-  override def undeployActions(solved: Boolean): Option[String] = {
-    super.undeployActions(solved)
-    //test xmls
-
-    try {
-      val xml1 = ObjectAddress(nisperonConfiguration.bucket, "map/" + mockSamples.ss2 + ".1.1/out.blast")
-      val xml2 = ObjectAddress(nisperonConfiguration.bucket, "map/" + mockSamples.ss2 + ".1.2/out.blast")
-
-      val query5 = "Query_5"
-      val query6 = "Query_6"
-      val query25 = "Query_25"
-      val query26 = "Query_26"
-
-      println("checking " + xml1)
-      check(aws.s3.readWholeObject(xml1).contains(query25), "BLAST xml " + xml1 + " should contain " + query25)
-      check(!aws.s3.readWholeObject(xml1).contains(query26), "BLAST xml " + xml1 + " shouldn't contain " + query26)
-      check(aws.s3.readWholeObject(xml2).contains(query5), "BLAST xml " + xml2 + " should contain " + query5)
[…]

### 2015-09-25 metapasta/src/main/scala/ohnosequences/formats/DynamoValue.scala

@@ -1,30 +1,0 @@
-package ohnosequences.formats
-
-import com.amazonaws.services.dynamodbv2.model.AttributeValue
-
-case class DynamoValue(value: String, tpe: String) {
-  def attributeValue = tpe match {
-    case DynamoValue.Num => new AttributeValue().withN(value)
-    case DynamoValue.Str => new AttributeValue().withS(value)
-    case _ => throw new Error("unknown type for DynamoValue: " + tpe)
-  }
-}
-
-object DynamoValue {
-  val Str = "s"
-  val Num = "n"
-
-  def long(n: Long): DynamoValue = new DynamoValue(n.toString, Num)
-  def string(s: String): DynamoValue = new DynamoValue(s, Str)
-
-  def fromAttributeValue(v: AttributeValue): DynamoValue = {
-    if(v.getN != null) {
-      DynamoValue(v.getN, Num)
-    } else if (v.getS != null) {
-      DynamoValue(v.getS, Str)
-    } else {
-      throw new Error("unknown type for AttributeValue: " + v)
-    }
-
-  }
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/formats/FASTAUtils.scala

@@ -1,29 +1,0 @@
-package ohnosequences.formats
-
-
-
-object FASTAUtils {
-  def main2(args: Array[String]) {
-    val fasta = io.Source.fromFile(args(0))
-    val output = new java.io.PrintWriter(args(1))
-    var c = 0
-    var totalSeq = 0L
-    var totalLen = 0L 
-    fasta.getLines().foreach { line =>
-      c += 1
-      if (c % 10000 == 0) {
-        println("processed " + c + " lines")
-      }
-      if (line.startsWith(">")) {
-        output.println("\n" + line)
-        totalSeq += 1
-      } else {
-        totalLen += line.size
-        output.print(line)
-      }
-    }
-    output.close()
-    println("total seq: " + totalSeq)
-    println("total len: " + totalLen)
-  }
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/Assigner.scala

@@ -1,236 +1,0 @@
-package ohnosequences.metapasta
-
-import scala.collection.mutable
-import ohnosequences.metapasta.databases.{GIMapper, Database16S}
-import ohnosequences.formats.{RawHeader, FASTQ}
-import ohnosequences.nisperon.{AWS, MapMonoid}
-import ohnosequences.nisperon.logging.{Logger, S3Logger}
-
-
-case class ReadId(readId: String)
-case class RefId(refId: String)
-case class Hit(readId: ReadId, refId: RefId, score: Double)
-
-
-trait AssignerAlgorithm {
-
-
-  def assignAll(taxonomyTree: Tree[Taxon], hits: List[Hit], reads: List[FASTQ[RawHeader]], getTaxIds: (List[Hit], Logger) =>  (List[(Hit, Taxon)], Set[RefId]), logger: Logger): (mutable.HashMap[ReadId, Assignment], mutable.HashSet[RefId]) = {
-    val result = new mutable.HashMap[ReadId, Assignment]()
-
-    val wrongRefIdsAll = new mutable.HashSet[RefId]()
-
-    val hitsPerReads: Map[ReadId, List[Hit]] = hits.groupBy(_.readId)
-
-    for ( (readId, hits) <- hitsPerReads) {
-      val (hitsTaxa, wrongRefIds) = getTaxIds(hits, logger)
-      wrongRefIdsAll ++= wrongRefIds
-      if (hitsTaxa.isEmpty) {
-        result.put(readId, NoTaxIdAssignment(hits.map(_.refId).toSet))
-      } else {
-        result.put(readId, assign(taxonomyTree, hitsTaxa, logger))
-      }
-
-    }
-    (result, wrongRefIdsAll)
-
-  }
-
-  def assign(taxonomyTree: Tree[Taxon], hitsTaxa: List[(Hit, Taxon)], logger: Logger): Assignment
-}
-
-class LCAAlgorithm(assignmentConfiguration: AssignmentConfiguration) extends AssignerAlgorithm {
-  override def assign(taxonomyTree: Tree[Taxon], hitsTaxa: List[(Hit, Taxon)], logger: Logger): Assignment = {
-
-    val maxScore = hitsTaxa.map(_._1.score).max
-
-    hitsTaxa.filter { case (hit, taxon) =>
-      hit.score >= assignmentConfiguration.bitscoreThreshold &&
-      hit.score >= assignmentConfiguration.p * maxScore
[…]

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/Factory.scala

@@ -1,5 +1,0 @@
-package ohnosequences.metapasta
-
-trait Factory[C, +T] {
-  def build(ctx: C): T
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/FastasWriter.scala

@@ -1,84 +1,0 @@
-package ohnosequences.metapasta
-
-import ohnosequences.formats.{RawHeader, FASTQ}
-import scala.collection.mutable
-import ohnosequences.nisperon.logging.{Logger, S3Logger}
-import ohnosequences.nisperon.AWS
-import ohnosequences.awstools.s3.ObjectAddress
-import ohnosequences.metapasta.reporting.SampleId
-
-
-case class ChunkId(sample: SampleId, start: Long, end: Long)
-
-object ChunkId {
-  def apply(chunk: MergedSampleChunk): ChunkId = ChunkId(SampleId(chunk.sample), chunk.range._1, chunk.range._2)
-}
-
-class FastasWriter(aws: AWS, readsDirectory: ObjectAddress, nodeRetriever: NodeRetriever) {
-  val noHitFasta = new mutable.StringBuilder()
-  val noTaxIdFasta = new mutable.StringBuilder()
-  val notAssignedFasta = new mutable.StringBuilder()
-  val assignedFasta = new mutable.StringBuilder()
-
-  def fastaHeader(sampleId: String, taxon: Taxon, refIds: Set[RefId], reason: Option[String]): String = {
-
-    val (taxname, taxonId, rank) = reason match {
-      case Some(reason) => {
-        (reason, "", "")
-      }
-      case None => {
-        try {
-          val node = nodeRetriever.nodeRetriever.getNCBITaxonByTaxId(taxon.taxId)
-          (node.getScientificName(), taxon.taxId, node.getRank())
-        } catch {
-          case t: Throwable => ("", taxon.taxId, "")
-        }
-      }
-    }
-
-
-    sampleId + "|" + taxname + "|" + taxonId + "|" + rank + "|" + refIds.foldLeft("")(_ + "_" + _.refId) + "|"
-  }
-
-
-
-  def write(sample: SampleId, read: FASTQ[RawHeader], readId: ReadId, assignment: Assignment) {
-    assignment match {
-      case TaxIdAssignment(taxon, refIds, _, _, _) => {
-        assignedFasta.append(read.toFasta(fastaHeader(sample.id, taxon, refIds, None)))
-        assignedFasta.append(System.lineSeparator())
-      }
-      case NoTaxIdAssignment(refIds) => {
[…]

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/ReadsInfo.scala

@@ -1,47 +1,0 @@
-package ohnosequences.metapasta
-
-import com.amazonaws.services.dynamodbv2.model.{ScalarAttributeType, AttributeDefinition, AttributeValue}
-
-case class ReadInfo(readId: String, gi: String, sequence: String, quality: String, sample: String, chunk: String, tax: String) {
-
-  import ReadInfo._
-
-  def chunkId(c: Int) = c + "-" + chunk
-
-  def toDynamoItem(c: Int): java.util.Map[String, AttributeValue] = {
-    val r = new java.util.HashMap[String, AttributeValue]()
-
-    r.put(idAttr, new AttributeValue().withS(readId))
-    r.put(sequenceAttr, new AttributeValue().withS(sequence))
-    r.put(qualityAttr, new AttributeValue().withS(quality))
-    r.put(chunkAttr, new AttributeValue().withS(chunkId(c)))
-    if(!gi.isEmpty) {
-      r.put(giAttr, new AttributeValue().withS(gi))
-    } else {
-      r.put(giAttr, new AttributeValue().withS(unassigned))
-    }
-
-    if(!tax.isEmpty) {
-      r.put(taxAttr, new AttributeValue().withS(tax))
-    } else {
-      r.put(taxAttr, new AttributeValue().withS(unassigned))
-    }
-    r
-
-  }
-}
-
-object ReadInfo {
-  val unassigned = "unassigned"
-
-  val idAttr = "header"
-  val sequenceAttr = "seq"
-  val qualityAttr = "qual"
-  val giAttr = "gi"
-  val taxAttr = "tax"
-  val chunkAttr = "chunk"
-
-  val hash = new AttributeDefinition().withAttributeName(chunkAttr).withAttributeType(ScalarAttributeType.S)
-  val range = new AttributeDefinition().withAttributeName(idAttr).withAttributeType(ScalarAttributeType.S)
-
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/S3Paths.scala

@@ -1,57 +1,0 @@
-package ohnosequences.metapasta
-
-import ohnosequences.awstools.s3.ObjectAddress
-import ohnosequences.metapasta.reporting.SampleId
-
-
-object S3Paths {
-
-  def mergedFastq(resultsObject: ObjectAddress, sample: String): ObjectAddress = resultsObject / sample / "reads" / (sample + ".merged.fastq")
-
-  def notMergedFastq(resultsObject: ObjectAddress, sample: String): (ObjectAddress, ObjectAddress) = {
-    (resultsObject / sample / "reads" / (sample + ".notMerged1.fastq"),  resultsObject / sample / "reads" / (sample + ".notMerged2.fastq"))
-  }
-
-  def mergedNoHitFasta(resultsObject: ObjectAddress, sample: String) : ObjectAddress =    resultsObject / sample / "reads" / (sample + ".noHit.fasta")
-  def mergedNoTaxIdFasta(resultsObject: ObjectAddress, sample: String, assignmentType: AssignmentType) : ObjectAddress =    resultsObject / sample / "reads" / (sample +"." + assignmentType + ".noTaxId.fasta")
-
-  def mergedAssignedFasta(resultsObject: ObjectAddress, sample: String, assignmentType: AssignmentType) : ObjectAddress = resultsObject / sample / "reads" / (sample + "." + assignmentType + ".assigned.fasta")
-  def mergedNotAssignedFasta(resultsObject: ObjectAddress, sample: String, assignmentType: AssignmentType):ObjectAddress =resultsObject / sample / "reads" / (sample + "." + assignmentType + ".notAssigned.fasta")
-
-
-  def noHitFastas(readsObject: ObjectAddress, sample: String) : ObjectAddress = readsObject / sample / "noHit"
-
-  def noTaxIdFastas(readsObject: ObjectAddress, sample: String, assignmentType: AssignmentType) : ObjectAddress = readsObject / (sample + "###" + assignmentType) / "noTaxId"
-  def notAssignedFastas(readsObject: ObjectAddress, sample: String, assignmentType: AssignmentType) : ObjectAddress = readsObject / (sample + "###" + assignmentType) / "notAssigned"
-  def assignedFastas(readsObject: ObjectAddress, sample: String, assignmentType: AssignmentType): ObjectAddress = readsObject / (sample + "###" + assignmentType) / "assigned"
[…]

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/databases/BlastDatabase.scala

@@ -1,32 +1,0 @@
-package ohnosequences.metapasta.databases
-
-import ohnosequences.nisperon.AWS
-import ohnosequences.metapasta.Factory
-import org.clapper.avsl.Logger
-import ohnosequences.awstools.s3.{LoadingManager, ObjectAddress}
-import java.io.File
-
-object Blast16SFactory extends DatabaseFactory[BlastDatabase16S] {
-
-
-  val logger = Logger(this.getClass)
-
-  class BlastDatabase extends BlastDatabase16S {
-    val name: String = "nt.march.14.blast/nt.march.14.fasta"
-
-    //gi|313494140|gb|GU939576.1|
-    val re = """\Qgi|\E(\d+)[^\d]+.*""".r
-    def parseGI(refId: String) : Option[String] = {
-      refId match {
-        case re(id) => Some(id)
-        case s => None
-      }
-    }
-  }
-
-  override def build(loadingManager: LoadingManager): BlastDatabase = {
-    logger.info("downloading database")
-    loadingManager.downloadDirectory(ObjectAddress("metapasta", "nt.march.14.blast"), new File("."))
-    new BlastDatabase()
-  }
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/databases/GiMapper.scala

@@ -1,40 +1,0 @@
-package ohnosequences.metapasta.databases
-
-import java.io.File
-import ohnosequences.awstools.s3.{LoadingManager, ObjectAddress}
-import scala.collection.mutable
-import org.clapper.avsl.Logger
-import ohnosequences.metapasta.{Taxon, Factory}
-
-/**
- * todo it will moved to bio4j
- */
-trait GIMapper {
-  def getTaxIdByGi(gi: String): Option[Taxon]
-}
-
-class InMemoryGIMapperFactory() extends DatabaseFactory[GIMapper] {
-
-  val logger = Logger(this.getClass)
-
-  class InMemoryGIMapper(map: mutable.HashMap[String, Taxon]) extends GIMapper {
-    override def getTaxIdByGi(gi: String): Option[Taxon] = map.get(gi)
-  }
-
-
-  override def build(loadingManager: LoadingManager): GIMapper = {
-    val mapping = new mutable.HashMap[String, Taxon]()
-    val mappingFile = new File("gi.map")
-    loadingManager.download(ObjectAddress("metapasta", "gi.map"), mappingFile)
-    val giP = """(\d+)\s+(\d+).*""".r
-    for(line <- io.Source.fromFile(mappingFile).getLines()) {
-      line match {
-        case giP(gi, tax) => mapping.put(gi, Taxon(tax))
-        case l => logger.error("can't parse " + l)
-      }
-    }
-
-    new InMemoryGIMapper(mapping)
-  }
-
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/databases/LastDatabase.scala

@@ -1,30 +1,0 @@
-package ohnosequences.metapasta.databases
-
-import ohnosequences.nisperon.AWS
-import ohnosequences.metapasta.Factory
-import ohnosequences.awstools.s3.{LoadingManager, ObjectAddress}
-import java.io.File
-import org.clapper.avsl.Logger
-
-
-object Last16SFactory extends DatabaseFactory[LastDatabase16S] {
-
-  val logger = Logger(this.getClass)
-
-  class LastDatabase extends LastDatabase16S {
-    val name: String = "nt.march.14.last/nt.march.14"
-    //gi|313494140|gb|GU939576.1|
-    val re = """\Qgi|\E(\d+)[^\d]+.*""".r
-    def parseGI(refId: String): Option[String] = {
-      refId match {
-        case re(id) => Some(id)
-        case s => None
-      }
-    }
-  }
-  override def build(loadingManager: LoadingManager): LastDatabase16S = {
-    logger.info("downloading database")
-    loadingManager.downloadDirectory(ObjectAddress("metapasta", "nt.march.14.last"), new File("."))
-    new LastDatabase()
-  }
-}

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/instructions/BlastInstructions.scala

@@ -1,143 +1,0 @@
-package ohnosequences.metapasta.instructions
-
-import ohnosequences.nisperon._
-import org.clapper.avsl.Logger
-import java.io.{File, PrintWriter}
-import scala.collection.mutable.ListBuffer
-import ohnosequences.nisperon.logging.S3Logger
-import ohnosequences.metapasta._
-import ohnosequences.metapasta.databases.{DatabaseFactory, InMemoryGIMapperFactory, BlastDatabase16S}
-import ohnosequences.nisperon.bundles.NisperonMetadataBuilder
-import ohnosequences.metapasta.MergedSampleChunk
-import ohnosequences.parsers.S3ChunksReader
-import ohnosequences.formats.RawHeader
-import ohnosequences.formats.FASTQ
-import ohnosequences.metapasta.AssignTable
-import ohnosequences.awstools.s3.ObjectAddress
-
-
-class BlastInstructions(
-                       aws: AWS,
-                       metadataBuilder: NisperonMetadataBuilder, //for bio4j
-                       assignmentConfiguration: AssignmentConfiguration,
-                       databaseFactory: DatabaseFactory[BlastDatabase16S],
-                       blastCommandTemplate: String = """blastn -task megablast -db $db$ -query $input$ -out $output$ -max_target_seqs 1 -num_threads 1 -outfmt 6 -show_gis""",
-                       useXML: Boolean,
-                       logging: Boolean,
-                       resultDirectory: ObjectAddress,
-                       readsDirectory: ObjectAddress
-                       ) extends
-   MapInstructions[List[MergedSampleChunk],  (AssignTable, Map[(String, AssignmentType), ReadsStats])] {
-
-
-  case class BlastContext(nodeRetriever: NodeRetriever, database: BlastDatabase16S, blast: Blast, assigner: Assigner)
-  override type Context = BlastContext
-
-  val logger = Logger(this.getClass)
-
-  //todo think about this space
-  def extractHeader(s: String) = s.replace("@", "").split("\\s")(0)
-
-
-  override def prepare() = {
-    val lm = aws.s3.createLoadingManager()
-    val nodeRetreiver = new BundleNodeRetrieverFactory().build(metadataBuilder)
-    val blastDatabase = databaseFactory.build(lm)
[…]

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/reporting/Ranks.scala

@@ -1,39 +1,0 @@
-package ohnosequences.metapasta.reporting
-
-
-trait TaxonomyRank
-
-
-object TaxonomyRank {
-  val ranks = List(Genus, Phylum, Species, Class, Order, Superkingdom)
-}
-
-case object Genus extends TaxonomyRank {
-  override def toString: String = "genus"
-}
-
-case object Phylum extends TaxonomyRank {
-  override def toString: String = "phylum"
-}
-
-case object Species extends TaxonomyRank {
-  override def toString: String = "species"
-}
-
-case object Class extends TaxonomyRank {
-  override def toString: String = "class"
-}
-
-case object Order extends TaxonomyRank {
-  override def toString: String = "order"
-}
-
-case object Superkingdom extends TaxonomyRank {
-  override def toString: String = "superkingdom"
-}
-
-case object NoRank extends TaxonomyRank {
-  override def toString: String = "no rank"
-}
-
-

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/reporting/Reporter.scala

@@ -1,232 +1,0 @@
-package ohnosequences.metapasta.reporting
-
-import ohnosequences.metapasta._
-import org.clapper.avsl.Logger
-import ohnosequences.nisperon._
-import ohnosequences.awstools.s3.{S3, ObjectAddress}
-import ohnosequences.metapasta.ReadsStats
-import ohnosequences.metapasta.AssignTable
-import scala.collection.mutable
-import ohnosequences.metapasta.reporting.spreadsheeet.CSVExecutor
-import scala.collection.mutable.ListBuffer
-import java.io.{PrintWriter, File}
-
-
-case class PerSampleData(direct: Long, cumulative: Long)
-
-case class TaxonInfo(scientificName: String, rank: String)
-
-case class SampleId(id: String)
-
-case class SampleTag(tag: String)
-
-class Reporter(aws: AWS,
-               tablesAddresses: List[ObjectAddress],
-               statsAddresses: List[ObjectAddress],
-               tags: Map[SampleId, List[SampleTag]],
-               nodeRetriever: NodeRetriever,
-               destination: ObjectAddress,
-               projectName: String
-               ) {
-  val assignmentTableSerializer = ohnosequences.metapasta.assignTableSerializer
-  val statsMonoid = new MapMonoid[(String, AssignmentType), ReadsStats](readsStatsMonoid)
-  val statsSerializer = readsStatsSerializer
-
-  def read[T](address: ObjectAddress, serializer: Serializer[T]): T = {
-     serializer.fromString(aws.s3.readWholeObject(address))
-  }
-
-  val logger = Logger(this.getClass)
-
-  def generate() {
-    logger.info("reading assignment tables")
-
-    var table: AssignTable = assignTableMonoid.unit
-    for (obj <- tablesAddresses) {
-      logger.info("reading from " + obj)
-      val t = read[AssignTable](obj, assignmentTableSerializer)
-      table = assignTableMonoid.mult(table, t)
-    }
-
-    var stats: Map[(String, AssignmentType), ReadsStats] = statsMonoid.unit
-
-    logger.info("reading reads stats")
-    for (obj <- statsAddresses) {
-      logger.info("reading from " + obj)
-      val t = read[Map[(String, AssignmentType), ReadsStats]](obj, statsSerializer)
-      stats = statsMonoid.mult(stats, t)
-    }
-
-    val samplesS: Set[String] = stats.keySet.map(_._1)
[…]

### 2015-09-25 metapasta/src/main/scala/ohnosequences/metapasta/reporting/spreadsheeet/Attribute.scala

@@ -1,225 +1,0 @@
-package ohnosequences.metapasta.reporting.spreadsheeet
-
-import ohnosequences.nisperon.{longMonoid, doubleMonoid, intMonoid, Monoid}
-import scala.collection.mutable
-
-
-trait Context {
-  def get[A <: AnyAttribute](attribute: A, index: Int): attribute.Type
-  def getTotal[A <: AnyAttribute](attribute: A): attribute.Type
-  def set[A <: AnyAttribute](attribute: A, index: Int)(value: attribute.Type)
-}
-
-class ListContext(val attributes: List[AnyAttribute]) extends Context {
-
-  val map = new mutable.HashMap[(String, Int), Any]()
-
-  val totals = new mutable.HashMap[String, Any]()
-
-  for (attribute <- attributes) {
-    totals.put(attribute.name, attribute.monoid.unit)
-  }
-
-  override def set[A <: AnyAttribute](attribute: A, index: Int)(value: attribute.Type): Unit = {
-    val cur = totals.getOrElse(attribute.name, attribute.monoid.unit).asInstanceOf[attribute.Type]
-    totals.put(attribute.name, attribute.monoid.mult(cur, value))
-    map.put((attribute.name, index), value)
-  }
-
-  override def getTotal[A <: AnyAttribute](attribute: A): attribute.Type = {
-    totals.getOrElse(attribute.name, attribute.monoid.unit).asInstanceOf[attribute.Type]
-  }
-
-  override def get[A <: AnyAttribute](attribute: A, index: Int): attribute.Type = {
-    map.getOrElse((attribute.name, index), attribute.monoid.unit).asInstanceOf[attribute.Type]
-  }
-}
-
-case class StringConstantMonoid(c: String) extends Monoid[String] {
-  override def mult(x: String, y: String): String = c
-
-  override def unit: String = c
-}
-
-trait AnyAttribute {
-  type Type
-
-  val monoid: Monoid[Type]
-  val name: String
-  type Item
-
-  val hidden: Boolean
-  def execute(item: Item, index: Int, context: Context): Type
-
-  def printTotal(total: Type): String = total.toString
-
-}
-
-object AnyAttribute {
-  type For[T] = AnyAttribute { type Item = T }
-
-}
-
-abstract class LongAttribute[I](val name: String, val monoid: Monoid[Long], val hidden: Boolean = false) extends AnyAttribute {
-  override type Type = Long
-  override type Item = I
-}
-
[…]

### 2015-09-25 metapasta/src/test/scala/AssignerTests.scala

@@ -1,156 +1,0 @@
-
-import ohnosequences.formats.FASTQ
-import ohnosequences.formats.RawHeader
-import ohnosequences.formats.{RawHeader, FASTQ}
-import ohnosequences.metapasta._
-import ohnosequences.metapasta.AssignmentConfiguration
-import ohnosequences.metapasta.databases.{GIMapper, Blast16SFactory}
-
-import ohnosequences.metapasta.Hit
-import ohnosequences.metapasta.NoTaxIdAssignment
-import ohnosequences.metapasta.ReadId
-import ohnosequences.metapasta.RefId
-import ohnosequences.metapasta.reporting.SampleId
-import ohnosequences.metapasta.reporting.SampleId
-import ohnosequences.metapasta.Taxon
-import ohnosequences.nisperon.logging.ConsoleLogger
-import org.junit.Test
-import org.junit.Assert._
-import scala.Some
-
-
-class AssignerTests {
-  val blast16s = new Blast16SFactory.BlastDatabase()
-
-  val fakeTaxonomiTree = new MapTree(Map(
-    Taxon("2") -> Taxon("1"),
-    Taxon("3") -> Taxon("2"),
-    Taxon("4") -> Taxon("3"),
-    Taxon("4l") -> Taxon("3"),
-    Taxon("5l") -> Taxon("4"),
-    Taxon("5") -> Taxon("4"),
-    Taxon("5r") -> Taxon("4")
-  ), Taxon("1"))
-
-
-
-  //we use fake 16s and fake taxonomy where taxid = gi
-  val idGIMapper = new GIMapper {
-    override def getTaxIdByGi(gi: String): Option[Taxon] = {
-      if (fakeTaxonomiTree.map.contains(Taxon(gi)) || fakeTaxonomiTree.root.taxId.equals(gi) || gi.equals("20142015")) {
-        Some(Taxon(gi))
-      } else {
-        None
-      }
-    }
-  }
-
-  def extractHeader(header: String) = header
-
-  @Test
-  def assignmentTest1() {
-    val assignmentConfiguration = AssignmentConfiguration(100, 0.8)
-    val assigner = new Assigner(
-      taxonomyTree = fakeTaxonomiTree,
-      database = blast16s,
-      giMapper = idGIMapper,
-      assignmentConfiguration = assignmentConfiguration,
-      extractHeader,
-      None
-    )
-
-    val testSample = "test"
[…]

### 2015-09-25 metapasta/src/test/scala/OldTestCase.scala

@@ -1,53 +1,0 @@
-import java.io.File
-import ohnosequences.awstools.s3.{S3, ObjectAddress}
-
-import ohnosequences.nisperon.AWS
-import org.junit.Test
-import org.junit.Assert._
-
-class S3Splitter(s3: S3, address: ObjectAddress, chunksSize: Long) {
-
-  def objectSize(): Long = {
-    s3.s3.getObjectMetadata(address.bucket, address.key).getContentLength
-  }
-
-  def chunks(): List[(Long, Long)] = {
-
-    val size = objectSize()
-
-    val starts = 0L until size by chunksSize
-    val ends = starts.map { start =>
-      math.min(start + chunksSize - 1, size - 1)
-    }
-
-    starts.zip(ends).toList
-  }
-}
-
-class OldTestCase {
-
-
-
-//  @Test
-  def test {
-  val aws = new AWS(new File(System.getProperty("user.home"), "nispero.credentials"))
-    val s3 = aws.s3
-
-    val file = ObjectAddress("metapasta-test", "microtest.fastq")
-    val chunks = new S3Splitter(s3, file, 100000).chunks()
-    import ohnosequences.formats._
-    import ohnosequences.parsers._
-    val reader = S3ChunksReader(s3, file)
-    var left = chunks.size
-    var size = 0
-    for (chunk <- chunks) {
-      println(left + " chunks left")
-      left -= 1
-      val parsed: List[FASTQ[RawHeader]] = reader.parseChunk[RawHeader](chunk._1, chunk._2)._1
-      size += parsed.size
-    }
-    //print 10001 instead 10000 ( = amount of reads in microtest.fastq)
-    assertEquals(10000, size)
-    println(size)
-  }
-}

### 2015-09-25 metapasta/src/test/scala/ParserTest.scala

@@ -1,33 +1,0 @@
-package ohnosequences.metapasta.tests
-
-import org.junit.Test
-import org.junit.Assert._
-import ohnosequences.awstools.s3._
-import ohnosequences.formats._
-import ohnosequences.parsers._
-
-class ParserTest {
-
-  import TestReads._
-
-  //@Test
-  def duplicatesTest() = {
-    def noDuplicates[T](reads: List[T]): Boolean = {
-      val readsNumber = reads.length
-      val distinctNumber = reads.distinct.length
-      if (readsNumber != distinctNumber)
-        println(s"Of ${readsNumber} reads only ${distinctNumber} are distinct")
-      readsNumber == distinctNumber
-    }
-   // assert(noDuplicates(reads1000))
-   // assert(noDuplicates(reads10000))
-    assert(noDuplicates(reads100000))
-  }
-
-  //@Test // checking that its independent from the chunk size
-  def sameReadsNumberTest() = {
-    assert(reads1000.length == reads10000.length)
-    assert(reads100000.length == reads100000.length)
-  }
-
-}

### 2015-09-25 metapasta/src/test/scala/ohnosequences/metapasta/automatic/Serializers.scala

@@ -1,20 +1,0 @@
-package ohnosequences.metapasta.automatic
-
-import ohnosequences.metapasta._
-import ohnosequences.metapasta.reporting.{FileType, TaxonInfo}
-import org.scalacheck.Prop._
-import org.scalacheck.{Gen, Properties}
-
-import scala.collection.mutable
-
-object Serializers extends Properties("Serializers") {
-  property("assign table") = forAll (Gen.choose(1, 100)) { n =>
-    val assignTableRaw = new mutable.HashMap[Taxon, TaxInfo]()
-    assignTableRaw.put(NoTaxId.taxon, TaxInfo(n, n))
-    assignTableRaw.put(Taxon("uuuuu"), TaxInfo(n, n))
-
-    val assignTable = AssignTable(Map( ("sample1" -> LCA) -> assignTableRaw.toMap))
-
-    assignTableSerializer.fromString(assignTableSerializer.toString(assignTable)).equals(assignTable)
-  }
-}

### 2015-09-25 compota/src/main/scala/ohnosequences/nisperon/MetaManager.scala
remove legacy compota code

@@ -1,178 +1,0 @@
-package ohnosequences.nisperon
-
-import org.clapper.avsl.Logger
-import ohnosequences.nisperon.queues.{DefaultQueueMerger, QueueMerger, SQSQueue}
-import ohnosequences.nisperon.logging.FailTable
-import ohnosequences.nisperon.console.Server
-
-
-class MetaManager(nisperon: Nisperon) {
-  import nisperon._
-
-  val logger = Logger(this.getClass)
-
-  def printURL(domain: String, port: Int = 443): String = port match {
-    case 443 => "https://" + domain
-    case 80 => "http://" + domain
-    case p => domain + ":" + port
-  }
-
-  def currentAddress: String = {
-    aws.ec2.getCurrentInstance.flatMap {_.getPublicDNS()}.getOrElse("<undefined>")
-  }
-
-
-  //todo remove notification topic
- // todo use sync queue reader!!!
-  def run() {
-
-    val failTable = new FailTable(nisperon.aws, nisperon.nisperonConfiguration.errorTable)
-
-    try {
-      logger.info("starting console")
-      new Thread("console") {
-        override def start() {
-          val message = new StringBuilder()
-          message.append(nisperonConfiguration.id + " started\n")
-          message.append("\n")
-          message.append("console address: " + printURL(currentAddress) + "\n")
-          message.append("user: nispero\n")
-          message.append("password: " + nisperonConfiguration.password + "\n")
-          nisperon.notification(nisperonConfiguration.id + " started", message.toString())
-          new Server(nisperon).start()
-
-
-        }
-      }.start()
-    } catch {
-      case t: Throwable =>   Nisperon.reportFailure(nisperon.aws, nisperon.nisperonConfiguration, "metamanager", t, true, failTable, "console")
-    }
-
-    try {
-
-      logger.info("metamanager started")
-
-      val queueName = nisperon.nisperonConfiguration.metamanagerQueue
-
-      //todo is it needed?
-      val controlTopic = nisperon.aws.sns.createTopic(nisperon.nisperonConfiguration.controlTopic)
-
-      val queueWrap = nisperon.aws.sqs.createQueue(queueName)
-
-
-      val controlQueue = new SQSQueue[List[MetaManagerCommand0]](aws.sqs.sqs, queueName, new JsonSerializer[List[MetaManagerCommand0]]())
-      val reader = controlQueue.getSyncReader()
-
[…]

### 2015-09-25 compota/src/main/scala/ohnosequences/nisperon/TerminationDaemon.scala

@@ -1,66 +1,0 @@
-package ohnosequences.nisperon
-
-import org.clapper.avsl.Logger
-
-
-class TerminationDaemon(nisperon: Nisperon) extends Thread {
-  val logger = Logger(this.getClass)
-
-  val initLaunchTime = {
-
-    var ititTime= 0L
-    var attempt = 10
-    while (ititTime== 0 && attempt > 0) {
-      try {
-        attempt -= 1
-        ititTime = nisperon.launchTime
-
-      } catch {
-        case t: Throwable => t.printStackTrace()
-      }
-
-    }
-    ititTime
-
-  }
-  val t0 = System.currentTimeMillis() / 1000
-
-  def launchTime(): Long = {
-    val t1 = System.currentTimeMillis() / 1000
-    initLaunchTime + (t1 - t0)
-  }
-
-  override def run() {
-
-    logger.info("termination daemon started")
-    var stopped = false
-
-      while (!stopped) {
-
-
-        if (nisperon.nisperonConfiguration.autoTermination) {
-          logger.info("checking queues")
-          nisperon.checkQueues() match {
-            case Left(queue) => logger.info(queue.name + " isn't empty")
-            case Right(queues) => {
-              logger.info("all queues are empty. terminating")
-              nisperon.sendUndeployCommand("solved", force = false)
-              stopped = true
-            }
-          }
-        }
-
-        if (launchTime() > nisperon.nisperonConfiguration.timeout) {
-          logger.info("terminating due to timeout")
-          //?????
-          nisperon.sendUndeployCommand("timeout " + nisperon.nisperonConfiguration.timeout + " sec", force = true)
-          stopped = true
-        }
-
-        Thread.sleep(10000)
-      }
-
-
-
-  }
-}

### 2015-09-25 compota/src/main/scala/ohnosequences/nisperon/logging/FailTable.scala

@@ -1,127 +1,0 @@
-package ohnosequences.nisperon.logging
-
-import ohnosequences.nisperon.AWS
-import com.amazonaws.services.dynamodbv2.model._
-import java.util
-import java.text.SimpleDateFormat
-import java.util.Date
-import ohnosequences.awstools.ddb.Utils
-import org.clapper.avsl.Logger
-
-case class Failure(taskId: String, time: Long, instanceId: String, message: String) {
-  val format = new SimpleDateFormat("yyyy.MM.dd HH:mm:ss")
-
-  def formattedTime(): String = {
-    format.format(new Date(time))
-  }
-}
-
-//todo add sqs queue for this
-//todo add table with all success (or jsut retrive it)
-case class FailTable(aws: AWS, name: String) {
-  val logger = Logger(this.getClass)
-
-  val hashKeyName = "h"
-  val rangeKeyName = "r"
-  val bodyKeyName = "b"
-
-  val hashAttribute = new AttributeDefinition().withAttributeName(hashKeyName).withAttributeType(ScalarAttributeType.S)
-  val rangeAttribute = new AttributeDefinition().withAttributeName(rangeKeyName).withAttributeType(ScalarAttributeType.S)
-
-  def create() {
-    Utils.createTable(
-      ddb = aws.ddb,
-      tableName = name,
-      hash = hashAttribute,
-      range = Some(rangeAttribute),
-      logger = logger,
-      waitForCreation = false
-    )
-  }
-
-  //todo add call for this
-  def delete() {
-    Utils.deleteTable(aws.ddb, name)
-  }
-
-  def hashKey(taskId: String): AttributeValue = {
-    new AttributeValue().withS(taskId)
-  }
-
-  def rangeKey(failure: Failure): AttributeValue = {
-    //val timeStamp = System.currentTimeMillis()
-    new AttributeValue().withS(failure.instanceId + ":" + failure.time)
-  }
-
-
-
-  def fail(taskId: String, instanceId: String, message: String) = {
-    val failure = Failure(taskId, System.currentTimeMillis(), instanceId, message)
-    val item = new util.HashMap[String, AttributeValue]()
-    item.put(hashKeyName, hashKey(failure.taskId))
-    item.put(rangeKeyName, rangeKey(failure))
-    item.put(bodyKeyName,  new AttributeValue().withS(message))
-    aws.ddb.putItem(name, item)
-  }
-
[…]

### 2015-09-26 src/main/scala/metagenomica/configuration.scala
Moved things from blast loquat to the global config

@@ -51,53 +51,52 @@ case object configuration {
 
     // BLAST
 
-    type BlastRecord <: AnyBlastOutputRecord
-    val  blastRecord: BlastRecord
+    type BlastOutRec <: AnyBlastOutputRecord
+    val  blastOutRec: BlastOutRec
 
+    type BlastExprType >: BlastExpressionType[blastn, BlastOutRec]
+                       <: BlastExpressionType[blastn, BlastOutRec]
+    val  blastExprType: BlastExprType
 
-    type BlastOutput <: AnyBlastOutput
-    // {
-    //   type BlastExpressionType <: AnyBlastExpressionType {
-    //     type OutputRecord = BlastRecord
-    //   }
-    // }
-    val  blastOutput: BlastOutput
+    type BlastOutType >: BlastOutputType[BlastExprType]
+                      <: BlastOutputType[BlastExprType]
+    val  blastOutType: BlastOutType
+
+    def blastExpr(args: ValueOf[blastn.Arguments]): BlastExpression[BlastExprType]
+
+    type BlastOut >: BlastOutput[BlastOutType]
+                  <: BlastOutput[BlastOutType]
+    val  blastOut: BlastOut
   }
 
-  // abstract class MetagenomicaData [
-  //   RT <: AnyReadsType { type EndType = pairedEndType }
-  // ](val readsType: RT) extends AnyMetagenomicaData {
-  //
-  //   type ReadsType = RT
-  //
-  //   // case object reads1_ extends PairedEnd1Fastq(readsType, "reads1.fastq.gz")
-  //   // case object reads2_ extends PairedEnd2Fastq(readsType, "reads2.fastq.gz")
-  //
-  //   // type Reads1 = reads1_.type
-  //   // val  reads1 = reads1_
-  //   //
-  //   // type Reads2 = reads2_.type
-  //   // val  reads2 = reads2_
-  //
-  //   // val  flashInput = reads1 :^: reads2 :^: DNil
-  //
-  //   // case object merged_ extends MergedReads(readsType, reads1, reads2, flashOptions)
-  //   // type Merged = merged_.type
-  //   // val  merged = merged_
-  //   //
-  //   // case object stats_ extends MergedReadsStats[Merged](merged)
-  //   // type Stats = stats_.type
-  //   // val  stats = stats_
-  //   //
-  //   // val  flashOutput = merged :^: stats :^: DNil
-  //
[…]

### 2015-09-26 src/main/scala/metagenomica/loquats/blast/dataProcessing.scala

@@ -22,36 +22,10 @@ import java.io.{ BufferedWriter, FileWriter, File }
 import sys.process._
 
 
-case object blastDataProcessing {
+trait AnyBlastDataProcessing extends AnyDataProcessingBundle {
 
-  // TODO with great power comes great responsibility. Move to conf
-  case object outRec extends BlastOutputRecord(
-    qseqid    :&:
-    qlen      :&:
-    qstart    :&:
-    qend      :&:
-    sseqid    :&:
-    slen      :&:
-    sstart    :&:
-    send      :&:
-    bitscore  :&:
-    sgi       :&: □
-  )
-  case object blastExprType extends BlastExpressionType(blastn)(outRec)
-  case object blastOutputType extends BlastOutputType(blastExprType, "blastn.blablabla")
-
-  private def blastExpr(args: ValueOf[blastn.Arguments]): BlastExpression[blastExprType.type] = {
-    BlastExpression(blastExprType)(
-      argumentValues  = args,
-      // TODO whatever
-      optionValues    = blastn.defaults update (
-        num_threads(1) :~:
-        max_target_seqs(10) :~:
-        ohnosequences.blast.api.evalue(0.001)  :~:
-        blastn.task(blastn.megablast) :~: ∅
-      )
-    )
-  }
+  type MD <: AnyMetagenomicaData
+  val md: MD
 
   private def writeFastaToFile(v: ValueOf[FASTA], file: File): File = {
 
@@ -75,103 +49,80 @@ case object blastDataProcessing {
   }
 
 
-  trait AnyBlastDataProcessing extends AnyDataProcessingBundle {
+  def instructions: AnyInstructions = say("Let the blasting begin!")
 
-    def instructions: AnyInstructions = say("Let the blasting begin!")
-
-    val bundleDependencies: List[AnyBundle] = List[AnyBundle](
-      bundles.blast,
-      bundles.blast16s
-    )
-
-    // TODO: more precise type
-    type FastqInput <: AnyData
-    val  fastqInput: FastqInput
+  val bundleDependencies: List[AnyBundle] = List[AnyBundle](
+    bundles.blast,
+    bundles.blast16s
+  )
 
 
-    // FIXME: we should use this instead of blastCmd, but .cmd requires implicits "/
-    // type BlastExpr <: AnyBlastExpression
-    // val  blastExpr: BlastExpr
-    // val blastCmd: Seq[String]
-
-    type BlastOutput <: AnyBlastOutput
-    val  blastOutput: BlastOutput
[…]

### 2015-09-26 src/main/scala/metagenomica/loquats/flash/dataProcessing.scala

@@ -21,7 +21,7 @@ import java.io.File
 
 trait AnyFlashDataProcessing extends AnyDataProcessingBundle {
 
-  type MD <: AnyMetagenomicaData;
+  type MD <: AnyMetagenomicaData
   val md: MD
 
   val bundleDependencies: List[AnyBundle] = List( bundles.flash )

### 2015-09-26 src/main/scala/metagenomica/configuration.scala
Restored assignment loquat; moved things to the glob conf

@@ -10,12 +10,32 @@ import ohnosequences.flash.data._
 import ohnosequences.blast._, api._, data._, outputFields._
 
 import java.io.File
+import scala.util.Try
+
 
 case object configuration {
 
   // TODO: move it to datasets
   implicit def genericParser[D <: AnyData](implicit d: D): DenotationParser[D, FileDataLocation, File] =
     new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
+
+
+  case object CSVDataType extends AnyDataType { val label = "fastq" }
+  case object lcaCSV extends Data(CSVDataType, "lca.csv")
+  case object bbhCSV extends Data(CSVDataType, "bbh.csv")
+
+  type ID = String
+  type GI = ID
+  type TaxID = ID
+  type ReadID = ID
+  type NodeID = ID
+
+  type LCA = Option[NodeID]
+  type BBH = Option[NodeID]
+
+  // TODO: move it somewhere up for global use
+  def parseInt(str: String): Option[Int] = Try(str.toInt).toOption
+
 
 
   trait AnyMetagenomicaData {

### 2015-09-26 src/main/scala/metagenomica/loquats/assignment/dataProcessing.scala

@@ -1,140 +1,126 @@
-// package ohnosequences.metagenomica.loquats.assignment
-//
-// import  ohnosequences.metagenomica._
-//
-// import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
-//
-// import ohnosequences.loquat.dataProcessing._
-// import ohnosequences.statika.bundles._
-// import ohnosequences.statika.instructions._
-// import ohnosequencesBundles.statika.Blast
-// import ohnosequences.blast._, api._, data._
-// import ohnosequences.cosas._, typeSets._, types._, properties._
-// import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-// import java.io.{ BufferedWriter, FileWriter, File }
-// import ohnosequences.fastarious._, fasta._, fastq._
-// import ohnosequences.blast._, api._, data._, outputFields._
-//
-// import scala.util.Try
-//
-//
-// case object dataProcessing {
-//
-//   // TODO: move it to the config
-//   case object CSVDataType extends AnyDataType { val label = "fastq" }
-//   case object lcaCSV extends Data(CSVDataType, "lca.csv")
-//   case object bbhCSV extends Data(CSVDataType, "bbh.csv")
-//
-//   // type BlastOutput = loquats.blast.blastTest.blastOutput.type
-//   val  blastOutput = loquats.blast.blastTest.blastOutput
-//
-//   type BlastRecord = loquats.blast.blastDataProcessing.outRec.type
-//   val blastRecord: BlastRecord = loquats.blast.blastDataProcessing.outRec
-//
-//   val headers: Seq[String] = blastRecord.properties.mapToList(typeLabel)
-//
-//   // TODO: move it somewhere up for global use
-//   type ID = String
-//
-//   type GI = ID
-//   type TaxID = ID
-//   type ReadID = ID
-//   type NodeID = ID
-//
-//   type LCA = Option[NodeID]
-//   type BBH = Option[NodeID]
-//
-//   // TODO: move it somewhere up for global use
-//   def parseInt(str: String): Option[Int] = Try(str.toInt).toOption
-//
-//   // this method looks up particular column by its header
-//   def column(row: Seq[String], header: AnyOutputField): Option[String] =
-//     headers.zip(row).toMap.get(header.label)
-//
-//   // FIXME: this thing should be used globally (ideally it should be in datasets)
[…]

### 2015-09-26 src/main/scala/metagenomica/configuration.scala
Restored counting loquat; LOL: it doesn't depend on the global conf at all :sweat_smile:

@@ -20,9 +20,13 @@ case object configuration {
     new DenotationParser(d, d.label)({ f: File => Some(FileDataLocation(f)) })
 
 
-  case object CSVDataType extends AnyDataType { val label = "fastq" }
+  case object CSVDataType extends AnyDataType { val label = "csv" }
+
   case object lcaCSV extends Data(CSVDataType, "lca.csv")
   case object bbhCSV extends Data(CSVDataType, "bbh.csv")
+
+  case object lcaCountsCSV extends Data(CSVDataType, "lca.counts.csv")
+  case object bbhCountsCSV extends Data(CSVDataType, "bbh.counts.csv")
 
   type ID = String
   type GI = ID

### 2015-09-26 src/main/scala/metagenomica/loquats/counting/dataProcessing.scala

@@ -1,114 +1,98 @@
-// package ohnosequences.metagenomica.loquats.counting
-//
-// import ohnosequences.metagenomica._
-// import ohnosequences.metagenomica.loquats.assignment.dataProcessing._
-//
-// import ohnosequences.metagenomica.bio4j._, taxonomyTree._, titanTaxonomyTree._
-//
-// import ohnosequences.loquat.dataProcessing._
-// import ohnosequences.loquat.utils._
-//
-// import ohnosequences.statika.bundles._
-// import ohnosequences.statika.instructions._
-// import ohnosequencesBundles.statika.Blast
-// import ohnosequences.blast._, api._, data._
-// import ohnosequences.cosas._, typeSets._, types._, properties._
-// import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
-// import java.io.{ BufferedWriter, FileWriter, File }
-// import ohnosequences.fastarious._, fasta._, fastq._
-// import ohnosequences.blast._, api._, data._, outputFields._
-//
-// import scala.util.Try
-//
-//
-// case object dataProcessing {
-//
-//   case object lcaCountsCSV extends Data(CSVDataType, "lca.counts.csv")
-//   case object bbhCountsCSV extends Data(CSVDataType, "bbh.counts.csv")
-//
-//   // returns count of the given id and a filtered list (without that id)
-//   def count(id: ID, list: List[ID]): (Int, List[ID]) =
-//     list.foldLeft( (0, List[ID]()) ) { case ((count, rest), next) =>
-//       if (next == id) (count + 1, rest)
-//       else (count, next :: rest)
-//     }
-//
-//   // TODO: make it tailrec
-//   def directCounts(taxIds: List[TaxID]): Map[TaxID, Int] = {
-//
-//     @scala.annotation.tailrec
-//     def rec(list: List[TaxID], acc: Map[TaxID, Int]): Map[TaxID, Int] =
-//       list match{
-//         case Nil => acc
-//         case h :: t => {
-//           val (n, rest) = count(h, t)
-//           rec(rest, acc.updated(h, n + 1))
-//         }
-//       }
-//
-//     rec(taxIds, Map[TaxID, Int]())
-//   }
-//
-//   // TODO: figure out some more effective algorithm
-//   // Caution: it uses bio4j bundle!
[…]

### 2015-09-27 docs/mg7-preprint/out/out.sh
add PDF generation structure and fix format -- preprint

@@ -1,0 +1,7 @@
+#!/bin/sh
+set -x
+ME=`dirname $0`
+pandoc --toc --template=${ME}/ohnosequences.report.tex ${ME}/../paper.md --smart -s --biblatex --bibliography ${ME}/../refs.bib -o ${ME}/paper.tex
+xelatex -output-directory ${ME} -halt-on-error ${ME}/paper.tex
+biber ${ME}/paper
+xelatex -output-directory ${ME} -halt-on-error ${ME}/paper.tex

### 2015-09-27 docs/mg7-preprint/out/paper.tex

--- 2015-09-24 docs/paper/out/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Version 3.1 Generated 2015/22/05 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-
-\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
-%\documentclass{frontiersHLTH} % for Health articles
-%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
-
-%\setcitestyle{square}
-\usepackage{url,hyperref,lineno,microtype}
-\usepackage[onehalfspacing]{setspace}
-\linenumbers
-
-% fix for pandoc 1.14
+% \documentclass[a4paper]{article}
+\documentclass[fontsize=8pt,paper=a4,pagesize,abstracton]{scrartcl}
+
+% to prevent funny stuff happening with old packages (I think)
+\usepackage{ifxetex}
+% xetex font-related packages
+\usepackage{fontspec,xltxtra,xunicode}
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% I don't know why this is here
+\usepackage{fixltx2e}
+
+%%% color definitions %%%
+\usepackage[usenames,dvipsnames,xetex]{xcolor}
+% old cyan variants %
+\definecolor{OldCyan}{RGB}{62,127,137}
+\definecolor{OldCyan-Dark}{RGB}{38,104,113}
+\definecolor{OldCyan-Light}{RGB}{86,156,166}
+\definecolor{OldCyan-Matte}{RGB}{62,113,120}
+\definecolor{OldCyan-MatteLight}{RGB}{95,158,166}
+% light amber variants %
+\definecolor{LightAmber}{RGB}{223,184,98}
+\definecolor{LightAmber-Dark}{RGB}{184,145,61}
+\definecolor{LightAmber-Light}{RGB}{231,195,118}
+\definecolor{LightAmber-Matte}{RGB}{195,165,100}
+\definecolor{LightAmber-MatteLight}{RGB}{231,199,131}
+% salmon variants %
+\definecolor{Salmon}{RGB}{223,107,98}
+\definecolor{Salmon-Dark}{RGB}{184,70,61}
+\definecolor{Salmon-Light}{RGB}{231,126,118}
+\definecolor{Salmon-Matte}{RGB}{195,107,100}
+\definecolor{Salmon-MatteLight}{RGB}{231,138,131}
+\definecolor{Grey}{RGB}{88,88,88}
+\definecolor{Grey-Light}{RGB}{129,129,129}
+\definecolor{Grey-Dark}{RGB}{65,65,65}
+% SOLARIZED
+\definecolor{solarized@base03}{HTML}{002B36}
+\definecolor{solarized@base02}{HTML}{073642}
+\definecolor{solarized@base01}{HTML}{586e75}
+\definecolor{solarized@base00}{HTML}{657b83}
+\definecolor{solarized@base0}{HTML}{839496}
+\definecolor{solarized@base1}{HTML}{93a1a1}
+\definecolor{solarized@base2}{HTML}{EEE8D5}
+\definecolor{solarized@base3}{HTML}{FDF6E3}
+\definecolor{solarized@yellow}{HTML}{B58900}
+\definecolor{solarized@orange}{HTML}{CB4B16}
+\definecolor{solarized@red}{HTML}{DC322F}
+\definecolor{solarized@magenta}{HTML}{D33682}
+\definecolor{solarized@violet}{HTML}{6C71C4}
+\definecolor{solarized@blue}{HTML}{268BD2}
+\definecolor{solarized@cyan}{HTML}{2AA198}
+\definecolor{solarized@green}{HTML}{859900}
+
+% math symbols and such
+\usepackage{amsmath}
+\usepackage{unicode-math}
+
+% english support
+\usepackage{polyglossia}
+\setmainlanguage{english}
+
+% paper size and lead
+\usepackage[]{geometry}
+\geometry{a4paper}
+\usepackage{setspace}
+% default lead is 1.216666
+\setstretch{1.15}
+
+% Activate to begin paragraphs with an empty line rather than an indent
+\usepackage{parskip}
+
+% references
+
+% for natbib
+
+% for biblatex
+\usepackage[
+  backend=biber,
+  texencoding=utf8,
+  bibencoding=utf8,
+  hyperref=true,
+  maxbibnames=15,
+  isbn=false,
+  dateabbrev=false,
+  style=ieee
+]{biblatex}
+
+\renewcommand*{\newunitpunct}{\space}
+\renewcommand{\bibfont}{\small}
+\setlength\bibitemsep{5pt}
+
+\renewbibmacro*{volume+number+eid}{%
+  \printfield{volume}%
+  \setunit{\,}% DELETED
+  \printfield{number}%
+  \setunit{.\space}%
+  \printfield{eid}}
+
+\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{volume}{\mkbibbold{#1}}
+\DeclareFieldFormat{date}{#1}
+\DeclareFieldFormat{pages}{#1\adddot}
+\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{number}{\mkbibparens{#1}\addcolon}
+
+\DeclareFieldFormat{doi}{\newline DOI: \href{https://dx.doi.org/#1}{#1}}
+\DeclareFieldFormat{url}{}
+\DeclareFieldFormat[unpublished]{url}{\newline URL: \href{#1}{#1}}
+\DeclareFieldFormat
+  [article,inbook,incollection,inproceedings,patent,thesis,unpublished]
+  {journaltitle}{\newline\emph{#1}}
+
+\DeclareFieldFormat
+  [article,inbook,incollection,inproceedings,patent,thesis,unpublished]
+  {title}{\textbf{#1}}
+
+\bibliography{./out/../refs.bib}
+
+\usepackage{longtable}
+\usepackage{ctable}
+\usepackage{float}
+\usepackage[normalem]{ulem}
+% typography settings
+% default conf applied
+
+% Open Sans, a beautiful fontface
+\setmainfont[Mapping = tex-text, Color = Grey, BoldFont = {* Semibold}, BoldItalicFont = {* Semibold Italic}, ItalicFont = {* Italic}]{Open Sans}
+% sort of like Cambria Math; somehow matches a sans body text
+\setmathfont{Asana Math} % math font
+% font for sections: Bitter
+\newfontfamily\sectfont
+  [BoldFont={* Bold},
+   Numbers={OldStyle,Proportional},
+   Color = Grey,
+   Mapping = tex-text
+  ]
+  {Bitter}
+% the title font
+\newfontfamily\titlefont
+  [BoldFont={Open Sans},
+   Numbers={OldStyle,Proportional},
+   Mapping = tex-text
+  ]
+  {Open Sans Light}
+
+% for changing the sections font
+\usepackage{sectsty}
+\allsectionsfont{\sectfont\textmd}
+
+% verbatim and code highlighting: Solarized Light
+\usepackage{fancyvrb}
+\DefineShortVerb[commandchars=\\\{\}]{\|}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{fontsize=\footnotesize,%
+commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+% this is solarized light
+\definecolor{shadecolor}{RGB}{253,246,227} % solarized@base3
+\newenvironment{Shaded}{\vspace{\baselineskip}\begin{shaded}}{\end{shaded}\vspace{\baselineskip}}
+% colored backgrd for verb
+\let\oldverbatim=\verbatim
+\let\endoldverbatim=\endverbatim
+\renewenvironment{verbatim}[1]{%
+  \vspace{\baselineskip}
+  \scriptsize
+  \par\setstretch{1}
+  \begin{shaded}
+  \begin{oldverbatim}{#1}%
+}%
+{%
+  \end{oldverbatim}%
+  \end{shaded}
+  \vspace{\baselineskip}
+}
+
+\newcommand{\KeywordTok}[1]{\textbf{\textcolor{solarized@green}{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor{solarized@blue}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor{solarized@violet}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor{solarized@violet}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor{solarized@violet}{#1}}
+\newcommand{\CharTok}[1]{\textcolor{solarized@cyan}{#1}}
+\newcommand{\StringTok}[1]{\textcolor{solarized@cyan}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor{solarized@base1}{\textit{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor{solarized@base00}{#1}}
+\newcommand{\AlertTok}[1]{\textcolor{solarized@yellow}{\textbf{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor{solarized@blue}{#1}}
+\newcommand{\RegionMarkerTok}[1]{\textcolor{solarized@base1}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor{solarized@red}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{\textcolor{solarized@base00}{#1}}
+
+% monospace also grey; Source Code Pro
+\setmonofont[Mapping=tex-ansi,Color=solarized@base00,Numbers=Lining,Ligatures=NoRequired]{Source Code Pro}
+
+% allow to break lines more easily on tt text
+% http://tex.stackexchange.com/questions/52850/temporarily-increase-the-limit-on-space-size
+\let\OldTexttt\texttt
+\renewcommand{\texttt}[1]{ \emergencystretch=2em \OldTexttt{#1} }
+% maybe?
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+
+% color fixes
+\makeatletter
+\newcommand{\globalcolor}[1]{%
+  \color{#1}\global\let\default@color\current@color
+}
+\makeatother
+
+\AtBeginDocument{\globalcolor{Grey}}
+\renewcommand{\textcolor}[2]{{\addfontfeature{Color=#1}#2}}
+
+% Lists formatting:
+% note sure what 'fancy enums' are; something to do with lists,
+% as the further comment suggests:
+
+% tables
+\usepackage{tabu}
+\taburulecolor{Salmon-Dark}
+\let\quoteOld\quote
+\let\endquoteOld\endquote
+\renewenvironment{quote}{\vspace{\baselineskip}\itshape\quoteOld\begin{tabu}{|[2pt]p{.9\textwidth}}}{\end{tabu}\endquoteOld\vspace{\baselineskip}}
+
+% longtable needed by latest pandoc version
+\usepackage{longtable}
+
+\setupctable{
+captionskip=0pt, framerule=0pt, nostar,
+center, framesep=0pt, pos=tbp,
+continued=(continued), maxwidth=0pt, super,
+doinside={}, mincapwidth=0pt, table,
+framebg=1 1 1, nonotespar, topcap,
+framefg=.3 .1 .68, nosideways, width=0pt
+}
+
+% math diagrams
+\usepackage[all]{xy}
+\xyoption{pdf}  % improves rendering
+
+\usepackage[breaklinks=true]{hyperref}
+\hypersetup{
+  setpagesize=false, % page size defined by xetex
+  unicode=false, % unicode breaks when used with xetex
+  xetex,
+  pdfnewwindow,
+  colorlinks,%
+  citecolor=OldCyan,%
+  filecolor=OldCyan,%
+  linkcolor=OldCyan,%
+  urlcolor=OldCyan,
+  urlbordercolor=OldCyan
+}
+
+\let\OldHref\href
+\renewcommand{\href}[2]{\OldHref[pdfnewwindow]{#1}{{#2}}}
+
+% ugly fix for latest pandoc
 \providecommand{\tightlist}{%
@@
 
-% Leave a blank line between paragraphs instead of using \\
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Alexey Alekhin {et~al.}} %use et al only if is more than 1 author
-\def\Authors{
- Alexey Alekhin\(^\dagger\)\,$^{1}$  Evdokim Kovach\(^\dagger\)\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja-Tobes\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
-}% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{$^{1}$Oh no sequences! Research Group, Era7 Bioinformatics, Granada, Spain}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Eduardo Pareja-Tobes}
-\def\corrAddress{Oh no sequences! Research Group, Era7 Bioinformatics, Plaza Campo Verde 3, Granada, 18001, Spain}
-\def\corrEmail{eparejatobes@ohnosequences.com}
+% ugly fix for link colors in xetex
+\makeatletter
+\def\HyColor@@@@UseColor#1\@nil{\addfontfeatures{Color=#1}}
+\makeatother
+
+% images
+\usepackage{graphicx}
+% -- We will generate all images so they have a width \maxwidth. This means
+% -- that they will get their normal width if they fit onto the page, but
+% -- are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+
+% this is wrong
+\setcounter{secnumdepth}{0}
+
+  \title{\LARGE \bfseries MG7: Configurable and scalable 16S metagenomics data analysis}
+
+% authors
+\usepackage{authblk}
+  \author[1]{\normalsize Alexey Alekhin\(^\dagger\)}
+  \author[1]{\normalsize Evdokim Kovach\(^\dagger\)}
+  \author[1]{\normalsize Marina Manrique}
+  \author[1]{\normalsize Pablo Pareja-Tobes}
+  \author[1]{\normalsize Eduardo Pareja}
+  \author[1]{\normalsize Raquel Tobes}
+  \author[1]{\normalsize Eduardo Pareja-Tobes}
+
+\affil[1]{\small
+  {\itshape\href{http://ohnosequences.com}{oh no sequences!}} research group, \href{http://www.era7bioinformatics.com}{Era7 bioinformatics}
+}
+
+\date{\normalsize \today}
+
+% page style!
+% spaces
+\setlength{\footskip}{3\baselineskip}
+\setlength{\headsep}{3\baselineskip}
+\setlength{\skip\footins}{0cm}
+% \setlength{\voffset}{-10pt}
+\setlength{\topmargin}{0pt}
+% \setlength{\textheight}{612pt}
+
+\makeatletter
+\let\Title\@title
+\let\Author\@author
+\makeatother
+
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+
+\renewcommand{\sectionmark}[1]{%
+\markboth{#1}{}}
+% \fancyheadoffset{0cm}
+\renewcommand{\headrulewidth}{.5pt}
+\renewcommand{\headrule}{\vbox to 0pt{\hbox
+to\headwidth{\dotfill}\vss}}
+
+\lhead{\titlefont MG7: Configurable and scalable 16S metagenomics data analysis}
+% \chead{\titlefont {\color{Grey-Light}\leftmark}}
+\chead{}
+\rhead{\titlefont \thepage}
+
+\lfoot{\titlefont {{\color{Salmon-Dark}oh}{\color{LightAmber-Dark}no}{\color{Grey}sequences}{\color{Salmon-Dark}!} }}
+\cfoot{}
+\rfoot{\titlefont {\color{Grey-Light} era7 bioinformatics R\&D group}}
+% footer rule
+\renewcommand{\footruleskip}{0.66\baselineskip}
+\renewcommand{\footrulewidth}{.5pt}
+\renewcommand{\footrule}{{\color{LightAmber-MatteLight}%
+\vskip-\footruleskip\vskip-\footrulewidth
+\hrule width\headwidth height\footrulewidth\vskip\footruleskip}}
+
+\renewcommand{\footrule}{\vbox to 0pt{\hbox
+to\headwidth{\color{Salmon}{\dotfill}}\vss}}
 
 \begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[MG7: Configurable and scalable 16S metagenomics data analysis]{MG7: Configurable and scalable 16S metagenomics data analysis -- new
-methods optimized for massive cloud computing}
-
-\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
 
@@
 
-\begin{abstract}
-
+\begin{abstract} % \footnotesize
 As part of the Cambrian explosion of omics data, metagenomics brings to
@@
 sustainable update system, and the possibility of project-driven
-personalization.
-
+personalization. \newline\newline
 \(^\dagger\) The first and second authors contributed equally to this
 work.
-
-\tiny
- \keyFont{ \section{Keywords:} Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
-Cloud computing, NGS, Genomic big data, Microbiome, Environmental, 16S
-Database } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \end{abstract}
+
+\thispagestyle{empty}
+
+% \setcounter{tocdepth}{3}
+\tableofcontents
+
+\newpage
 
@@
 derived decrease in sequencing costs), and the wide impact of
-metagenomics studies \citep{oulas2015metagenomics}, especially in human
-health (diagnostics, treatments, drug response or prevention)
-\citep{bikel2015combining}. We should also mention what could be called
-the microbiome explosion: all kind of microbiomes (gut, mouth, skin,
-urinary tract, airway, milk, bladder) are now routinely sequenced in
-different conditions of health and disease, or after different
+metagenomics studies \autocite{oulas2015metagenomics}, especially in
+human health (diagnostics, treatments, drug response or prevention)
+\autocite{bikel2015combining}. We should also mention what could be
+called the microbiome explosion: all kind of microbiomes (gut, mouth,
+skin, urinary tract, airway, milk, bladder) are now routinely sequenced
+in different conditions of health and disease, or after different
 treatments. The impact of Metagenomics is also being felt in
-environmental sciences \citep{ufarte2015discovery}, crop sciences, the
-agrifood sector \citep{coughlan2015biotechnological} and biotechnology
-in general \citep[\citet{kodzius2015marine}]{cowan2015metagenomics}.
-These new possibilities for exploring the diversity of micro-organisms
-in the most varied environments are opening new research areas, and
-drastically changing the existing ones.
+environmental sciences \autocite{ufarte2015discovery}, crop sciences,
+the agrifood sector \autocite{coughlan2015biotechnological} and
+biotechnology in general \autocite{cowan2015metagenomics}
+\autocite{kodzius2015marine}. These new possibilities for exploring the
+diversity of micro-organisms in the most varied environments are opening
+new research areas, and drastically changing the existing ones.
 
@@
 data acquisition to data analysis: the amount of data is expected to be
-overwhelming in a very short time \citep{stephens2015big}.
+overwhelming in a very short time \autocite{stephens2015big}.
 
 Genome researchers have raised the alarm over big data in the past
-\citep{hayden2015genome}, but even a more serious challenge might be
+\autocite{hayden2015genome}, but even a more serious challenge might be
 faced with the metagenomics boom. If we compare metagenomics data with
@@
 studies, serial sampling from the same patient
-\citep{faust2015metagenomics} along several weeks (or years) is being
+\autocite{faust2015metagenomics} along several weeks (or years) is being
 used for the follow up of some intestinal pathologies, for studying the
 evolution of the gut microbiome after antibiotic treatment, or for colon
-cancer early detection
-\citep[\citet{garrett2015cancer}]{zeller2014potential}. This need of
-sampling across time adds more complexity to metagenomics data storage
-and demands adapted algorithms to detect state variations across time as
-well as idiosyncratic commonalities of the microbiome of each individual
-\citep{franzosa2015identifying}. In addition to the intra-individual
+cancer early detection \autocite{zeller2014potential}
+\autocite{garrett2015cancer}. This need of sampling across time adds
+more complexity to metagenomics data storage and demands adapted
+algorithms to detect state variations across time as well as
+idiosyncratic commonalities of the microbiome of each individual
+\autocite{franzosa2015identifying}. In addition to the intra-individual
 sampling-time dependence, metagenomic clinical test results vary
@@
 big data challenge before. A key difference is the existence of
-standards for data processing \citep{stephens2015big}; in metagenomics
-global standards for converting raw sequence data into processed data
-are not yet well defined, and there are shortcomings derived from the
-fact that most bioinformatics methodologies used for metagenomics data
-analysis were designed for scenarios very different from the current
-one. These are some of the aspects that have suffered crucial changes
-and advances with a direct impact in metagenomics data analysis:
+standards for data processing \autocite{stephens2015big}; in
+metagenomics global standards for converting raw sequence data into
+processed data are not yet well defined, and there are shortcomings
+derived from the fact that most bioinformatics methodologies used for
+metagenomics data analysis were designed for scenarios very different
+from the current one. These are some of the aspects that have suffered
+crucial changes and advances with a direct impact in metagenomics data
+analysis:
 
@@
 an extensible record system with concatenation on disjoint labels, in
-the spirit of \citep[\citet{harper1991record}]{harper1990extensible}.
-For that \emph{Datasets} uses
+the spirit of \autocite{harper1990extensible}
+\autocite{harper1991record}. For that \emph{Datasets} uses
 \href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
@@
 \href{https://github.com/ohnosequences/flash}{FLASh}
-\citep{magovc2011flash} and
+\autocite{magovc2011flash} and
 \href{https://github.com/ohnosequences/blast}{BLAST}
-\citep{camacho2009blast} expressions and their execution.
+\autocite{camacho2009blast} expressions and their execution.
 
@@
 
-\citep[Bio4j][]{pareja2015bio4j} is a data platform integrating data
+Bio4j \autocite{pareja2015bio4j} is a data platform integrating data
 from different resources such as UniProt or GO in a graph data paradigm.
@@
 sequences downloaded from the \textbf{RDP database}
-\citep{cole2013ribosomal}. RDP unaligned sequences were used to capture
-new 16S RNA sequences from \textbf{nt} using BLAST similarity search
-strategies and then, performing additional curation steps to remove
-sequences with poor taxonomic assignments to taxonomic nodes close to
-the root of the taxonomy tree. All the nucleotide sequences included in
-\textbf{nt} database has a taxonomic assignment provided by the
-\textbf{Genbank} sequence submitter. NCBI provides a table (available at
-ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
-Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
-based on a crowdsourced submitter-maintained taxonomic annotation system
-for reference sequences. It supposes a sustainable system able to face
-the expected number of reference sequences that will populate the public
-global nucleotide databases in the near future. Another advantageous
-point is that we are based on NCBI taxonomy, the \emph{de-facto}
-standard taxonomic classification for biomolecular data
-\citep{cochrane20102010}. NCBI taxonomy is, undoubtedly, the most used
-taxonomy all over the world and the most similar to the official
-taxonomies of each specific field. This is a crucial point because all
-the type-culture and tissue databanks follow this official taxonomical
-classification and, in addition, all the knowledge accumulated during
-last decades is referred to this taxonomy. In addition NCBI provides a
-direct connection between taxonomical formal names and the physical
-specimens that serve as exemplars for the species
-\citep{federhen2014type}.
+\autocite{cole2013ribosomal}. RDP unaligned sequences were used to
+capture new 16S RNA sequences from \textbf{nt} using BLAST similarity
+search strategies and then, performing additional curation steps to
+remove sequences with poor taxonomic assignments to taxonomic nodes
+close to the root of the taxonomy tree. All the nucleotide sequences
+included in \textbf{nt} database has a taxonomic assignment provided by
+the \textbf{Genbank} sequence submitter. NCBI provides a table
+(available at ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the
+mapping of any Genbank Identifier (GI) to its Taxonomy Identifier
+(TaxID). Thus, we are based on a crowdsourced submitter-maintained
+taxonomic annotation system for reference sequences. It supposes a
+sustainable system able to face the expected number of reference
+sequences that will populate the public global nucleotide databases in
+the near future. Another advantageous point is that we are based on NCBI
+taxonomy, the \emph{de-facto} standard taxonomic classification for
+biomolecular data \autocite{cochrane20102010}. NCBI taxonomy is,
+undoubtedly, the most used taxonomy all over the world and the most
+similar to the official taxonomies of each specific field. This is a
+crucial point because all the type-culture and tissue databanks follow
+this official taxonomical classification and, in addition, all the
+knowledge accumulated during last decades is referred to this taxonomy.
+In addition NCBI provides a direct connection between taxonomical formal
+names and the physical specimens that serve as exemplars for the species
+\autocite{federhen2014type}.
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figure-1.jpg}
+\caption{\textbf{MG7 analysis workflow.} The paired reads in fastq
+format are merged resulting in only one sequence per read pair. The next
+step is a parallelized BLASTN of every merged sequence against the 16S
+reference database 16S-DB7. Then, the mapping of the detected similar
+sequences in the database to the taxonomy node to which they belong is
+carried out. This is done using Bio4j that includes a module with all
+the NCBI taxonomy in a graph connected with the Gene Ontology, Uniprot,
+and RefSeq graphs. Then the taxonomic assignment is done for each
+sequence following two different approaches: LCA and BBH, and finally
+the abundances corresponding to direct and cumulative assignments for
+each node in percentage and absolute counts are provided for each
+assignment mode.}
+\end{figure}
+
 \subsubsection{Joining reads of each pair using
@@
 that yields pairs of reads with an overlapping region between them, are
-assembled using FLASh \citep{magovc2011flash}. FLASh is designed to
+assembled using FLASh \autocite{magovc2011flash}. FLASh is designed to
 merge pairs of reads when the original DNA fragments are shorter than
@@
 considered in different reviews of metagenomics analysis methodologies
-\citep[\citet{morgan2012chapter}]{segata2013computational} as a very
-exhaustive method for assignment. Some methods of assignment compare the
-sequences only against the 16S genes from available complete bacterial
-genomes or avoid computational cost clustering or binning the sequences
-first, and then doing the assignments only for the representative
-sequence of each cluster. MG7 carries out an exhaustive comparison of
-all the reads under analysis and it does not applies any binning
-strategy. Every read is specifically compared with all the sequences of
-the 16S database. We select the best BLAST hits (10 hits by default)
-obtained for each read to do the taxonomic assignment.
+\autocite{segata2013computational} \autocite{morgan2012chapter} as a
+very exhaustive method for assignment. Some methods of assignment
+compare the sequences only against the 16S genes from available complete
+bacterial genomes or avoid computational cost clustering or binning the
+sequences first, and then doing the assignments only for the
+representative sequence of each cluster. MG7 carries out an exhaustive
+comparison of all the reads under analysis and it does not applies any
+binning strategy. Every read is specifically compared with all the
+sequences of the 16S database. We select the best BLAST hits (10 hits by
+default) obtained for each read to do the taxonomic assignment.
 
@@
 the assignment to the most specific taxon.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figure-2.jpg}
+\caption{\textbf{Lowest Common Ancestor algorithm for taxonomic
+assignment.} The Left panel displays an example of the application of
+LCA algorithm in a \emph{sensu stricto} mode. A, B, C and D represent
+taxonomy tree nodes with assigned reads. Right panel displays the
+\emph{in line} mode of assignment which is an exception for the
+\emph{sensu stricto} mode of application of LCA algorithm. The \emph{in
+line} mode is used when all the nodes are located in a line without
+bifurcations. In that case the taxon assigned is the most specific (the
+most distant from the root).}
+\end{figure}
 
@@
 and then they would be classified as reads with no hits. Advanced
-metagenomics analysis approaches \citep{huson2012microbial} have adopted
-LCA assignment algorithms because it provides fine and trusted
-taxonomical assignment.
+metagenomics analysis approaches \autocite{huson2012microbial} have
+adopted LCA-based assignment algorithms because it provides fine and
+trusted taxonomical assignment.
 
@@
 taxonomic assignment because, in some cases, it can provide information
-about the sequences that adds information to that obtained using LCA
-algorithm. Using LCA algorithm, when some reference sequences with BLAST
-alignments over the required thresholds map to a not sufficiently
+about the sequences that adds information to that obtained using the LCA
+algorithm. With the LCA algorithm, when some reference sequences with
+BLAST alignments over the required thresholds map to a not sufficiently
 specific taxID, the read can be assigned to an unspecific taxon near to
@@
 taxonomy tree as a graph using the graph database paradigm. Previously
-we developed Bio4j \textbf{{[}Pareja-Tobes-2015{]}}, a platform for the
+we developed Bio4j \autocite{pareja2015bio4j}, a platform for the
 integration of semantically rich biological data using typed graph
@@
 visualizations are static. We are working in the \emph{Biographika}
-project \citep{tobes2015biographika}, to provide interactive rich
+project \autocite{tobes2015biographika}, to provide interactive rich
 visualizations on the web for Bio4j data. The development of
@@
   \textbf{EK} developed \emph{nispero} (a prototype for \emph{Loquat}
-  \citep{kovach2014nispero}) and \emph{aws-scala-tools}.
+  \autocite{kovach2014nispero}) and \emph{aws-scala-tools}.
 \item
@@
 
-\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
-\bibliography{refs}
+  \printbibliography
 

### 2015-09-27 src/main/scala/metagenomica/configuration.scala
WIP: common testing configuration

@@ -1,6 +1,7 @@
 package ohnosequences.metagenomica
 
 import ohnosequences.cosas._, typeSets._, types._
+import ops.typeSets._
 
 import ohnosequences.datasets._, dataSets._, fileLocations._, illumina._, reads._
 
@@ -78,49 +79,52 @@ case object configuration {
     type BlastOutRec <: AnyBlastOutputRecord
     val  blastOutRec: BlastOutRec
 
+    // implicitly:
+    val validBlastRecord: BlastOutRec#Properties CheckForAll ValidOutputRecordFor[blastn]
+
     type BlastExprType >: BlastExpressionType[blastn, BlastOutRec]
                        <: BlastExpressionType[blastn, BlastOutRec]
-    val  blastExprType: BlastExprType
+    lazy val blastExprType: BlastExprType = new BlastExpressionType(blastn)(blastOutRec)(validBlastRecord)
 
     type BlastOutType >: BlastOutputType[BlastExprType]
                       <: BlastOutputType[BlastExprType]
-    val  blastOutType: BlastOutType
+    val  blastOutType: BlastOutType = new BlastOutputType(blastExprType, "useless label")
 
+    // type BlastExpr <: AnyBlastExpression {
+    //   type Tpe <: AnyBlastExpressionType {
+    //     type OutputRecord = BlastOutRec
+    //   }
+    // }
     def blastExpr(args: ValueOf[blastn.Arguments]): BlastExpression[BlastExprType]
 
     type BlastOut >: BlastOutput[BlastOutType]
                   <: BlastOutput[BlastOutType]
-    val  blastOut: BlastOut
+    val  blastOut: BlastOut = new BlastOutput(blastOutType, "blast.out.csv")
   }
 
-  // case object outRec extends BlastOutputRecord(
-  //   qseqid    :&:
-  //   qlen      :&:
-  //   qstart    :&:
-  //   qend      :&:
-  //   sseqid    :&:
-  //   slen      :&:
-  //   sstart    :&:
-  //   send      :&:
-  //   bitscore  :&:
-  //   sgi       :&: □
-  // )
+  abstract class MetagenomicaData[
+    RT <: AnyReadsType { type EndType = pairedEndType },
+    BR <: AnyBlastOutputRecord
+  ](val readsType: RT,
+    val blastOutRec: BR
+  )(implicit
[…]

### 2015-09-27 src/main/scala/metagenomica/test.scala

@@ -1,0 +1,68 @@
+package ohnosequences.metagenomica
+
+import ohnosequences.metagenomica.configuration._
+
+import ohnosequences.datasets._, dataSets._, fileLocations._, s3Locations._, illumina._, reads._
+
+import ohnosequences.cosas._, typeSets._, types._, records._, properties._
+import ohnosequences.cosas.ops.typeSets._
+
+import ohnosequences.loquat._, utils._, configs._, dataMappings._, dataProcessing._
+
+import ohnosequences.statika.bundles._
+import ohnosequences.statika.aws._, api._, amazonLinuxAMIs._
+
+import ohnosequences.awstools._, regions.Region._, ec2.InstanceType._
+import ohnosequences.awstools.s3._
+import ohnosequences.awstools.autoscaling._
+
+import ohnosequences.flash.api._
+import ohnosequences.flash.data._
+
+import ohnosequences.blast.api._, outputFields._
+import ohnosequences.blast.data._
+
+import era7.project.loquats._
+
+import java.io.File
+
+
+case object test {
+
+  case object outRec extends BlastOutputRecord(
+    qseqid    :&:
+    qlen      :&:
+    qstart    :&:
+    qend      :&:
+    sseqid    :&:
+    slen      :&:
+    sstart    :&:
+    send      :&:
+    bitscore  :&:
+    sgi       :&: □
+  )
+
+  // case object blastExprType extends BlastExpressionType(blastn)(outRec)
+  // case object blastOutputType extends BlastOutputType(blastExprType, "blastn.blablabla")
+  // case object blastOutput extends BlastOutput(blastOutputType, "blast.csv")
+
+  case object testData extends MetagenomicaData(
+    readsType = illumina.PairedEnd(bp300, InsertSize(3000)),
+    blastOutRec = outRec
+  ) {
+
+    def blastExpr(args: ValueOf[blastn.Arguments]): BlastExpression[BlastExprType] = {
+      BlastExpression(blastExprType)(
+        argumentValues = args,
+        optionValues   = blastn.defaults update (
+          num_threads(1) :~:
+          max_target_seqs(10) :~:
+          ohnosequences.blast.api.evalue(0.001)  :~:
+          blastn.task(blastn.megablast) :~: ∅
+        )
+      )
+    }
+
+  }
+
+}

### 2015-10-11 docs/paper/out/paper.tex

@@
 
+\section{Figure 1: MG7 analysis
+workflow.}\label{figure-1-mg7-analysis-workflow.}
+
+The paired reads in fastq format are merged resulting in only one
+sequence per read pair. The next step is a parallelized BLASTN of every
+merged sequence against the 16S reference database 16S-DB7. Then, the
+mapping of the detected similar sequences in the database to the
+taxonomy node to which they belong is carried out. This is done using
+Bio4j that includes a module with all the NCBI taxonomy in a graph
+connected with the Gene Ontology, Uniprot, and RefSeq graphs. Then the
+taxonomic assignment is done for each sequence following two different
+approaches: LCA and BBH, and finally the abundances corresponding to
+direct and cumulative assignments for each node in percentage and
+absolute counts are provided for each assignment mode.
+
+\section{Figure 2: Lowest Common Ancestor algorithm for taxonomic
+assignment.}\label{figure-2-lowest-common-ancestor-algorithm-for-taxonomic-assignment.}
+
+The Left panel displays an example of the application of LCA algorithm
+in a \emph{sensu stricto} mode. A, B, C and D represent taxonomy tree
+nodes with assigned reads. Right panel displays the \emph{in line} mode
+of assignment which is an exception for the \emph{sensu stricto} mode of
+application of LCA algorithm. The \emph{in line} mode is used when all
+the nodes are located in a line without bifurcations. In that case the
+taxon assigned is the most specific (the most distant from the root).
+
 \bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations

### 2016-07-22 docs/paper/out/out.sh
remove old Frontiers paper sources

@@ -1,8 +1,0 @@
-#!/bin/sh
-set -x
-ME=`dirname $0`
-pandoc --natbib --template=${ME}/frontiers.tex ${ME}/../paper.md --smart -s -o ${ME}/paper.tex
-cd ${ME}
-pdflatex -halt-on-error paper.tex
-bibtex paper
-pdflatex -halt-on-error paper.tex

### 2016-07-22 docs/paper/out/watch.sh

@@ -1,4 +1,0 @@
-#!/bin/sh
-set -x
-ME=`dirname $0`
-ls ${ME}/../paper.md ${ME}/refs.bib | entr -c ${ME}/out.sh

### 2016-07-22 docs/mg7-paper/out/paper.tex
add preprint as a starting point for the paper, set up infrastructure

--- 2015-09-24 docs/paper/out/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[fleqn,10pt,lineno]{wlpeerj}
 
-%%% Version 3.1 Generated 2015/22/05 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
+\definecolor{OldCyan}{RGB}{62,127,137}
+% hyperref
+\usepackage[breaklinks=true]{hyperref}
+\hypersetup{
+  % setpagesize=false, % page size defined by xetex
+  % unicode=true, % unicode breaks when used with xetex
+  pdfnewwindow,
+  colorlinks,%
+  citecolor=OldCyan,%
+  filecolor=OldCyan,%
+  linkcolor=OldCyan,%
+  urlcolor=OldCyan,
+  urlbordercolor=OldCyan
+}
+\let\OldHref\href
+\renewcommand{\href}[2]{\OldHref[pdfnewwindow]{#1}{{#2}}}
 
-\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
-%\documentclass{frontiersHLTH} % for Health articles
-%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
-
-%\setcitestyle{square}
-\usepackage{url,hyperref,lineno,microtype}
-\usepackage[onehalfspacing]{setspace}
-\linenumbers
-
-% fix for pandoc 1.14
+% ugly fix for latest pandoc
 \providecommand{\tightlist}{%
@@
 
-% Leave a blank line between paragraphs instead of using \\
+\title{MG7: Configurable and scalable 16S metagenomics data analysis}
 
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Alexey Alekhin {et~al.}} %use et al only if is more than 1 author
-\def\Authors{
- Alexey Alekhin\(^\dagger\)\,$^{1}$  Evdokim Kovach\(^\dagger\)\,$^{1}$  Marina Manrique\,$^{1}$  Pablo Pareja-Tobes\,$^{1}$  Eduardo Pareja\,$^{1}$  Raquel Tobes\,$^{1}$  and Eduardo Pareja-Tobes $^{1,*}$
-}% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{$^{1}$Oh no sequences! Research Group, Era7 Bioinformatics, Granada, Spain}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Eduardo Pareja-Tobes}
-\def\corrAddress{Oh no sequences! Research Group, Era7 Bioinformatics, Plaza Campo Verde 3, Granada, 18001, Spain}
-\def\corrEmail{eparejatobes@ohnosequences.com}
+\author[1]{Alexey Alekhin\(^\dagger\)} \affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
+\author[2]{Evdokim Kovach\(^\dagger\)} \affil[2]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
+\author[3]{Marina Manrique} \affil[3]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
+\author[4]{Pablo Pareja-Tobes} \affil[4]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
+\author[5]{Eduardo Pareja} \affil[5]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
+\author[6]{Raquel Tobes} \affil[6]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
+\author[7]{Eduardo Pareja-Tobes} \affil[7]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
 
-\begin{document}
-\onecolumn
-\firstpage{1}
+\keywords{ Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
+ Cloud computing, NGS, Genomic big data, Microbiome, Environmental, 16S
+ Database}
 
-\title[MG7: Configurable and scalable 16S metagenomics data analysis]{MG7: Configurable and scalable 16S metagenomics data analysis -- new
-methods optimized for massive cloud computing}
+% PeerJ doesn't like numbered sections
+\setcounter{secnumdepth}{0}
 
-\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
-
-\maketitle
+% \renewcommand\refname{}
 
 \begin{abstract}
-
 As part of the Cambrian explosion of omics data, metagenomics brings to
@@
 sustainable update system, and the possibility of project-driven
-personalization.
+personalization. \(^\dagger\) The first and second authors contributed
+equally to this work.
+\end{abstract}
 
-\(^\dagger\) The first and second authors contributed equally to this
-work.
+\begin{document}
 
-\tiny
- \keyFont{ \section{Keywords:} Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
-Cloud computing, NGS, Genomic big data, Microbiome, Environmental, 16S
-Database } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
-\end{abstract}
+\flushbottom
+\maketitle
+\thispagestyle{empty}
 
@@
 agrifood sector \citep{coughlan2015biotechnological} and biotechnology
-in general \citep[\citet{kodzius2015marine}]{cowan2015metagenomics}.
+in general \citep{cowan2015metagenomics} \citep{kodzius2015marine}.
 These new possibilities for exploring the diversity of micro-organisms
@@
 evolution of the gut microbiome after antibiotic treatment, or for colon
-cancer early detection
-\citep[\citet{garrett2015cancer}]{zeller2014potential}. This need of
-sampling across time adds more complexity to metagenomics data storage
-and demands adapted algorithms to detect state variations across time as
-well as idiosyncratic commonalities of the microbiome of each individual
+cancer early detection \citep{zeller2014potential}
+\citep{garrett2015cancer}. This need of sampling across time adds more
+complexity to metagenomics data storage and demands adapted algorithms
+to detect state variations across time as well as idiosyncratic
+commonalities of the microbiome of each individual
 \citep{franzosa2015identifying}. In addition to the intra-individual
@@
 an extensible record system with concatenation on disjoint labels, in
-the spirit of \citep[\citet{harper1991record}]{harper1990extensible}.
-For that \emph{Datasets} uses
+the spirit of \citep{harper1990extensible} \citep{harper1991record}. For
+that \emph{Datasets} uses
 \href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
@@
 
-\citep[Bio4j][]{pareja2015bio4j} is a data platform integrating data
-from different resources such as UniProt or GO in a graph data paradigm.
-In the assignment phase we use a subgraph containing the NCBI Taxonomy,
+Bio4j \citep{pareja2015bio4j} is a data platform integrating data from
+different resources such as UniProt or GO in a graph data paradigm. In
+the assignment phase we use a subgraph containing the NCBI Taxonomy,
 wrapping in Scala its Java API in a tree algebraic data type.
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{./Figure-1.jpg}
+\caption{\textbf{MG7 analysis workflow.} The paired reads in fastq
+format are merged resulting in only one sequence per read pair. The next
+step is a parallelized BLASTN of every merged sequence against the 16S
+reference database 16S-DB7. Then, the mapping of the detected similar
+sequences in the database to the taxonomy node to which they belong is
+carried out. This is done using Bio4j that includes a module with all
+the NCBI taxonomy in a graph connected with the Gene Ontology, Uniprot,
+and RefSeq graphs. Then the taxonomic assignment is done for each
+sequence following two different approaches: LCA and BBH, and finally
+the abundances corresponding to direct and cumulative assignments for
+each node in percentage and absolute counts are provided for each
+assignment mode.}
+\end{figure}
+
 \subsubsection{Joining reads of each pair using
@@
 considered in different reviews of metagenomics analysis methodologies
-\citep[\citet{morgan2012chapter}]{segata2013computational} as a very
+\citep{segata2013computational} \citep{morgan2012chapter} as a very
 exhaustive method for assignment. Some methods of assignment compare the
@@
 the assignment to the most specific taxon.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{./Figure-2.jpg}
+\caption{\textbf{Lowest Common Ancestor algorithm for taxonomic
+assignment.} The Left panel displays an example of the application of
+LCA algorithm in a \emph{sensu stricto} mode. A, B, C and D represent
+taxonomy tree nodes with assigned reads. Right panel displays the
+\emph{in line} mode of assignment which is an exception for the
+\emph{sensu stricto} mode of application of LCA algorithm. The \emph{in
+line} mode is used when all the nodes are located in a line without
+bifurcations. In that case the taxon assigned is the most specific (the
+most distant from the root).}
+\end{figure}
 
@@
 metagenomics analysis approaches \citep{huson2012microbial} have adopted
-LCA assignment algorithms because it provides fine and trusted
+LCA-based assignment algorithms because it provides fine and trusted
 taxonomical assignment.
@@
 taxonomy tree as a graph using the graph database paradigm. Previously
-we developed Bio4j \textbf{{[}Pareja-Tobes-2015{]}}, a platform for the
+we developed Bio4j \citep{pareja2015bio4j}, a platform for the
 integration of semantically rich biological data using typed graph
@@
 
-\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
-\bibliography{refs}
+\bibliography{./out/../refs}
 

### 2016-07-23 src/main/scala/mg7/package.scala
change format and add comments to mg7 package object

@@ -4,12 +4,10 @@ import ohnosequences.mg7.bio4j.taxonomyTree._
 import ohnosequences.cosas._, types._, klists._, typeUnions._
 import ohnosequences.blast.api._
 
-import better.files._
-
 package object mg7 {
 
-  type ID = String
-  type TaxID = ID
+  type ID     = String
+  type TaxID  = ID
   type ReadID = ID
   type NodeID = ID
 
@@ -23,17 +21,18 @@ package object mg7 {
   def parseLong(str: String): Option[Long] = util.Try(str.toLong).toOption
   def parseDouble(str: String): Option[Double] = util.Try(str.toDouble).toOption
 
+  // TODO why this?
   def lookup[A, B](a: A, m: Map[A, B]): (A, B) = a -> m.apply(a)
 
-  def maximums[T, X](s: Iterable[T])(f: T => X)
-    (implicit cmp: Ordering[X]): List[T] =
-      s.foldLeft(List[T]()) {
-        case (a :: acc, t) if (    cmp.lt(f(t), f(a)) ) => a :: acc
-        case (a :: acc, t) if ( cmp.equiv(f(t), f(a)) ) => t :: a :: acc
-        // either acc is empty or t is the new maximum
-        case (_, t) => List(t)
-      }
-
+  // TODO why not create an ordering from `f` and use std library?
+  def maximums[T, X](s: Iterable[T])(f: T => X)(implicit cmp: Ordering[X]): List[T] =
+    s.foldLeft(List[T]()) {
+      case (a :: acc, t) if (    cmp.lt(f(t), f(a)) ) => a :: acc
+      case (a :: acc, t) if ( cmp.equiv(f(t), f(a)) ) => t :: a :: acc
+      // either acc is empty or t is the new maximum
+      case (_, t) => List(t)
+    }
+  // TODO why not as ops?
   def averageOf(vals: Seq[Double]): Double = vals.sum / vals.length
 
 
@@ -43,22 +42,21 @@ package object mg7 {
     (out.type   := out.Raw)   ::
     *[AnyDenotation]
 
-
   // We set here all options explicitly
   val defaultBlastnOptions: blastn.Options := blastn.OptionsVals =
     blastn.options(
       /* This actually depends on the workers instance type */
-      num_threads(4) ::
-      blastn.task(blastn.blastn) ::
-      evalue(BigDecimal(1E-100)) ::
+      num_threads(4)              ::
+      blastn.task(blastn.blastn)  ::
+      evalue(BigDecimal(1E-100))  ::
[…]

### 2016-07-23 src/main/scala/mg7/package.scala
move blast defaults to their own scope

@@ -35,30 +35,9 @@ package object mg7 {
   // TODO why not as ops?
   def averageOf(vals: Seq[Double]): Double = vals.sum / vals.length
 
-
   type BlastArgumentsVals =
     (db.type    := db.Raw)    ::
     (query.type := query.Raw) ::
     (out.type   := out.Raw)   ::
     *[AnyDenotation]
-
-  // We set here all options explicitly
-  val defaultBlastnOptions: blastn.Options := blastn.OptionsVals =
-    blastn.options(
-      /* This actually depends on the workers instance type */
-      num_threads(4)              ::
-      blastn.task(blastn.blastn)  ::
-      evalue(BigDecimal(1E-100))  ::
-      /* We're going to use all hits to do global sample-coherent assignment. But not now, so no reason for this to be huge */
-      max_target_seqs(150)        ::
-      strand(Strands.both)        ::
-      word_size(46)               ::
-      show_gis(false)             ::
-      ungapped(false)             ::
-      penalty(-2)                 ::
-      reward(1)                   ::
-      /* 95% is a reasonable minimum. If it does not work, be more stringent with read preprocessing */
-      perc_identity(95.0) ::
-      *[AnyDenotation]
-    )
 }

### 2016-07-23 src/main/scala/mg7/parameters.scala

@@ -10,7 +10,6 @@ import ohnosequences.blast.api.{ outputFields => out }
 
 import better.files._
 
-
 sealed trait SplitInputFormat
 case object FastaInput extends SplitInputFormat
 case object FastQInput extends SplitInputFormat
@@ -58,9 +57,7 @@ trait AnyMG7Parameters {
   implicit val argValsToSeq: BlastOptionsToSeq[BlastArgumentsVals]
   implicit val optValsToSeq: BlastOptionsToSeq[BlastCommand#OptionsVals]
 
-  def blastExpr(inFile: File, outFile: File):
-    BlastExpression[BlastCommand, BlastOutputRecord[BlastOutRecKeys]] = {
-
+  def blastExpr(inFile: File, outFile: File): BlastExpression[BlastCommand, BlastOutputRecord[BlastOutRecKeys]] =
     BlastExpression(blastCommand)(
       outputRecord = blastOutRec,
       argumentValues =
@@ -70,7 +67,6 @@ trait AnyMG7Parameters {
         *[AnyDenotation],
       optionValues = blastOptions
     )(argValsToSeq, optValsToSeq)
-  }
 
 
   // The minimal set of the output fields neccessary for the MG7 generic code
@@ -102,8 +98,8 @@ abstract class MG7Parameters[
   val splitInputFormat: SplitInputFormat = FastQInput,
   val splitChunkSize: Int = 10,
   val blastCommand: BC = blastn,
-  val blastOutRec: BlastOutputRecord[BK]  = defaultBlastOutRec,
-  val blastOptions: BC#OptionsVals        = defaultBlastnOptions.value,
+  val blastOutRec: BlastOutputRecord[BK]  = defaults.blastnOutputRecord,
+  val blastOptions: BC#OptionsVals        = defaults.blastnOptions.value,
   val referenceDBs: Set[AnyReferenceDB]
 )(implicit
   val argValsToSeq: BlastOptionsToSeq[BC#ArgumentsVals],
@@ -120,27 +116,49 @@ abstract class MG7Parameters[
   type BlastOutRecKeys = BK
 }
 
+case object defaults {
 
-case object defaultBlastOutRec extends BlastOutputRecord(
-  // query
-  out.qseqid      :×:
-  out.qstart      :×:
-  out.qend        :×:
-  out.qlen        :×:
-  // reference
-  out.sseqid      :×:
-  out.sstart      :×:
-  out.send        :×:
[…]

### 2016-07-23 src/main/scala/mg7/parameters.scala
note about default parameters

@@ -98,6 +98,7 @@ abstract class MG7Parameters[
   val splitInputFormat: SplitInputFormat = FastQInput,
   val splitChunkSize: Int = 10,
   val blastCommand: BC = blastn,
+  // TODO why default values? not good
   val blastOutRec: BlastOutputRecord[BK]  = defaults.blastnOutputRecord,
   val blastOptions: BC#OptionsVals        = defaults.blastnOptions.value,
   val referenceDBs: Set[AnyReferenceDB]

### 2016-07-23 src/main/scala/mg7/dataflow.scala
minor formatting in dataflow

@@ -7,33 +7,30 @@ import ohnosequences.awstools.s3._
 
 trait AnyDataflow {
 
-  // param-pam-pam
   type Params <: AnyMG7Parameters
   val  params: Params
 
   /* The essential steps of any MG7 dataflow are */
-
-  val splitDataMappings: List[DataMapping[splitDataProcessing]]
-
+  val splitDataMappings : List[DataMapping[splitDataProcessing]]
   /* - BLAST */
-  val blastDataMappings: List[DataMapping[blastDataProcessing[Params]]]
-
+  val blastDataMappings : List[DataMapping[blastDataProcessing[Params]]]
   /* - Assignment */
   val assignDataMappings: List[DataMapping[assignDataProcessing[Params]]]
-
   /* - Merge */
-  val mergeDataMappings: List[DataMapping[mergeDataProcessing.type]]
-
+  val mergeDataMappings : List[DataMapping[mergeDataProcessing.type]]
   /* - Counting */
   lazy val countDataMappings: List[DataMapping[countDataProcessing.type]] =
     mergeDataMappings.map { case mergeDM =>
+
       val sampleId = mergeDM.label
 
       DataMapping(sampleId, countDataProcessing)(
+
         remoteInput = Map(
           lookup(data.lcaCSV, mergeDM.remoteOutput),
           lookup(data.bbhCSV, mergeDM.remoteOutput)
         ),
+
         remoteOutput = Map(
           data.lcaDirectCountsCSV     -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.lca.direct.absolute.counts.csv"),
           data.lcaAccumCountsCSV      -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.lca.accum.absolute.counts.csv"),
@@ -46,5 +43,4 @@ trait AnyDataflow {
         )
       )
     }
-
 }

### 2016-07-23 src/main/scala/mg7/csv.scala
minor formatting in csv

@@ -1,16 +1,16 @@
 package ohnosequences.mg7
 
 import ohnosequences.cosas._, types._, klists._, typeUnions._
-
 import com.github.tototoshi.csv._
 import better.files._
 
 // Some minimal CSV utils, will be replaced by a specialized library based on cosas
-
 case object csv {
 
   case object UnixCSVFormat extends DefaultCSVFormat {
-    override val lineTerminator: String = "\n"
+
+    override val lineTerminator: String =
+      "\n"
   }
 
   def newWriter(file: File, append: Boolean = true): CSVWriter =
@@ -50,27 +50,28 @@ case object csv {
   case class Row[Hs <: AnyProductType](
     val header: Hs,
     val values: Seq[String]
-  ) {
+  )
+  {
 
-    def toMap: Map[AnyType, String] = header.types.asList.zip(values).toMap
+    def toMap: Map[AnyType, String] =
+      header.types.asList.zip(values).toMap
 
-    def select[C <: AnyType](column: C)(implicit
-      check: C isOneOf Hs#Types#AllTypes
-    ): String =
+    def select[C <: AnyType](column: C)(implicit check: C isOneOf Hs#Types#AllTypes): String =
       this.toMap.apply(column)
   }
 
   case class Reader[Hs <: AnyProductType](val header: Hs, val csvReader: CSVReader) {
 
-    def rows: Iterator[Row[Hs]] = csvReader.iterator.map { Row(header, _) }
+    def rows: Iterator[Row[Hs]] =
+      csvReader.iterator.map { Row(header, _) }
   }
 
-  implicit def toCSVReader[Hs <: AnyProductType](r: Reader[Hs]): CSVReader = r.csvReader
+  implicit def toCSVReader[Hs <: AnyProductType](r: Reader[Hs]): CSVReader =
+    r.csvReader
 
   case object Reader {
 
     def apply[Hs <: AnyProductType](header: Hs, file: File): Reader[Hs] =
       Reader(header, CSVReader.open(file.toJava)(UnixCSVFormat))
   }
-
 }

### 2016-07-23 src/main/scala/mg7/csv.scala
note about column names

@@ -24,6 +24,7 @@ case object csv {
 
     val Lineage = "Lineage"
     val ReadID  = "Read-ID"
+    // TODO better Taxa/Taxon
     val TaxID   = "Tax-ID"
     val TaxName = "Tax-name"
     val TaxRank = "Tax-rank"

### 2016-07-23 src/main/scala/mg7/dataflows/full.scala
improve std dataflow description

@@ -16,16 +16,17 @@ import ohnosequences.awstools.autoscaling._
 import ohnosequences.awstools.regions.Region._
 import com.amazonaws.auth._, profile._
 
+// TODO this description is outdated right? assignment is done per chunk IIRC
 /* ## Standard Dataflow
 
-  Standard dataflow consists of all steps of the MG7 pipeline:
+  Standard dataflow, which will perform the following steps:
 
-  1. flash: merging paired end reads
-  2. split: splitting each dataset of reads on small chunks
-  3. blast: processing each chunk of reads with blast
-  4. merge: merging blast chunks into complete results per original reads datasets
-  5. assign: assigning taxons (LCA and BBH)
-  6. count: count assigns
+  1. **flash** merge paired-end reads
+  2. **split** split each reads dataset into smaller chunks
+  3. **blast** blast each chunk of reads against the reference database
+  4. **merge** merge blast results for each reads dataset
+  5. **assign** assign reads to taxa (LCA and BBH)
+  6. **count** count assignments
 */
 trait AnyFullDataflow extends AnyNoFlashDataflow {
 

### 2016-07-23 src/main/scala/mg7/dataflows/full.scala
minor formatting in std dataflow

@@ -32,69 +32,68 @@ trait AnyFullDataflow extends AnyNoFlashDataflow {
 
   val flashInputs: Map[SampleID, (S3Resource, S3Resource)]
 
-  lazy val flashDataMappings: List[AnyDataMapping] = flashInputs.toList.map {
-    case (sampleId, (reads1S3Resource, reads2S3Resource)) =>
+  lazy val flashDataMappings: List[AnyDataMapping] =
+    flashInputs.toList.map {
+      case (sampleId, (reads1S3Resource, reads2S3Resource)) =>
+        DataMapping(sampleId, flashDataProcessing(params))(
+          remoteInput = Map(
+            data.pairedReads1 -> reads1S3Resource,
+            data.pairedReads2 -> reads2S3Resource
+          ),
+          remoteOutput = Map(
+            data.mergedReads    -> S3Resource(params.outputS3Folder(sampleId, "flash") / s"${sampleId}.merged.fastq"),
+            data.pair1NotMerged -> S3Resource(params.outputS3Folder(sampleId, "flash") / s"${sampleId}.pair1.not-merged.fastq"),
+            data.pair2NotMerged -> S3Resource(params.outputS3Folder(sampleId, "flash") / s"${sampleId}.pair2.not-merged.fastq"),
+            data.flashHistogram -> S3Resource(params.outputS3Folder(sampleId, "flash") / s"${sampleId}.hist")
+          )
+        )
+    }
 
-      DataMapping(sampleId, flashDataProcessing(params))(
+  val splitInputs: Map[SampleID, S3Resource] =
+    flashDataMappings.map { flashDM => flashDM.label -> flashDM.remoteOutput(data.mergedReads) }.toMap
+
+  lazy val statsDataMappings: List[AnyDataMapping] =
+    List[List[AnyDataMapping]](
+      flashDataMappings,
+      mergeDataMappings,
+      assignDataMappings
+    ).flatten
+    .groupBy { _.label }
+    .map {
+      case (sampleID: String, dms: List[AnyDataMapping]) =>
+        val outputs: Map[AnyData, S3Resource] =
+          dms.map{ _.remoteOutput }.foldLeft(Map[AnyData, S3Resource]()){ _ ++ _ }
+
+        DataMapping(sampleID, statsDataProcessing)(
+          remoteInput = Map[AnyData, AnyRemoteResource](
+            data.sampleID       -> MessageResource(sampleID),
[…]

### 2016-07-23 src/main/scala/mg7/dataflows/noFlash.scala
improve no flash dataflow description

@@ -16,31 +16,32 @@ import ohnosequences.awstools.autoscaling._
 import ohnosequences.awstools.regions.Region._
 import com.amazonaws.auth._, profile._
 
+// TODO this description is outdated right? assignment is done per chunk IIRC
 /* ## No-flash Dataflow
 
-  No-flash dataflow consists of the following steps of the MG7 pipeline:
+  No-flash dataflow does the same as the full one *but* the first flash-based merging step:
 
-  2. split: splitting each dataset of reads on small chunks
-  3. blast: processing each chunk of reads with blast
-  4. merge: merging blast chunks into complete results per original reads datasets
-  5. assign: assigning taxons (LCA and BBH)
-  6. count: count assigns
+  1. **split** split each reads dataset into smaller chunks
+  2. **blast** blast each chunk of reads against the reference database
+  3. **merge** merge blast results for each reads dataset
+  4. **assign** assign reads to taxa (LCA and BBH)
+  5. **count** count assignments
 */
 trait AnyNoFlashDataflow extends AnyDataflow {
 
   val splitInputs: Map[SampleID, S3Resource]
 
-  lazy val splitDataMappings = splitInputs.toList.map { case (sampleId, readsS3Resource) =>
-
-    DataMapping(sampleId, splitDataProcessing(params))(
-      remoteInput = Map(
-        data.mergedReads -> readsS3Resource
-      ),
-      remoteOutput = Map(
-        data.fastaChunks -> S3Resource(params.outputS3Folder(sampleId, "split"))
+  lazy val splitDataMappings =
+    splitInputs.toList.map { case (sampleId, readsS3Resource) =>
+      DataMapping(sampleId, splitDataProcessing(params))(
+        remoteInput = Map(
+          data.mergedReads -> readsS3Resource
+        ),
+        remoteOutput = Map(
+          data.fastaChunks -> S3Resource(params.outputS3Folder(sampleId, "split"))
+        )
       )
-    )
-  }
+    }
 
   lazy val blastDataMappings = splitDataMappings.flatMap { splitDM =>
     val sampleId = splitDM.label

### 2016-07-23 src/main/scala/mg7/dataflows/noFlash.scala
review no flash dataflow

@@ -43,82 +43,90 @@ trait AnyNoFlashDataflow extends AnyDataflow {
       )
     }
 
-  lazy val blastDataMappings = splitDataMappings.flatMap { splitDM =>
-    val sampleId = splitDM.label
+  // TODO what's this supposed to be doing?
+  lazy val blastDataMappings =
+    splitDataMappings.flatMap { splitDM =>
+      val sampleId = splitDM.label
 
-    lazy val s3 = S3.create(
-      new AWSCredentialsProviderChain(
-        new InstanceProfileCredentialsProvider(),
-        new ProfileCredentialsProvider()
+      lazy val s3 = S3.create(
+        new AWSCredentialsProviderChain(
+          new InstanceProfileCredentialsProvider(),
+          new ProfileCredentialsProvider()
+        )
       )
-    )
 
-    lazy val chunksS3Folder: AnyS3Address = splitDM.remoteOutput(data.fastaChunks).resource
-    lazy val chunks: List[S3Object] = s3.listObjects(chunksS3Folder.bucket, chunksS3Folder.key)
+      lazy val chunksS3Folder: AnyS3Address = splitDM.remoteOutput(data.fastaChunks).resource
+      lazy val chunks: List[S3Object] = s3.listObjects(chunksS3Folder.bucket, chunksS3Folder.key)
 
-    chunks.zipWithIndex.map { case (chunkS3Obj, n) =>
+      chunks.zipWithIndex.map { case (chunkS3Obj, n) =>
 
-      DataMapping(s"${sampleId}.${n}", blastDataProcessing(params))(
+        DataMapping(s"${sampleId}.${n}", blastDataProcessing(params))(
+          remoteInput = Map(
+            data.fastaChunk -> S3Resource(chunkS3Obj)
+          ),
+          remoteOutput = Map(
+            data.blastChunk -> S3Resource(params.outputS3Folder(sampleId, "blast") / "chunks" / s"blast.${n}.csv"),
+            data.noHitsChunk -> S3Resource(params.outputS3Folder(sampleId, "blast") / "no-hits" / s"no-hits.${n}.fa")
+          )
+        )
+      }
+    }
+
+  lazy val assignDataMappings =
+    splitInputs.keys.toList.flatMap {
+      // TODO there are a lot of similar anonymous functions like this one which should use braces, so that the scope is clear
+      case sampleId => {
+
[…]

### 2016-07-23 src/main/scala/mg7/loquats/1.flash.scala
review flash loquat

@@ -1,28 +1,25 @@
 package ohnosequences.mg7.loquats
 
+// TODO in general I prefer imports to occupy less space
 import ohnosequences.mg7._
-
 import ohnosequences.loquat._
-
 import ohnosequences.statika._
-
 import ohnosequences.{ flash => f }, f.api._
-
 import ohnosequences.cosas._, types._, klists._
-
 import ohnosequences.datasets._
-
 import better.files._
 
+// TODO move all these "tool" bundles to a separate file
 case object flashBundle extends ohnosequencesBundles.statika.Flash("1.2.11")
 
 case class flashDataProcessing[MD <: AnyMG7Parameters](val md: MD)
 extends DataProcessingBundle(
   flashBundle
 )(
-  input = data.flashInput,
-  output = data.flashOutput
-) {
+  input   = data.flashInput,
+  output  = data.flashOutput
+)
+{
 
   def instructions: AnyInstructions = say("I'll be fast as a flash!")
 
@@ -54,10 +51,9 @@ extends DataProcessingBundle(
       md.flashOptions
     )
 
-    // run expression, hope for the best
     cmd("gunzip")(reads1gz.path.toString) -&-
     cmd("gunzip")(reads2gz.path.toString) -&-
-    seqToInstructions(flashExpr.cmd) -&-
+    seqToInstructions(flashExpr.cmd)      -&-
     success(
       "FLASh merged reads, much success so fast",
       data.mergedReads(flashOutput.mergedReads) ::

### 2016-07-23 src/main/scala/mg7/loquats/4.assign.scala
review assign loquat

@@ -1,29 +1,27 @@
 package ohnosequences.mg7.loquats
 
-import ohnosequences.mg7._
-import ohnosequences.mg7.bio4j._, taxonomyTree._, titanTaxonomyTree._
+import ohnosequences.mg7._, bio4j._, taxonomyTree._, titanTaxonomyTree._
 import ohnosequences.loquat._
 import ohnosequences.statika._
 import ohnosequences.cosas._, types._, klists._
 import ohnosequences.datasets._
 import ohnosequences.blast.api._, outputFields._
-
 import com.bio4j.titan.model.ncbiTaxonomy.TitanNCBITaxonomyGraph
-
 import java.io.{ BufferedWriter, FileWriter, File }
 import scala.util.Try
-
 import com.github.tototoshi.csv._
 
 
-case class assignDataProcessing[MD <: AnyMG7Parameters](val md: MD)
-extends DataProcessingBundle()(
+case class assignDataProcessing[MD <: AnyMG7Parameters](val md: MD) extends DataProcessingBundle()(
   input  = data.assignInput,
   output = data.assignOutput
-) {
+)
+{
+
   // For the output fields implicits
   import md._
 
+  // TODO why override val here? why this is not a param?
   override val bundleDependencies: List[AnyBundle] =
     bio4j.taxonomyBundle :: md.referenceDBs.toList
 
@@ -33,6 +31,7 @@ extends DataProcessingBundle()(
 
   // This iterates over reference DBs and merges their id2taxa tables in one Map
   private lazy val referenceMap: Map[ID, Seq[TaxID]] = {
+
     val refMap: scala.collection.mutable.Map[ID, Seq[TaxID]] = scala.collection.mutable.Map()
 
     md.referenceDBs.foreach { refDB =>
@@ -51,11 +50,12 @@ extends DataProcessingBundle()(
     refMap.toMap
   }
 
-  private def taxIDsFor(id: ID): Seq[TaxID] = referenceMap.get(id).getOrElse(Seq())
-
+  private def taxIDsFor(id: ID): Seq[TaxID] =
+    referenceMap.get(id).getOrElse(Seq())
 
   def instructions: AnyInstructions = say("Let's see who is who!")
 
+  // TODO this is too big. Factor BBH and LCA into methods
   def process(context: ProcessingContext[Input]): AnyInstructions { type Out <: OutputFiles } = {
 
     val blastReader = csv.Reader(md.blastOutRec.keys, context.inputFile(data.blastChunk))

### 2016-07-25 docs/mg7-paper/out/paper.tex
remove equal contribution remark

@@
 
-\author[1]{Alexey Alekhin\(^\dagger\)} \affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\author[1]{Alexey Alekhin} \affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
 \href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[2]{Evdokim Kovach\(^\dagger\)} \affil[2]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\author[2]{Evdokim Kovach} \affil[2]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
 \href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
@@
 sustainable update system, and the possibility of project-driven
-personalization. \(^\dagger\) The first and second authors contributed
-equally to this work.
+personalization.
 \end{abstract}
@@
 
-\bibliography{./out/../refs}
+\bibliography{out/../refs}
 

### 2016-07-25 docs/mg7-paper/out/paper.tex
no 16S metagenomics, only 16S data analysis
fix keywords

@@
 
-\title{MG7: Configurable and scalable 16S metagenomics data analysis}
+\title{MG7: Configurable and scalable 16S data analysis}
 
@@
 
-\keywords{ Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
- Cloud computing, NGS, Genomic big data, Microbiome, Environmental, 16S
+\keywords{ Metagenomics, 16S, taxonomic profiling, Bio4j, Graph databases, Cloud
+ computing, NGS, Genomics, big data, Microbiome, Environmental, 16S
  Database}

### 2016-07-25 docs/mg7-paper/out/paper.tex
add ref to metagenomics and bioenergy

@@
 environmental sciences \citep{ufarte2015discovery}, crop sciences, the
-agrifood sector \citep{coughlan2015biotechnological} and biotechnology
-in general \citep{cowan2015metagenomics} \citep{kodzius2015marine}.
-These new possibilities for exploring the diversity of micro-organisms
-in the most varied environments are opening new research areas, and
-drastically changing the existing ones.
+agrifood sector \citep{coughlan2015biotechnological}, bioenergy
+\citep{yang2016discovery}, and biotechnology in general
+\citep{cowan2015metagenomics} \citep{kodzius2015marine}. These new
+possibilities for exploring the diversity of micro-organisms in the most
+varied environments are opening new research areas, and drastically
+changing the existing ones.
 

### 2016-07-25 docs/mg7-paper/out/paper.tex
change metagenomics to microbiome analysis

@@
 different conditions of health and disease, or after different
-treatments. The impact of Metagenomics is also being felt in
+treatments. The impact of microbiome analysis is also being felt in
 environmental sciences \citep{ufarte2015discovery}, crop sciences, the

### 2016-07-25 docs/mg7-paper/out/paper.tex
make all explicit links named

@@
   Static reproducible specification of dependencies and behavior of the
-  different components using \emph{Statika} and \emph{Datasets}
+  different components using \emph{\protect\hyperlink{statika}{Statika}}
+  and \emph{\protect\hyperlink{datasets}{Datasets}}
 \item
@@
 \item
-  Definition of complex workflows using \emph{Loquat}, a composable
-  system for scaling/parallelizing stateless computations especially
-  designed for AWS
+  Definition of complex workflows using
+  \emph{\protect\hyperlink{loquat}{Loquat}}, a composable system for
+  scaling/parallelizing stateless computations especially designed for
+  AWS
 \item
@@
 authors on top of which MG7 is built. All MG7 code is written in
-\href{http://www.scala-lang.org/}{Scala}, a hybrid object-functional
-programming language. Scala was chosen based on the possibility of using
-certain advanced programming styles, and Java interoperability, which
-let us build on the vast number of existing Java libraries; we take
-advantage of this when using Bio4j as an API for the NCBI taxonomy. It
-has support for type-level programming, type-dependent types (through
-type members) and singleton types, which permits a restricted form of
-dependent types where types can depend essentially on values determined
-at compile time (through their corresponding singleton types).
-Conversely, through implicits one can retrieve the value corresponding
-to a singleton type.
+\protect\hyperlink{scala}{Scala}, a hybrid object-functional programming
+language. Scala was chosen based on the possibility of using certain
+advanced programming styles, and Java interoperability, which let us
+build on the vast number of existing Java libraries; we take advantage
+of this when using Bio4j as an API for the NCBI taxonomy. It has support
+for type-level programming, type-dependent types (through type members)
+and singleton types, which permits a restricted form of dependent types
+where types can depend essentially on values determined at compile time
+(through their corresponding singleton types). Conversely, through
+implicits one can retrieve the value corresponding to a singleton type.
 
@@
 
-\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
-library developed by the first and last authors which serves as a way of
-defining and composing machine behaviors statically. The main component
-are \textbf{bundles}. Each bundle declares a sequence of computations
-(its behavior) which will be executed in an \textbf{environment}. A
-bundle can \emph{depend} on other bundles, and when being executed by an
-environment, its DAG (Directed Acyclic Graph) of dependencies is
-linearized and run in sequence. In our use, bundles correspond to what
-an EC2 instance should do and an environment to an AMI (Amazon Machine
-Image) which prepares the basic configuration, downloads the Scala code
-and runs it.
+\protect\hyperlink{statika}{Statika} is a Scala library developed by the
+first and last authors which serves as a way of defining and composing
+machine behaviors statically. The main component are \textbf{bundles}.
+Each bundle declares a sequence of computations (its behavior) which
+will be executed in an \textbf{environment}. A bundle can \emph{depend}
+on other bundles, and when being executed by an environment, its DAG
+(Directed Acyclic Graph) of dependencies is linearized and run in
+sequence. In our use, bundles correspond to what an EC2 instance should
+do and an environment to an AMI (Amazon Machine Image) which prepares
+the basic configuration, downloads the Scala code and runs it.
 
@@
 
-\href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
-library developed by the first and last authors with the goal of being a
-Scala-embedded mini-language for datasets and their locations.
-\textbf{Data} is represented as type-indexed fields: keys are modeled as
-singleton types, and values correspond to what could be called a
-denotation of the key: a value of type \texttt{Location} tagged with the
-key type. Then a \textbf{Dataset} is essentially a collection of data,
-which are guaranteed statically to be different through type-level
-predicates, making use of the value--type correspondence which can be
-established through singleton types and implicits. A dataset location is
-then just a list of locations formed by locations of each dataset key.
-All this is based on what could be described as an embedding in Scala of
-an extensible record system with concatenation on disjoint labels, in
-the spirit of \citep{harper1990extensible} \citep{harper1991record}. For
-that \emph{Datasets} uses
-\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
+\protect\hyperlink{datasets}{Datasets} is a Scala library developed by
+the first and last authors with the goal of being a Scala-embedded
+mini-language for datasets and their locations. \textbf{Data} is
+represented as type-indexed fields: keys are modeled as singleton types,
+and values correspond to what could be called a denotation of the key: a
+value of type \texttt{Location} tagged with the key type. Then a
+\textbf{Dataset} is essentially a collection of data, which are
+guaranteed statically to be different through type-level predicates,
+making use of the value--type correspondence which can be established
+through singleton types and implicits. A dataset location is then just a
+list of locations formed by locations of each dataset key. All this is
+based on what could be described as an embedding in Scala of an
+extensible record system with concatenation on disjoint labels, in the
+spirit of \citep{harper1990extensible} \citep{harper1991record}. For
+that \emph{Datasets} uses the {[}ohnosequences/cosas{]}{[}cosas{]}
 library.
@@
 
-\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
-developed by the first, second and last authors designed for the
-execution of embarrassingly parallel tasks using S3, SQS and EC2 Amazon
-services.
+\protect\hyperlink{loquat}{Loquat} is a library developed by the first,
+second and last authors designed for the execution of embarrassingly
+parallel tasks using S3, SQS and EC2 Amazon services.
 
@@
 We developed our own Scala-based type-safe eDSLs (embedded Domain
-Specific Languages) for
-\href{https://github.com/ohnosequences/flash}{FLASh}
-\citep{magovc2011flash} and
-\href{https://github.com/ohnosequences/blast}{BLAST}
-\citep{camacho2009blast} expressions and their execution.
+Specific Languages) for {[}FLASh{]}{[}flash{]} \citep{magovc2011flash}
+and {[}BLAST{]}{[}blast{]} \citep{camacho2009blast} expressions and
+their execution.
 
@@
 
-\subsection{Scala}\label{scala}
+\hypertarget{scala}{\subsection{Scala}\label{scala}}
 
@@
 
-\subsection{Statika}\label{statika}
+\hypertarget{statika}{\subsection{Statika}\label{statika}}
 
@@
 
-\subsection{Datasets}\label{datasets}
+\hypertarget{datasets}{\subsection{Datasets}\label{datasets}}
 
@@
 
-\subsection{Loquat}\label{loquat}
+\hypertarget{loquat}{\subsection{Loquat}\label{loquat}}
 
@@
 
+\protect\hyperlink{statika}{statika} :
+https://github.com/ohnosequences/statika
+\protect\hyperlink{datasets}{datasets} :
+https://github.com/ohnosequences/datasets
+\protect\hyperlink{loquat}{loquat} :
+https://github.com/ohnosequences/loquat {[}flash{]} :
+https://github.com/ohnosequences/flash-api {[}blast{]} :
+https://github.com/ohnosequences/blast-api {[}cosas{]} :
+https://github.com/ohnosequences/cosas/ \protect\hyperlink{scala}{scala}
+: http://www.scala-lang.org/
+
 \bibliography{out/../refs}

### 2016-07-25 docs/mg7-paper/out/paper.tex
cite bio4j again

@@
   Modeling of the taxonomy tree using the new paradigm of graph
-  databases (Bio4j). It facilitates the taxonomic assignment tasks and
-  the calculation of the taxa abundance values considering the
-  hierarchic structure of taxonomy tree (cumulative values)
+  databases (Bio4j, \citep{pareja2015bio4j}). It facilitates the
+  taxonomic assignment tasks and the calculation of the taxa abundance
+  values considering the hierarchic structure of taxonomy tree
 \item

### 2016-07-25 docs/mg7-paper/out/paper.tex
switch to initials in credits

@@
 
-\protect\hyperlink{statika}{Statika} is a Scala library developed by the
-first and last authors which serves as a way of defining and composing
-machine behaviors statically. The main component are \textbf{bundles}.
-Each bundle declares a sequence of computations (its behavior) which
-will be executed in an \textbf{environment}. A bundle can \emph{depend}
-on other bundles, and when being executed by an environment, its DAG
-(Directed Acyclic Graph) of dependencies is linearized and run in
-sequence. In our use, bundles correspond to what an EC2 instance should
-do and an environment to an AMI (Amazon Machine Image) which prepares
-the basic configuration, downloads the Scala code and runs it.
+\protect\hyperlink{statika}{Statika} is a Scala library developed by
+\emph{AA} and \emph{EP-T} which serves as a way of defining and
+composing machine behaviors statically. The main component are
+\textbf{bundles}. Each bundle declares a sequence of computations (its
+behavior) which will be executed in an \textbf{environment}. A bundle
+can \emph{depend} on other bundles, and when being executed by an
+environment, its DAG (Directed Acyclic Graph) of dependencies is
+linearized and run in sequence. In our use, bundles correspond to what
+an EC2 instance should do and an environment to an AMI (Amazon Machine
+Image) which prepares the basic configuration, downloads the Scala code
+and runs it.
 
@@
 \protect\hyperlink{datasets}{Datasets} is a Scala library developed by
-the first and last authors with the goal of being a Scala-embedded
+\emph{AA} and \emph{EP-T} with the goal of being a Scala-embedded
 mini-language for datasets and their locations. \textbf{Data} is
@@
 
-\protect\hyperlink{loquat}{Loquat} is a library developed by the first,
-second and last authors designed for the execution of embarrassingly
+\protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
+\emph{EK} and \emph{EP-T} designed for the execution of embarrassingly
 parallel tasks using S3, SQS and EC2 Amazon services.

### 2016-07-25 docs/mg7-paper/out/paper.tex
add ncbi taxonomy to bio4j data
bio4j taxonomy contains ncbi taxonomy

@@
 Bio4j \citep{pareja2015bio4j} is a data platform integrating data from
-different resources such as UniProt or GO in a graph data paradigm. In
-the assignment phase we use a subgraph containing the NCBI Taxonomy,
-wrapping in Scala its Java API in a tree algebraic data type.
+different resources such as UniProt, the NCBI taxonomy, or GO, in a
+graph data paradigm. In the assignment phase we use a subgraph
+containing the NCBI Taxonomy, wrapping in Scala its Java API in a tree
+algebraic data type.
 
@@
 analysis and especially for biological data. MG7 works based on the
-Bio4j taxonomy module. It opens the possibility to connect the taxonomic
-profiling data obtained with MG7 to all the biological knowledge
-associated to each taxon. Using the information available in Bio4j for
-all the proteins assigned to each taxon we are connected to all the
-functional data available in Uniprot related with it.
+Bio4j taxonomy module, which contains all the NCBI taxonomy data. It
+opens the possibility to connect the taxonomic profiling data obtained
+with MG7 to all the biological knowledge associated to each taxon. Using
+the information available in Bio4j for all the proteins assigned to each
+taxon we are connected to all the functional data available in Uniprot
+related with it.
 

### 2016-07-25 docs/mg7-paper/out/paper.tex
update funding
update author contributions

@@
 \protect\hyperlink{statika}{Statika} is a Scala library developed by
-\emph{AA} and \emph{EP-T} which serves as a way of defining and
-composing machine behaviors statically. The main component are
-\textbf{bundles}. Each bundle declares a sequence of computations (its
-behavior) which will be executed in an \textbf{environment}. A bundle
-can \emph{depend} on other bundles, and when being executed by an
-environment, its DAG (Directed Acyclic Graph) of dependencies is
-linearized and run in sequence. In our use, bundles correspond to what
-an EC2 instance should do and an environment to an AMI (Amazon Machine
-Image) which prepares the basic configuration, downloads the Scala code
-and runs it.
+\emph{AA} and \emph{EPT} which serves as a way of defining and composing
+machine behaviors statically. The main component are \textbf{bundles}.
+Each bundle declares a sequence of computations (its behavior) which
+will be executed in an \textbf{environment}. A bundle can \emph{depend}
+on other bundles, and when being executed by an environment, its DAG
+(Directed Acyclic Graph) of dependencies is linearized and run in
+sequence. In our use, bundles correspond to what an EC2 instance should
+do and an environment to an AMI (Amazon Machine Image) which prepares
+the basic configuration, downloads the Scala code and runs it.
 
@@
 \protect\hyperlink{datasets}{Datasets} is a Scala library developed by
-\emph{AA} and \emph{EP-T} with the goal of being a Scala-embedded
+\emph{AA} and \emph{EPT} with the goal of being a Scala-embedded
 mini-language for datasets and their locations. \textbf{Data} is
@@
 \protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
-\emph{EK} and \emph{EP-T} designed for the execution of embarrassingly
+\emph{EK} and \emph{EPT} designed for the execution of embarrassingly
 parallel tasks using S3, SQS and EC2 Amazon services.
@@
 
+Partially funded by ITN INTERCROSSING (Grant 289974) and Cardiobiome
+project ITC-20151148.
+
 \section{Author Contributions}\label{author-contributions}
@@
 \item
-  \textbf{AA} developed \emph{MG7}, \emph{Statika}, \emph{Datasets}, and
-  \emph{aws-scala-tools}; wrote the paper;
+  \textbf{AA} developed \emph{MG7}, \emph{Loquat}, \emph{Statika},
+  \emph{Datasets}, and \emph{aws-scala-tools}; wrote the paper;
 \item
@@
 \item
-  \textbf{PPT} curation, design, data extraction code of the
-  \emph{16S-DB7} reference database.
+  \textbf{PPT} design and development of the first \emph{MG7} prototype
 \item
@@
   \textbf{EPT} developed \emph{MG7}, \emph{Statika}, \emph{Datasets},
-  \emph{FLASh/BLAST eDSLs}; data analysis approach and design; wrote the
-  paper.
+  \emph{FLASh/BLAST eDSLs}; data analysis approach and design; reference
+  database automated curation and filtering; wrote the paper.
 \end{itemize}
@@
 All authors have read and approved the final manuscript.
-
-\section{Acknowledgements}\label{acknowledgements}
-
-\emph{Funding:} The two first authors are funded by INTERCROSSING (Grant
-289974).
 

### 2016-07-25 docs/mg7-paper/out/paper.tex
fix weird conflict with paper sources

--- 2016-07-25 docs/mg7-paper/out/paper.tex
@@
 
-\title{MG7: Configurable and scalable 16S data analysis}
+\title{MG7: Configurable and scalable 16S metagenomics data analysis}
 
-\author[1]{Alexey Alekhin} \affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\author[1]{Alexey Alekhin\(^\dagger\)} \affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
 \href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[2]{Evdokim Kovach} \affil[2]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+\author[2]{Evdokim Kovach\(^\dagger\)} \affil[2]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
 \href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
@@
 
-\keywords{ Metagenomics, 16S, taxonomic profiling, Bio4j, Graph databases, Cloud
- computing, NGS, Genomics, big data, Microbiome, Environmental, 16S
+\keywords{ Metagenomics, 16S, Bacterial diversity profile, Bio4j, Graph databases,
+ Cloud computing, NGS, Genomic big data, Microbiome, Environmental, 16S
  Database}
@@
 sustainable update system, and the possibility of project-driven
-personalization.
+personalization. \(^\dagger\) The first and second authors contributed
+equally to this work.
 \end{abstract}
@@
 different conditions of health and disease, or after different
-treatments. The impact of microbiome analysis is also being felt in
+treatments. The impact of Metagenomics is also being felt in
 environmental sciences \citep{ufarte2015discovery}, crop sciences, the
-agrifood sector \citep{coughlan2015biotechnological}, bioenergy
-\citep{yang2016discovery}, and biotechnology in general
-\citep{cowan2015metagenomics} \citep{kodzius2015marine}. These new
-possibilities for exploring the diversity of micro-organisms in the most
-varied environments are opening new research areas, and drastically
-changing the existing ones.
+agrifood sector \citep{coughlan2015biotechnological} and biotechnology
+in general \citep{cowan2015metagenomics} \citep{kodzius2015marine}.
+These new possibilities for exploring the diversity of micro-organisms
+in the most varied environments are opening new research areas, and
+drastically changing the existing ones.
 
@@
   Static reproducible specification of dependencies and behavior of the
-  different components using \emph{\protect\hyperlink{statika}{Statika}}
-  and \emph{\protect\hyperlink{datasets}{Datasets}}
+  different components using \emph{Statika} and \emph{Datasets}
 \item
@@
 \item
-  Definition of complex workflows using
-  \emph{\protect\hyperlink{loquat}{Loquat}}, a composable system for
-  scaling/parallelizing stateless computations especially designed for
-  AWS
+  Definition of complex workflows using \emph{Loquat}, a composable
+  system for scaling/parallelizing stateless computations especially
+  designed for AWS
 \item
@@
   Modeling of the taxonomy tree using the new paradigm of graph
-  databases (Bio4j, \citep{pareja2015bio4j}). It facilitates the
-  taxonomic assignment tasks and the calculation of the taxa abundance
-  values considering the hierarchic structure of taxonomy tree
+  databases (Bio4j). It facilitates the taxonomic assignment tasks and
+  the calculation of the taxa abundance values considering the
+  hierarchic structure of taxonomy tree (cumulative values)
 \item
@@
 authors on top of which MG7 is built. All MG7 code is written in
-\protect\hyperlink{scala}{Scala}, a hybrid object-functional programming
-language. Scala was chosen based on the possibility of using certain
-advanced programming styles, and Java interoperability, which let us
-build on the vast number of existing Java libraries; we take advantage
-of this when using Bio4j as an API for the NCBI taxonomy. It has support
-for type-level programming, type-dependent types (through type members)
-and singleton types, which permits a restricted form of dependent types
-where types can depend essentially on values determined at compile time
-(through their corresponding singleton types). Conversely, through
-implicits one can retrieve the value corresponding to a singleton type.
+\href{http://www.scala-lang.org/}{Scala}, a hybrid object-functional
+programming language. Scala was chosen based on the possibility of using
+certain advanced programming styles, and Java interoperability, which
+let us build on the vast number of existing Java libraries; we take
+advantage of this when using Bio4j as an API for the NCBI taxonomy. It
+has support for type-level programming, type-dependent types (through
+type members) and singleton types, which permits a restricted form of
+dependent types where types can depend essentially on values determined
+at compile time (through their corresponding singleton types).
+Conversely, through implicits one can retrieve the value corresponding
+to a singleton type.
 
@@
 
-\protect\hyperlink{statika}{Statika} is a Scala library developed by
-\emph{AA} and \emph{EP-T} which serves as a way of defining and
-composing machine behaviors statically. The main component are
-\textbf{bundles}. Each bundle declares a sequence of computations (its
-behavior) which will be executed in an \textbf{environment}. A bundle
-can \emph{depend} on other bundles, and when being executed by an
+\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
+library developed by the first and last authors which serves as a way of
+defining and composing machine behaviors statically. The main component
+are \textbf{bundles}. Each bundle declares a sequence of computations
+(its behavior) which will be executed in an \textbf{environment}. A
+bundle can \emph{depend} on other bundles, and when being executed by an
 environment, its DAG (Directed Acyclic Graph) of dependencies is
@@
 
-\protect\hyperlink{datasets}{Datasets} is a Scala library developed by
-\emph{AA} and \emph{EP-T} with the goal of being a Scala-embedded
-mini-language for datasets and their locations. \textbf{Data} is
-represented as type-indexed fields: keys are modeled as singleton types,
-and values correspond to what could be called a denotation of the key: a
-value of type \texttt{Location} tagged with the key type. Then a
-\textbf{Dataset} is essentially a collection of data, which are
-guaranteed statically to be different through type-level predicates,
-making use of the value--type correspondence which can be established
-through singleton types and implicits. A dataset location is then just a
-list of locations formed by locations of each dataset key. All this is
-based on what could be described as an embedding in Scala of an
-extensible record system with concatenation on disjoint labels, in the
-spirit of \citep{harper1990extensible} \citep{harper1991record}. For
-that \emph{Datasets} uses the {[}ohnosequences/cosas{]}{[}cosas{]}
+\href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
+library developed by the first and last authors with the goal of being a
+Scala-embedded mini-language for datasets and their locations.
+\textbf{Data} is represented as type-indexed fields: keys are modeled as
+singleton types, and values correspond to what could be called a
+denotation of the key: a value of type \texttt{Location} tagged with the
+key type. Then a \textbf{Dataset} is essentially a collection of data,
+which are guaranteed statically to be different through type-level
+predicates, making use of the value--type correspondence which can be
+established through singleton types and implicits. A dataset location is
+then just a list of locations formed by locations of each dataset key.
+All this is based on what could be described as an embedding in Scala of
+an extensible record system with concatenation on disjoint labels, in
+the spirit of \citep{harper1990extensible} \citep{harper1991record}. For
+that \emph{Datasets} uses
+\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
 library.
@@
 
-\protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
-\emph{EK} and \emph{EP-T} designed for the execution of embarrassingly
-parallel tasks using S3, SQS and EC2 Amazon services.
+\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
+developed by the first, second and last authors designed for the
+execution of embarrassingly parallel tasks using S3, SQS and EC2 Amazon
+services.
 
@@
 We developed our own Scala-based type-safe eDSLs (embedded Domain
-Specific Languages) for {[}FLASh{]}{[}flash{]} \citep{magovc2011flash}
-and {[}BLAST{]}{[}blast{]} \citep{camacho2009blast} expressions and
-their execution.
+Specific Languages) for
+\href{https://github.com/ohnosequences/flash}{FLASh}
+\citep{magovc2011flash} and
+\href{https://github.com/ohnosequences/blast}{BLAST}
+\citep{camacho2009blast} expressions and their execution.
 
@@
 
-\hypertarget{scala}{\subsection{Scala}\label{scala}}
+\subsection{Scala}\label{scala}
 
@@
 
-\hypertarget{statika}{\subsection{Statika}\label{statika}}
+\subsection{Statika}\label{statika}
 
@@
 
-\hypertarget{datasets}{\subsection{Datasets}\label{datasets}}
+\subsection{Datasets}\label{datasets}
 
@@
 
-\hypertarget{loquat}{\subsection{Loquat}\label{loquat}}
+\subsection{Loquat}\label{loquat}
 
@@
 
-\protect\hyperlink{statika}{statika} :
-https://github.com/ohnosequences/statika
-\protect\hyperlink{datasets}{datasets} :
-https://github.com/ohnosequences/datasets
-\protect\hyperlink{loquat}{loquat} :
-https://github.com/ohnosequences/loquat {[}flash{]} :
-https://github.com/ohnosequences/flash-api {[}blast{]} :
-https://github.com/ohnosequences/blast-api {[}cosas{]} :
-https://github.com/ohnosequences/cosas/ \protect\hyperlink{scala}{scala}
-: http://www.scala-lang.org/
-
 \bibliography{out/../refs}

### 2016-07-25 docs/mg7-paper/out/paper.tex

--- 2016-07-25 docs/mg7-paper/out/paper.tex
@@
 
-\author[1]{Alexey Alekhin} \affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
-\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[2]{Evdokim Kovach} \affil[2]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
-\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[3]{Marina Manrique} \affil[3]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
-\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[4]{Pablo Pareja-Tobes} \affil[4]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
-\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[5]{Eduardo Pareja} \affil[5]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
-\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[6]{Raquel Tobes} \affil[6]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
-\href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}
-\author[7]{Eduardo Pareja-Tobes} \affil[7]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
+% modified for this paper
+\author[1]{Alexey Alekhin}
+\author[1]{Evdokim Kovach}
+\author[1]{Marina Manrique}
+\author[1]{Pablo Pareja-Tobes}
+\author[1]{Eduardo Pareja}
+\author[1]{Raquel Tobes}
+\author[1]{Eduardo Pareja-Tobes}
+\affil[1]{\emph{\href{http://ohnosequences.com}{oh no sequences!}} research group,
 \href{http://www.era7bioinformatics.com}{Era7 bioinformatics}}

### 2016-07-25 docs/mg7-paper/out/paper.tex
fix named links in paper

@@
   Static reproducible specification of dependencies and behavior of the
-  different components using \emph{\protect\hyperlink{statika}{Statika}}
-  and \emph{\protect\hyperlink{datasets}{Datasets}}
+  different components using
+  \emph{\href{https://github.com/ohnosequences/statika}{Statika}} and
+  \emph{\href{https://github.com/ohnosequences/datasets}{Datasets}}
 \item
@@
   Definition of complex workflows using
-  \emph{\protect\hyperlink{loquat}{Loquat}}, a composable system for
-  scaling/parallelizing stateless computations especially designed for
-  AWS
+  \emph{\href{https://github.com/ohnosequences/loquat}{Loquat}}, a
+  composable system for scaling/parallelizing stateless computations
+  especially designed for AWS
 \item
@@
 
-\protect\hyperlink{statika}{Statika} is a Scala library developed by
-\emph{AA} and \emph{EPT} which serves as a way of defining and composing
-machine behaviors statically. The main component are \textbf{bundles}.
-Each bundle declares a sequence of computations (its behavior) which
-will be executed in an \textbf{environment}. A bundle can \emph{depend}
-on other bundles, and when being executed by an environment, its DAG
-(Directed Acyclic Graph) of dependencies is linearized and run in
-sequence. In our use, bundles correspond to what an EC2 instance should
-do and an environment to an AMI (Amazon Machine Image) which prepares
-the basic configuration, downloads the Scala code and runs it.
+\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
+library developed by \emph{AA} and \emph{EPT} which serves as a way of
+defining and composing machine behaviors statically. The main component
+are \textbf{bundles}. Each bundle declares a sequence of computations
+(its behavior) which will be executed in an \textbf{environment}. A
+bundle can \emph{depend} on other bundles, and when being executed by an
+environment, its DAG (Directed Acyclic Graph) of dependencies is
+linearized and run in sequence. In our use, bundles correspond to what
+an EC2 instance should do and an environment to an AMI (Amazon Machine
+Image) which prepares the basic configuration, downloads the Scala code
+and runs it.
 
@@
 
-\protect\hyperlink{datasets}{Datasets} is a Scala library developed by
-\emph{AA} and \emph{EPT} with the goal of being a Scala-embedded
-mini-language for datasets and their locations. \textbf{Data} is
-represented as type-indexed fields: keys are modeled as singleton types,
-and values correspond to what could be called a denotation of the key: a
-value of type \texttt{Location} tagged with the key type. Then a
-\textbf{Dataset} is essentially a collection of data, which are
-guaranteed statically to be different through type-level predicates,
-making use of the value--type correspondence which can be established
-through singleton types and implicits. A dataset location is then just a
-list of locations formed by locations of each dataset key. All this is
-based on what could be described as an embedding in Scala of an
-extensible record system with concatenation on disjoint labels, in the
-spirit of \citep{harper1990extensible} \citep{harper1991record}. For
-that \emph{Datasets} uses the {[}ohnosequences/cosas{]}{[}cosas{]}
+\href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
+library developed by \emph{AA} and \emph{EPT} with the goal of being a
+Scala-embedded mini-language for datasets and their locations.
+\textbf{Data} is represented as type-indexed fields: keys are modeled as
+singleton types, and values correspond to what could be called a
+denotation of the key: a value of type \texttt{Location} tagged with the
+key type. Then a \textbf{Dataset} is essentially a collection of data,
+which are guaranteed statically to be different through type-level
+predicates, making use of the value--type correspondence which can be
+established through singleton types and implicits. A dataset location is
+then just a list of locations formed by locations of each dataset key.
+All this is based on what could be described as an embedding in Scala of
+an extensible record system with concatenation on disjoint labels, in
+the spirit of \citep{harper1990extensible} \citep{harper1991record}. For
+that \emph{Datasets} uses the
+\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
 library.
@@
 
-\protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
-\emph{EK} and \emph{EPT} designed for the execution of embarrassingly
-parallel tasks using S3, SQS and EC2 Amazon services.
+\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
+developed by \emph{AA}, \emph{EK} and \emph{EPT} designed for the
+execution of embarrassingly parallel tasks using S3, SQS and EC2 Amazon
+services.
 
@@
 We developed our own Scala-based type-safe eDSLs (embedded Domain
-Specific Languages) for {[}FLASh{]}{[}flash{]} \citep{magovc2011flash}
-and {[}BLAST{]}{[}blast{]} \citep{camacho2009blast} expressions and
-their execution.
+Specific Languages) for
+\href{https://github.com/ohnosequences/flash-api}{FLASh}
+\citep{magovc2011flash} and
+\href{https://github.com/ohnosequences/blast-api}{BLAST}
+\citep{camacho2009blast} expressions and their execution.
 
@@
 
+The starting point for our 16S reference database is RNA Central
+\citep{rnacentral2014rnacentral},
+\href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
+5}.
+
+\begin{quote}
 Our 16S-DB7 Reference Database is a curated subset of sequences from the
@@
 sequences with poor taxonomic assignments to taxonomic nodes close to
-the root of the taxonomy tree. All the nucleotide sequences included in
-\textbf{nt} database has a taxonomic assignment provided by the
-\textbf{Genbank} sequence submitter. NCBI provides a table (available at
+the root of the taxonomy tree.
+
+All the nucleotide sequences included in \textbf{nt} database has a
+taxonomic assignment provided by the \textbf{Genbank} sequence
+submitter. NCBI provides a table (available at
 ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
@@
 personalization.
+\end{quote}
 

### 2016-07-25 docs/mg7-paper/out/paper.tex
add a description of the first two steps in ref db generation

@@
 \href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
-5}.
+5}. RNACentral was chosen being the most up to date, comprehensive RNA
+sequence repository, including among others all RNA data from Silva,
+GreenGenes, RDP, ENA, and RefSeq. First we take those sequences which
 
-\begin{quote}
-Our 16S-DB7 Reference Database is a curated subset of sequences from the
-NCBI nucleotide database \textbf{nt}. The sequences included were
-selected by similarity with the bacterial and archaeal reference
-sequences downloaded from the \textbf{RDP database}
-\citep{cole2013ribosomal}. RDP unaligned sequences were used to capture
-new 16S RNA sequences from \textbf{nt} using BLAST similarity search
-strategies and then, performing additional curation steps to remove
-sequences with poor taxonomic assignments to taxonomic nodes close to
-the root of the taxonomy tree.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  are annotated as being of \texttt{rRNA} type\footnote{We are aware of
+    the existence of a gene annotation corresponding to 16S, that we are
+    \textbf{not using} due to a significant amount of 16S sequences
+    lacking the corresponding annotation}
+\item
+  their length is at least \texttt{1300}\footnote{Note that 16S
+    sequences are sometimes part of an entry corresponding to whole
+    small subunits; that's why we do not set a maximum length threshold}
+\item
+  have at least one taxonomic assignment to a descendant of
+  \emph{Bacteria} or \emph{Archaea}
+\item
+  their lineage does not contain a set of taxa deemed
+  uninformative\footnote{for example: ``unclassified Bacteria
+    (miscellaneous)'' or ``unclassified''. Assigning a read to an
+    ``unclassified'' taxon defeats the first and foremost goal of
+    taxonomic profiling: \emph{classification}.}
+\end{enumerate}
 
-All the nucleotide sequences included in \textbf{nt} database has a
-taxonomic assignment provided by the \textbf{Genbank} sequence
-submitter. NCBI provides a table (available at
-ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
-Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
-based on a crowdsourced submitter-maintained taxonomic annotation system
-for reference sequences. It supposes a sustainable system able to face
-the expected number of reference sequences that will populate the public
-global nucleotide databases in the near future. Another advantageous
-point is that we are based on NCBI taxonomy, the \emph{de-facto}
-standard taxonomic classification for biomolecular data
-\citep{cochrane20102010}. NCBI taxonomy is, undoubtedly, the most used
-taxonomy all over the world and the most similar to the official
-taxonomies of each specific field. This is a crucial point because all
-the type-culture and tissue databanks follow this official taxonomical
-classification and, in addition, all the knowledge accumulated during
-last decades is referred to this taxonomy. In addition NCBI provides a
-direct connection between taxonomical formal names and the physical
-specimens that serve as exemplars for the species
-\citep{federhen2014type}.
+After this first step, we drop redundant assignments: if sequences
+\(S_1 \subseteq S_2\) share an assignment \(T\), it gets dropped from
+\(S_1\); sequences which as a result end up having no assignments are
+removed.
 
-Certainly, if metagenomics results are easily integrated with the
-theoretical and experimental knowledge of each specific area, the impact
-of metagenomics will be higher than if it progresses as a disconnected
-research branch. Considering that metagenomics data interoperability,
-which is especially critical in clinical environments, requires a stable
-taxonomy to be used as reference, we decided to rely on the most widely
-used taxonomy: the NCBI taxonomy. In addition, the biggest global
-sequence database GenBank follows this taxonomy to register the origin
-of all their submitted sequences. Our 16S database building strategy
-allows the substitution of the 16S database by any other subset of
-\textbf{nt}, even by the complete \textbf{nt} database if it would be
-needed, for example, for analyzing shotgun metagenomics data. This
-possibility of changing the reference database provides flexibility to
-the system enabling it for easy updating and project-driven
-personalization.
-\end{quote}
+Once we have non-redundant set of RNA sequences containing all those
+corresponding to 16S, we can apply our procedure for checking internal
+consistency of taxonomic assignments.
 

### 2016-07-25 docs/mg7-paper/out/paper.tex
add a note for what to add in the assignment correction phase

@@
   are annotated as being of \texttt{rRNA} type\footnote{We are aware of
-    the existence of a gene annotation corresponding to 16S, that we are
-    \textbf{not using} due to a significant amount of 16S sequences
-    lacking the corresponding annotation}
+    the existence of a gene annotation corresponding to 16S in
+    RNACentra, that we are \textbf{not using} due to a significant
+    amount of 16S sequences lacking it}
 \item
@@
 consistency of taxonomic assignments.
+
+We run MG7 using this set of sequences as query, and with reference all
+but the one we are trying to assign. Explain
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  which kind of misassignments can be detected
+\item
+  how
+\item
+  why it is OK for sequences assigned to strange taxa, with no close 16S
+  seq (they will get no hits)
+\item
+  we analyze every assignment independently
+\end{enumerate}
 

### 2016-07-25 src/main/scala/mg7/bio4j/taxonomyTree.scala
Addressed some comments from @eparejatobes (WIP)

@@ -44,8 +44,8 @@ case object taxonomyTree {
       case "species subgroup" => 17
       case "species"          => 18
       case "subspecies"       => 19
-      // "no rank"
-      // TODO why this?? +1??
+      // "no rank" case
+      // NOTE: if a node has no rank, we take the rank of its parent and add 1 as it should be more specific
       case _ => this.parent.map(_.rankNumber).getOrElse(0) + 1
     }
 

### 2016-07-25 src/main/scala/mg7/dataflows/full.scala

@@ -16,7 +16,6 @@ import ohnosequences.awstools.autoscaling._
 import ohnosequences.awstools.regions.Region._
 import com.amazonaws.auth._, profile._
 
-// TODO this description is outdated right? assignment is done per chunk IIRC
 /* ## Standard Dataflow
 
   Standard dataflow, which will perform the following steps:
@@ -24,8 +23,8 @@ import com.amazonaws.auth._, profile._
   1. **flash** merge paired-end reads
   2. **split** split each reads dataset into smaller chunks
   3. **blast** blast each chunk of reads against the reference database
-  4. **merge** merge blast results for each reads dataset
-  5. **assign** assign reads to taxa (LCA and BBH)
+  4. **assign** assign reads from each chunk to taxa (LCA and BBH)
+  5. **merge** merge results for each reads dataset
   6. **count** count assignments
 */
 trait AnyFullDataflow extends AnyNoFlashDataflow {
@@ -92,7 +91,7 @@ trait AnyFullDataflow extends AnyNoFlashDataflow {
     )
 }
 
-case class FullDataflow[P <: AnyMG7Parameters](val params: P)(val flashInputs: Map[SampleID, (S3Resource, S3Resource)]) 
+case class FullDataflow[P <: AnyMG7Parameters](val params: P)(val flashInputs: Map[SampleID, (S3Resource, S3Resource)])
 extends AnyFullDataflow {
 
   type Params = P

### 2016-07-25 src/main/scala/mg7/dataflows/noFlash.scala

@@ -16,16 +16,9 @@ import ohnosequences.awstools.autoscaling._
 import ohnosequences.awstools.regions.Region._
 import com.amazonaws.auth._, profile._
 
-// TODO this description is outdated right? assignment is done per chunk IIRC
 /* ## No-flash Dataflow
 
-  No-flash dataflow does the same as the full one *but* the first flash-based merging step:
-
-  1. **split** split each reads dataset into smaller chunks
-  2. **blast** blast each chunk of reads against the reference database
-  3. **merge** merge blast results for each reads dataset
-  4. **assign** assign reads to taxa (LCA and BBH)
-  5. **count** count assignments
+  No-flash dataflow does the same as the full one *but* the first flash-based merging step. See AnyFullDataflow.
 */
 trait AnyNoFlashDataflow extends AnyDataflow {
 
@@ -43,7 +36,7 @@ trait AnyNoFlashDataflow extends AnyDataflow {
       )
     }
 
-  // TODO what's this supposed to be doing?
+  // Here we generate tasks/data mappings for the blast loquat one per each S3 object generated by the split loquat
   lazy val blastDataMappings =
     splitDataMappings.flatMap { splitDM =>
       val sampleId = splitDM.label

### 2016-07-25 src/main/scala/mg7/loquats/1.flash.scala

@@ -1,6 +1,5 @@
 package ohnosequences.mg7.loquats
 
-// TODO in general I prefer imports to occupy less space
 import ohnosequences.mg7._
 import ohnosequences.loquat._
 import ohnosequences.statika._
@@ -9,12 +8,10 @@ import ohnosequences.cosas._, types._, klists._
 import ohnosequences.datasets._
 import better.files._
 
-// TODO move all these "tool" bundles to a separate file
-case object flashBundle extends ohnosequencesBundles.statika.Flash("1.2.11")
 
 case class flashDataProcessing[MD <: AnyMG7Parameters](val md: MD)
 extends DataProcessingBundle(
-  flashBundle
+  bundles.flash
 )(
   input   = data.flashInput,
   output  = data.flashOutput

### 2016-07-25 src/main/scala/mg7/loquats/3.blast.scala

@@ -12,16 +12,14 @@ import better.files._
 import sys.process._
 
 
-case object blastBundle extends ohnosequencesBundles.statika.Blast("2.2.31")
 
 case class blastDataProcessing[MD <: AnyMG7Parameters](val md: MD)
-extends DataProcessingBundle()(
+extends DataProcessingBundle(
+  (bundles.blast +: md.referenceDBs.toSeq): _*
+)(
   input  = data.blastInput,
   output = data.blastOutput
 ) {
-
-  override val bundleDependencies: List[AnyBundle] =
-    blastBundle :: md.referenceDBs.toList
 
   def instructions: AnyInstructions = say("Let the blasting begin!")
 

### 2016-07-25 src/main/scala/mg7/loquats/4.assign.scala

@@ -12,18 +12,14 @@ import scala.util.Try
 import com.github.tototoshi.csv._
 
 
-case class assignDataProcessing[MD <: AnyMG7Parameters](val md: MD) extends DataProcessingBundle()(
+case class assignDataProcessing[MD <: AnyMG7Parameters](val md: MD) extends DataProcessingBundle(
+  (bio4j.taxonomyBundle :: md.referenceDBs.toList): _*
+)(
   input  = data.assignInput,
   output = data.assignOutput
-)
-{
-
+) {
   // For the output fields implicits
   import md._
-
-  // TODO why override val here? why this is not a param?
-  override val bundleDependencies: List[AnyBundle] =
-    bio4j.taxonomyBundle :: md.referenceDBs.toList
 
   private lazy val taxonomyGraph: TitanNCBITaxonomyGraph = bio4j.taxonomyBundle.graph
 

### 2016-07-25 src/main/scala/mg7/loquats/5.merge.scala

@@ -18,11 +18,9 @@ case object mergeDataProcessing extends DataProcessingBundle()(
   // TODO: use streams, file-writers, etc. stuff
   // TODO no default arguments please
   def mergeChunks(dir: File, out: File, header: Option[String] = None): Unit = {
-    // TODO foreach on option :|
     header.foreach { out.appendLine }
     // only one level in depth:
-    // TODO as in other places, use {} here
-    dir.list foreach { chunkFile =>
+    dir.list.foreach { chunkFile =>
       out.append( chunkFile.contentAsString )
       chunkFile.delete()
     }

### 2016-07-25 src/main/scala/mg7/loquats/6.count.scala

@@ -102,7 +102,6 @@ case object countDataProcessing extends DataProcessingBundle(
       val assignedReadsNumber: Double = taxIDs.length
       def frequency(absolute: Int): Double = absolute / assignedReadsNumber
 
-      // TODO why mutable Map??
       val lineageMap: scala.collection.mutable.Map[TaxID, Seq[TaxID]] = scala.collection.mutable.Map()
 
       def getLineage(id: TaxID): Seq[TaxID] =

### 2016-07-25 src/main/scala/mg7/loquats/7.stats.scala

@@ -21,11 +21,7 @@ case object statsDataProcessing extends DataProcessingBundle()(
     parser: Iterator[String] => Iterator[Any],
     file: File
   ): Integer = {
-    // TODO why io.Source here??
-    val source = io.Source.fromFile( file.toJava )
-    val readsNumber = parser( source.getLines ).length
-    source.close()
-    readsNumber
+    parser( file.lines ).length
   }
 
   def instructions: ohnosequences.statika.AnyInstructions = say("Running stats loquat")

### 2016-07-25 src/main/scala/mg7/loquats/8.summary.scala

@@ -17,11 +17,7 @@ case object summaryDataProcessing extends DataProcessingBundle()(
 {
 
   def countReads(file: File): Integer = {
-    // TODO why io.Source??
-    val source = io.Source.fromFile( file.toJava )
-    val readsNumber = fasta.parseMap( source.getLines ).length
-    source.close()
-    readsNumber
+    fasta.parseMap( file.lines ).length
   }
 
   def countLines(file: File): Integer = { file.lines.length }

### 2016-07-25 src/main/scala/mg7/package.scala

@@ -21,7 +21,7 @@ package object mg7 {
   def parseLong(str: String): Option[Long] = util.Try(str.toLong).toOption
   def parseDouble(str: String): Option[Double] = util.Try(str.toDouble).toOption
 
-  // TODO why this?
+  // Used in dataMappings to get value pairs from another maps
   def lookup[A, B](a: A, m: Map[A, B]): (A, B) = a -> m.apply(a)
 
   // TODO why not create an ordering from `f` and use std library?
@@ -32,7 +32,7 @@ package object mg7 {
       // either acc is empty or t is the new maximum
       case (_, t) => List(t)
     }
-  // TODO why not as ops?
+
   def averageOf(vals: Seq[Double]): Double = vals.sum / vals.length
 
   type BlastArgumentsVals =

### 2016-07-25 src/main/scala/mg7/parameters.scala

@@ -97,10 +97,9 @@ abstract class MG7Parameters[
   val readsLength: illumina.Length,
   val splitInputFormat: SplitInputFormat = FastQInput,
   val splitChunkSize: Int = 10,
-  val blastCommand: BC = blastn,
-  // TODO why default values? not good
-  val blastOutRec: BlastOutputRecord[BK]  = defaults.blastnOutputRecord,
-  val blastOptions: BC#OptionsVals        = defaults.blastnOptions.value,
+  val blastCommand: BC,
+  val blastOutRec: BlastOutputRecord[BK],
+  val blastOptions: BC#OptionsVals,
   val referenceDBs: Set[AnyReferenceDB]
 )(implicit
   val argValsToSeq: BlastOptionsToSeq[BC#ArgumentsVals],

### 2016-07-26 docs/mg7-paper/out/paper.tex
note about why this misassignment correction strategy, reusable for other data
start describing the mg7 database self-correction step

@@
 \(S_1 \subseteq S_2\) share an assignment \(T\), it gets dropped from
-\(S_1\); sequences which as a result end up having no assignments are
-removed.
+\(S_1\); sequences which as a result of this process end up having no
+assignments are removed.
 
@@
 corresponding to 16S, we can apply our procedure for checking internal
-consistency of taxonomic assignments.
-
-We run MG7 using this set of sequences as query, and with reference all
-but the one we are trying to assign. Explain
+consistency of taxonomic assignments. We run MG7 using this set of
+sequences as query, with reference all but the one we are trying to
+assign. Under the (reasonable) assumption of 16S sequence similarity
+being correlated with the taxonomy tree topology, the MG7 result should
+either
 
@@
 \item
-  which kind of misassignments can be detected
+  be a taxon \emph{close} to the original assignment, if there are
+  sequences similar enough in the reference database
 \item
-  how
-\item
-  why it is OK for sequences assigned to strange taxa, with no close 16S
-  seq (they will get no hits)
-\item
-  we analyze every assignment independently
+  be empty, due to this sequence having no similar sequences in the
+  reference database
 \end{enumerate}
+
+In any other case, under our assumptions, this assignment should be
+discarded.
+
+All these steps are automated, and integrated with every release of our
+reference database. The whole process can be repeated and adapted to
+other subsets of RNACentral; we do this for a 18S reference database.
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
add the definition of close, a note for a figure

@@
 In any other case, under our assumptions, this assignment should be
-discarded.
+discarded. What we consider as close is: the parent of the MG7
+assignment should be contained in the lineage of the original
+assignment.
 
@@
 reference database. The whole process can be repeated and adapted to
-other subsets of RNACentral; we do this for a 18S reference database.
+other subsets of RNACentral; we do this for an 18S reference database.
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
note for LCA abbrv

--- 2016-07-25 docs/mg7-paper/out/paper.tex
@@
   Static reproducible specification of dependencies and behavior of the
-  different components using
-  \emph{\href{https://github.com/ohnosequences/statika}{Statika}} and
-  \emph{\href{https://github.com/ohnosequences/datasets}{Datasets}}
+  different components using \emph{\protect\hyperlink{statika}{Statika}}
+  and \emph{\protect\hyperlink{datasets}{Datasets}}
 \item
@@
   Definition of complex workflows using
-  \emph{\href{https://github.com/ohnosequences/loquat}{Loquat}}, a
-  composable system for scaling/parallelizing stateless computations
-  especially designed for AWS
+  \emph{\protect\hyperlink{loquat}{Loquat}}, a composable system for
+  scaling/parallelizing stateless computations especially designed for
+  AWS
 \item
@@
   Exhaustive per-read taxonomic assignment using two complementary
-  assignment algorithms Lowest Common Ancestor and Best BLAST Hit
+  assignment algorithms: Lowest Common Ancestor (LCA) and Best BLAST Hit
 \item
@@
 
-\href{https://github.com/ohnosequences/statika}{Statika} is a Scala
-library developed by \emph{AA} and \emph{EPT} which serves as a way of
-defining and composing machine behaviors statically. The main component
-are \textbf{bundles}. Each bundle declares a sequence of computations
-(its behavior) which will be executed in an \textbf{environment}. A
-bundle can \emph{depend} on other bundles, and when being executed by an
-environment, its DAG (Directed Acyclic Graph) of dependencies is
-linearized and run in sequence. In our use, bundles correspond to what
-an EC2 instance should do and an environment to an AMI (Amazon Machine
-Image) which prepares the basic configuration, downloads the Scala code
-and runs it.
+\protect\hyperlink{statika}{Statika} is a Scala library developed by
+\emph{AA} and \emph{EPT} which serves as a way of defining and composing
+machine behaviors statically. The main component are \textbf{bundles}.
+Each bundle declares a sequence of computations (its behavior) which
+will be executed in an \textbf{environment}. A bundle can \emph{depend}
+on other bundles, and when being executed by an environment, its DAG
+(Directed Acyclic Graph) of dependencies is linearized and run in
+sequence. In our use, bundles correspond to what an EC2 instance should
+do and an environment to an AMI (Amazon Machine Image) which prepares
+the basic configuration, downloads the Scala code and runs it.
 
@@
 
-\href{https://github.com/ohnosequences/datasets}{Datasets} is a Scala
-library developed by \emph{AA} and \emph{EPT} with the goal of being a
-Scala-embedded mini-language for datasets and their locations.
-\textbf{Data} is represented as type-indexed fields: keys are modeled as
-singleton types, and values correspond to what could be called a
-denotation of the key: a value of type \texttt{Location} tagged with the
-key type. Then a \textbf{Dataset} is essentially a collection of data,
-which are guaranteed statically to be different through type-level
-predicates, making use of the value--type correspondence which can be
-established through singleton types and implicits. A dataset location is
-then just a list of locations formed by locations of each dataset key.
-All this is based on what could be described as an embedding in Scala of
-an extensible record system with concatenation on disjoint labels, in
-the spirit of \citep{harper1990extensible} \citep{harper1991record}. For
-that \emph{Datasets} uses the
-\href{https://github.com/ohnosequences/cosas/}{ohnosequences/cosas}
+\protect\hyperlink{datasets}{Datasets} is a Scala library developed by
+\emph{AA} and \emph{EPT} with the goal of being a Scala-embedded
+mini-language for datasets and their locations. \textbf{Data} is
+represented as type-indexed fields: keys are modeled as singleton types,
+and values correspond to what could be called a denotation of the key: a
+value of type \texttt{Location} tagged with the key type. Then a
+\textbf{Dataset} is essentially a collection of data, which are
+guaranteed statically to be different through type-level predicates,
+making use of the value--type correspondence which can be established
+through singleton types and implicits. A dataset location is then just a
+list of locations formed by locations of each dataset key. All this is
+based on what could be described as an embedding in Scala of an
+extensible record system with concatenation on disjoint labels, in the
+spirit of \citep{harper1990extensible} \citep{harper1991record}. For
+that \emph{Datasets} uses the {[}ohnosequences/cosas{]}{[}cosas{]}
 library.
@@
 
-\href{https://github.com/ohnosequences/loquat}{Loquat} is a library
-developed by \emph{AA}, \emph{EK} and \emph{EPT} designed for the
-execution of embarrassingly parallel tasks using S3, SQS and EC2 Amazon
-services.
+\protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
+\emph{EK} and \emph{EPT} designed for the execution of embarrassingly
+parallel tasks using S3, SQS and EC2 Amazon services.
 
@@
 We developed our own Scala-based type-safe eDSLs (embedded Domain
-Specific Languages) for
-\href{https://github.com/ohnosequences/flash-api}{FLASh}
-\citep{magovc2011flash} and
-\href{https://github.com/ohnosequences/blast-api}{BLAST}
-\citep{camacho2009blast} expressions and their execution.
+Specific Languages) for {[}FLASh{]}{[}flash{]} \citep{magovc2011flash}
+and {[}BLAST{]}{[}blast{]} \citep{camacho2009blast} expressions and
+their execution.
 
@@
 
-The starting point for our 16S reference database is RNA Central
-\citep{rnacentral2014rnacentral},
-\href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
-5}. RNACentral was chosen being the most up to date, comprehensive RNA
-sequence repository, including among others all RNA data from Silva,
-GreenGenes, RDP, ENA, and RefSeq. First we take those sequences which
+Our 16S-DB7 Reference Database is a curated subset of sequences from the
+NCBI nucleotide database \textbf{nt}. The sequences included were
+selected by similarity with the bacterial and archaeal reference
+sequences downloaded from the \textbf{RDP database}
+\citep{cole2013ribosomal}. RDP unaligned sequences were used to capture
+new 16S RNA sequences from \textbf{nt} using BLAST similarity search
+strategies and then, performing additional curation steps to remove
+sequences with poor taxonomic assignments to taxonomic nodes close to
+the root of the taxonomy tree. All the nucleotide sequences included in
+\textbf{nt} database has a taxonomic assignment provided by the
+\textbf{Genbank} sequence submitter. NCBI provides a table (available at
+ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
+Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
+based on a crowdsourced submitter-maintained taxonomic annotation system
+for reference sequences. It supposes a sustainable system able to face
+the expected number of reference sequences that will populate the public
+global nucleotide databases in the near future. Another advantageous
+point is that we are based on NCBI taxonomy, the \emph{de-facto}
+standard taxonomic classification for biomolecular data
+\citep{cochrane20102010}. NCBI taxonomy is, undoubtedly, the most used
+taxonomy all over the world and the most similar to the official
+taxonomies of each specific field. This is a crucial point because all
+the type-culture and tissue databanks follow this official taxonomical
+classification and, in addition, all the knowledge accumulated during
+last decades is referred to this taxonomy. In addition NCBI provides a
+direct connection between taxonomical formal names and the physical
+specimens that serve as exemplars for the species
+\citep{federhen2014type}.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  are annotated as being of \texttt{rRNA} type\footnote{We are aware of
-    the existence of a gene annotation corresponding to 16S in
-    RNACentra, that we are \textbf{not using} due to a significant
-    amount of 16S sequences lacking it}
-\item
-  their length is at least \texttt{1300}\footnote{Note that 16S
-    sequences are sometimes part of an entry corresponding to whole
-    small subunits; that's why we do not set a maximum length threshold}
-\item
-  have at least one taxonomic assignment to a descendant of
-  \emph{Bacteria} or \emph{Archaea}
-\item
-  their lineage does not contain a set of taxa deemed
-  uninformative\footnote{for example: ``unclassified Bacteria
-    (miscellaneous)'' or ``unclassified''. Assigning a read to an
-    ``unclassified'' taxon defeats the first and foremost goal of
-    taxonomic profiling: \emph{classification}.}
-\end{enumerate}
-
-After this first step, we drop redundant assignments: if sequences
-\(S_1 \subseteq S_2\) share an assignment \(T\), it gets dropped from
-\(S_1\); sequences which as a result end up having no assignments are
-removed.
-
-Once we have non-redundant set of RNA sequences containing all those
-corresponding to 16S, we can apply our procedure for checking internal
-consistency of taxonomic assignments.
-
-We run MG7 using this set of sequences as query, and with reference all
-but the one we are trying to assign. Explain
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  which kind of misassignments can be detected
-\item
-  how
-\item
-  why it is OK for sequences assigned to strange taxa, with no close 16S
-  seq (they will get no hits)
-\item
-  we analyze every assignment independently
-\end{enumerate}
+Certainly, if metagenomics results are easily integrated with the
+theoretical and experimental knowledge of each specific area, the impact
+of metagenomics will be higher than if it progresses as a disconnected
+research branch. Considering that metagenomics data interoperability,
+which is especially critical in clinical environments, requires a stable
+taxonomy to be used as reference, we decided to rely on the most widely
+used taxonomy: the NCBI taxonomy. In addition, the biggest global
+sequence database GenBank follows this taxonomy to register the origin
+of all their submitted sequences. Our 16S database building strategy
+allows the substitution of the 16S database by any other subset of
+\textbf{nt}, even by the complete \textbf{nt} database if it would be
+needed, for example, for analyzing shotgun metagenomics data. This
+possibility of changing the reference database provides flexibility to
+the system enabling it for easy updating and project-driven
+personalization.
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
note about fixing LCA figure

@@
 BLAST Hit based taxonomic assignment (BBH). Figure 2 displays
-schematically the LCA algorithm applied sensu stricto (left panel) and
-the called `in line' exception (right panel) designed in order to gain
-specificity in the assignments in the cases in which the topology of the
-taxonomical nodes corresponding to the BLAST hits support sufficiently
-the assignment to the most specific taxon.
+schematically the LCA algorithm.
 
@@
 \caption{\textbf{Lowest Common Ancestor algorithm for taxonomic
-assignment.} The Left panel displays an example of the application of
-LCA algorithm in a \emph{sensu stricto} mode. A, B, C and D represent
-taxonomy tree nodes with assigned reads. Right panel displays the
-\emph{in line} mode of assignment which is an exception for the
-\emph{sensu stricto} mode of application of LCA algorithm. The \emph{in
-line} mode is used when all the nodes are located in a line without
-bifurcations. In that case the taxon assigned is the most specific (the
-most distant from the root).}
+assignment.} \textbf{TODO UPDATE THIS FIGURE!} The Left panel displays
+an example of the application of LCA algorithm in a \emph{sensu stricto}
+mode. A, B, C and D represent taxonomy tree nodes with assigned reads.
+Right panel displays the \emph{in line} mode of assignment which is an
+exception for the \emph{sensu stricto} mode of application of LCA
+algorithm. The \emph{in line} mode is used when all the nodes are
+located in a line without bifurcations. In that case the taxon assigned
+is the most specific (the most distant from the root).}
 \end{figure}
 
-\paragraph{Lowest Common Ancestor based Taxonomic
-Assignment}\label{lowest-common-ancestor-based-taxonomic-assignment}
+\paragraph{LCA Assignment}\label{lca-assignment}
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
explain the notion of best hits for a query

@@
 
+All the assignment procedures work by default with input the set of
+BLAST hits which have maximum bitscore among those covering the whole
+query sequence; we will call this the set of best hits of a query. Note
+that the criteria which determines this set of best hits is fully
+configurable, adjusted to conform to its meaning in each particular
+dataset: those hits which we should take into account as valid for
+assignment. This is of course experiment-dependent, and we can account
+for read quality or technology-specific error profiles through
+configuring these criteria for the set of best hits.
+
 \paragraph{LCA Assignment}\label{lca-assignment}

### 2016-07-26 docs/mg7-paper/out/paper.tex
remove older LCA text, explain why LCA a bit

@@
 
-For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
-over a threshold of similarity. To evaluate similarity for this first
-filtering of hits we use the Expect value (by default
-\(evalue \leq e^{-15}\)) that describes the number of hits one can
-``expect'' to see by chance when searching a database of a particular
-size. In a second filtering step we filtering those hits that are not
-sufficiently good comparing them with the best one. We select the best
-HSP (High Similarity Pair) per reference sequence and then choose the
-best HSP (that with lowest E-value) between all the selected ones. The
-bitscore of this best HSP (called S) is used as reference to filter the
-rest of HSPs. All the HSPs with bitscore below the product \(p S\) are
-filtered. p is a coefficient fixed by the user to define the bitscore
-required, e.g.~if \(p=0.9\) and \(S=700\) the required bitscore
-threshold would be \(630\). Once we have the definitive HSPs selected,
-we obtain their corresponding taxonomic nodes using the taxonomic
-assignments that NCBI provides for all the nt database sequences. Now we
-have to analyze the topological distribution of these nodes in the
-taxonomy tree: i. If all the nodes forms a line in the taxonomy tree
-(are located in a not branched lineage to the tree root) we should
-choose the most specific taxID as the final assignment for that read. We
-call to this kind of assignment the `in line' exception (see Figure 2
-right panel). ii. If not, we should search for the \emph{sensu stricto}
-Lowest Common Ancestor (LCA) of all the selected taxonomic nodes (See
-Figure 2 left panel). In this approach we decided to use the bitscore
-for evaluating the similarity because it is a value that increases when
-similarity is higher and depends a lot on the length of the HSP. Some
-reads could not find sequences with enough similarity in the database
-and then they would be classified as reads with no hits. Advanced
-metagenomics analysis approaches \citep{huson2012microbial} have adopted
-LCA-based assignment algorithms because it provides fine and trusted
-taxonomical assignment.
+The LCA assignment simply computes the lowest common ancestor of the set
+of best hits. Other metagenomics analysis approaches
+\citep{huson2012microbial} have adopted LCA-based assignment algorithms,
+as it provides the most precise assignment for a set of taxa, with no
+further information available\footnote{the lowest common ancestor of a
+  set of nodes in a tree is just their product in the tree viewed as a
+  poset, with the ``parent of'' partial order}; as our reference
+database is already heavily preprocessed and checked for internal
+consistency, there is no need for altering the lowest common ancestor in
+any way, as is customarily done in other cases.
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
reword best hits to valid hits

@@
 BLAST hits which have maximum bitscore among those covering the whole
-query sequence; we will call this the set of best hits of a query. Note
-that the criteria which determines this set of best hits is fully
-configurable, adjusted to conform to its meaning in each particular
-dataset: those hits which we should take into account as valid for
-assignment. This is of course experiment-dependent, and we can account
-for read quality or technology-specific error profiles through
-configuring these criteria for the set of best hits.
+query sequence; we will call this the set of valid hits of a query. Note
+that the criteria which determines this set of valid hits is fully
+configurable, and should be adjusted to conform to its meaning in each
+particular dataset: those hits which we should take into account as
+valid for assignment. This is of course experiment-dependent, and the
+criteria for valid hits should account for read quality or
+technology-specific error profiles.
 
@@
 The LCA assignment simply computes the lowest common ancestor of the set
-of best hits. Other metagenomics analysis approaches
+of valid hits. Other metagenomics analysis approaches
 \citep{huson2012microbial} have adopted LCA-based assignment algorithms,
@@
 consistency, there is no need for altering the lowest common ancestor in
-any way, as is customarily done in other cases.
+any way, as is customarily done in other cases. MG7 operates under the
+assumption that hits which fulfill the user-defined criteria for being
+valid hits should be treated on an equal footing, and in that case
+taking their LCA is the coherent assignment choice with minimum
+information loss.
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
no more select the best 10 hits
fix named links

@@
 strategy. Every read is specifically compared with all the sequences of
-the 16S database. We select the best BLAST hits (10 hits by default)
-obtained for each read to do the taxonomic assignment.
+the 16S database.
 

### 2016-07-26 docs/mg7-paper/out/paper.tex
fix named links in paper

@@
   Exhaustive per-read taxonomic assignment using two complementary
-  assignment algorithms: Lowest Common Ancestor (LCA) and Best BLAST Hit
+  assignment algorithms Lowest Common Ancestor and Best BLAST Hit
 \item
@@
 strategy. Every read is specifically compared with all the sequences of
-the 16S database.
+the 16S database. We select the best BLAST hits (10 hits by default)
+obtained for each read to do the taxonomic assignment.
 
@@
 BLAST Hit based taxonomic assignment (BBH). Figure 2 displays
-schematically the LCA algorithm.
+schematically the LCA algorithm applied sensu stricto (left panel) and
+the called `in line' exception (right panel) designed in order to gain
+specificity in the assignments in the cases in which the topology of the
+taxonomical nodes corresponding to the BLAST hits support sufficiently
+the assignment to the most specific taxon.
 
@@
 \caption{\textbf{Lowest Common Ancestor algorithm for taxonomic
-assignment.} \textbf{TODO UPDATE THIS FIGURE!} The Left panel displays
-an example of the application of LCA algorithm in a \emph{sensu stricto}
-mode. A, B, C and D represent taxonomy tree nodes with assigned reads.
-Right panel displays the \emph{in line} mode of assignment which is an
-exception for the \emph{sensu stricto} mode of application of LCA
-algorithm. The \emph{in line} mode is used when all the nodes are
-located in a line without bifurcations. In that case the taxon assigned
-is the most specific (the most distant from the root).}
+assignment.} The Left panel displays an example of the application of
+LCA algorithm in a \emph{sensu stricto} mode. A, B, C and D represent
+taxonomy tree nodes with assigned reads. Right panel displays the
+\emph{in line} mode of assignment which is an exception for the
+\emph{sensu stricto} mode of application of LCA algorithm. The \emph{in
+line} mode is used when all the nodes are located in a line without
+bifurcations. In that case the taxon assigned is the most specific (the
+most distant from the root).}
 \end{figure}
 
-All the assignment procedures work by default with input the set of
-BLAST hits which have maximum bitscore among those covering the whole
-query sequence; we will call this the set of valid hits of a query. Note
-that the criteria which determines this set of valid hits is fully
-configurable, and should be adjusted to conform to its meaning in each
-particular dataset: those hits which we should take into account as
-valid for assignment. This is of course experiment-dependent, and the
-criteria for valid hits should account for read quality or
-technology-specific error profiles.
+\paragraph{Lowest Common Ancestor based Taxonomic
+Assignment}\label{lowest-common-ancestor-based-taxonomic-assignment}
 
-\paragraph{LCA Assignment}\label{lca-assignment}
-
-The LCA assignment simply computes the lowest common ancestor of the set
-of valid hits. Other metagenomics analysis approaches
-\citep{huson2012microbial} have adopted LCA-based assignment algorithms,
-as it provides the most precise assignment for a set of taxa, with no
-further information available\footnote{the lowest common ancestor of a
-  set of nodes in a tree is just their product in the tree viewed as a
-  poset, with the ``parent of'' partial order}; as our reference
-database is already heavily preprocessed and checked for internal
-consistency, there is no need for altering the lowest common ancestor in
-any way, as is customarily done in other cases. MG7 operates under the
-assumption that hits which fulfill the user-defined criteria for being
-valid hits should be treated on an equal footing, and in that case
-taking their LCA is the coherent assignment choice with minimum
-information loss.
+For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
+over a threshold of similarity. To evaluate similarity for this first
+filtering of hits we use the Expect value (by default
+\(evalue \leq e^{-15}\)) that describes the number of hits one can
+``expect'' to see by chance when searching a database of a particular
+size. In a second filtering step we filtering those hits that are not
+sufficiently good comparing them with the best one. We select the best
+HSP (High Similarity Pair) per reference sequence and then choose the
+best HSP (that with lowest E-value) between all the selected ones. The
+bitscore of this best HSP (called S) is used as reference to filter the
+rest of HSPs. All the HSPs with bitscore below the product \(p S\) are
+filtered. p is a coefficient fixed by the user to define the bitscore
+required, e.g.~if \(p=0.9\) and \(S=700\) the required bitscore
+threshold would be \(630\). Once we have the definitive HSPs selected,
+we obtain their corresponding taxonomic nodes using the taxonomic
+assignments that NCBI provides for all the nt database sequences. Now we
+have to analyze the topological distribution of these nodes in the
+taxonomy tree: i. If all the nodes forms a line in the taxonomy tree
+(are located in a not branched lineage to the tree root) we should
+choose the most specific taxID as the final assignment for that read. We
+call to this kind of assignment the `in line' exception (see Figure 2
+right panel). ii. If not, we should search for the \emph{sensu stricto}
+Lowest Common Ancestor (LCA) of all the selected taxonomic nodes (See
+Figure 2 left panel). In this approach we decided to use the bitscore
+for evaluating the similarity because it is a value that increases when
+similarity is higher and depends a lot on the length of the HSP. Some
+reads could not find sequences with enough similarity in the database
+and then they would be classified as reads with no hits. Advanced
+metagenomics analysis approaches \citep{huson2012microbial} have adopted
+LCA-based assignment algorithms because it provides fine and trusted
+taxonomical assignment.
 

### 2016-08-17 docs/mg7-paper/out/paper.tex
address @rtobes review in #88

@@
 
-Our 16S-DB7 Reference Database is a curated subset of sequences from the
-NCBI nucleotide database \textbf{nt}. The sequences included were
-selected by similarity with the bacterial and archaeal reference
-sequences downloaded from the \textbf{RDP database}
-\citep{cole2013ribosomal}. RDP unaligned sequences were used to capture
-new 16S RNA sequences from \textbf{nt} using BLAST similarity search
-strategies and then, performing additional curation steps to remove
-sequences with poor taxonomic assignments to taxonomic nodes close to
-the root of the taxonomy tree. All the nucleotide sequences included in
-\textbf{nt} database has a taxonomic assignment provided by the
-\textbf{Genbank} sequence submitter. NCBI provides a table (available at
-ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
-Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
-based on a crowdsourced submitter-maintained taxonomic annotation system
-for reference sequences. It supposes a sustainable system able to face
-the expected number of reference sequences that will populate the public
-global nucleotide databases in the near future. Another advantageous
-point is that we are based on NCBI taxonomy, the \emph{de-facto}
-standard taxonomic classification for biomolecular data
-\citep{cochrane20102010}. NCBI taxonomy is, undoubtedly, the most used
-taxonomy all over the world and the most similar to the official
-taxonomies of each specific field. This is a crucial point because all
-the type-culture and tissue databanks follow this official taxonomical
-classification and, in addition, all the knowledge accumulated during
-last decades is referred to this taxonomy. In addition NCBI provides a
-direct connection between taxonomical formal names and the physical
-specimens that serve as exemplars for the species
-\citep{federhen2014type}.
+The starting point for our 16S reference database is RNA Central
+\citep{rnacentral2014rnacentral},
+\href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
+5}. RNACentral was chosen being the most up to date, comprehensive RNA
+sequence repository, including among others all RNA data from Silva,
+GreenGenes, RDP, ENA, and RefSeq. First we take those sequences which
 
-Certainly, if metagenomics results are easily integrated with the
-theoretical and experimental knowledge of each specific area, the impact
-of metagenomics will be higher than if it progresses as a disconnected
-research branch. Considering that metagenomics data interoperability,
-which is especially critical in clinical environments, requires a stable
-taxonomy to be used as reference, we decided to rely on the most widely
-used taxonomy: the NCBI taxonomy. In addition, the biggest global
-sequence database GenBank follows this taxonomy to register the origin
-of all their submitted sequences. Our 16S database building strategy
-allows the substitution of the 16S database by any other subset of
-\textbf{nt}, even by the complete \textbf{nt} database if it would be
-needed, for example, for analyzing shotgun metagenomics data. This
-possibility of changing the reference database provides flexibility to
-the system enabling it for easy updating and project-driven
-personalization.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  are annotated as being of \texttt{rRNA} type\footnote{We are aware of
+    the existence of a gene annotation corresponding to 16S in
+    RNACentra, that we are \textbf{not using} due to a significant
+    amount of 16S sequences lacking it}
+\item
+  their length is at least \texttt{1300}\footnote{Note that 16S
+    sequences are sometimes part of an entry corresponding to whole
+    small subunits; that's why we do not set a maximum length threshold}
+\item
+  have at least one taxonomic assignment to a descendant of
+  \emph{Bacteria} or \emph{Archaea}
+\item
+  their lineage does not contain a set of taxa deemed
+  uninformative\footnote{for example: ``unclassified Bacteria
+    (miscellaneous)'' or ``unclassified''. Assigning a read to an
+    ``unclassified'' taxon defeats the first and foremost goal of
+    taxonomic profiling: \emph{classification}.}
+\end{enumerate}
+
+After this first step, we drop redundant assignments: if sequences
+\(S_1 \subseteq S_2\) share an assignment \(T\), it gets dropped from
+\(S_1\); sequences which as a result of this process end up having no
+assignments are removed.
+
+Once we have a non-redundant set of ribosomal RNA sequences (containing
+all those corresponding to 16S), we can apply our procedure for checking
+internal consistency of taxonomic assignments. We run MG7 using this set
+of sequences as query, with reference all but the one we are trying to
+assign. Under the (reasonable) assumption of 16S sequence similarity
+being correlated with the taxonomy tree topology, the resulting MG7
+taxonomic assignment should either
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  be a taxon \emph{close} to the original assignment, if there are
+  sequences similar enough in the reference database; what we consider
+  as close is: the parent of the MG7 assignment should be contained in
+  the lineage of the original assignment.
+\item
+  be empty, due to this sequence having no similar sequences in the
+  reference database, thus making impossible to do any taxonomic
+  assignment.
+\end{enumerate}
+
+In any other case, under our assumptions, this assignment should be
+discarded.
+
+All these steps are automated, and integrated with every release of our
+reference database. The whole process can be repeated and adapted to
+other subsets of RNACentral; we do this for an 18S reference database.
 

### 2016-08-17 src/main/scala/mg7/dataflow.scala
Refactored code related to lca/bbh counts: data and counting loquat files juggling

@@ -24,6 +24,10 @@ trait AnyDataflow {
 
       val sampleId = mergeDM.label
 
+      def outputFor(d: FileData): (FileData, S3Resource) = {
+        d -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.${d.baseName}.csv")
+      }
+
       DataMapping(sampleId, countDataProcessing)(
 
         remoteInput = Map(
@@ -32,14 +36,14 @@ trait AnyDataflow {
         ),
 
         remoteOutput = Map(
-          data.lcaDirectCountsCSV     -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.lca.direct.absolute.counts.csv"),
-          data.lcaAccumCountsCSV      -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.lca.accum.absolute.counts.csv"),
-          data.lcaDirectFreqCountsCSV -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.lca.direct.frequency.counts.csv"),
-          data.lcaAccumFreqCountsCSV  -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.lca.accum.frequency.counts.csv"),
-          data.bbhDirectCountsCSV     -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.bbh.direct.absolute.counts.csv"),
-          data.bbhAccumCountsCSV      -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.bbh.accum.absolute.counts.csv"),
-          data.bbhDirectFreqCountsCSV -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.bbh.direct.frequency.counts.csv"),
-          data.bbhAccumFreqCountsCSV  -> S3Resource(params.outputS3Folder(sampleId, "count") / s"${sampleId}.bbh.accum.frequency.counts.csv")
+          outputFor(data.lca.direct.absolute),
+          outputFor(data.lca.accum.absolute),
+          outputFor(data.lca.direct.relative),
+          outputFor(data.lca.accum.relative),
+          outputFor(data.bbh.direct.absolute),
+          outputFor(data.bbh.accum.absolute),
+          outputFor(data.bbh.direct.relative),
+          outputFor(data.bbh.accum.relative)
         )
       )
     }

### 2016-08-18 docs/mg7-paper/out/paper.tex

@@
 \href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
-5}. RNACentral was chosen being the most up to date, comprehensive RNA
+5}. RNAcentral was chosen being the most up to date, comprehensive RNA
 sequence repository, including among others all RNA data from Silva,
@@
     the existence of a gene annotation corresponding to 16S in
-    RNACentra, that we are \textbf{not using} due to a significant
+    RNAcentral, that we are \textbf{not using} due to a significant
     amount of 16S sequences lacking it}
@@
 reference database. The whole process can be repeated and adapted to
-other subsets of RNACentral; we do this for an 18S reference database.
+other subsets of RNAcentral; we do this for an 18S reference database.
 

### 2016-08-18 docs/mg7-paper/out/paper.tex

@@
 \caption{\textbf{Lowest Common Ancestor algorithm for taxonomic
-assignment.} The Left panel displays an example of the application of
-LCA algorithm in a \emph{sensu stricto} mode. A, B, C and D represent
-taxonomy tree nodes with assigned reads. Right panel displays the
-\emph{in line} mode of assignment which is an exception for the
-\emph{sensu stricto} mode of application of LCA algorithm. The \emph{in
-line} mode is used when all the nodes are located in a line without
-bifurcations. In that case the taxon assigned is the most specific (the
-most distant from the root).}
+assignment.} \textbf{TODO UPDATE THIS FIGURE!} The Left panel displays
+an example of the application of LCA algorithm in a \emph{sensu stricto}
+mode. A, B, C and D represent taxonomy tree nodes with assigned reads.
+Right panel displays the \emph{in line} mode of assignment which is an
+exception for the \emph{sensu stricto} mode of application of LCA
+algorithm. The \emph{in line} mode is used when all the nodes are
+located in a line without bifurcations. In that case the taxon assigned
+is the most specific (the most distant from the root).}
 \end{figure}
 
-\paragraph{Lowest Common Ancestor based Taxonomic
-Assignment}\label{lowest-common-ancestor-based-taxonomic-assignment}
+All the assignment procedures work by default with input the set of
+BLAST hits which have maximum bitscore among those covering the whole
+query sequence; we will call this the set of valid hits of a query. Note
+that the criteria which determines this set of valid hits is fully
+configurable, and should be adjusted to conform to its meaning in each
+particular dataset: those hits which we should take into account as
+valid for assignment. This is of course experiment-dependent, and the
+criteria for valid hits should account for read quality or
+technology-specific error profiles.
 
-For each read, first, we select the BEST BLAST HITs (by default 10 Hits)
-over a threshold of similarity. To evaluate similarity for this first
-filtering of hits we use the Expect value (by default
-\(evalue \leq e^{-15}\)) that describes the number of hits one can
-``expect'' to see by chance when searching a database of a particular
-size. In a second filtering step we filtering those hits that are not
-sufficiently good comparing them with the best one. We select the best
-HSP (High Similarity Pair) per reference sequence and then choose the
-best HSP (that with lowest E-value) between all the selected ones. The
-bitscore of this best HSP (called S) is used as reference to filter the
-rest of HSPs. All the HSPs with bitscore below the product \(p S\) are
-filtered. p is a coefficient fixed by the user to define the bitscore
-required, e.g.~if \(p=0.9\) and \(S=700\) the required bitscore
-threshold would be \(630\). Once we have the definitive HSPs selected,
-we obtain their corresponding taxonomic nodes using the taxonomic
-assignments that NCBI provides for all the nt database sequences. Now we
-have to analyze the topological distribution of these nodes in the
-taxonomy tree: i. If all the nodes forms a line in the taxonomy tree
-(are located in a not branched lineage to the tree root) we should
-choose the most specific taxID as the final assignment for that read. We
-call to this kind of assignment the `in line' exception (see Figure 2
-right panel). ii. If not, we should search for the \emph{sensu stricto}
-Lowest Common Ancestor (LCA) of all the selected taxonomic nodes (See
-Figure 2 left panel). In this approach we decided to use the bitscore
-for evaluating the similarity because it is a value that increases when
-similarity is higher and depends a lot on the length of the HSP. Some
-reads could not find sequences with enough similarity in the database
-and then they would be classified as reads with no hits. Advanced
-metagenomics analysis approaches \citep{huson2012microbial} have adopted
-LCA-based assignment algorithms because it provides fine and trusted
-taxonomical assignment.
+\paragraph{LCA Assignment}\label{lca-assignment}
+
+The LCA assignment simply computes the lowest common ancestor of the set
+of valid hits. Other metagenomics analysis approaches
+\citep{huson2012microbial} have adopted LCA-based assignment algorithms,
+as it provides the most precise assignment for a set of taxa, with no
+further information available\footnote{the lowest common ancestor of a
+  set of nodes in a tree is just their product in the tree viewed as a
+  poset, with the ``parent of'' partial order}; as our reference
+database is already heavily preprocessed and checked for internal
+consistency, there is no need for altering the lowest common ancestor in
+any way, as is customarily done in other cases. MG7 operates under the
+assumption that hits which fulfill the user-defined criteria for being
+valid hits should be treated on an equal footing, and in that case
+taking their LCA is the coherent assignment choice with minimum
+information loss.
 

### 2016-08-18 docs/mg7-paper/out/paper.tex

@@
 database. The taxonomic assignment for each read is based on BLASTN of
-each read against the 16S database.Assignment based on direct similarity
-of each read one by one compared against a sufficiently wide database is
-considered in different reviews of metagenomics analysis methodologies
-\citep{segata2013computational} \citep{morgan2012chapter} as a very
-exhaustive method for assignment. Some methods of assignment compare the
-sequences only against the 16S genes from available complete bacterial
-genomes or avoid computational cost clustering or binning the sequences
-first, and then doing the assignments only for the representative
-sequence of each cluster. MG7 carries out an exhaustive comparison of
-all the reads under analysis and it does not applies any binning
-strategy. Every read is specifically compared with all the sequences of
-the 16S database.
+each read against the 16S database. Assignment based on direct
+similarity of each read one by one compared against a sufficiently wide
+database is considered in different reviews of metagenomics analysis
+methodologies \citep{segata2013computational} \citep{morgan2012chapter}
+as a very exhaustive method for assignment. Some methods of assignment
+compare the sequences only against the 16S genes from available complete
+bacterial genomes or avoid computational cost clustering or binning the
+sequences first, and then doing the assignments only for the
+representative sequence of each cluster. MG7 carries out an exhaustive
+comparison of all the reads under analysis and it does not apply any
+binning strategy. Every read is specifically compared with all the
+sequences of the 16S database.
 
@@
 
-\paragraph{LCA Assignment}\label{lca-assignment}
+\paragraph{LCA assignment}\label{lca-assignment}
 
@@
 
-\paragraph{Best BLAST hit taxonomic
-assignment}\label{best-blast-hit-taxonomic-assignment}
+\paragraph{Best BLAST hit assignment}\label{best-blast-hit-assignment}
 
